---
title: "RDA analysis"
author: "Ariana S Huffmyer, Serena Hackerott, E5 RoL Team"
date: "07/18/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("vegan")) install.packages("vegan")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(tidyverse)
library(vegan)
library(cowplot)
```


# Load dataframes

```{r}
#Load in ITS2 dataframe generated from its2_analysis.Rmd.  
its2_rel_abund<-read.csv("Output/ITS2_rel_abund_matrix.csv")

#Load in master dataframe generated from 1_assemble_data.Rmd.  
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

#reorder site levels 
master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

Add in haplotype information
```{r}
haplotypes<-read_csv("Species_ID/master_colony_metadata_speciesID.csv")

master$haplotype<-haplotypes$Genus.Species[match(master$colony_id_corr, haplotypes$colony_id_corr)]
```

# Holobiont responses   
```{r}
data_all<-master%>%
  select(colony_id_corr, haplotype, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)
```

#Log+1 transform all responses
```{r}
data_all_log <-data_all
data_all_log[,-c(1:5)]<-log(data_all_log[,-c(1:5)]+1)

#change timepoint to a factor 
data_all_log$timepoint<-as.factor(data_all_log$timepoint)

#change species to a factor 
data_all_log$species<-as.factor(data_all_log$species)

#change haplotype to a factor 
data_all_log$haplotype<-as.factor(data_all_log$haplotype)
```


#Merge physiology and symbiont composition and subset by species
```{r}
data_community<-merge(data_all_log, its2_rel_abund)

data_community<-data_community[complete.cases(data_community), ]

data_community$tp<-"TP1"
data_community$tp[which(data_community$timepoint=="timepoint2")]<-"TP2"
data_community$tp[which(data_community$timepoint=="timepoint3")]<-"TP3"
data_community$tp[which(data_community$timepoint=="timepoint4")]<-"TP4"

data_community$sample_id<-paste(data_community$colony_id_corr, data_community$tp, sep="_")
rownames(data_community)<-data_community$sample_id

acr_data_community<-data_community%>%
  filter(species=="Acropora")

por_data_community<-data_community%>%
  filter(species=="Porites")

poc_data_community<-data_community%>%
  filter(species=="Pocillopora")
```


#Acropora RDA- Host metrics 
```{r}
names(acr_data_community)
acr.comm.rda.host<-dbrda(vegdist(acr_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, 
                       data=acr_data_community, dist="bray")

##Check model significance
anova(acr.comm.rda.host)
# Model: dbrda(formula = vegdist(acr_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, data = acr_data_community, distance = "bray")
#          Df SumOfSqs      F Pr(>F)
# Model     5   0.3807 0.6385  0.751
# Residual 94  11.2112  

plot(acr.comm.rda.host)

anova(acr.comm.rda.host, by="terms")

```

#Acropora RDA- Symbiont metrics 

```{r}
names(acr_data_community)
acr.comm.rda.sym<-dbrda(vegdist(acr_data_community[, c(24:63)], "bray") ~ Symbiont_Density + Symbiont_Biomass + Total_Chl + Total_Chl_cell + Am + AQY + Ik + Ic + S_H_Biomass_Ratio, data=acr_data_community, dist="bray")


##Check model significance
anova(acr.comm.rda.sym)
# Model: dbrda(formula = vegdist(acr_data_community[, c(24:63)], "bray") ~ Symbiont_Density + Symbiont_Biomass + Total_Chl + Total_Chl_cell + Am + AQY + Ik + Ic + S_H_Biomass_Ratio, data = acr_data_community, distance = "bray")
#          Df SumOfSqs      F Pr(>F)
# Model     9   0.9007 0.8424  0.624
# Residual 90  10.6913  

plot(acr.comm.rda.sym)

anova(acr.comm.rda.sym, by="terms")

```

#Porites RDA- Host metrics 
```{r}
names(por_data_community)
por.comm.rda.host<-dbrda(vegdist(por_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, 
                       data=por_data_community, dist="bray")

##Check model significance
anova(por.comm.rda.host)
# Model: dbrda(formula = vegdist(por_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, data = por_data_community, distance = "bray")
#           Df SumOfSqs      F Pr(>F)    
# Model      5    6.684 4.2585  0.001 ***
# Residual 129   40.493 

plot(por.comm.rda.host)

##Check variance explained by model
(summary(por.comm.rda.host)$constr.chi/summary(por.comm.rda.host)$tot.chi)*100
#14.17% of the variance in Symbiont community composition is constrained by Host physiological metrics 

RsquareAdj(por.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 10.84% of the variation in Symbiont community composition

##Check variance explained by each axis
anova(por.comm.rda.host, by="axis")
# Model: dbrda(formula = vegdist(por_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, data = por_data_community, distance = "bray")
#           Df SumOfSqs       F Pr(>F)    
# dbRDA1     1    6.034 19.2228  0.001 ***
# dbRDA2     1    0.367  1.1705  0.938    
# dbRDA3     1    0.157  0.4988  0.998    
# dbRDA4     1    0.095  0.3022  1.000    
# dbRDA5     1    0.031  0.0980           
# Residual 129   40.493     

summary(por.comm.rda.host)$cont
#dbRDA1 explains 12.79% of total variance 
#dbRDA2 explains 0.78% of total variance 

summary(por.comm.rda.host)$concont
#dbRDA1 explains 90.28% of constrained variance 
#dbRDA2 explains 5.50% of constrained variance 

##Check variance explained by each Host physiological metric
anova(por.comm.rda.host, by="terms")
#                 Df SumOfSqs       F Pr(>F)    
# Antiox_Capacity   1    4.067 12.9555  0.001 ***
# Host_Biomass      1    1.304  4.1542  0.002 ** 
# Host_Protein      1    0.636  2.0249  0.077 .  
# Rd                1    0.297  0.9446  0.428    
# Calc              1    0.381  1.2131  0.269    
# Residual        129   40.493  

```

#Porites RDA- Symbiont metrics 
```{r}
names(por_data_community)
por.comm.rda.sym<-dbrda(vegdist(por_data_community[, c(24:63)], "bray") ~ Symbiont_Density + Symbiont_Biomass + Total_Chl + Total_Chl_cell + Am + AQY + Ik + Ic + S_H_Biomass_Ratio, data=por_data_community, dist="bray")

##Check model significance
anova(por.comm.rda.sym)
# Model: dbrda(formula = vegdist(por_data_community[, c(24:63)], "bray") ~ Symbiont_Density + Symbiont_Biomass + Total_Chl + Total_Chl_cell + Am + AQY + Ik + Ic + S_H_Biomass_Ratio, data = por_data_community, distance = "bray")
#           Df SumOfSqs      F Pr(>F)    
# Model      9   10.655 4.0519  0.001 ***
# Residual 125   36.522  

plot(por.comm.rda.sym)

##Check variance explained by model
(summary(por.comm.rda.sym)$constr.chi/summary(por.comm.rda.sym)$tot.chi)*100
#22.58% of the variance in Symbiont community composition is constrained by symbiont physiological metrics 

RsquareAdj(por.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 17.01% of the variation in Symbiont community composition

##Check variance explained by each axis
anova(por.comm.rda.sym, by="axis")
#  Df SumOfSqs       F Pr(>F)    
# dbRDA1     1    8.538 29.2228  0.001 ***
# dbRDA2     1    0.852  2.9160  0.731    
# dbRDA3     1    0.590  2.0202  0.910    
# dbRDA4     1    0.296  1.0121  1.000    
# dbRDA5     1    0.202  0.6926           
# dbRDA6     1    0.072  0.2464           
# dbRDA7     1    0.058  0.1979           
# dbRDA8     1    0.026  0.0884           
# dbRDA9     1    0.021  0.0709           
# Residual 125   36.522   

summary(por.comm.rda.sym)$cont
#dbRDA1 explains 18.1% of total variance 
#dbRDA2 explains 1.81% of total variance 

summary(por.comm.rda.sym)$concont
#dbRDA1 explains 80.13% of constrained variance 
#dbRDA2 explains 7.99% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(por.comm.rda.sym, by="terms")
#                    Df SumOfSqs       F Pr(>F)    
# Symbiont_Density    1    2.888  9.8831  0.002 ** 
# Symbiont_Biomass    1    3.651 12.4975  0.001 ***
# Total_Chl           1    0.288  0.9862  0.423    
# Total_Chl_cell      1    0.565  1.9332  0.101    
# Am                  1    0.772  2.6419  0.038 *  
# AQY                 1    1.520  5.2040  0.002 ** 
# Ik                  1    0.299  1.0228  0.354    
# Ic                  1    0.402  1.3759  0.214    
# S_H_Biomass_Ratio   1    0.270  0.9228  0.438    
# Residual          125   36.522   

```

#Pocillopora RDA- Host metrics 
```{r}
names(poc_data_community)
poc.comm.rda.host<-dbrda(vegdist(poc_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, 
                       data=poc_data_community, dist="bray")

##Check model significance
anova(poc.comm.rda.host)
# Model: dbrda(formula = vegdist(poc_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, data = poc_data_community, distance = "bray")
#           Df SumOfSqs      F Pr(>F)
# Model      5    1.291 0.6516  0.886
# Residual 120   47.551     

plot(poc.comm.rda.host)

##Check variance explained by model
(summary(poc.comm.rda.host)$constr.chi/summary(poc.comm.rda.host)$tot.chi)*100
#2.64% of the variance in Symbiont community composition is constrained by symbiont physiological metrics 

RsquareAdj(poc.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains <0% of the variation in Symbiont community composition

##Check variance explained by each axis
anova(poc.comm.rda.host, by="axis")
# Model: dbrda(formula = vegdist(poc_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc + haplotype, data = poc_data_community, distance = "bray")
# Model: dbrda(formula = vegdist(poc_data_community[, c(24:63)], "bray") ~ Antiox_Capacity + Host_Biomass + Host_Protein + Rd + Calc, data = poc_data_community, distance = "bray")
#           Df SumOfSqs      F Pr(>F)
# dbRDA1     1    0.824 2.0784  0.755
# dbRDA2     1    0.251 0.6324  1.000
# dbRDA3     1    0.175 0.4417       
# dbRDA4     1    0.031 0.0771       
# dbRDA5     1    0.011 0.0286       
# Residual 120   47.551 

summary(poc.comm.rda.host)$cont
#dbRDA1 explains 1.67% of total variance 
#dbRDA2 explains 0.51% of total variance 

summary(poc.comm.rda.host)$concont
#dbRDA1 explains 63.79% of constrained variance 
#dbRDA2 explains 19.41% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(poc.comm.rda.host, by="terms")
#                  Df SumOfSqs      F Pr(>F)  
# Antiox_Capacity   1    0.050 0.1272  0.992  
# Host_Biomass      1    0.168 0.4239  0.791  
# Host_Protein      1    0.147 0.3701  0.826  
# Rd                1    0.742 1.8714  0.096 .
# Calc              1    0.184 0.4655  0.826  
# Residual        120   47.551  

```

#Pocillopora RDA- Symbiont metrics 
```{r}
names(poc_data_community)
poc.comm.rda.sym<-dbrda(vegdist(poc_data_community[, c(24:63)], "bray") ~ Symbiont_Density + Symbiont_Biomass + Total_Chl + Total_Chl_cell + Am + AQY + Ik + Ic + S_H_Biomass_Ratio, data=poc_data_community, dist="bray")

##Check model significance
anova(poc.comm.rda.sym)
#           Df SumOfSqs      F Pr(>F)    
# Model      9    6.882 2.1141  0.001 ***
# Residual 116   41.959  

plot(poc.comm.rda.sym)

##Check variance explained by model
(summary(poc.comm.rda.sym)$constr.chi/summary(poc.comm.rda.sym)$tot.chi)*100
#14.09% of the variance in Symbiont community composition is constrained by Host physiological metrics 

RsquareAdj(poc.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 7.43% of the variation in Symbiont community composition

##Check variance explained by each axis
anova(poc.comm.rda.sym, by="axis")
#           Df SumOfSqs       F Pr(>F)    
# dbRDA1     1    4.173 11.5355  0.001 ***
# dbRDA2     1    1.444  3.9919  0.486    
# dbRDA3     1    0.962  2.6584  0.794    
# dbRDA4     1    0.183  0.5065  1.000    
# dbRDA5     1    0.094  0.2603           
# dbRDA6     1    0.016  0.0438           
# dbRDA7     1    0.009  0.0240           
# dbRDA8     1    0.002  0.0052           
# dbRDA9     1    0.000  0.0013           
# Residual 116   41.959  

summary(poc.comm.rda.sym)$cont
#dbRDA1 explains 8.54% of total variance 
#dbRDA2 explains 2.96% of total variance 

summary(poc.comm.rda.sym)$concont
#dbRDA1 explains 60.63% of constrained variance 
#dbRDA2 explains 20.98% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(poc.comm.rda.sym, by="terms")

#                    Df SumOfSqs      F Pr(>F)    
# Symbiont_Density    1    0.305 0.8425  0.514    
# Symbiont_Biomass    1    1.171 3.2363  0.014 *  
# Total_Chl           1    1.789 4.9445  0.001 ***
# Total_Chl_cell      1    0.841 2.3251  0.046 *  
# Am                  1    0.096 0.2646  0.921    
# AQY                 1    0.177 0.4887  0.757    
# Ik                  1    0.488 1.3483  0.254    
# Ic                  1    1.015 2.8063  0.024 *  
# S_H_Biomass_Ratio   1    1.002 2.7705  0.033 *  
# Residual          116   41.959     
```
