---
title: "RDA analysis of environmental data"
author: "Ariana S Huffmyer, Serena Hackerott, E5 RoL Team"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("vegan")) install.packages("vegan")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(tidyverse)
library(vegan)
library(cowplot)
```


# Load dataframes

```{r}
#Load in dataframe of temperature information   
enviro<-read.csv("Output/environment_characteristics_RDA.csv")%>%rename(solar_mean=mean_solar_rad_kwpm2_mean, solar_sd=mean_solar_rad_kwpm2_sd, rain_mean=cumulative_rainfall_mm_mean, rain_sd=cumulative_rainfall_mm_sd, temp_mean=mean_Temp_mean, temp_sd=mean_Temp_sd)
  
#Load in master dataframe generated from 1_assemble_data.Rmd.  
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

#reorder site levels 
master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

Add in haplotype information
```{r}
haplotypes<-read_csv("Species_ID/master_colony_metadata_speciesID.csv")

master$haplotype<-haplotypes$Genus.Species[match(master$colony_id_corr, haplotypes$colony_id_corr)]
```

# Prepare data frame   
```{r}
data_all<-master%>%
  select(colony_id_corr, haplotype, timepoint, species, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)
```

# Log+1 transform all responses
```{r}
data_all_log <-data_all
data_all_log[,-c(1:4)]<-log(data_all_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_all_log$timepoint<-as.factor(data_all_log$timepoint)

#change species to a factor 
data_all_log$species<-as.factor(data_all_log$species)

#change haplotype to a factor 
data_all_log$haplotype<-as.factor(data_all_log$haplotype)
```

# Merge datasets 
```{r}
data_temp<-merge(data_all_log, enviro)

data_temp<-data_temp[complete.cases(data_temp), ]

data_temp$tp<-"TP1"
data_temp$tp[which(data_temp$timepoint=="timepoint2")]<-"TP2"
data_temp$tp[which(data_temp$timepoint=="timepoint3")]<-"TP3"
data_temp$tp[which(data_temp$timepoint=="timepoint4")]<-"TP4"

data_temp$sample_id<-paste(data_temp$colony_id_corr, data_temp$tp, sep="_")
rownames(data_temp)<-data_temp$sample_id

acr_data_temp<-data_temp%>%
  filter(species=="Acropora")

por_data_temp<-data_temp%>%
  filter(species=="Porites")

poc_data_temp<-data_temp%>%
  filter(species=="Pocillopora")
```

Test for autocorrelation in environmental characteristics. 
```{r}
library("corrplot")
library("Hmisc")

correlation<-enviro%>%
  select(solar_mean, solar_sd, rain_mean, rain_sd, temp_mean, temp_sd)

M<-cor(correlation)

corrplot(M, method="number", diag=FALSE)

cor(M, method = "pearson")

cor2<-rcorr(M)
cor2$r
cor2$P
```

Rain and solar are correlated, temperature is not correlated with rain or solar. Keep temperature and solar. 

# Acropora RDA- Host metrics 
```{r}
names(acr_data_temp)
acr.comm.rda.host<-rda(acr_data_temp[, c(5,8,11,13,18)] ~ solar_mean + temp_mean, 
                       data=acr_data_temp, scale=TRUE)

##Check model significance
anova(acr.comm.rda.host)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = acr_data_temp[, c(5, 8, 11, 13, 18)] ~ solar_mean + temp_mean, data = acr_data_temp, scale = TRUE)
#          Df Variance      F Pr(>F)    
# Model     2   0.9276 9.2249  0.001 ***
# Residual 81   4.0724       

plot(acr.comm.rda.host)

##Check variance explained by model
(summary(acr.comm.rda.host)$constr.chi/summary(acr.comm.rda.host)$tot.chi)*100
#18.55% of the variance in host physiology is constrained by metrics 

RsquareAdj(acr.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 16.54% of the variation in host physiology 

summary(acr.comm.rda.host)$cont
#RDA1 explains 15.16% of total variance 
#RDA2 explains 3.39% of total variance 

summary(acr.comm.rda.host)$concont
#RDA1 explains 81.73% of constrained variance 
#RDA2 explains 18.27% of constrained variance 

anova(acr.comm.rda.host, by="terms")

# Permutation test for rda under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = acr_data_temp[, c(5, 8, 11, 13, 18)] ~ solar_mean + temp_mean, data = acr_data_temp, scale = TRUE)
#            Df Variance       F Pr(>F)    
# solar_mean  1   0.6557 13.0412  0.001 ***
# temp_mean   1   0.2719  5.4087  0.002 ** 
# Residual   81   4.0724  

##Run varpart function with responses and explanatory variables of interest
varpart_acr.comm.rda.host<-varpart(acr_data_temp[, c(5,8,11,13,18)], acr_data_temp$temp_mean, acr_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_acr.comm.rda.host$part
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.03190       0.02009     TRUE
# [b+c] = X2            1   0.08411       0.07294     TRUE
# [a+b+c] = X1+X2       2   0.11719       0.09539     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.02245     TRUE
# [b] = X2|X1           1                 0.07529     TRUE
# [c]                   0                -0.00235    FALSE
# [d] = Residuals                         0.90461    FALSE

acr_host_temp<-2.25
acr_host_light<-7.53
```

# Acropora RDA- Symbiont metrics 

```{r}
names(acr_data_temp)

acr.comm.rda.sym<-rda(acr_data_temp[, c(6,7,9,11,12,14,15,17,18)] ~ temp_mean + solar_mean, 
                       data=acr_data_temp)

##Check model significance
anova(acr.comm.rda.sym)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = acr_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean, data = acr_data_temp)
#          Df Variance     F Pr(>F)    
# Model     2  0.40747 35.69  0.001 ***
# Residual 81  0.46239                 

plot(acr.comm.rda.sym)

##Check variance explained by model
(summary(acr.comm.rda.sym)$constr.chi/summary(acr.comm.rda.sym)$tot.chi)*100
#46.84% of the variance in symbiont physiology is constrained by metrics 

RsquareAdj(acr.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 45.53% of the variation in symbiont physiology 

summary(acr.comm.rda.sym)$cont
#RDA1 explains 44.34% of total variance 
#RDA2 explains 2.50% of total variance 

summary(acr.comm.rda.sym)$concont
#RDA1 explains 94.66% of constrained variance 
#RDA2 explains 0.53% of constrained variance 

anova(acr.comm.rda.sym, by="terms")

# Permutation test for rda under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = acr_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean, data = acr_data_temp)
#            Df Variance      F Pr(>F)    
# temp_mean   1  0.10705 18.753  0.001 ***
# solar_mean  1  0.30042 52.627  0.001 ***
# Residual   81  0.46239 

##Run varpart function with responses and explanatory variables of interest
varpart_acr.comm.rda.sym<-varpart(acr_data_temp[, c(6,7,9,11,12,14,15,17,18)], acr_data_temp$temp_mean, acr_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_acr.comm.rda.sym$part

# No. of explanatory tables: 2 
# Total variation (SS): 72.198 
#             Variance: 0.86986 
# No. of observations: 84 
# 
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.12307       0.11237     TRUE
# [b+c] = X2            1   0.42112       0.41406     TRUE
# [a+b+c] = X1+X2       2   0.46843       0.45531     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.04124     TRUE
# [b] = X2|X1           1                 0.34294     TRUE
# [c]                   0                 0.07113    FALSE
# [d] = Residuals                         0.54469    FALSE

acr_sym_temp<-4.12
acr_sym_light<-34.29
```

# Porites RDA- Host metrics 

```{r}
names(por_data_temp)
por.comm.rda.host<-rda(por_data_temp[, c(5,8,10,13,16)] ~ temp_mean + solar_mean, 
                       data=por_data_temp)

##Check model significance
anova(por.comm.rda.host)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = por_data_temp[, c(5, 8, 10, 13, 16)] ~ temp_mean + solar_mean, data = por_data_temp)
#          Df Variance      F Pr(>F)    
# Model     2 0.056414 17.378  0.001 ***
# Residual 99 0.160687

plot(por.comm.rda.host)

##Check variance explained by model
(summary(por.comm.rda.host)$constr.chi/summary(por.comm.rda.host)$tot.chi)*100
#25.99% of the variance in host physiology is constrained by metrics 

RsquareAdj(por.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 24.49% of the variation in host physiology 

summary(por.comm.rda.host)$cont
#RDA1 explains 22.86% of total variance 
#RDA2 explains 3.12% of total variance 

summary(por.comm.rda.host)$concont
#RDA1 explains 87.97% of constrained variance 
#RDA2 explains 12.03% of constrained variance 

##Check variance explained by each metric
anova(por.comm.rda.host, by="terms")
# Permutation test for rda under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = por_data_temp[, c(5, 8, 10, 13, 16)] ~ temp_mean + solar_mean, data = por_data_temp)
#            Df Variance       F Pr(>F)    
# temp_mean   1 0.049629 30.5768  0.001 ***
# solar_mean  1 0.006785  4.1803  0.007 ** 
# Residual   99 0.160687 

##Run varpart function with responses and explanatory variables of interest
varpart_por.comm.rda.host<-varpart(por_data_temp[, c(5,8,10,13,16)], por_data_temp$temp_mean, por_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_por.comm.rda.host$part

# No. of explanatory tables: 2 
# Total variation (SS): 21.927 
#             Variance: 0.2171 
# No. of observations: 102 
# 
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.22860       0.22089     TRUE
# [b+c] = X2            1   0.15348       0.14501     TRUE
# [a+b+c] = X1+X2       2   0.25985       0.24490     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.09989     TRUE
# [b] = X2|X1           1                 0.02401     TRUE
# [c]                   0                 0.12100    FALSE
# [d] = Residuals                         0.75510    FALSE

por_host_temp<-9.99
por_host_light<-2.40
```

# Porites RDA- Symbiont metrics 
```{r}
names(por_data_temp)
por.comm.rda.sym<-rda(por_data_temp[, c(6,7,9,11,12,14,15,17,18)] ~ temp_mean + solar_mean + Condition(haplotype), 
                       data=por_data_temp)

##Check model significance
anova(por.comm.rda.sym)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = por_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean + Condition(haplotype), data = por_data_temp)
#          Df Variance      F Pr(>F)    
# Model     2  0.17045 14.922  0.001 ***
# Residual 98  0.55969    

plot(por.comm.rda.sym)

##Check variance explained by model
(summary(por.comm.rda.sym)$constr.chi/summary(por.comm.rda.sym)$tot.chi)*100
#19.85% of the variance in symbiont physiology is constrained by metrics

RsquareAdj(por.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 18.71% of the variation in Symbiont physiology

summary(por.comm.rda.sym)$cont
#RDA1 explains 18.43% of total variance 
#RDA2 explains 4.92% of total variance 

summary(por.comm.rda.sym)$concont
#RDA1 explains 78.94% of constrained variance 
#RDA2 explains 21.96% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(por.comm.rda.sym, by="terms")
# Permutation test for rda under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = por_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean + Condition(haplotype), data = por_data_temp)
#            Df Variance       F Pr(>F)    
# temp_mean   1  0.05239  9.1728  0.001 ***
# solar_mean  1  0.11806 20.6721  0.001 ***
# Residual   98  0.55969    

##Run varpart function with responses and explanatory variables of interest
varpart_por.comm.rda.sym<-varpart(por_data_temp[, c(6,7,9,11,12,14,15,17,18)], por_data_temp$temp_mean, por_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_por.comm.rda.sym$part
# No. of explanatory tables: 2 
# Total variation (SS): 86.707 
#             Variance: 0.85849 
# No. of observations: 102 
# 
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.06619       0.05685     TRUE
# [b+c] = X2            1   0.14112       0.13253     TRUE
# [a+b+c] = X1+X2       2   0.20421       0.18813     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.05560     TRUE
# [b] = X2|X1           1                 0.13128     TRUE
# [c]                   0                 0.00125    FALSE
# [d] = Residuals                         0.81187    FALSE

por_sym_temp<-5.56
por_sym_light<-13.13
```

# Pocillopora RDA- Host metrics 
```{r}
names(poc_data_temp)
poc.comm.rda.host<-rda(poc_data_temp[, c(5,8,10,13,16)] ~ temp_mean + solar_mean + Condition(haplotype), 
                       data=poc_data_temp)

##Check model significance
anova(poc.comm.rda.host)
# Permutation test for rda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: rda(formula = poc_data_temp[, c(5, 8, 10, 13, 16)] ~ temp_mean + solar_mean + Condition(haplotype), data = poc_data_temp)
#           Df Variance      F Pr(>F)    
# Model      2 0.014262 14.136  0.001 ***
# Residual 101 0.050950   

plot(poc.comm.rda.host)

##Check variance explained by model
(summary(poc.comm.rda.host)$constr.chi/summary(poc.comm.rda.host)$tot.chi)*100
#21.40% of the variance in host physiology is constrained by metrics 

RsquareAdj(poc.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 20.08% of the variation in host physiology 

summary(poc.comm.rda.host)$cont
#RDA1 explains 19.77% of total variance 
#RDA2 explains 2.10% of total variance 

summary(poc.comm.rda.host)$concont
#RDA1 explains 90.40% of constrained variance 
#RDA2 explains 9.59% of constrained variance 

##Check variance explained by each Host physiological metric
anova(poc.comm.rda.host, by="terms")
# Model: rda(formula = poc_data_temp[, c(5, 8, 10, 13, 16)] ~ temp_mean + solar_mean + Condition(haplotype), data = poc_data_temp)
#             Df Variance      F Pr(>F)    
# temp_mean    1 0.009585 19.000  0.001 ***
# solar_mean   1 0.004677  9.272  0.001 ***
# Residual   101 0.050950 

##Run varpart function with responses and explanatory variables of interest
varpart_poc.comm.rda.host<-varpart(poc_data_temp[, c(5,8,10,13,16)], poc_data_temp$temp_mean, poc_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_poc.comm.rda.host$part
# No. of explanatory tables: 2 
# Total variation (SS): 6.9317 
#             Variance: 0.066651 
# No. of observations: 105 
# 
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.15039       0.14214     TRUE
# [b+c] = X2            1   0.20283       0.19509     TRUE
# [a+b+c] = X1+X2       2   0.22574       0.21056     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.01547     TRUE
# [b] = X2|X1           1                 0.06842     TRUE
# [c]                   0                 0.12668    FALSE
# [d] = Residuals                         0.78944    FALSE

poc_host_temp<-1.55
poc_host_light<-6.84
```

# Pocillopora RDA- Symbiont metrics 

```{r}
names(poc_data_temp)
poc.comm.rda.sym<-rda(poc_data_temp[, c(6,7,9,11,12,14,15,17,18)] ~ temp_mean + solar_mean + Condition(haplotype), 
                       data=poc_data_temp)

##Check model significance
anova(poc.comm.rda.sym)
# Model: rda(formula = poc_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean + Condition(haplotype), data = poc_data_temp)
#           Df Variance      F Pr(>F)    
# Model      2  0.24775 25.311  0.001 ***
# Residual 101  0.49430        

plot(poc.comm.rda.sym)

##Check variance explained by model
(summary(poc.comm.rda.sym)$constr.chi/summary(poc.comm.rda.sym)$tot.chi)*100
#32.61% of the variance in symbiont physiology is constrained by metrics

RsquareAdj(poc.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 31.63% of the variation in Symbiont physiology

summary(poc.comm.rda.sym)$cont
#RDA1 explains 26.53% of total variance 
#RDA2 explains 6.85% of total variance 

summary(poc.comm.rda.sym)$concont
#RDA1 explains 79.47% of constrained variance 
#RDA2 explains 20.53% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(poc.comm.rda.sym, by="terms")
# Model: rda(formula = poc_data_temp[, c(6, 7, 9, 11, 12, 14, 15, 17, 18)] ~ temp_mean + solar_mean + Condition(haplotype), data = poc_data_temp)
#             Df Variance      F Pr(>F)    
# temp_mean    1  0.05891 12.038  0.001 ***
# solar_mean   1  0.18883 38.584  0.001 ***
# Residual   101  0.49430    

##Run varpart function with responses and explanatory variables of interest
varpart_poc.comm.rda.sym<-varpart(poc_data_temp[, c(6,7,9,11,12,14,15,17,18)], poc_data_temp$temp_mean, poc_data_temp$solar_mean)

##Check variance explained by individual fractions
varpart_poc.comm.rda.sym$part

# No. of explanatory tables: 2 
# Total variation (SS): 79.006 
#             Variance: 0.75967 
# No. of observations: 105 
# 
# Partition table:
#                      Df R.squared Adj.R.squared Testable
# [a+c] = X1            1   0.08078       0.07185     TRUE
# [b+c] = X2            1   0.19759       0.18980     TRUE
# [a+b+c] = X1+X2       2   0.33607       0.32305     TRUE
# Individual fractions                                    
# [a] = X1|X2           1                 0.13326     TRUE
# [b] = X2|X1           1                 0.25120     TRUE
# [c]                   0                -0.06140    FALSE
# [d] = Residuals                         0.67695    FALSE

poc_sym_temp<-13.33
poc_sym_light<-25.12
```

# Plot ordination plots 

```{r}
ordiplot(acr.comm.rda.host, scaling = 2, main = "ACR Host RDA")
```

```{r}
ordiplot(acr.comm.rda.sym, scaling = 2, main = "ACR Symbiont RDA")
```

```{r}
ordiplot(por.comm.rda.host, scaling = 2, main = "POR Host RDA")
```

```{r}
ordiplot(por.comm.rda.sym, scaling = 2, main = "POR Symbiont RDA")
```

```{r}
ordiplot(poc.comm.rda.host, scaling = 2, main = "POC Host RDA")
```

```{r}
ordiplot(poc.comm.rda.sym, scaling = 2, main = "POC Symbiont RDA")
```

# Plot variance explained by temp and light 

Assemble data from above analyses.  
```{r}
var_data <- data.frame(
  genus = c("Acropora", "Acropora", "Acropora", "Acropora", "Pocillopora", "Pocillopora", "Pocillopora", "Pocillopora", "Porites", "Porites", "Porites", "Porites"),
  fraction = c("Host", "Host", "Symbiont", "Symbiont", "Host", "Host", "Symbiont", "Symbiont", "Host", "Host", "Symbiont", "Symbiont"),
  metric   = c("Temperature", "Light", "Temperature", "Light", "Temperature", "Light", "Temperature", "Light", "Temperature", "Light", "Temperature", "Light"),
  value   = c(acr_host_temp, acr_host_light, acr_sym_temp, acr_sym_light, poc_host_temp, poc_host_light, poc_sym_temp, poc_sym_light, por_host_temp, por_host_light, por_sym_temp, por_sym_light)
)

var_data
```

Plot 
```{r}
var_plot<-var_data%>%
  ggplot(aes(x=fraction, y=value, shape=metric))+
  facet_wrap(~genus)+
  scale_shape_manual(name="Metric", values=c(21,19))+
  geom_point(size=4)+
  ylab("Variance Explained (%)")+
  xlab("")+
  theme_classic()+
  theme(
    panel.border = element_rect(fill=NA, colour="black"), 
    text = element_text(size=14)
  );var_plot

ggsave(var_plot, filename="Figures/Multivariate/environment_variance_partioning.png", width=10, height=4)
```



# Plot physiological metric correlations with light and temperature 

Plot correlations between host metrics and light and temperature. 

```{r}
library(Hmisc)

res1.r<-outer(data_temp[, c(19,23)], data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(data_temp[, c(19,23)], data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/allspecies-host-correlation-light-temp.pdf", width=6, height=4)

host_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```

Plot correlations between symbiont metrics and light and temperature. 

```{r}

res1.r<-outer(data_temp[, c(19,23)], data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(data_temp[, c(19,23)], data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/allspecies-symbiont-correlation-light-temp.pdf", width=8, height=4)

sym_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```

## Plot for Acropora 

```{r}
res1.r<-outer(acr_data_temp[, c(19,23)], acr_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(acr_data_temp[, c(19,23)], acr_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/acr-host-correlation-light-temp.pdf", width=6, height=4)

acr_host_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```

Plot correlations between symbiont metrics and light and temperature. 

```{r}

res1.r<-outer(acr_data_temp[, c(19,23)], acr_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(acr_data_temp[, c(19,23)], acr_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/acr-sym-correlation-light-temp.pdf", width=8, height=4)

acr_sym_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```

## Plot for Pocillopora 

```{r}
res1.r<-outer(poc_data_temp[, c(19,23)], poc_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(poc_data_temp[, c(19,23)], poc_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/poc-host-correlation-light-temp.pdf", width=6, height=4)

poc_host_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()

```

Plot correlations between symbiont metrics and light and temperature. 

```{r}

res1.r<-outer(poc_data_temp[, c(19,23)], poc_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(poc_data_temp[, c(19,23)], poc_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/poc-sym-correlation-light-temp.pdf", width=8, height=4)

poc_sym_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```

## Plot for Porites 

```{r}
res1.r<-outer(por_data_temp[, c(19,23)], por_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(por_data_temp[, c(19,23)], por_data_temp[, c(5,8,10,13,16)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/por-host-correlation-light-temp.pdf", width=6, height=4)

por_host_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))
dev.off()

```

Plot correlations between symbiont metrics and light and temperature. 

```{r}

res1.r<-outer(por_data_temp[, c(19,23)], por_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$estimate,
           X, Y)
})

res1.p<-outer(por_data_temp[, c(19,23)], por_data_temp[, c(6,7,9,11,12,14,15,17,18)], function(X, Y){
    mapply(function(...) cor.test(..., na.action = "na.exclude")$p.value,
           X, Y)
})

pdf("Figures/Correlations/por-sym-correlation-light-temp.pdf", width=8, height=4)

por_sym_plot<-corrplot(res1.r, method="color", hclust.method="complete", insig="blank", p.mat = res1.p, sig.level = 0.05, tl.col = "black", tl.srt = 45, addCoef.col ='black', col=colorRampPalette(c("blue","white","red"))(50))

dev.off()
```


