---
title: "RDA analysis of environmental data"
author: "Ariana S Huffmyer, Serena Hackerott, E5 RoL Team"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("vegan")) install.packages("vegan")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(tidyverse)
library(vegan)
library(cowplot)
```


# Load dataframes

```{r}
#Load in dataframe of temperature information   
temp<-read.csv("Output/Clean_Environmental_Temp_Means.csv")

temp$site_num<-temp$site

temp<-temp%>%
  mutate(site=if_else(temp$site_num=="site1", "Manava", 
                   if_else(temp$site_num=="site2", "Mahana", 
                           if_else(temp$site_num=="site3", "Hilton", NA))))%>%
  select(!site_num)%>%
  rename(minimum=FullTemp_min, maximum=FullTemp_max, mean=FullTemp_mean, SD=FullTemp_stdev, range=FullTemp_range)
  
#Load in master dataframe generated from 1_assemble_data.Rmd.  
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

#reorder site levels 
master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

Add in haplotype information
```{r}
haplotypes<-read_csv("Species_ID/master_colony_metadata_speciesID.csv")

master$haplotype<-haplotypes$Genus.Species[match(master$colony_id_corr, haplotypes$colony_id_corr)]
```

# Holobiont responses   
```{r}
data_all<-master%>%
  select(colony_id_corr, haplotype, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)
```

#Log+1 transform all responses
```{r}
data_all_log <-data_all
data_all_log[,-c(1:5)]<-log(data_all_log[,-c(1:5)]+1)

#change timepoint to a factor 
data_all_log$timepoint<-as.factor(data_all_log$timepoint)

#change species to a factor 
data_all_log$species<-as.factor(data_all_log$species)

#change haplotype to a factor 
data_all_log$haplotype<-as.factor(data_all_log$haplotype)
```

#Merge physiology and symbiont composition and subset by species
```{r}
data_temp<-merge(data_all_log, temp)

data_temp<-data_temp[complete.cases(data_temp), ]

data_temp$tp<-"TP1"
data_temp$tp[which(data_temp$timepoint=="timepoint2")]<-"TP2"
data_temp$tp[which(data_temp$timepoint=="timepoint3")]<-"TP3"
data_temp$tp[which(data_temp$timepoint=="timepoint4")]<-"TP4"

data_temp$sample_id<-paste(data_temp$colony_id_corr, data_temp$tp, sep="_")
rownames(data_temp)<-data_temp$sample_id

acr_data_temp<-data_temp%>%
  filter(species=="Acropora")

por_data_temp<-data_temp%>%
  filter(species=="Porites")

poc_data_temp<-data_temp%>%
  filter(species=="Pocillopora")
```


#Acropora RDA- Host metrics 
```{r}
names(acr_data_temp)
acr.comm.rda.host<-rda(acr_data_temp[, c(6,9,11,14,17)] ~ mean + maximum + minimum + SD, 
                       data=acr_data_temp)

##Check model significance
anova(acr.comm.rda.host)
# Permutation test for dbrda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: dbrda(formula = acr_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD, data = acr_data_temp)
#          Df SumOfSqs      F Pr(>F)    
# Model     4   2.3901 8.9301  0.001 ***
# Residual 97   6.4904    

plot(acr.comm.rda.host)

##Check variance explained by model
(summary(acr.comm.rda.host)$constr.chi/summary(acr.comm.rda.host)$tot.chi)*100
#26.91% of the variance in host physiology is constrained by temperature metrics 

RsquareAdj(acr.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 23.90% of the variation in host physiology 

summary(acr.comm.rda.host)$cont
#dbRDA1 explains 20.96% of total variance 
#dbRDA2 explains 4.40% of total variance 

summary(acr.comm.rda.host)$concont
#dbRDA1 explains 77.86% of constrained variance 
#dbRDA2 explains 16.35% of constrained variance 

anova(acr.comm.rda.host, by="terms")

# Model: dbrda(formula = acr_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD + range, data = acr_data_temp)
#          Df SumOfSqs       F Pr(>F)    
# mean      1   1.3737 20.5304  0.001 ***
# maximum   1   0.6318  9.4430  0.001 ***
# minimum   1   0.2352  3.5154  0.014 *  
# SD        1   0.1493  2.2315  0.062 .  
# Residual 97   6.4904 
```

#Acropora RDA- Symbiont metrics 

```{r}
names(acr_data_temp)

acr.comm.rda.sym<-rda(acr_data_temp[, c(7,8,10,12,13,15,16,18,19)] ~ mean + maximum + minimum + SD, 
                       data=acr_data_temp)

##Check model significance
anova(acr.comm.rda.sym)
# Permutation test for dbrda under reduced model
# Permutation: free
# Number of permutations: 999
# 
# Model: dbrda(formula = acr_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD + range, data = acr_data_temp)
#          Df SumOfSqs      F Pr(>F)    
# Model     4   35.341 14.917  0.001 ***
# Residual 97   57.454    

plot(acr.comm.rda.sym)

##Check variance explained by model
(summary(acr.comm.rda.sym)$constr.chi/summary(acr.comm.rda.sym)$tot.chi)*100
#38.08% of the variance in symbiont physiology is constrained by temperature metrics 

RsquareAdj(acr.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 35.53% of the variation in symbiont physiology 

summary(acr.comm.rda.sym)$cont
#dbRDA1 explains 33.23% of total variance 
#dbRDA2 explains 3.96% of total variance 

summary(acr.comm.rda.sym)$concont
#dbRDA1 explains 87.25% of constrained variance 
#dbRDA2 explains 10.40% of constrained variance 

anova(acr.comm.rda.sym, by="terms")

# Model: dbrda(formula = acr_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD, data = acr_data_temp)
#          Df SumOfSqs       F Pr(>F)    
# mean      1   15.111 25.5131  0.001 ***
# maximum   1   12.498 21.0999  0.001 ***
# minimum   1    5.526  9.3304  0.001 ***
# SD        1    2.205  3.7231  0.014 *  
# Residual 97   57.454 

```

#Porites RDA- Host metrics 
```{r}
names(por_data_temp)
por.comm.rda.host<-rda(por_data_temp[, c(6,9,11,14,17)] ~ mean + maximum + minimum + SD, 
                       data=por_data_temp)

##Check model significance
anova(por.comm.rda.host)
# Model: rda(formula = por_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD, data = por_data_temp)
#           Df Variance      F Pr(>F)    
# Model      4 0.062431 13.652  0.001 ***
# Residual 134 0.153202  

plot(por.comm.rda.host)

##Check variance explained by model
(summary(por.comm.rda.host)$constr.chi/summary(por.comm.rda.host)$tot.chi)*100
#28.95% of the variance in host physiology is constrained by temperature metrics 

RsquareAdj(por.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 26.83% of the variation in host physiology 

summary(por.comm.rda.host)$cont
#dbRDA1 explains 24.38% of total variance 
#dbRDA2 explains 3.49% of total variance 

summary(por.comm.rda.host)$concont
#dbRDA1 explains 84.22% of constrained variance 
#dbRDA2 explains 12.05% of constrained variance 

##Check variance explained by each Host physiological metric
anova(por.comm.rda.host, by="terms")
# Model: rda(formula = por_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD, data = por_data_temp)
#           Df Variance       F Pr(>F)    
# mean       1 0.044662 39.0641  0.001 ***
# maximum    1 0.013423 11.7409  0.001 ***
# minimum    1 0.002465  2.1562  0.098 .  
# SD         1 0.001881  1.6451  0.188    
# Residual 134 0.153202  

```

#Porites RDA- Symbiont metrics 
```{r}
names(por_data_temp)
por.comm.rda.sym<-rda(por_data_temp[, c(7,8,10,12,13,15,16,18,19)] ~ mean + maximum + minimum + SD + Condition(haplotype), 
                       data=por_data_temp)

##Check model significance
anova(por.comm.rda.sym)
# Model: rda(formula = por_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD + Condition(haplotype), data = por_data_temp)
#           Df Variance      F Pr(>F)    
# Model      4  0.17355 9.4521  0.001 ***
# Residual 133  0.61050    

plot(por.comm.rda.sym)

##Check variance explained by model
(summary(por.comm.rda.sym)$constr.chi/summary(por.comm.rda.sym)$tot.chi)*100
#18.77% of the variance in symbiont physiology is constrained by temperature

RsquareAdj(por.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 16.91% of the variation in Symbiont physiology

summary(por.comm.rda.sym)$cont
#dbRDA1 explains 15.19% of total variance 
#dbRDA2 explains 5.93 of total variance 

summary(por.comm.rda.sym)$concont
#dbRDA1 explains 68.64% of constrained variance 
#dbRDA2 explains 26.78% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(por.comm.rda.sym, by="terms")
# Model: dbrda(formula = por_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD + Condition(haplotype), data = por_data_temp)
#           Df Variance       F Pr(>F)    
# mean       1  0.05473 11.9226  0.001 ***
# maximum    1  0.05334 11.6214  0.001 ***
# minimum    1  0.03224  7.0226  0.001 ***
# SD         1  0.03324  7.2417  0.001 ***
# Residual 133  0.61050    

```

#Pocillopora RDA- Host metrics 
```{r}
names(poc_data_temp)
poc.comm.rda.host<-rda(poc_data_temp[, c(6,9,11,14,17)] ~ mean + maximum + minimum + SD + Condition(haplotype), 
                       data=poc_data_temp)

##Check model significance
anova(poc.comm.rda.host)
# Model: rda(formula = poc_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD + Condition(haplotype), data = poc_data_temp)
#           Df Variance      F Pr(>F)    
# Model      4 0.017712 11.336  0.001 ***
# Residual 127 0.049605    

plot(poc.comm.rda.host)

##Check variance explained by model
(summary(poc.comm.rda.host)$constr.chi/summary(poc.comm.rda.host)$tot.chi)*100
#25.94% of the variance in host physiology is constrained by temperature metrics 

RsquareAdj(poc.comm.rda.host)$adj.r.squared*100
#Adjusted: The model explains 23.84% of the variation in host physiology 

summary(poc.comm.rda.host)$cont
#dbRDA1 explains 22.66% of total variance 
#dbRDA2 explains 3.21% of total variance 

summary(poc.comm.rda.host)$concont
#dbRDA1 explains 86.13% of constrained variance 
#dbRDA2 explains 12.18% of constrained variance 

##Check variance explained by each Host physiological metric
anova(poc.comm.rda.host, by="terms")
# Model: dbrda(formula = poc_data_temp[, c(6, 9, 11, 14, 17)] ~ mean + maximum + minimum + SD + Condition(haplotype), data = poc_data_temp)
#           Df SumOfSqs       F Pr(>F)    
# mean       1   1.8829 36.5190  0.001 ***
# maximum    1   0.0528  1.0242  0.367    
# minimum    1   0.0506  0.9819  0.370    
# SD         1   0.3516  6.8203  0.002 ** 
# Residual 127   6.5479 

```

#Pocillopora RDA- Symbiont metrics 
```{r}
names(poc_data_temp)
poc.comm.rda.sym<-rda(poc_data_temp[, c(7,8,10,12,13,15,16,18,19)] ~ mean + maximum + minimum + SD + Condition(haplotype), 
                       data=poc_data_temp)

##Check model significance
anova(poc.comm.rda.sym)
# Model: rda(formula = poc_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD + Condition(haplotype), data = poc_data_temp)
#           Df Variance      F Pr(>F)    
# Model      4  0.26262 14.584  0.001 ***
# Residual 127  0.57176    

plot(poc.comm.rda.sym)

##Check variance explained by model
(summary(poc.comm.rda.sym)$constr.chi/summary(poc.comm.rda.sym)$tot.chi)*100
#31.16% of the variance in symbiont physiology is constrained by temperature

RsquareAdj(poc.comm.rda.sym)$adj.r.squared*100
#Adjusted: The model explains 29.02% of the variation in Symbiont physiology

summary(poc.comm.rda.sym)$cont
#dbRDA1 explains 23.14% of total variance 
#dbRDA2 explains 7.85% of total variance 

summary(poc.comm.rda.sym)$concont
#dbRDA1 explains 74.26% of constrained variance 
#dbRDA2 explains 25.20% of constrained variance 

##Check variance explained by each Symbiont physiological metric
anova(poc.comm.rda.sym, by="terms")
# Model: dbrda(formula = poc_data_temp[, c(7, 8, 10, 12, 13, 15, 16, 18, 19)] ~ mean + maximum + minimum + SD, data = poc_data_temp)
#           Df SumOfSqs       F Pr(>F)    
# mean       1    9.346 15.6422  0.001 ***
# maximum    1   10.684 17.8812  0.001 ***
# minimum    1    2.426  4.0601  0.008 ** 
# SD         1   12.172 20.3716  0.001 ***
# Residual 128   76.482  
```

#Plot results 

ACR: 
- Significant effect of mean, max, min on host phys 
- Significant effect of mean, max, min, and SD on sym phys 

POR: 
- Significant effect of mean and max on host phys 
- Significant effect of mean, max, min, and SD on sym phys

POC:   
- Significant effect of mean and SD on host phys 
- Significant effect of mean, max, min, SD on sym phys

```{r}
ordiplot(acr.comm.rda.host, scaling = 2, main = "ACR Host RDA")
```

```{r}
ordiplot(acr.comm.rda.sym, scaling = 2, main = "ACR Symbiont RDA")
```

```{r}
ordiplot(por.comm.rda.host, scaling = 2, main = "POR Host RDA")
```

```{r}
ordiplot(por.comm.rda.sym, scaling = 2, main = "POR Symbiont RDA")
```

```{r}
ordiplot(poc.comm.rda.host, scaling = 2, main = "POC Host RDA")
```

```{r}
ordiplot(poc.comm.rda.sym, scaling = 2, main = "POC Symbiont RDA")
```

Try a fancy plot. 

ACR host 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(acr.comm.rda.host)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(acr.comm.rda.host, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(acr.comm.rda.host, display="bp", choices=c(1,2), scaling=1)
sc_sp <- scores(acr.comm.rda.host, display="species", choices=c(1,2), scaling=1)

scores(acr.comm.rda.host)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(acr.comm.rda.host,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-2,2), 
     ylim = c(-2,2),
     # label the plot (title, and axes)
     main = "Acropora Host RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

ACR sym 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(acr.comm.rda.sym)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(acr.comm.rda.sym, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(acr.comm.rda.sym, display="bp", choices=c(1,2), scaling=1)
sc_sp<- scores(acr.comm.rda.sym, display="species", choices=c(1,2), scaling=1)

scores(acr.comm.rda.sym)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(acr.comm.rda.sym,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-3,1.5), 
     ylim = c(-2,1.5),
     # label the plot (title, and axes)
     main = "Acropora Symbiont RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

POR host 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(por.comm.rda.host)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(por.comm.rda.host, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(por.comm.rda.host, display="bp", choices=c(1,2), scaling=1)
sc_sp <- scores(por.comm.rda.host, display="species", choices=c(1,2), scaling=1)

scores(por.comm.rda.host)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(por.comm.rda.host,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-2,3), 
     ylim = c(-2,3),
     # label the plot (title, and axes)
     main = "Porites Host RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

POR sym 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(por.comm.rda.sym)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(por.comm.rda.sym, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(por.comm.rda.sym, display="bp", choices=c(1,2), scaling=1)
sc_sp <- scores(por.comm.rda.sym, display="species", choices=c(1,2), scaling=1)

scores(por.comm.rda.sym)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(por.comm.rda.sym,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-3.5,1), 
     ylim = c(-3,1),
     # label the plot (title, and axes)
     main = "Porites Symbiont RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

POC host 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(poc.comm.rda.host)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(poc.comm.rda.host, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(poc.comm.rda.host, display="bp", choices=c(1,2), scaling=1)
sc_sp <- scores(poc.comm.rda.host, display="species", choices=c(1,2), scaling=1)

scores(poc.comm.rda.host)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(poc.comm.rda.host,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-2,2), 
     ylim = c(-2,2),
     # label the plot (title, and axes)
     main = "Porites Host RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

POC sym 
```{r}
# Custom triplot code!

## extract % explained by the first 2 axes
perc <- round(100*(summary(poc.comm.rda.sym)$cont$importance[2, 1:2]), 2)

## extract scores - these are coordinates in the RDA space
sc_si <- scores(poc.comm.rda.sym, display="sites", choices=c(1,2), scaling=1)
sc_bp <- scores(poc.comm.rda.sym, display="bp", choices=c(1,2), scaling=1)
sc_sp <- scores(poc.comm.rda.sym, display="species", choices=c(1,2), scaling=1)

scores(poc.comm.rda.sym)

## Custom triplot, step by step

# Set up a blank plot with scaling, axes, and labels
plot(poc.comm.rda.sym,
     scaling = 2, # set scaling type 
     type = "none", # this excludes the plotting of any points from the results
     frame = FALSE,
     # set axis limits
     xlim = c(-2,2), 
     ylim = c(-2,2),
     # label the plot (title, and axes)
     main = "Porites Symbiont RDA",
     xlab = paste0("RDA1 (", perc[1], "%)"), 
     ylab = paste0("RDA2 (", perc[2], "%)") 
)
# add points for site scores
points(sc_si, 
       pch = 21, # set shape (here, circle with a fill colour)
       col = "darkgray", # outline colour
       bg = "darkgray", # fill colour
       cex = 1.2) # size
# add points for species scores
points(sc_sp, 
       pch = 22, # set shape (here, square with a fill colour)
       col = "black",
       bg = "#f2bd33", 
       cex = 1.2)
# add text labels for species abbreviations
text(sc_sp + c(0.03, 0.09), # adjust text coordinates to avoid overlap with points 
     labels = rownames(sc_sp), 
     col = "grey40", 
     font = 2, # bold
     cex = 0.6)

# add arrows for effects of the expanatory variables
arrows(0,0, # start them from (0,0)
       sc_bp[,1], sc_bp[,2], # end them at the score value
       col = "red", 
       lwd = 3)
# add text labels for arrows
text(x = sc_bp[,1] +0.1, # adjust text coordinate to avoid overlap with arrow tip
     y = sc_bp[,2] - 0.03, 
     labels = rownames(sc_bp), 
     col = "red", 
     cex = 1, 
     font = 2)
```

