---
title: "CSR strategy analysis"
author: "Hollie Putnam, Ariana S Huffmyer"
date: "2025"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Trialing analysis of CSR strategies via Grimes et al. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(tidyr)
library(ggtern)

```

# Single time point (TP1) trial with one metric 

## load data and normalize metrics
```{r}
df <- read.csv("output/master_timeseries.csv")

df <- df %>%
  filter(timepoint=="timepoint1")

df <- df %>%
  select(!colony_id)%>%
  select(!species)%>%
  select(!timepoint)%>%
  select(!month)%>%
  select(!site)%>%
  select(!site_code)%>%
  select(!nutrient)%>%
  select(!code)%>%
  select(!haplotype)

# Normalizing all metric columns (excluding 'ids')
metrics <- names(df)[!names(df) %in% "colony_id_corr"]
df_norm <- df
str(df_norm)

for (m in metrics) {
  min_val <- min(df[[m]], na.rm = TRUE)
  max_val <- max(df[[m]], na.rm = TRUE)
  df_norm[[m]] <- (df[[m]] - min_val) / (max_val - min_val)
}


# Extract first 3 letters from colony_id_corr
df_norm$Group <- substr(df_norm$colony_id_corr, 1, 3)

# Optional: convert to factor for consistent coloring
df_norm$Group <- as.factor(df_norm$Group)


ggplot(df_norm, aes(x = Am, fill = Group)) +
  geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
  theme_minimal() +
  labs(title = "Histogram of Am by Species", x = "Normalized Am", y = "Count") +
  scale_fill_brewer(palette = "Set2")

ggplot(df_norm, aes(x = Am, fill = Group)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "black") +
  facet_wrap(~ Group, scales = "free_y") +
  theme_minimal() +
  labs(title = "Histogram of Am by Species", x = "Normalized Am", y = "Count") +
  scale_fill_brewer(palette = "Set2")
```

## Set up Triangular membership function
```{r}
# Triangular membership function
#triangular <- function(x, a, b, c) {
#  ifelse(
#    x <= a | x >= c, 0,
#    ifelse(x < b, (x - a) / (b - a), (c - x) / (c - b))
#  )
#}


# Use the safer triangular function
triangular <- function(x, a, b, c) {
  sapply(x, function(xi) {
    if (xi <= a || xi >= c) return(0)
    if (xi == b) return(1)
    if (xi < b) return(ifelse(b != a, (xi - a) / (b - a), 1))
    return(ifelse(c != b, (c - xi) / (c - b), 1))
  })
}


```

🟦 Competative (Am_C): highest membership when Am is very low

🟩 Ruderal (Am_R): highest membership when Am is mid-range

🟥 Stress Tolerant (Am_S): highest membership when Am is very high
```{r}
# Custom fuzzy coding for Am
# Example Assignment of one metric CRS membership
df_norm$Am_C <- triangular(df_norm$Am, a = 0.00, b = 0.20, c = 0.60)
df_norm$Am_R <- triangular(df_norm$Am, a = 0.20, b = 0.50, c = 0.80)
df_norm$Am_S <- triangular(df_norm$Am, a = 0.40, b = 0.80, c = 1.00)


#Use percentiles for wider fuzzy spans
qs <- quantile(df_norm$Am, probs = c(0.10, 0.5, 0.90), na.rm = TRUE)

df_norm$Am_C <- triangular(df_norm$Am, a = 0.00, b = qs[1], c = qs[2])
df_norm$Am_R <- triangular(df_norm$Am, a = qs[1], b = qs[2], c = qs[3])
df_norm$Am_S <- triangular(df_norm$Am, a = qs[2], b = qs[3], c = 1.00)


```


## Plot triangle plot
```{r}


# Select fuzzy membership columns
df_tern <- df_norm[, c("colony_id_corr", "Am_C", "Am_R", "Am_S")]

# Optional: renaming for plot clarity
names(df_tern) <- c("colony_id", "Am_C", "Am_R", "Am_S")

# Extract first 3 letters from colony_id_corr
df_tern$Group <- substr(df_tern$colony_id, 1, 3)

# Optional: convert to factor for consistent coloring
df_tern$Group <- as.factor(df_tern$Group)


ggtern(data = df_tern, aes(x = Am_C, y = Am_R, z = Am_S, color = Group)) +
  geom_point(alpha = 0.8, size = 3) +
  theme_bw() +
  labs(
    T = "Competative",
    L = "Ruderal",
    R = "Stress Tolerant",
    color = "Species"
  ) +
  theme(legend.position = "right")

```


# Variance 

Generate a list of responses 
```{r}
responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2", "cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

```{r}
# Load full time series data
df <- read.csv("output/master_timeseries.csv")

# Keep only relevant metadata + metrics
df <- df %>%
  select(colony_id_corr, timepoint, species, all_of(responses)) # adjust starts_with if needed

# Get metric columns (exclude ID and metadata)
metrics <- names(df)[names(df) %in% responses]

```

Normalize across time points. 

```{r}
df_norm <- df
for (m in metrics) {
  min_val <- min(df[[m]], na.rm = TRUE)
  max_val <- max(df[[m]], na.rm = TRUE)
  df_norm[[m]] <- (df[[m]] - min_val) / (max_val - min_val)
}
```

Compute CV for each metric 
```{r}

# Compute CV (coefficient of variation) per colony per metric
df_cv <- df_norm %>%
  pivot_longer(cols = all_of(metrics), names_to = "metric", values_to = "value") %>%
  group_by(colony_id_corr, metric) %>%
  summarize(mean_val = mean(value, na.rm = TRUE),
            sd_val = sd(value, na.rm = TRUE),
            CV = sd_val / mean_val, .groups = "drop")


# Pivot wider so each metric has its CV column
df_cv_wide <- df_cv %>%
  select(colony_id_corr, metric, CV) %>%
  pivot_wider(names_from = metric, values_from = CV)

#remove colonies with NA for some metrics 
df_cv_wide <- df_cv_wide %>% drop_na()
```

Make a membership function. 
```{r}
# Triangular membership function
triangular <- function(x, a, b, c) {
  sapply(x, function(xi) {
    if (xi <= a || xi >= c) return(0)
    if (xi == b) return(1)
    if (xi < b) return(ifelse(b != a, (xi - a) / (b - a), 1))
    return(ifelse(c != b, (c - xi) / (c - b), 1))
  })
}

# Apply fuzzy membership for all metrics
csr_memberships <- list()
for (m in metrics) {
  metric_vals <- df_cv_wide[[m]]
  qs <- quantile(metric_vals, probs = c(0.10, 0.5, 0.90), na.rm = TRUE)
  csr_memberships[[paste0(m, "_C")]] <- triangular(metric_vals, a = 0, b = qs[1], c = qs[2])
  csr_memberships[[paste0(m, "_R")]] <- triangular(metric_vals, a = qs[1], b = qs[2], c = qs[3])
  csr_memberships[[paste0(m, "_S")]] <- triangular(metric_vals, a = qs[2], b = qs[3], c = 1)
}

# Combine with CV table
df_csr <- bind_cols(df_cv_wide, as_tibble(csr_memberships))

# Add group ID
df_csr$Group <- substr(df_csr$colony_id_corr, 1, 3)
df_csr$Group <- as.factor(df_csr$Group)
```

Summarise for each colony - calculate mean CSR value from all metric calculations. 
```{r}
# Select only C/R/S columns
csr_cols <- df_csr %>%
  select(colony_id_corr, Group, ends_with("_C"), ends_with("_R"), ends_with("_S"))

# Pivot to long format to separate metric and strategy type
df_csr_long <- csr_cols %>%
  pivot_longer(
    cols = -c(colony_id_corr, Group),
    names_to = "name",
    values_to = "value"
  ) %>%
  separate(name, into = c("metric", "strategy"), sep = "_(?=[^_]+$)")

# Average across all metrics per colony for each strategy
df_csr_summary <- df_csr_long %>%
  group_by(colony_id_corr, Group, strategy) %>%
  summarize(mean_score = mean(value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = strategy, values_from = mean_score)  # gives columns C, R, S
```

Add groupings and make plots. 
```{r}
# Plot ternary CSR triangle
ggtern(data = df_csr_summary, aes(x = C, y = R, z = S, color = Group)) +
  geom_point(size = 3, alpha = 0.85) +
  theme_bw() +
  labs(
    title = "Summary CSR Strategy per Colony (All Metrics)",
    T = "Competitor",
    L = "Ruderal",
    R = "Stress Tolerant"
  ) +
  theme(legend.position = "right")
```











# Calculate for each time point using all metrics with trait specific assignments  

Generate a list of responses 
```{r}
responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2", "cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

```{r}
# Load full time series data
df <- read.csv("output/master_timeseries.csv")

# Keep only relevant metadata + metrics
df <- df %>%
  select(colony_id_corr, timepoint, species, all_of(responses)) # adjust starts_with if needed

# Get metric columns (exclude ID and metadata)
metrics <- names(df)[names(df) %in% responses]

```

Normalize across time points. 

```{r}
df_norm <- df
for (m in metrics) {
  min_val <- min(df[[m]], na.rm = TRUE)
  max_val <- max(df[[m]], na.rm = TRUE)
  df_norm[[m]] <- (df[[m]] - min_val) / (max_val - min_val)
}
```

Assign traits 
```{r}
# 3. Define trait-to-CSR category logic
csr_trait_map <- list(
  "Am" = c("C", "R", "S"),
  "calc.umol.cm2.hr" = c("S", "R", "C"),
  "Host_AFDW.mg.cm2" = c("C", "R", "S"),
  "cells.mgAFDW" = c("S", "R", "C"),
  "Rd" = c("C", "R", "S"),
  "prot_mg.mgafdw" = c("C", "R", "S"),
  "cre.umol.mgafdw" = c("C", "R", "S"),
  "Sym_AFDW.mg.cm2" = c("S", "R", "C"),
  "Total_Chl_cell" = c("C", "R", "S"),
  "Total_Chl" = c("C", "R", "S"),
  "AQY" = c("C", "R", "S"),
  "Ik" = c("C", "R", "S"),
  "Ic" = c("C", "R", "S"),
  "Ratio_AFDW.mg.cm2" = c("C", "R", "S")
)

```

Compute metrics and assign strategy. 
```{r}
# 4. Compute quantile-based memberships and assign strategy per metric
csr_rows <- list()

for (m in responses) {
  strategy_order <- csr_trait_map[[m]]
  vals <- df_norm[[m]]
  qs <- quantile(vals, probs = c(0.33, 0.66), na.rm = TRUE)
  
  # Assign C, R, or S based on normalized value
  assignment <- case_when(
    vals <= qs[1] ~ strategy_order[1],
    vals <= qs[2] ~ strategy_order[2],
    TRUE ~ strategy_order[3]
  )
  
  df_norm[[paste0(m, "_CSR")]] <- assignment
}
```

```{r}
# 5. Count C/R/S membership per colony across all metrics at each timepoint
df_long <- df_norm %>%
  select(colony_id_corr, timepoint, ends_with("_CSR")) %>%
  pivot_longer(-c(colony_id_corr, timepoint), names_to = "metric", values_to = "strategy") %>%
  mutate(metric = sub("_CSR$", "", metric))
```

```{r}
# 6. Compute mean CSR strategy per colony across metrics
df_csr_summary <- df_long %>%
  group_by(colony_id_corr, strategy) %>%
  summarize(n = n(), .groups = "drop") %>%
  pivot_wider(names_from = strategy, values_from = n, values_fill = 0)

```

```{r}
# Normalize counts to proportions
df_csr_summary <- df_csr_summary %>%
  rowwise() %>%
  mutate(total = sum(c_across(c("C", "R", "S")), na.rm = TRUE)) %>%
  mutate(C = C / total, R = R / total, S = S / total) %>%
  select(-total)
```

```{r}
# Add Group (species) info for plotting
df_csr_summary <- df_csr_summary %>%
  left_join(df %>% select(colony_id_corr, species) %>% distinct(), by = "colony_id_corr") %>%
  rename(Group = species)
```

```{r}
# 7. Plot
ggtern(data = df_csr_summary, aes(x = C, y = R, z = S, color = Group)) +
  geom_point(size = 3, alpha = 0.85) +
  theme_bw() +
  scale_color_manual(values=c("darkgray", "orange", "purple"))+
  labs(
    title = "Summary CSR Strategy per Colony (Trait-Based)",
    T = "Competitor",
    L = "Ruderal",
    R = "Stress Tolerant"
  ) +
  theme(legend.position = "right")
```

# Make a plot for each season 

Load and normalize data 

```{r}
# 1. Set metric list
responses <- c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic",
               "Ratio_AFDW.mg.cm2", "cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")

# 2. Load data
df <- read.csv("output/master_timeseries.csv")

# 3. Keep relevant columns
df <- df %>%
  select(colony_id_corr, timepoint, species, all_of(responses))

# 4. Normalize across *all* timepoints
df_norm <- df
for (m in responses) {
  min_val <- min(df[[m]], na.rm = TRUE)
  max_val <- max(df[[m]], na.rm = TRUE)
  df_norm[[m]] <- (df[[m]] - min_val) / (max_val - min_val)
}

```


Map traits 

```{r}
# 5. Trait-to-CSR mapping
csr_trait_map <- list(
  "Am" = c("C", "R", "S"),
  "calc.umol.cm2.hr" = c("S", "R", "C"),
  "Host_AFDW.mg.cm2" = c("C", "R", "S"),
  "cells.mgAFDW" = c("S", "R", "C"),
  "Rd" = c("C", "R", "S"),
  "prot_mg.mgafdw" = c("C", "R", "S"),
  "cre.umol.mgafdw" = c("C", "R", "S"),
  "Sym_AFDW.mg.cm2" = c("S", "R", "C"),
  "Total_Chl_cell" = c("C", "R", "S"),
  "Total_Chl" = c("C", "R", "S"),
  "AQY" = c("C", "R", "S"),
  "Ik" = c("C", "R", "S"),
  "Ic" = c("C", "R", "S"),
  "Ratio_AFDW.mg.cm2" = c("C", "R", "S")
)

# 6. Assign CSR per metric
for (m in responses) {
  strategy_order <- csr_trait_map[[m]]
  vals <- df_norm[[m]]
  qs <- quantile(vals, probs = c(0.33, 0.66), na.rm = TRUE)
  
  df_norm[[paste0(m, "_CSR")]] <- case_when(
    vals <= qs[1] ~ strategy_order[1],
    vals <= qs[2] ~ strategy_order[2],
    TRUE ~ strategy_order[3]
  )
}
```

Change to long format
```{r}
# 7. Long format CSR assignments
df_long <- df_norm %>%
  select(colony_id_corr, timepoint, ends_with("_CSR")) %>%
  pivot_longer(-c(colony_id_corr, timepoint), names_to = "metric", values_to = "strategy") %>%
  mutate(metric = sub("_CSR$", "", metric))
```

Create a dataset for each time point. 

```{r}
# 8. Loop over each timepoint and make plots
timepoints <- unique(df$timepoint)

plot_list <- list()

for (tp in timepoints) {
  df_tp <- df_long %>% filter(timepoint == tp)
  
  # Summarize C/R/S counts per colony
  df_summary <- df_tp %>%
    group_by(colony_id_corr, strategy) %>%
    summarize(n = n(), .groups = "drop") %>%
    pivot_wider(names_from = strategy, values_from = n, values_fill = 0) %>%
    rowwise() %>%
    mutate(total = sum(c_across(c("C", "R", "S")), na.rm = TRUE)) %>%
    mutate(C = C / total, R = R / total, S = S / total) %>%
    select(-total)
  
  # Add species
  df_summary <- df_summary %>%
    left_join(df %>% select(colony_id_corr, species) %>% distinct(), by = "colony_id_corr") %>%
    rename(Group = species)
  
   # Create plot
  p <- ggtern(data = df_summary, aes(x = C, y = R, z = S, color = Group)) +
    geom_point(size = 3, alpha = 0.85) +
    theme_bw() +
    scale_color_manual(values = c("darkgray", "orange", "purple")) +
    labs(
      title = paste("CSR Strategy per Colony -", tp),
      T = "Competitor",
      L = "Ruderal",
      R = "Stress Tolerant"
    ) +
    theme(legend.position = "right")
  
  plot_list[[tp]] <- p
}
```

View plots
```{r}
# View plots one by one
plot_list[["timepoint1"]]
plot_list[["timepoint2"]]
plot_list[["timepoint3"]]
plot_list[["timepoint4"]]
```

show all at once

```{r}
df_all <- bind_rows(lapply(timepoints, function(tp) {
  df_tp <- df_long %>% filter(timepoint == tp)
  df_summary <- df_tp %>%
    group_by(colony_id_corr, strategy) %>%
    summarize(n = n(), .groups = "drop") %>%
    pivot_wider(names_from = strategy, values_from = n, values_fill = 0) %>%
    rowwise() %>%
    mutate(total = sum(c_across(c("C", "R", "S")), na.rm = TRUE)) %>%
    mutate(C = C / total, R = R / total, S = S / total) %>%
    select(-total) %>%
    left_join(df %>% select(colony_id_corr, species) %>% distinct(), by = "colony_id_corr") %>%
    mutate(timepoint = tp)
}))

ggtern(df_all, aes(x = C, y = R, z = S, color = species)) +
  geom_point(size = 2.5, alpha = 0.85) +
  facet_wrap(~timepoint) +
  scale_color_manual(values = c("darkgray", "orange", "purple")) +
  theme_bw() +
  labs(
    title = "CSR Strategy Across Timepoints",
    T = "Competitor",
    L = "Ruderal",
    R = "Stress Tolerant"
  )
```



