---
title: "Assemble master data frame for time series analysis"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "03/15/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 
 
 
# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
```
 
# Load and manipulate data  

## Loading data files     

Load all .csv files from output of all timepoints for each biological response     

Note: I had to remake some of these files in each timepoint because some files were saving with capitol letters (?) and therefore not loading into the master dataframe. 
 
```{r}
biomass_files<-list.files("../", pattern = "biomass.csv", recursive=T, full.names=T)
antioxidant_files<-list.files("../", pattern = "antioxidant_capacity.csv", recursive=T, full.names=T)
pi_curve_pars_files<-list.files("../", pattern = "pi_curve_pars_nls.csv", recursive=T, full.names=T)
surface_area_files<-list.files("../", pattern = "surface_area.csv", recursive=T, full.names=T)
protein_files<-list.files("../", pattern = "protein.csv", recursive=T, full.names=T)
symb_densities_files<-list.files("../", pattern = "symb_densities.csv", recursive=T, full.names=T)
chlorophyll_files<-list.files("../", pattern = "chlorophyll.csv", recursive=T, full.names=T)
calcification_files<-list.files("../", pattern = "calcification_output.csv", recursive=T, full.names=T)
```
 
## Read data files 
 
Load all data frames.  

```{r}
#biomass 
biomass_dataset <- data.frame()

for (i in 1:length(biomass_files)){
  biomass_df <- read.csv(biomass_files[i]) #each file will be read in
  biomass_dataset <- rbind(biomass_dataset, biomass_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#antioxidant capacity 
antioxidant_dataset <- data.frame()

for (i in 1:length(antioxidant_files)){
  antioxidant_df <- read.csv(antioxidant_files[i]) #each file will be read in
  antioxidant_dataset <- rbind(antioxidant_dataset, antioxidant_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#PI curve pars
pi_curve_pars_dataset <- data.frame()

for (i in 1:length(pi_curve_pars_files)){
  pi_curve_pars_df <- read.csv(pi_curve_pars_files[i]) #each file will be read in
  pi_curve_pars_dataset <- rbind(pi_curve_pars_dataset, pi_curve_pars_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#surface area 
surface_area_dataset <- data.frame()

for (i in 1:length(surface_area_files)){
  surface_area_df <- read.csv(surface_area_files[i]) #each file will be read in
  surface_area_dataset <- rbind(surface_area_dataset, surface_area_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#protein 
protein_dataset <- data.frame()

for (i in 1:length(protein_files)){
  protein_df <- read.csv(protein_files[i]) #each file will be read in
  protein_dataset <- rbind(protein_dataset, protein_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#symb densities 
symb_densities_dataset <- data.frame()

for (i in 1:length(symb_densities_files)){
  symb_densities_df <- read.csv(symb_densities_files[i]) #each file will be read in
  symb_densities_dataset <- rbind(symb_densities_dataset, symb_densities_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#chlorophyll files 
chlorophyll_dataset <- data.frame()

for (i in 1:length(chlorophyll_files)){
  chlorophyll_df <- read.csv(chlorophyll_files[i]) #each file will be read in
  chlorophyll_dataset <- rbind(chlorophyll_dataset, chlorophyll_df) #for each iteration, bind the new data to the building dataset
}

```


```{r}
#calcification files 
calcification_dataset <- data.frame()

for (i in 1:length(calcification_files)){
  calcification_df <- read.csv(calcification_files[i]) #each file will be read in
  calcification_dataset <- rbind(calcification_dataset, calcification_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#remove files that are not needed from loops
rm(list = ls(pattern = "*_df"))
```


## Generate master data frame    

Read in tag metadata frame. 
```{r}
#Load tag metadata sheet 
tags<-read.csv("../metadata/coral_id_metadata.csv")
```

Prepare datasets for merging by renaming columns and spreading dataframes. Each file needs to have one line per colony per response variable in order to merge and should be without site/species columns as these will be added from metadata sheet.   
```{r}
#antioxidant data 
antioxidant_dataset<-antioxidant_dataset%>%
  select(colony_id, cre.umol.mgprot, timepoint)

#biomass data 
biomass_dataset<- biomass_dataset %>%
  filter(!colony_id=="1xPBS")%>%
  nest(value_col = c(AFDW.mg.cm2, DW.mg.cm2)) %>%
  spread(key = partner, value = value_col) %>%
  unnest(Host, Sym, .sep = '_')%>%
  select(colony_id, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Host_DW.mg.cm2, Sym_DW.mg.cm2, timepoint)

#protein data
protein_dataset<-protein_dataset%>%
  select(colony_id, prot_ug.cm2, timepoint)
  
#symbiont densitites
symb_densities_dataset<-symb_densities_dataset%>%
  select(colony_id, cells.cm2, timepoint)

```

Join all data frames together.  

```{r}
#can also use "join_all"
master <- full_join(full_join(full_join(full_join(full_join(full_join(full_join(
  antioxidant_dataset,
  biomass_dataset, by=c("colony_id", "timepoint")),
  chlorophyll_dataset, by=c("colony_id", "timepoint")),
  protein_dataset, by=c("colony_id", "timepoint")),
  surface_area_dataset, by=c("colony_id", "timepoint")), 
  symb_densities_dataset, by=c("colony_id", "timepoint")), 
  pi_curve_pars_dataset, by=c("colony_id", "timepoint")), 
  calcification_dataset, by=c("colony_id", "timepoint"))

head(master)
```

Identify rows that have no data for any metric (typos, blanks, missing colonies). This helps with cleaning the data set.   

```{r}
master<-master%>%
  relocate(timepoint, .after = colony_id)%>%
  mutate(master, "RowSum" = rowSums(is.na(master[ , 3:18])))%>%
  filter(!RowSum>=12) %>% #remove rows with all na's (rowSums=11)
  select(!RowSum)
```

Correct for colony ID's that changed over time.  

Control for colony id, site name, and timepoint names. 
```{r}
#Add column for month of year for each timepoint
master$month[master$timepoint == "timepoint1"] <- "January 2020"
master$month[master$timepoint == "timepoint2"] <- "March 2020"
master$month[master$timepoint == "timepoint3"] <- "September 2020"
master$month[master$timepoint == "timepoint4"] <- "November 2020"
```


```{r}
#replace colony id with the original tag number if the tag was changed during the timeseries, as shown by a new tag id at the end of the time series
master$colony_id_corr<-tags$Original_Tag[match(master$colony_id, tags$Tag_ID_Nov)] 

#if there is an na (because the colony tag id did not change over the timesereis), populate with the colonyid number from colony_id column
master$colony_id_corr[is.na(master$colony_id_corr)] <- master$colony_id[is.na(master$colony_id_corr)]

#reorder columns
master<-master%>%
    relocate(month, .after = timepoint)%>%
   relocate(colony_id_corr, .after = colony_id)

#change the one colony id manually that was changed in the middle of the series
master<-master%>%
  mutate(colony_id_corr=as.character(colony_id_corr))%>%
  mutate(colony_id_corr=if_else(colony_id=="POR-206", "POR-251", colony_id_corr))%>%
  mutate(colony_id_corr=as.factor(colony_id_corr))

#match site name by site number and species from tag sheet 
master$site<-tags$Site[match(master$colony_id_corr, tags$Original_Tag)]
master$species<-tags$Coral_Species[match(master$colony_id_corr, tags$Original_Tag)]

#add a nutrient level for each site  
master<-master%>%
  mutate(nutrient = if_else(site == "Hilton", "Medium", 
                            if_else(site == "Mahana", "Low", 
                                    if_else(site == "Manava", "High", "NA"))))%>%
  mutate(site_code = paste(site, nutrient))

master$site_code <- factor(master$site_code, levels = c("Mahana Low", "Hilton Medium", "Manava High"))

#rearrange columns
master<-master%>%
  relocate(site, .after = month)%>%
  relocate(species, .after = colony_id_corr)%>%
  relocate(nutrient, .after = site)%>%
  relocate(site_code, .after = nutrient)
  
```

140 colonies in this dataset. This number is expected.     
```{r}
length(levels(master$colony_id_corr))
```

Check for duplicates, will return TRUE if there are no duplicates.    

```{r}
master<-master%>%
  mutate(code=as.factor(paste(colony_id_corr, "-", timepoint)))

length(unique(master$code)) == nrow(master)
length(unique(master$code))
nrow(master)

master<-master%>%
  unique()

length(unique(master$code)) == nrow(master)
length(unique(master$code))
nrow(master)

```

Export list of duplicate values to examine.  

```{r}
duplicates<-master[duplicated(master$code),]
duplicates<-duplicates%>%
   droplevels()
list<-levels(duplicates$code)
 
 master%>%
   filter(code %in% list)%>%
   write_csv(file="output/timeseries_duplicates_qc.csv")
```

There are no duplicates.  

# Normalization and write master file  

In order to examine the effect of normalizer on all response variables, normalize each response to: 
(1) Host protein  (ug/cm2)  
(2) Host AFDW  (mg/cm2)   
(3) Host surface area (cm2)  

To do this, back-normalize variables that are already normalized and divide by each new normalizer.     
```{r}
master<-master%>%
  mutate(mgprot = (prot_ug.cm2*surface.area.cm2)/1000)%>% #get absolute values of normaliers for each coral sample, change from ug to mg
  mutate(mgafdw = Host_AFDW.mg.cm2*surface.area.cm2)%>% #get absolute values of normaliers for each coral sample
  mutate(cells = cells.cm2 * surface.area.cm2)%>% #get absolute number of symbiont cells
  mutate(cells.mgprot = (cells.cm2 * surface.area.cm2)/mgprot)%>% #convert cells per cm2 to cells per mg protein
  mutate(cells.mgAFDW = (cells.cm2 * surface.area.cm2)/mgafdw)%>% #converts cells per cm2 to cells per mg afdw
  mutate(chla.ug.mgprot = (chla.ug.cm2 * surface.area.cm2)/mgprot)%>% #normalize chla to protein
  mutate(chla.ug.mgAFDW = (chla.ug.cm2 * surface.area.cm2)/mgafdw)%>% #normalize chla to afdw
  mutate(chlc2.ug.mgprot = (chlc2.ug.cm2 * surface.area.cm2)/mgprot)%>% #normalize chlc2 to protein
  mutate(chlc2.ug.mgAFDW = (chlc2.ug.cm2 * surface.area.cm2)/mgafdw)%>% #normalize chlc2 to afdw
  mutate(prot_mg.cm2=prot_ug.cm2/1000)%>% #change protein to mg
  mutate(prot_mg.mgafdw=(prot_mg.cm2*surface.area.cm2)/mgafdw)%>% #normalize protein to afdw
  mutate(cre.umol.mgafdw= (cre.umol.mgprot*mgprot)/mgafdw)%>% #normalize tac to afdw
  mutate(cre.umol.cm2= (cre.umol.mgprot*mgprot)/surface.area.cm2)%>% #normalize tac to surface area
  mutate(chla.ug.cell = (chla.ug.cm2 * surface.area.cm2)/cells) %>% #normalize to chl per cell
  mutate(chlc2.ug.cell = (chlc2.ug.cm2 * surface.area.cm2)/cells) %>% #normalize to chl per cell
  rename(calc.umol.cm2.hr = umol.cm2.hr)%>%
  mutate(calc.umol.mgprot.hr = (calc.umol.cm2.hr * surface.area.cm2)/mgprot)%>%
  mutate(calc.umol.mgAFDW.hr = (calc.umol.cm2.hr * surface.area.cm2)/mgafdw)
```

Calculate additional metrics of host:symbiont biomass ratio and total cholorophyll (a + c2).  

```{r}
master <- master %>% 
  mutate(Ratio_AFDW.mg.cm2=Sym_AFDW.mg.cm2/(Sym_AFDW.mg.cm2+Host_AFDW.mg.cm2))%>%
  mutate(Total_Chl=chla.ug.mgAFDW+chlc2.ug.mgAFDW)%>%
  mutate(Total_Chl_cell=chla.ug.cell+chlc2.ug.cell)

head(master)
```

Subset data for D. Conetta Acropora colonies for separate analysis.  

```{r}
conetta_colonies<-c("ACR-140", "ACR-150", "ACR-173", "ACR-185", "ACR-187", "ACR-225", "ACR-228", "ACR-234", "ACR-237", "ACR-244", "ACR-343", "ACR-368", "ACR-379", "ACR-389", "ACR-390", "ACR-193", "ACR-180", "ACR-178", "ACR-243", "ACR-256", "ACR-374", "ACR-350", "ACR-351", "ACR-363")
conetta<-master%>%
  filter(colony_id_corr %in% conetta_colonies)%>%
  arrange(colony_id_corr)%>%
  write.csv(., "Output/conetta_data.csv")
```

 
Write master file to csv.  
```{r}
write_csv(master, "Output/master_timeseries.csv")
```

Generate list of colonies found at time point 2 for molecular analyses.  

```{r}
tp2_list<-master%>%
  filter(timepoint=="timepoint2")%>%
  select(colony_id, colony_id_corr) #generate list of colonies in tp2 

all_list<-master%>%
  select(colony_id, colony_id_corr)%>%
  unique()

#show colonies that are not in tp2 that are present in the other datasets shared values
others<-setdiff(all_list, tp2_list)
```


# Generating summary by groups   

Summarize all response values by group (species + timepoint + site).       

```{r}
summary_colony<-master%>%
  select(., -c(colony_id, timepoint, site_code))%>%
  group_by(colony_id_corr, month, site, nutrient, species)%>%
  summarise(across(.cols=everything(), ~mean(.x, na.rm = TRUE)))%>%
  drop_na("site")%>%
  write_csv(., "Output/Summary_Colony_Responses.csv")

summary_group<-master%>%
  select(., -c(colony_id, timepoint, site_code, colony_id_corr))%>%
  group_by(month, site, nutrient, species)%>%
  summarise(across(.cols=everything(), ~mean(.x, na.rm = TRUE)))%>%
  drop_na("site")%>%
  write_csv(., "Output/Summary_Responses.csv")
           
```











