---
title: "Univariate analysis of E5 time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "03/15/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("directlabels")) install.packages("directlabels")
if (!require("broom")) install.packages("broom")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(directlabels)
library(broom)
library(emmeans)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")

master$month<-as.factor(master$month)
master$month<-fct_relevel(master$month, "January 2020", "March 2020", "September 2020", "November 2020")

master<-master%>%
  mutate(site=if_else(site=="Mahana", "Orovau",
                      if_else(site=="Hilton", "Vaipahu", 
                              if_else(site=="Manava", "Matotia", NA))))

master$site<-fct_relevel(master$site, "Orovau", "Vaipahu", "Matotia")
```

# Filter to colonies that are present throughout entire time series

Filter our data set to only analyze colonies that had all physiological metrics measured. This will show univariate responses that match the same data sets that are analyzed in our multivariate work. The full data set including colonies that did not have all metrics measured can be run by removing this filtering step. 

Read in list of colonies present in multivariate time series by filtering the master data frame for complete cases.  

```{r}
str(master)
```
There are 448 observations before filtering. 

View sample size for each group before filtering. 

```{r}
n_before<-master%>%
  group_by(species, timepoint, site)%>%
  summarise(length=length(unique(colony_id_corr)));n_before
```

```{r}
master<-master%>%
  filter(complete.cases(.)) 

str(master)
```
There are 379 observations remaining after complete cases filtering. 

View sample size for each group after filtering. 

```{r}
n_after<-master%>%
  group_by(species, timepoint, site)%>%
  summarise(length=length(unique(colony_id_corr)));n_after
```

Some groups have low sample size (mainly ACR at certain sites and times) due to mortality and loss of tags.  

# Univariate Responses  

Plot univariate response plots and then generate panel of all plots at the end of this section. Individual plots will not be displayed.

Haplotype will be included as a random effect in models. We will include haplotype as a random effect with colony ID nested within haplotype (each colony is assigned to only one haplotype always and always within a species). This will be denoted as `+(1|species/haplotype/colony_id_corr)`. Multiple haplotypes only occur in POC and POR, so models analyzing individual species will not include haplotype for ACR. 

## Antioxidant capacity  

### Plot: CRE umol mgprot 

View by site and timepoint for each species.   
```{r}
tacplot_prot<-master %>%
  filter(!is.na(cre.umol.mgprot)) %>% 
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cre.umol.mgprot)%>%
  
  ggplot(., aes(x = site, y = log(1+cre.umol.mgprot), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Antioxidant Capacity (", mu, "mol CRE mg protein"^-1,")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );tacplot_prot
```

### Plot: CRE umol mgafdw 

View by site and timepoint for each species.   
```{r}
tacplot_afdw_box<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cre.umol.mgafdw)%>%
  
  ggplot(., aes(x = site, y = log(1+cre.umol.mgafdw), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Antioxidant Capacity (", mu, "mol CRE mg afdw"^-1,")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );tacplot_afdw_box
```

View by means
```{r}
tacplot_afdw<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, cre.umol.mgafdw)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(cre.umol.mgafdw, na.rm=TRUE), se=sd(cre.umol.mgafdw, na.rm=TRUE)/sqrt(length(cre.umol.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
    geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  

  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    xlab("Month") + 
    ylab(expression(bold(paste("Antioxidant Capacity (", mu, "mol CRE mg afdw"^-1,")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );tacplot_afdw
```

### Plot: CRE umol cm2 

View by site and timepoint for each species.   
```{r}
tacplot_cm2<-master %>%
  filter(!is.na(cre.umol.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cre.umol.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+cre.umol.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
     scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Antioxidant Capacity (", mu, "mol CRE mg cm"^-2,")"))))+
    theme_classic() + 
    theme(
      legend.position="right",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );tacplot_cm2

```

Join all plots for each normalization together.  

```{r}
tac_figure<-plot_grid(tacplot_prot, tacplot_afdw_box, tacplot_cm2, labels = c("", "", ""), ncol=3, nrow=1, rel_widths = c(.8, .8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/TAC_Figure.pdf", plot=tac_figure, dpi=300, width=24, height=5, units="in")
```

### Analysis 

Build a mixed model for univariate analysis and examine data distribution of biomass normalized antioxidant capacity. 

```{r}
antiox_model<-lmer(cre.umol.mgafdw~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(antiox_model))
#hist(residuals(antiox_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
antiox_model<-lmer(log(1+cre.umol.mgafdw)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(antiox_model))
#hist(residuals(antiox_model))
```

Generate a Type III Anova of model.    

```{r}
anova(antiox_model, type="III")
summary(antiox_model)
```

Since species:time is significant, run a model for site and time effects within each species.  

*Acropora*:

```{r}
antiox_model_acr<-lmer(log(1+cre.umol.mgafdw)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(antiox_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(antiox_model_acr, type="III")
summary(antiox_model_acr)
rand(antiox_model_acr)

emm<-emmeans(antiox_model_acr, ~ site | timepoint)
pairs(emm)

emm<-emmeans(antiox_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(antiox_model_acr, ~ timepoint)
pairs(emm)
```

*Pocillopora*: 

```{r}
antiox_model_poc<-lmer(log(1+cre.umol.mgafdw)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(antiox_model_poc))
#hist(residuals(antiox_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(antiox_model_poc, type="III")
summary(antiox_model_poc)
rand(antiox_model_poc)

emm<-emmeans(antiox_model_poc, ~ timepoint)
pairs(emm)
```

*Porites*:  

```{r}
antiox_model_por<-lmer(log(1+cre.umol.mgafdw)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(antiox_model_por))
#hist(residuals(antiox_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(antiox_model_por, type="III")
summary(antiox_model_por)
rand(antiox_model_por)

emm<-emmeans(antiox_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(antiox_model_por, ~ site)
pairs(emm)
```

## Symbiont Densities  

### Plot: Surface Area    
 
View by site and timepoint for each species.   
```{r}
symbplot_cm2<-master %>%
  filter(!is.na(cells.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cells.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+cells.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Symbiont Cells cm"^-2))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );symbplot_cm2
```

### Plot: Protein   

View by site and timepoint for each species.   
```{r}
symbplot_prot<-master %>%
  filter(!is.na(cells.mgprot)) %>%
   filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cells.mgprot)%>%
  
  ggplot(., aes(x = site, y = log(1+cells.mgprot), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Symbiont Cells mg protein"^-1))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );symbplot_prot

```

### Plot: Biomass   

View by site and timepoint for each species.   
```{r}
symbplot_afdw_box<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
    filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, cells.mgAFDW)%>%
  
  ggplot(., aes(x = site, y = log(1+cells.mgAFDW), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Symbiont Cells mg afdw"^-1))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );symbplot_afdw_box
```

View by means
```{r}
symbplot_afdw<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, cells.mgAFDW)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(cells.mgAFDW, na.rm=TRUE), se=sd(cells.mgAFDW, na.rm=TRUE)/sqrt(length(cells.mgAFDW)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(1e+04, 9e+05)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Symbiont Density (cells mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );symbplot_afdw
```

Join all plots for each normalization together.  

```{r}
symbcells_figure<-plot_grid(symbplot_prot, symbplot_afdw_box, symbplot_cm2, labels = c("", "", ""), ncol=3, nrow=1, rel_widths = c(.8, .8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/CellDensity_Figure.pdf", plot=symbcells_figure, dpi=300, width=24, height=5, units="in")
```

### Analysis  

Build a mixed model for univariate analysis and examine data distribution.

```{r}
density_model<-lmer(cells.mgAFDW~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(cells.mgAFDW>0))
qqPlot(residuals(density_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

`density_model<-lmer(log(cells.mgAFDW)~timepoint*site*species+(1|colony_id_corr:haplotype)+(1|haplotype), na.action=na.omit, data=master)` 

```{r}
density_model<-lmer(log(cells.mgAFDW)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(cells.mgAFDW>0))
qqPlot(residuals(density_model))
```

Residuals are improved, but we need to revisit this model to improve fit.  

Generate a Type III Anova of model.    

```{r}
anova(density_model, type="III")

emm<-emmeans(density_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
density_model_acr<-lmer(log(cells.mgAFDW+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(cells.mgAFDW>0 & species=="Acropora"))
qqPlot(residuals(density_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(density_model_acr, type="III")
rand(density_model_acr)

emm<-emmeans(density_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(density_model_acr, ~ site | timepoint)
pairs(emm)

emm<-emmeans(density_model_acr, ~ timepoint | site)
pairs(emm)
```

*Pocillopora*  

```{r}
density_model_poc<-lmer(log(cells.mgAFDW+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(cells.mgAFDW>0 & species=="Pocillopora"))
qqPlot(residuals(density_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(density_model_poc, type="III")
rand(density_model_poc)

emm<-emmeans(density_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(density_model_poc, ~ site | timepoint)
pairs(emm)

emm<-emmeans(density_model_poc, ~ timepoint | site)
pairs(emm)
```

*Porites*  

```{r}
density_model_por<-lmer(log(cells.mgAFDW+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(cells.mgAFDW>0 & species=="Porites"))
qqPlot(residuals(density_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(density_model_por, type="III")
summary(density_model_por)
rand(density_model_por)

emm<-emmeans(density_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(density_model_por, ~ site)
pairs(emm)

emm<-emmeans(density_model_por, ~ site | timepoint)
pairs(emm)

emm<-emmeans(density_model_por, ~ timepoint | site)
pairs(emm)
```

## Biomass

### Plot Host Biomass  
 
View by site and timepoint for each species in the host.  
```{r} 
hAFDWplot_box<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  select(colony_id_corr, species, site, timepoint, Host_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+Host_AFDW.mg.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Host Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );hAFDWplot_box
```

View by means
```{r}
hAFDWplot<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Host_AFDW.mg.cm2)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Host_AFDW.mg.cm2, na.rm=TRUE), se=sd(Host_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Host_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  
    ylim(0, 12)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Host Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );hAFDWplot
```

### Plot Symbiont biomass  

View by site and timepoint for each species in the symbiont   
```{r} 
sAFDWplot_box<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  select(colony_id_corr, species, site, timepoint, Sym_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+Sym_AFDW.mg.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Sym. Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );sAFDWplot_box
```

View by means
```{r}
sAFDWplot<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Sym_AFDW.mg.cm2)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Sym_AFDW.mg.cm2, na.rm=TRUE), se=sd(Sym_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Sym_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 6)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Symbiont Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );sAFDWplot
```

### Plot S:H Ratio  

Calculate biomass as a ratio of host to symbiont and plot.  

```{r} 
ratioAFDWplot_box<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  select(colony_id_corr, species, site, timepoint, Ratio_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+Ratio_AFDW.mg.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Symbiont : Holobiont Biomass"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );ratioAFDWplot_box
```

View by means
```{r}
ratioAFDWplot<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  select(colony_id_corr, species, site, month, Ratio_AFDW.mg.cm2)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Ratio_AFDW.mg.cm2, na.rm=TRUE), se=sd(Ratio_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Ratio_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 0.5)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Symbiont : Host Biomass"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );ratioAFDWplot
```

Join all plots together.  

```{r}
afdw_figure<-plot_grid(hAFDWplot_box, sAFDWplot_box, ratioAFDWplot_box, labels = c("", "", ""), ncol=3, nrow=1, rel_widths = c(.8, .8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/Biomass_Figure.pdf", plot=afdw_figure, dpi=300, width=24, height=5, units="in")
```

### Analysis  

Build a mixed model for univariate analysis and examine data distribution.

#### Host Biomass:  

```{r}
hAFDW_model<-lmer(Host_AFDW.mg.cm2~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0))
qqPlot(residuals(hAFDW_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
hAFDW_model<-lmer(log(1+Host_AFDW.mg.cm2)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0))
qqPlot(residuals(hAFDW_model))
```

Residuals are improved.

Generate a Type III Anova of model.    

```{r}
anova(hAFDW_model, type="III")
summary(hAFDW_model)

#emm<-emmeans(hAFDW_model, ~ species)
pairs(emm)
```

*Acropora*  
 
`hAFDW_model_acr<-lmer(log(1+Host_AFDW.mg.cm2)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0 & species=="Acropora"))` 

```{r}
hAFDW_model_acr<-lmer(log(1+Host_AFDW.mg.cm2)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0 & species=="Acropora"))
qqPlot(residuals(hAFDW_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(hAFDW_model_acr, type="III")
summary(hAFDW_model_acr)
rand(hAFDW_model_acr)

emm<-emmeans(hAFDW_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(hAFDW_model_acr, ~ site)
pairs(emm)
```

*Pocillopora*  

```{r}
hAFDW_model_poc<-lmer(log(1+Host_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0 & species=="Pocillopora"))
qqPlot(residuals(hAFDW_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(hAFDW_model_poc, type="III")
summary(hAFDW_model_poc)
rand(hAFDW_model_poc)

emm<-emmeans(hAFDW_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(hAFDW_model_poc, ~ site)
pairs(emm)

emm<-emmeans(hAFDW_model_poc, ~ site | timepoint)
pairs(emm)

emm<-emmeans(hAFDW_model_poc, ~ timepoint | site)
pairs(emm)
```

*Porites*  

```{r}
hAFDW_model_por<-lmer(log(1+Host_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Host_AFDW.mg.cm2>0 & species=="Porites"))
qqPlot(residuals(hAFDW_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(hAFDW_model_por, type="III")
summary(hAFDW_model_por)
rand(hAFDW_model_por)

emm<-emmeans(hAFDW_model_por, ~ site)
pairs(emm)
```

#### Symbiont Biomass:  

```{r}
sAFDW_model<-lmer(Sym_AFDW.mg.cm2~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Sym_AFDW.mg.cm2>0))
qqPlot(residuals(sAFDW_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
sAFDW_model<-lmer(log(1+Sym_AFDW.mg.cm2)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Sym_AFDW.mg.cm2>0))
qqPlot(residuals(sAFDW_model))
```

Residuals are improved.  

Generate a Type III Anova of model.    

```{r}
anova(sAFDW_model, type="III")
summary(sAFDW_model)

emm<-emmeans(sAFDW_model, ~ species)
pairs(emm)
```

*Acropora* 

```{r}
sAFDW_model_acr<-lmer(log(1+Sym_AFDW.mg.cm2)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Sym_AFDW.mg.cm2>0 & species=="Acropora"))
qqPlot(residuals(sAFDW_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(sAFDW_model_acr, type="III")
summary(sAFDW_model_acr)
rand(sAFDW_model_acr)

emm<-emmeans(sAFDW_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(sAFDW_model_acr, ~ site)
pairs(emm)

emm<-emmeans(sAFDW_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(sAFDW_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora* 

```{r}
sAFDW_model_poc<-lmer(log(1+Sym_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Sym_AFDW.mg.cm2>0 & species=="Pocillopora"))
qqPlot(residuals(sAFDW_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(sAFDW_model_poc, type="III")
summary(sAFDW_model_poc)
rand(sAFDW_model_poc)

emm<-emmeans(sAFDW_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(sAFDW_model_poc, ~ timepoint | site)
pairs(emm)

emm<-emmeans(sAFDW_model_poc, ~ site | timepoint)
pairs(emm)
```

*Porites* 

```{r}
sAFDW_model_por<-lmer(log(1+Sym_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Sym_AFDW.mg.cm2>0 & species=="Porites"))
qqPlot(residuals(sAFDW_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(sAFDW_model_por, type="III")
summary(sAFDW_model_por)
rand(sAFDW_model_por)

emm<-emmeans(sAFDW_model_por, ~ site)
pairs(emm)
```

#### S:H Biomass: 

```{r}
shAFDW_model<-lmer(Ratio_AFDW.mg.cm2~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Ratio_AFDW.mg.cm2>0))
qqPlot(residuals(shAFDW_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
shAFDW_model<-lmer(log(1+Ratio_AFDW.mg.cm2)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Ratio_AFDW.mg.cm2>0))
qqPlot(residuals(shAFDW_model))
```

Residuals are improved. 

Generate a Type III Anova of model.    

```{r}
anova(shAFDW_model, type="III")
summary(shAFDW_model)

#emm<-emmeans(shAFDW_model, ~ species)
#pairs(emm)
```

*Acropora*   

```{r}
shAFDW_model_acr<-lmer(log(1+Ratio_AFDW.mg.cm2)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Ratio_AFDW.mg.cm2>0 & species=="Acropora"))
qqPlot(residuals(shAFDW_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(shAFDW_model_acr, type="III")
summary(shAFDW_model_acr)
rand(shAFDW_model_acr)

emm<-emmeans(shAFDW_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(shAFDW_model_acr, ~ site)
pairs(emm)

emm<-emmeans(shAFDW_model_acr, ~ site | timepoint)
pairs(emm)

emm<-emmeans(shAFDW_model_acr, ~ timepoint | site)
pairs(emm)
```

*Pocillopora*   

```{r}
shAFDW_model_poc<-lmer(log(1+Ratio_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Ratio_AFDW.mg.cm2>0 & species=="Pocillopora"))
qqPlot(residuals(shAFDW_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(shAFDW_model_poc, type="III")
summary(shAFDW_model_poc)
rand(shAFDW_model_poc)

emm<-emmeans(shAFDW_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(shAFDW_model_poc, ~ site)
pairs(emm)

emm<-emmeans(shAFDW_model_poc, ~ site | timepoint)
pairs(emm)

emm<-emmeans(shAFDW_model_poc, ~ timepoint | site)
pairs(emm)
```

*Porites*   

```{r}
shAFDW_model_por<-lmer(log(1+Ratio_AFDW.mg.cm2)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Ratio_AFDW.mg.cm2>0 & species=="Porites"))
qqPlot(residuals(shAFDW_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(shAFDW_model_por, type="III")
summary(shAFDW_model_por)
rand(shAFDW_model_por)

emm<-emmeans(shAFDW_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(shAFDW_model_por, ~ site)
pairs(emm)
```

## Chlorophyll  

### Plot: Surface area     

View by site and timepoint for each species in Chl a.    
```{r}
chlaplot_cm2<-master %>%
  filter(!is.na(chla.ug.cm2)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chla.ug.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+chla.ug.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll a (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlaplot_cm2
```

View by site and timepoint for each species in Chl c2.   
```{r}
chlc2plot_cm2<-master %>%
  filter(!is.na(chlc2.ug.cm2)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chlc2.ug.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+chlc2.ug.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll c2 (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlc2plot_cm2
```

### Plot: Protein    

View by site and timepoint for each species in Chl a.    
```{r}
chlaplot_prot<-master %>%
  filter(!is.na(chla.ug.mgprot)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chla.ug.mgprot)%>%
  
  ggplot(., aes(x = site, y = log(1+chla.ug.mgprot), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll a (", mu, "g mg protein"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlaplot_prot
```

View by site and timepoint for each species in Chl c2.   
```{r}
chlc2plot_prot<-master %>%
  filter(!is.na(chlc2.ug.mgprot)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chlc2.ug.mgprot)%>%
  
  ggplot(., aes(x = site, y = log(1+chlc2.ug.mgprot), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll c2 (", mu, "g mg protein"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlc2plot_prot
```

### Plot: Biomass    

View by site and timepoint for each species in Chl a.    
```{r}
chlaplot_afdw<-master %>%
  filter(!is.na(chla.ug.mgAFDW)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chla.ug.mgAFDW)%>%
  
  ggplot(., aes(x = site, y = log(1+chla.ug.mgAFDW), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll a (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlaplot_afdw
```

View by site and timepoint for each species in Chl c2.   
```{r}
chlc2plot_afdw<-master %>%
  filter(!is.na(chlc2.ug.mgAFDW)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chlc2.ug.mgAFDW)%>%
  
  ggplot(., aes(x = site, y = log(1+chlc2.ug.mgAFDW), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll c2 (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlc2plot_afdw
```

### Plot: Per cell    

View by site and timepoint for each species in Chl a.    
```{r}
chlaplot_cell<-master %>%
  filter(!is.na(chla.ug.cell)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chla.ug.cell)%>%
  
  ggplot(., aes(x = site, y = log(1+chla.ug.cell), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll a (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlaplot_cell
```

View by site and timepoint for each species in Chl c2.   
```{r}
chlc2plot_cell<-master %>%
  filter(!is.na(chlc2.ug.cell)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, chlc2.ug.cell)%>%
  
  ggplot(., aes(x = site, y = log(1+chlc2.ug.cell), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Chlorophyll c2 (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlc2plot_cell
```

### Plot: Total chl  

View by site and timepoint for each species for total chl per unit AFDW.     
```{r}
chlplot_total_box<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, Total_Chl)%>%
  filter(Total_Chl>0)%>%
  
  ggplot(., aes(x = site, y = log(1+Total_Chl), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Total Chlorophyll (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlplot_total_box
```

View by means
```{r}
chlplot_total<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl>0)%>%
  select(colony_id_corr, species, site, month, Total_Chl)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Total_Chl, na.rm=TRUE), se=sd(Total_Chl, na.rm=TRUE)/sqrt(length(Total_Chl)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 7)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total
```

View by site and timepoint for each species for total chl per cell.     
```{r}
chlplot_total_cell_box<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, Total_Chl_cell)%>%
  filter(Total_Chl_cell>0)%>%
  
  ggplot(., aes(x = site, y = log(1+Total_Chl_cell), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );chlplot_total_cell_box
```

View by means
```{r}
chlplot_total_cell<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl_cell>0)%>%
  select(colony_id_corr, species, site, month, Total_Chl_cell)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Total_Chl_cell, na.rm=TRUE), se=sd(Total_Chl_cell, na.rm=TRUE)/sqrt(length(Total_Chl_cell)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 3e-05)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total_cell
```

Join all plots for each normalization together.  

```{r}
chl_figure<-plot_grid(chlplot_total_box, chlaplot_prot, chlaplot_afdw, chlaplot_cell, chlaplot_cm2, chlplot_total_cell_box, chlc2plot_prot, chlc2plot_afdw, chlc2plot_cell, chlc2plot_cm2,  labels = c("", "", ""), ncol=5, nrow=2, rel_widths = c(.8, .8, .8, .8, 1, .8, .8, .8, .8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/Chlorophyll_Figure.pdf", plot=chl_figure, dpi=300, width=36, height=10, units="in")
```

### Analysis  

Build a mixed model for univariate analysis and examine data distribution.

#### Chlorophyll a: 

```{r}
chla_model<-lmer(chla.ug.mgAFDW~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(chla_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
chla_model<-lmer(log(1+chla.ug.mgAFDW)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chla.ug.mgAFDW>0))
qqPlot(residuals(chla_model))
```

Generate a Type III Anova of model.    

```{r}
anova(chla_model, type="III")
summary(chla_model)

emm<-emmeans(chla_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
chla_model_acr<-lmer(log(1+chla.ug.mgAFDW)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(chla.ug.mgAFDW>0 & species=="Acropora"))
qqPlot(residuals(chla_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(chla_model_acr, type="III")
summary(chla_model_acr)

emm<-emmeans(chla_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(chla_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chla_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora*  

```{r}
chla_model_poc<-lmer(log(1+chla.ug.mgAFDW)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chla.ug.mgAFDW>0 & species=="Pocillopora"))
qqPlot(residuals(chla_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(chla_model_poc, type="III")
summary(chla_model_poc)

emm<-emmeans(chla_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(chla_model_poc, ~ site)
pairs(emm)

emm<-emmeans(chla_model_poc, ~ site | timepoint)
pairs(emm)

emm<-emmeans(chla_model_poc, ~ timepoint | site)
pairs(emm)
```

*Porites*  

```{r}
chla_model_por<-lmer(log(1+chla.ug.mgAFDW)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chla.ug.mgAFDW>0 & species=="Porites"))
qqPlot(residuals(chla_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(chla_model_por, type="III")
summary(chla_model_por)

emm<-emmeans(chla_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(chla_model_por, ~ site)
pairs(emm)

emm<-emmeans(chla_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chla_model_por, ~ site | timepoint)
pairs(emm)
```


#### Chlorophyll c2: 

Build a mixed model for univariate analysis and examine data distribution.

```{r}
chlc2_model<-lmer(chlc2.ug.mgAFDW~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chlc2.ug.mgAFDW>0))
qqPlot(residuals(chlc2_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
chlc2_model<-lmer(log(1+chlc2.ug.mgAFDW)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chlc2.ug.mgAFDW>0))
qqPlot(residuals(chlc2_model))
```

Generate a Type III Anova of model.    

```{r}
anova(chlc2_model, type="III")
summary(chlc2_model)

emm<-emmeans(chlc2_model, ~ species)
pairs(emm)
```

Since species is significant, test for effects of site and time within species.  

*Acropora* 

```{r}
chlc2_model_acr<-lmer(log(1+chlc2.ug.mgAFDW)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(chlc2.ug.mgAFDW>0 & species=="Acropora"))
qqPlot(residuals(chlc2_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(chlc2_model_acr, type="III")
summary(chlc2_model_acr)

emm<-emmeans(chlc2_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(chlc2_model_acr, ~ site)
pairs(emm)

emm<-emmeans(chlc2_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chlc2_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora* 

```{r}
chlc2_model_poc<-lmer(log(1+chlc2.ug.mgAFDW)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chlc2.ug.mgAFDW>0 & species=="Pocillopora"))
qqPlot(residuals(chlc2_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(chlc2_model_poc, type="III")
summary(chlc2_model_poc)

emm<-emmeans(chlc2_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(chlc2_model_poc, ~ site)
pairs(emm)

emm<-emmeans(chlc2_model_poc, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chlc2_model_poc, ~ site | timepoint)
pairs(emm)
```

*Porites* 

```{r}
chlc2_model_por<-lmer(log(1+chlc2.ug.mgAFDW)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(chlc2.ug.mgAFDW>0 & species=="Porites"))
qqPlot(residuals(chlc2_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(chlc2_model_por, type="III")
summary(chlc2_model_por)

emm<-emmeans(chlc2_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(chlc2_model_por, ~ site)
pairs(emm)

emm<-emmeans(chlc2_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chlc2_model_por, ~ site | timepoint)
pairs(emm)
```

#### Total Chlorophyll (a + c2): 

Build a mixed model for univariate analysis and examine data distribution.

```{r}
chltotal_model<-lmer(Total_Chl~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl>0))
qqPlot(residuals(chltotal_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
chltotal_model<-lmer(log(1+Total_Chl)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl>0))
qqPlot(residuals(chltotal_model))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotal_model, type="III")
summary(chltotal_model)

emm<-emmeans(chltotal_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
chltotal_model_acr<-lmer(log(1+Total_Chl)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl>0 & species=="Acropora"))
qqPlot(residuals(chltotal_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotal_model_acr, type="III")
summary(chltotal_model_acr)
rand(chltotal_model_acr)

emm<-emmeans(chltotal_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(chltotal_model_acr, ~ site)
pairs(emm)

emm<-emmeans(chltotal_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chltotal_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora*  

```{r}
chltotal_model_poc<-lmer(log(1+Total_Chl)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl>0 & species=="Pocillopora"))
qqPlot(residuals(chltotal_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotal_model_poc, type="III")
summary(chltotal_model_poc)
rand(chltotal_model_poc)

emm<-emmeans(chltotal_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(chltotal_model_poc, ~ site)
pairs(emm)

emm<-emmeans(chltotal_model_poc, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chltotal_model_poc, ~ site | timepoint)
pairs(emm)
```

*Porites*  

```{r}
chltotal_model_por<-lmer(log(1+Total_Chl)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl>0 & species=="Porites"))
qqPlot(residuals(chltotal_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotal_model_por, type="III")
summary(chltotal_model_por)
rand(chltotal_model_por)

emm<-emmeans(chltotal_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(chltotal_model_por, ~ site)
pairs(emm)

emm<-emmeans(chltotal_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(chltotal_model_por, ~ site | timepoint)
pairs(emm)
```

#### Total Chlorophyll per cell (a + c2): 

Build a mixed model for univariate analysis and examine data distribution.

```{r}
chltotalcell_model<-lmer(Total_Chl_cell~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl_cell>0))
qqPlot(residuals(chltotalcell_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
chltotalcell_model<-lmer(log(1+Total_Chl_cell)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl_cell>0))
qqPlot(residuals(chltotalcell_model))
```

Return to check distribution of residuals.  

Generate a Type III Anova of model.    

```{r}
anova(chltotalcell_model, type="III")
summary(chltotalcell_model)

emm<-emmeans(chltotalcell_model, ~ species)
pairs(emm)
```

*Acropora*   

```{r}
chltotalcell_model_acr<-lmer(log(1+Total_Chl_cell)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl_cell>0 & species=="Acropora"))
qqPlot(residuals(chltotalcell_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotalcell_model_acr, type="III")
summary(chltotalcell_model_acr)
rand(chltotalcell_model_acr)

emm<-emmeans(chltotalcell_model_acr, ~ timepoint)
pairs(emm)
```

*Pocillopora*   

```{r}
chltotalcell_model_poc<-lmer(log(1+Total_Chl_cell)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl_cell>0 & species=="Pocillopora"))
qqPlot(residuals(chltotalcell_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotalcell_model_poc, type="III")
summary(chltotalcell_model_poc)
rand(chltotalcell_model_poc)

emm<-emmeans(chltotalcell_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(chltotalcell_model_poc, ~ site)
pairs(emm)
```

*Porites*   

```{r}
chltotalcell_model_por<-lmer(log(1+Total_Chl_cell)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(Total_Chl_cell>0 & species=="Porites"))
qqPlot(residuals(chltotalcell_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(chltotalcell_model_por, type="III")
summary(chltotalcell_model_por)
rand(chltotalcell_model_por)

emm<-emmeans(chltotalcell_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(chltotalcell_model_por, ~ site)
pairs(emm)
```

## Protein 

### Plot: Surface Area  

View by site and timepoint for each species.    
```{r}
protplot_cm2<-master %>%
  filter(!is.na(prot_mg.cm2)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, prot_mg.cm2)%>%
  
  ggplot(., aes(x = site, y = log(1+prot_mg.cm2), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Protein (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );protplot_cm2
```

### Plot: AFDW  

View by site and timepoint for each species.    
```{r}
protplot_afdw_box<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, prot_mg.mgafdw)%>%
  filter(prot_mg.mgafdw>0)%>%
  
  ggplot(., aes(x = site, y = log(1+prot_mg.mgafdw), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Protein (mg mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );protplot_afdw_box
```

View by means
```{r}
protplot_afdw<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, prot_mg.mgafdw)%>%
  filter(prot_mg.mgafdw>0)%>%
  filter(prot_mg.mgafdw<1)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(prot_mg.mgafdw, na.rm=TRUE), se=sd(prot_mg.mgafdw, na.rm=TRUE)/sqrt(length(prot_mg.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 0.75)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Protein (mg mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );protplot_afdw
```

Join all plots for each normalization together.  

```{r}
prot_figure<-plot_grid(protplot_afdw_box, protplot_cm2, labels = c("", "", ""), ncol=2, nrow=1, rel_widths = c(.8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/Protein_Figure.pdf", plot=prot_figure, dpi=300, width=16, height=5, units="in")
```

### Analysis  

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
prot_model<-lmer(prot_mg.mgafdw~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(prot_mg.mgafdw>0 & prot_mg.mgafdw<1))
qqPlot(residuals(prot_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
prot_model<-lmer(log(1+prot_mg.mgafdw)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(prot_mg.mgafdw>0 & prot_mg.mgafdw<1))
qqPlot(residuals(prot_model))
```

Generate a Type III Anova of model.    

```{r}
anova(prot_model, type="III")
summary(prot_model)

#emm<-emmeans(prot_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
prot_model_acr<-lmer(log(1+prot_mg.mgafdw)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(prot_mg.mgafdw>0 & prot_mg.mgafdw<1 & species=="Acropora"))
qqPlot(residuals(prot_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(prot_model_acr, type="III")
summary(prot_model_acr)
rand(prot_model_acr)

emm<-emmeans(prot_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(prot_model_acr, ~ site)
pairs(emm)

emm<-emmeans(prot_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(prot_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora*  

```{r}
prot_model_poc<-lmer(log(1+prot_mg.mgafdw)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(prot_mg.mgafdw>0 & prot_mg.mgafdw<1 & species=="Pocillopora"))
qqPlot(residuals(prot_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(prot_model_poc, type="III")
summary(prot_model_poc)
rand(prot_model_poc)

emm<-emmeans(prot_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(prot_model_poc, ~ site)
pairs(emm)

emm<-emmeans(prot_model_poc, ~ timepoint | site)
pairs(emm)

emm<-emmeans(prot_model_poc, ~ site | timepoint)
pairs(emm)
```

*Porites*  

```{r}
prot_model_por<-lmer(log(1+prot_mg.mgafdw)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(prot_mg.mgafdw>0 & prot_mg.mgafdw<1 & species=="Porites"))
qqPlot(residuals(prot_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(prot_model_por, type="III")
summary(prot_model_por)
rand(prot_model_por)

emm<-emmeans(prot_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(prot_model_por, ~ site)
pairs(emm)

emm<-emmeans(prot_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(prot_model_por, ~ site | timepoint)
pairs(emm)
```

## PI Curve Parameters   

### Plot: Am   

View by site and timepoint for each species.    
```{r}
am_plot_box<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, AQY, Am, Rd)%>%
  
  ggplot(., aes(x = site, y = log(1+Am), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log P" [Max], " (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );am_plot_box
```

View by means
```{r}
am_plot<-master %>%
  filter(!is.na(Am)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Am)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Am, na.rm=TRUE), se=sd(Am, na.rm=TRUE)/sqrt(length(Am)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 3)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("P" [Max], " (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );am_plot
```

### Plot: AQY  

View by site and timepoint for each species.    
```{r}
aqy_plot_box<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, AQY, Am, Rd)%>%
  
  ggplot(., aes(x = site, y = log(1+AQY), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Quantum Yield (alpha)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );aqy_plot_box
```

View by means
```{r}
aqy_plot<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, AQY)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(AQY, na.rm=TRUE), se=sd(AQY, na.rm=TRUE)/sqrt(length(AQY)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 0.012)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Apparent Quantum Yield (AQY)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );aqy_plot
```

### Plot: Rd  

View by site and timepoint for each species.    
```{r}
rd_plot_box<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, AQY, Am, Rd)%>%
  
  ggplot(., aes(x = site, y = log(1+Rd), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("log Respiration (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.position="none",
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );rd_plot_box
```

View by means
```{r}
rd_plot<-master %>%
  filter(!is.na(Rd)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Rd)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Rd, na.rm=TRUE), se=sd(Rd, na.rm=TRUE)/sqrt(length(Rd)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 1.25)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Respiration (R" [D], ") (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );rd_plot
```

### Plot: Ik  

View by site and timepoint for each species.    
```{r}
Ik_plot_box<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, AQY, Am, Rd, Ik, Ic)%>%
  
  ggplot(., aes(x = site, y = Ik, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("Ik (PAR)"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      legend.position="none",
      strip.text.x=element_text(face="italic", size=14)
      );Ik_plot_box
```

View by means
```{r}
Ik_plot<-master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Ik)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Ik, na.rm=TRUE), se=sd(Ik, na.rm=TRUE)/sqrt(length(Ik)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 800)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("I" [K], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ik_plot
```

View species means 
```{r}
master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Ik)%>%
  group_by(species)%>%
  summarise(mean=mean(Ik, na.rm=TRUE), se=sd(Ik, na.rm=TRUE)/sqrt(length(Ik)))

```

### Plot: Ic  

View by site and timepoint for each species.    
```{r}
Ic_plot_box<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species)) %>%
  select(colony_id_corr, species, site, timepoint, AQY, Am, Rd, Ik, Ic)%>%
  filter(Ic<1000)%>% #remove outliers
  
  ggplot(., aes(x = site, y = Ic, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    ylab(expression(bold(paste("Ic (PAR)"))))+
    theme_classic() + 
    theme(
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );Ic_plot_box
```

View by means
```{r}
Ic_plot<-master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Ic)%>%
  filter(Ic<1000)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(Ic, na.rm=TRUE), se=sd(Ic, na.rm=TRUE)/sqrt(length(Ic)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 800)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("I" [C], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ic_plot
```

View species means. 

```{r}
master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, Ic)%>%
  group_by(species)%>%
  summarise(mean=mean(Ic, na.rm=TRUE), se=sd(Ic, na.rm=TRUE)/sqrt(length(Ic)))

```

Join all plots together.  

```{r}
pi_figure<-plot_grid(am_plot_box, aqy_plot_box, rd_plot_box, Ik_plot_box, Ic_plot_box, labels = c("", "", "", "", ""), ncol=5, nrow=1, rel_widths = c(.8, .8, .8, .8, 1), label_size = 20, label_x = 0.1)

ggsave(filename="Figures/Univariate/PI_Figure.pdf", plot=pi_figure, dpi=300, width=30, height=5, units="in")
```

### Analysis of Am   

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
am_model<-lmer(Am~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(am_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
am_model<-lmer(log(Am+1)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(am_model))
```

Generate a Type III Anova of model.    

```{r}
anova(am_model, type="III")
summary(am_model)

emm<-emmeans(am_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
am_model_acr<-lmer(log(Am+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(am_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(am_model_acr, type="III")
summary(am_model_acr)
rand(am_model_acr)

emm<-emmeans(am_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(am_model_acr, ~ site)
pairs(emm)
```

*Pocillopora*  

```{r}
am_model_poc<-lmer(log(Am+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(am_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(am_model_poc, type="III")
summary(am_model_poc)
rand(am_model_poc)

emm<-emmeans(am_model_poc, ~ timepoint)
pairs(emm)
```

*Porites*  

```{r}
am_model_por<-lmer(log(Am+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(am_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(am_model_por, type="III")
summary(am_model_por)
rand(am_model_por)

emm<-emmeans(am_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(am_model_por, ~ site)
pairs(emm)
```

### Analysis of AQY   

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
aqy_model<-lmer(AQY~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(aqy_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
aqy_model<-lmer(log(AQY+1)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(aqy_model))
```

Generate a Type II Anova of model.    

```{r}
anova(aqy_model, type="II")
summary(aqy_model)

emm<-emmeans(aqy_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
aqy_model_acr<-lmer(log(AQY+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(aqy_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(aqy_model_acr, type="III")
summary(aqy_model_acr)
rand(aqy_model_acr)
```

*Pocillopora*  

```{r}
aqy_model_poc<-lmer(log(AQY+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(aqy_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(aqy_model_poc, type="III")
summary(aqy_model_poc)
rand(aqy_model_poc)

emm<-emmeans(aqy_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(aqy_model_poc, ~ site)
pairs(emm)
```

*Porites*  

```{r}
aqy_model_por<-lmer(log(AQY+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(aqy_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(aqy_model_por, type="III")
summary(aqy_model_por)
rand(aqy_model_por)

emm<-emmeans(aqy_model_por, ~ timepoint)
pairs(emm)
```

### Analysis of Rd   

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
rd_model<-lmer(Rd~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(rd_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
rd_model<-lmer(log(Rd+1)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(rd_model))
```

Generate a Type II Anova of model.    

```{r}
anova(rd_model, type="III")
summary(rd_model)

#emm<-emmeans(rd_model, ~ species)
#pairs(emm)
```

*Acropora*  

```{r}
rd_model_acr<-lmer(log(Rd+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(rd_model_acr))
```

Generate a Type II Anova of model.    

```{r}
anova(rd_model_acr, type="III")
summary(rd_model_acr)
rand(rd_model_acr)

emm<-emmeans(rd_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(rd_model_acr, ~ site)
pairs(emm)
```

*Pocillopora*  

```{r}
rd_model_poc<-lmer(log(Rd+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(rd_model_poc))
```

Generate a Type II Anova of model.    

```{r}
anova(rd_model_poc, type="III")
summary(rd_model_poc)
rand(rd_model_poc)

emm<-emmeans(rd_model_poc, ~ timepoint | site)
pairs(emm)

emm<-emmeans(rd_model_poc, ~ site)
pairs(emm)

emm<-emmeans(rd_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(rd_model_poc, ~ site | timepoint)
pairs(emm)
```

*Porites*  

```{r}
rd_model_por<-lmer(log(Rd+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(rd_model_por))
```

Generate a Type II Anova of model.    

```{r}
anova(rd_model_por, type="III")
summary(rd_model_por)
rand(rd_model_por)

emm<-emmeans(rd_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(rd_model_por, ~ site)
pairs(emm)
```

### Analysis of Ik   

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
Ik_model<-lmer(Ik~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(Ik_model))
```

Residuals are not normally distributed. Attempt with log transformation.      
 
```{r}
Ik_model<-lmer(log(Ik+1)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(Ik_model))
```

Generate a Type II Anova of model.    

```{r}
anova(Ik_model, type="III")
summary(Ik_model)

emm<-emmeans(Ik_model, ~ species)
pairs(emm)
```

Since species is significant, test for effects of site and time within species.  

*Acropora*  

```{r}
Ik_model_acr<-lmer(log(Ik+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(Ik_model_acr))
```

Generate a Type II Anova of model.    

```{r}
anova(Ik_model_acr, type="III")
summary(Ik_model_acr)
rand(Ik_model_acr)

emm<-emmeans(Ik_model_acr, ~ timepoint)
pairs(emm)
```

*Pocillopora*  

```{r}
Ik_model_poc<-lmer(log(Ik+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(Ik_model_poc))
```

Generate a Type II Anova of model.    

```{r}
anova(Ik_model_poc, type="III")
summary(Ik_model_poc)
rand(Ik_model_poc)

emm<-emmeans(Ik_model_poc, ~ timepoint)
pairs(emm)
```

*Porites*  

```{r}
Ik_model_por<-lmer(log(Ik+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(Ik_model_por))
```

Generate a Type II Anova of model.    

```{r}
anova(Ik_model_por, type="III")
summary(Ik_model_por)
rand(Ik_model_por)

emm<-emmeans(Ik_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(Ik_model_por, ~ site)
pairs(emm)

emm<-emmeans(Ik_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(Ik_model_por, ~ site | timepoint)
pairs(emm)
```

### Analysis of Ic   

Build a mixed model for univariate analysis and examine data distribution. 

```{r}
Ic_model<-lmer(Ic~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(Ic_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
Ic_model<-lmer(log(Ic+1)~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(Ic_model))
```

Generate a Type II Anova of model.    

```{r}
anova(Ic_model, type="III")
summary(Ic_model)

emm<-emmeans(Ic_model, ~ species)
pairs(emm)
```

*Acropora*  

```{r}
Ic_model_acr<-lmer(log(Ic+1)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(Ic_model_acr))
```

Generate a Type II Anova of model.    

```{r}
anova(Ic_model_acr, type="III")
summary(Ic_model_acr)
rand(Ic_model_acr)

emm<-emmeans(Ic_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(Ic_model_acr, ~ site)
pairs(emm)
```

*Pocillopora* 

```{r}
Ic_model_poc<-lmer(log(Ic+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(Ic_model_poc))
```

Generate a Type II Anova of model.    

```{r}
anova(Ic_model_poc, type="III")
summary(Ic_model_poc)
rand(Ic_model_poc)

emm<-emmeans(Ic_model_poc, ~ timepoint)
pairs(emm)

emm<-emmeans(Ic_model_poc, ~ site | timepoint)
pairs(emm)

emm<-emmeans(Ic_model_poc, ~ timepoint | site)
pairs(emm)
```

*Porites* 

```{r}
Ic_model_por<-lmer(log(Ic+1)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(Ic_model_por))
```

Generate a Type II Anova of model.    

```{r}
anova(Ic_model_por, type="III")
summary(Ic_model_por)
rand(Ic_model_por)

emm<-emmeans(Ic_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(Ic_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(Ic_model_por, ~ site | timepoint)
pairs(emm)
```

## Calcification    

### Plot: Calcification mgprot 

View by site and timepoint for each species.   
```{r}
calcplot_prot<-master %>%
  filter(!is.na(calc.umol.mgprot.hr)) %>% 
  filter(calc.umol.mgprot.hr<3)%>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, calc.umol.mgprot.hr)%>%
  
  ggplot(., aes(x = site, y = log(1+calc.umol.mgprot.hr), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Calcification (", mu,"mol Ca", CO[3], " mg protein"^-1, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );calcplot_prot
```

### Plot: Calcification mgafdw 

View by site and timepoint for each species.   
```{r}
calcplot_afdw<-master %>%
  filter(!is.na(calc.umol.mgAFDW.hr)) %>% 
  filter(calc.umol.mgAFDW.hr>-.06)%>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, calc.umol.mgAFDW.hr)%>%
  
  ggplot(., aes(x = site, y = log(1+calc.umol.mgAFDW.hr), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Calcification (", mu,"mol Ca", CO[3], " mg protein"^-1, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );calcplot_afdw
```

### Plot: Calcification umol cm2 

View by site and timepoint for each species.   
```{r}
calcplot_cm2_box<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>% 
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, timepoint, calc.umol.cm2.hr)%>%
  
  ggplot(., aes(x = site, y = log(1+calc.umol.cm2.hr), fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    scale_fill_manual(values = brewer.pal(4,"YlGnBu"), limits=c("timepoint1", "timepoint2", "timepoint3", "timepoint4"), labels=c("January 2020", "March 2020", "September 2020", "November 2020"))+
    labs(fill = "Month") +
    facet_wrap(~species) +
    xlab("Site") + 
    scale_x_discrete(labels=c("Mahana Low" = "Mahana \nLow", "Hilton Medium" = "Hilton \nMedium",
                              "Manava High" = "Manava \nHigh"))+
    ylab(expression(bold(paste("log Calcification (", mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );calcplot_cm2_box
```

View by means
```{r}
calcplot_cm2<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>%
  filter(calc.umol.cm2.hr>0)%>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, site, month, calc.umol.cm2.hr)%>%
  
  group_by(species, site, month)%>%
  summarise(mean=mean(calc.umol.cm2.hr, na.rm=TRUE), se=sd(calc.umol.cm2.hr, na.rm=TRUE)/sqrt(length(calc.umol.cm2.hr)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = site, group=interaction(site,month))) +
  geom_line(aes(group=site, colour=site), position=position_dodge(0.5), size=1, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.5)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.5))+
  
    labs(fill = "Site") +
    facet_wrap(~species) +
  
    scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
    scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  
    ylim(0, 1.5)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Calcification (", mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );calcplot_cm2
```

Join all plots for each normalization together.  

```{r}
calc_figure<-plot_grid(calcplot_prot, calcplot_afdw, calcplot_cm2_box, labels = c("", "", ""), ncol=3, nrow=1, rel_widths = c(.8, .8, 1), label_size = 20, label_x = 0.1);calc_figure

ggsave(filename="Figures/Univariate/Calcification_Figure.pdf", plot=calc_figure, dpi=300, width=24, height=5, units="in")
```

### Analysis 

Build a mixed model for univariate analysis and examine data distribution of surface area normalized calcification.. 

```{r}
calc_model<-lmer(calc.umol.mgAFDW.hr~timepoint*species*site+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(calc_model))
#hist(residuals(antiox_model))
```

Residuals are not normally distributed. Attempt with log transformation.      

```{r}
calc_model<-lmer(log(1+calc.umol.cm2.hr)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=master)
qqPlot(residuals(calc_model))
#hist(residuals(antiox_model))
```

Generate a Type III Anova of model.    

```{r}
anova(calc_model, type="III")
summary(calc_model)

emm<-emmeans(calc_model, ~ species)
pairs(emm)
```

*Acropora*:

```{r}
calc_model_acr<-lmer(log(1+calc.umol.cm2.hr)~timepoint*site+(1|colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Acropora"))
qqPlot(residuals(calc_model_acr))
```

Generate a Type III Anova of model.    

```{r}
anova(calc_model_acr, type="III")
summary(calc_model_acr)
rand(calc_model_acr)

emm<-emmeans(calc_model_acr, ~ timepoint)
pairs(emm)

emm<-emmeans(calc_model_acr, ~ timepoint | site)
pairs(emm)

emm<-emmeans(calc_model_acr, ~ site | timepoint)
pairs(emm)
```

*Pocillopora*: 

```{r}
calc_model_poc<-lmer(log(1+calc.umol.cm2.hr)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Pocillopora"))
qqPlot(residuals(calc_model_poc))
#hist(residuals(calc_model_poc))
```

Generate a Type III Anova of model.    

```{r}
anova(calc_model_poc, type="III")
summary(calc_model_poc)
rand(calc_model_poc)

emm<-emmeans(calc_model_poc, ~ timepoint)
pairs(emm)

rand(calc_model_poc)
```

*Porites*:  

```{r}
calc_model_por<-lmer(log(1+calc.umol.cm2.hr)~timepoint*site+(1|haplotype/colony_id_corr), na.action=na.omit, data=master, subset=c(species=="Porites"))
qqPlot(residuals(calc_model_por))
#hist(residuals(calc_model_por))
```

Generate a Type III Anova of model.    

```{r}
anova(calc_model_por, type="III")
summary(calc_model_por)
rand(calc_model_por)

emm<-emmeans(calc_model_por, ~ timepoint)
pairs(emm)

emm<-emmeans(calc_model_por, ~ site)
pairs(emm)

emm<-emmeans(calc_model_por, ~ timepoint | site)
pairs(emm)

emm<-emmeans(calc_model_por, ~ site | timepoint)
pairs(emm)
```

## Generating univariate response panels  

First, specify symbiont and holobiont responses for plotting in separate panels.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ratio_AFDW.mg.cm2", "Ik", "Ic") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

Generate a panel of responses for symbiont metrics.  

```{r}
#build a plot for the legend
legend_plot<-chlplot_total_cell+theme(legend.position="bottom")

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot + theme(legend.box.margin = margin(1,1,1,1), legend.position="right")
)

Ic_plot<-Ic_plot + theme(legend.position="none")

symbplot_afdw<-symbplot_afdw+theme(axis.text.x=element_blank())
sAFDWplot<-sAFDWplot+theme(axis.text.x=element_blank())
am_plot<-am_plot+theme(axis.text.x=element_blank())
aqy_plot<-aqy_plot+theme(axis.text.x=element_blank())
Ik_plot<-Ik_plot+theme(axis.text.x=element_blank())
Ic_plot<-Ic_plot+theme(axis.text.x=element_blank())

symbiont_univ<-plot_grid(symbplot_afdw, sAFDWplot, am_plot, aqy_plot, Ik_plot, Ic_plot, chlplot_total, chlplot_total_cell, ratioAFDWplot, ncol=3, nrow=3, rel_widths = c(1,1,1), rel_heights = c(0.8, 0.8, 1), align="v", label_y=1, labels=c("A", "B", "C", "D", "E", "F", "G", "H", "I"), label_size=20)

symbiont_univ2<-plot_grid(symbiont_univ, legend, ncol=2, nrow=1, rel_widths = c(1, .1))

ggsave(filename="Figures/Univariate/Symbiont_Univariate_Panel.png", plot=symbiont_univ2, dpi=150, width=20, height=15, units="in")
```
 
Generate a panel of responses for host/holobiont metrics.  

```{r}
#build a plot for the legend
legend_plot<-calcplot_cm2+theme(legend.position="bottom")
rd_plot<-rd_plot+theme(legend.position="none")
calcplot_cm2<-calcplot_cm2+theme(legend.position="none") 

tacplot_afdw<-tacplot_afdw+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
hAFDWplot<-hAFDWplot+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
protplot_afdw<-protplot_afdw+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
rd_plot<-rd_plot+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
calcplot_cm2<-calcplot_cm2+ theme(axis.title.y=element_text(size=12))
 
host_univ<-plot_grid(tacplot_afdw, hAFDWplot, protplot_afdw, rd_plot, calcplot_cm2, ncol=2, nrow=3, align="vh", label_y=1, labels=c("A", "B", "C", "D", "E"), label_size=20)

host_univ2<-plot_grid(host_univ, legend, ncol=2, nrow=1, rel_widths = c(1, .2))

ggsave(filename="Figures/Univariate/Host_Univariate_Panel.png", plot=host_univ2, dpi=150, width=14, height=14, units="in")
```

# Generate a panel of responses for Porites haplotypes 

## Antioxidant capacity  
  
View by means
```{r}
tacplot_afdw_haplotype<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, cre.umol.mgafdw)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(cre.umol.mgafdw, na.rm=TRUE), se=sd(cre.umol.mgafdw, na.rm=TRUE)/sqrt(length(cre.umol.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    xlab("") + 
    ylab(expression(bold(paste("Antioxidant Capacity (", mu, "mol CRE mg afdw"^-1,")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );tacplot_afdw_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, cre.umol.mgafdw)%>%
  
  aov(cre.umol.mgafdw ~ haplotype * month, data=.)

summary(model)
```

Analyze effect on Pocillopora 
```{r}
model<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, cre.umol.mgafdw)%>%
  
  aov(cre.umol.mgafdw ~ haplotype * month, data=.)

summary(model)
```

## Symbiont Densities  

### Plot: Biomass   

View by means
```{r}
symbplot_afdw_haplotype<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, cells.mgAFDW)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(cells.mgAFDW, na.rm=TRUE), se=sd(cells.mgAFDW, na.rm=TRUE)/sqrt(length(cells.mgAFDW)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(1e+04, 6e+05)+
  
    xlab("") + 
    ylab(expression(bold(paste("Symbiont Density (cells mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );symbplot_afdw_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, cells.mgAFDW)%>%
  
  aov(cells.mgAFDW ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, cells.mgAFDW)%>%
  
  aov(cells.mgAFDW ~ haplotype * month, data=.)

summary(model)
```

## Biomass

### Plot Host Biomass  

View by means
```{r}
hAFDWplot_haplotype<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Host_AFDW.mg.cm2)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Host_AFDW.mg.cm2, na.rm=TRUE), se=sd(Host_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Host_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 10)+
  
    xlab("") + 
    ylab(expression(bold(paste("Host Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );hAFDWplot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Host_AFDW.mg.cm2)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  
  aov(Host_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Host_AFDW.mg.cm2)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  
  aov(Host_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

### Plot Symbiont biomass  

View by means
```{r}
sAFDWplot_haplotype<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Sym_AFDW.mg.cm2)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Sym_AFDW.mg.cm2, na.rm=TRUE), se=sd(Sym_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Sym_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 5)+
  
    xlab("") + 
    ylab(expression(bold(paste("Symbiont Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );sAFDWplot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Sym_AFDW.mg.cm2)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  
  aov(Sym_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Sym_AFDW.mg.cm2)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Pocillopora")%>%
  
  aov(Sym_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

### Plot S:H Ratio  

View by means
```{r}
ratioAFDWplot_haplotype<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Ratio_AFDW.mg.cm2)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ratio_AFDW.mg.cm2, na.rm=TRUE), se=sd(Ratio_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Ratio_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) +  
  
    ylim(0, 0.6)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Symbiont : Host Biomass"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );ratioAFDWplot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Ratio_AFDW.mg.cm2)%>%
  
  aov(Ratio_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Ratio_AFDW.mg.cm2)%>%
  
  aov(Ratio_AFDW.mg.cm2 ~ haplotype * month, data=.)

summary(model)
```

## Chlorophyll  

### Plot: Total chl  

View by means
```{r}
chlplot_total_haplotype<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl>0)%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Total_Chl)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Total_Chl, na.rm=TRUE), se=sd(Total_Chl, na.rm=TRUE)/sqrt(length(Total_Chl)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 7)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl>0)%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Total_Chl)%>%
  
  aov(Total_Chl ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl>0)%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Total_Chl)%>%
  
  aov(Total_Chl ~ haplotype * month, data=.)

summary(model)
```

View by means
```{r}
chlplot_total_cell_haplotype<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl_cell>0)%>%
  filter(species=="Porites")%>%
  
  select(colony_id_corr, species, haplotype, month, Total_Chl_cell)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Total_Chl_cell, na.rm=TRUE), se=sd(Total_Chl_cell, na.rm=TRUE)/sqrt(length(Total_Chl_cell)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) +  
  
    ylim(0, 3e-05)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total_cell_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl_cell>0)%>%
  filter(species=="Porites")%>%
  
  aov(Total_Chl_cell ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl_cell>0)%>%
  filter(species=="Pocillopora")%>%
  
  aov(Total_Chl_cell ~ haplotype * month, data=.)

summary(model)
```

## Protein 

### Plot: AFDW  

View by means
```{r}
protplot_afdw_haplotype<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, prot_mg.mgafdw)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw>0)%>%
  filter(prot_mg.mgafdw<1)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(prot_mg.mgafdw, na.rm=TRUE), se=sd(prot_mg.mgafdw, na.rm=TRUE)/sqrt(length(prot_mg.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
  geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +

    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) +  
  
    ylim(0, 0.75)+
  
    xlab("") + 
    ylab(expression(bold(paste("Protein (mg mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );protplot_afdw_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, prot_mg.mgafdw)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw>0)%>%
  filter(prot_mg.mgafdw<1)%>%
  
  aov(prot_mg.mgafdw ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, prot_mg.mgafdw)%>%
  filter(species=="Pocillopora")%>%
  filter(prot_mg.mgafdw>0)%>%
  filter(prot_mg.mgafdw<1)%>%
  
  aov(prot_mg.mgafdw ~ haplotype * month, data=.)

summary(model)
```

## PI Curve Parameters   

### Plot: Am   

View by means
```{r}
am_plot_haplotype<-master %>%
  filter(!is.na(Am)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Am)%>%
  filter(species=="Porites")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Am, na.rm=TRUE), se=sd(Am, na.rm=TRUE)/sqrt(length(Am)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
  geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 4)+
  
    xlab("") + 
    ylab(expression(bold(paste("P" [Max], " (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );am_plot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Am)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Am)%>%
  filter(species=="Porites")%>%
  
  aov(Am ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Am)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Am)%>%
  filter(species=="Pocillopora")%>%
  
  aov(Am ~ haplotype * month, data=.)

summary(model)
```

### Plot: AQY  

View by means
```{r}
aqy_plot_haplotype<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, AQY)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(AQY, na.rm=TRUE), se=sd(AQY, na.rm=TRUE)/sqrt(length(AQY)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 0.02)+
  
    xlab("") + 
    ylab(expression(bold(paste("Apparent Quantum Yield (AQY)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );aqy_plot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, AQY)%>%
  
  aov(AQY ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, AQY)%>%
  
  aov(AQY ~ haplotype * month, data=.)

summary(model)
```

### Plot: Rd  

View by means
```{r}
rd_plot_haplotype<-master %>%
  filter(!is.na(Rd)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Rd)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Rd, na.rm=TRUE), se=sd(Rd, na.rm=TRUE)/sqrt(length(Rd)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 2)+
  
    xlab("") + 
    ylab(expression(bold(paste("Respiration (R" [D], ") (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );rd_plot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Rd)) %>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, Rd)%>%
  
  aov(Rd ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(Rd)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Rd)%>%
  
  aov(Rd ~ haplotype * month, data=.)

summary(model)
```

### Plot: Ik  

View by means
```{r}
Ik_plot_haplotype<-master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ik)%>%
  filter(species=="Porites")%>%
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ik, na.rm=TRUE), se=sd(Ik, na.rm=TRUE)/sqrt(length(Ik)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 800)+
  
    xlab("") + 
    ylab(expression(bold(paste("I" [K], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ik_plot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ik)%>%
  filter(species=="Porites")%>%
  
  aov(Ik ~ haplotype * month, data=.)

summary(model)
```

```{r}
model<-master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ik)%>%
  filter(species=="Pocillopora")%>%
  
  aov(Ik ~ haplotype * month, data=.)

summary(model)
```

### Plot: Ic  

View by means
```{r}
Ic_plot_haplotype<-master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ic)%>%
  filter(species=="Porites")%>%
  filter(Ic<1000)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ic, na.rm=TRUE), se=sd(Ic, na.rm=TRUE)/sqrt(length(Ic)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 800)+
  
    xlab("") + 
    ylab(expression(bold(paste("I" [C], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ic_plot_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ic)%>%
  filter(species=="Porites")%>%
  filter(Ic<1000)%>%
  
  aov(Ic ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ic)%>%
  filter(species=="Pocillopora")%>%
  filter(Ic<1000)%>%
  
  aov(Ic ~ haplotype * month, data=.)

summary(model)
```

## Calcification    

### Plot: Calcification umol cm2 

View by means
```{r}
calcplot_cm2_haplotype<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>%
  filter(calc.umol.cm2.hr>0)%>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, calc.umol.cm2.hr)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(calc.umol.cm2.hr, na.rm=TRUE), se=sd(calc.umol.cm2.hr, na.rm=TRUE)/sqrt(length(calc.umol.cm2.hr)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 2)+
  
    xlab("") + 
    ylab(expression(bold(paste("Calcification (", mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );calcplot_cm2_haplotype
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>%
  filter(calc.umol.cm2.hr>0)%>%
  filter(!is.na(species))%>%
  filter(species=="Porites")%>%
  select(colony_id_corr, species, haplotype, month, calc.umol.cm2.hr)%>%
  
  aov(calc.umol.cm2.hr ~ haplotype * month, data=.)

summary(model)
```

Analyze for effect of haplotype. 
```{r}
model<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>%
  filter(calc.umol.cm2.hr>0)%>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, calc.umol.cm2.hr)%>%
  
  aov(calc.umol.cm2.hr ~ haplotype * month, data=.)

summary(model)
```

## Generating univariate response panels  

First, specify symbiont and holobiont responses for plotting in separate panels.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ratio_AFDW.mg.cm2", "Ik", "Ic") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

Generate a panel of responses for symbiont metrics.  

```{r}
#build a plot for the legend
legend_plot_haplotype<-chlplot_total_cell_haplotype+theme(legend.position="bottom")

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1,1,1,1), legend.position="right")
)

Ic_plot<-Ic_plot + theme(legend.position="none")

symbplot_afdw_haplotype<-symbplot_afdw_haplotype+theme(axis.text.x=element_blank())
sAFDWplot_haplotype<-sAFDWplot_haplotype+theme(axis.text.x=element_blank())
am_plot_haplotype<-am_plot_haplotype+theme(axis.text.x=element_blank())
aqy_plot_haplotype<-aqy_plot_haplotype+theme(axis.text.x=element_blank())
Ik_plot_haplotype<-Ik_plot_haplotype+theme(axis.text.x=element_blank())
Ic_plot_haplotype<-Ic_plot_haplotype+theme(axis.text.x=element_blank())

symbiont_univ_haplotype<-plot_grid(symbplot_afdw_haplotype, sAFDWplot_haplotype, am_plot_haplotype, aqy_plot_haplotype, Ik_plot_haplotype, Ic_plot_haplotype, chlplot_total_haplotype, chlplot_total_cell_haplotype, ratioAFDWplot_haplotype, ncol=3, nrow=3, rel_widths = c(1,1,1), rel_heights = c(0.8, 0.8, 1), align="v", label_y=1, labels=c("A", "B", "C", "D", "E", "F", "G", "H", "I"), label_size=20)

symbiont_univ2_haplotype<-plot_grid(symbiont_univ_haplotype, legend_haplotype, ncol=2, nrow=1, rel_widths = c(1, .2))

ggsave(filename="Figures/Univariate/Symbiont_Univariate_Panel_HaplotypePOR.png", plot=symbiont_univ2_haplotype, dpi=150, width=13, height=15, units="in")
```
 
Generate a panel of responses for host/holobiont metrics.  

```{r}
#build a plot for the legend
legend_plot_haplotype<-calcplot_cm2_haplotype+theme(legend.position="bottom")
rd_plot_haplotype<-rd_plot_haplotype+theme(legend.position="none")
calcplot_cm2_haplotype<-calcplot_cm2_haplotype+theme(legend.position="none") 

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1.5,1.5,1.5,1.5), legend.position="right")
)

tacplot_afdw_haplotype<-tacplot_afdw_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
hAFDWplot_haplotype<-hAFDWplot_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
protplot_afdw_haplotype<-protplot_afdw_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
rd_plot_haplotype<-rd_plot_haplotype+theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
calcplot_cm2_haplotype<-calcplot_cm2_haplotype+ theme(axis.title.y=element_text(size=12))
 
host_univ_haplotype<-plot_grid(tacplot_afdw_haplotype, hAFDWplot_haplotype, protplot_afdw_haplotype, rd_plot_haplotype, calcplot_cm2_haplotype, ncol=3, nrow=2, align="vh", label_y=1, labels=c("A", "B", "C", "D", "E"), label_size=20)

host_univ2_haplotype<-plot_grid(host_univ_haplotype, legend_haplotype, ncol=2, nrow=1, rel_widths = c(1, .5))

ggsave(filename="Figures/Univariate/Host_Univariate_Panel_HaplotypePOR.png", plot=host_univ2_haplotype, dpi=150, width=18, height=11, units="in")
```

# Generate a panel of responses for Pocillopora haplotypes 

## Antioxidant capacity  
  
View by means
```{r}
tacplot_afdw_haplotype<-master %>%
  filter(!is.na(cre.umol.mgafdw)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, cre.umol.mgafdw)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(cre.umol.mgafdw, na.rm=TRUE), se=sd(cre.umol.mgafdw, na.rm=TRUE)/sqrt(length(cre.umol.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0,0.1)+
    xlab("") + 
    ylab(expression(bold(paste("Antioxidant Capacity (", mu, "mol CRE mg afdw"^-1,")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );tacplot_afdw_haplotype
```

## Symbiont Densities  

### Plot: Biomass   

View by means
```{r}
symbplot_afdw_haplotype<-master %>%
  filter(!is.na(cells.mgAFDW)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, cells.mgAFDW)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(cells.mgAFDW, na.rm=TRUE), se=sd(cells.mgAFDW, na.rm=TRUE)/sqrt(length(cells.mgAFDW)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(1e+04, 6e+05)+
  
    xlab("") + 
    ylab(expression(bold(paste("Symbiont Density (cells mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );symbplot_afdw_haplotype
```

## Biomass

### Plot Host Biomass  

View by means
```{r}
hAFDWplot_haplotype<-master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Host_AFDW.mg.cm2)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Host_AFDW.mg.cm2, na.rm=TRUE), se=sd(Host_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Host_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 10)+
  
    xlab("") + 
    ylab(expression(bold(paste("Host Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );hAFDWplot_haplotype
```


### Plot Symbiont biomass  

View by means
```{r}
sAFDWplot_haplotype<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Sym_AFDW.mg.cm2)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Pocillopora")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Sym_AFDW.mg.cm2, na.rm=TRUE), se=sd(Sym_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Sym_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+  
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 5)+
  
    xlab("") + 
    ylab(expression(bold(paste("Symbiont Biomass (mg afdw cm"^-2, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );sAFDWplot_haplotype
```


### Plot S:H Ratio  

View by means
```{r}
ratioAFDWplot_haplotype<-master %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(species))%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Ratio_AFDW.mg.cm2)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ratio_AFDW.mg.cm2, na.rm=TRUE), se=sd(Ratio_AFDW.mg.cm2, na.rm=TRUE)/sqrt(length(Ratio_AFDW.mg.cm2)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
     scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 0.6)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Symbiont : Host Biomass"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );ratioAFDWplot_haplotype
```

## Chlorophyll  

### Plot: Total chl  

View by means
```{r}
chlplot_total_haplotype<-master %>%
  filter(!is.na(Total_Chl)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl>0)%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Total_Chl)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Total_Chl, na.rm=TRUE), se=sd(Total_Chl, na.rm=TRUE)/sqrt(length(Total_Chl)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 7)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total_haplotype
```


View by means
```{r}
chlplot_total_cell_haplotype<-master %>%
  filter(!is.na(Total_Chl_cell)) %>%
  filter(!is.na(species))%>%
  filter(Total_Chl_cell>0)%>%
  filter(species=="Pocillopora")%>%
  
  select(colony_id_corr, species, haplotype, month, Total_Chl_cell)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Total_Chl_cell, na.rm=TRUE), se=sd(Total_Chl_cell, na.rm=TRUE)/sqrt(length(Total_Chl_cell)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 3e-05)+
  
    xlab("Month") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cell"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );chlplot_total_cell_haplotype
```

## Protein 

### Plot: AFDW  

View by means
```{r}
protplot_afdw_haplotype<-master %>%
  filter(!is.na(prot_mg.mgafdw)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, prot_mg.mgafdw)%>%
  filter(species=="Pocillopora")%>%
  filter(prot_mg.mgafdw>0)%>%
  filter(prot_mg.mgafdw<1)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(prot_mg.mgafdw, na.rm=TRUE), se=sd(prot_mg.mgafdw, na.rm=TRUE)/sqrt(length(prot_mg.mgafdw)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
  geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +

    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 0.75)+
  
    xlab("") + 
    ylab(expression(bold(paste("Protein (mg mg afdw"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );protplot_afdw_haplotype
```


## PI Curve Parameters   

### Plot: Am   

View by means
```{r}
am_plot_haplotype<-master %>%
  filter(!is.na(Am)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Am)%>%
  filter(species=="Pocillopora")%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Am, na.rm=TRUE), se=sd(Am, na.rm=TRUE)/sqrt(length(Am)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
  geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 4)+
  
    xlab("") + 
    ylab(expression(bold(paste("P" [Max], " (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );am_plot_haplotype
```

### Plot: AQY  

View by means
```{r}
aqy_plot_haplotype<-master %>%
  filter(!is.na(AQY)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, AQY)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(AQY, na.rm=TRUE), se=sd(AQY, na.rm=TRUE)/sqrt(length(AQY)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 0.02)+
  
    xlab("") + 
    ylab(expression(bold(paste("Apparent Quantum Yield (AQY)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );aqy_plot_haplotype
```

### Plot: Rd  

View by means
```{r}
rd_plot_haplotype<-master %>%
  filter(!is.na(Rd)) %>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, Rd)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Rd, na.rm=TRUE), se=sd(Rd, na.rm=TRUE)/sqrt(length(Rd)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 2)+
  
    xlab("") + 
    ylab(expression(bold(paste("Respiration (R" [D], ") (mol O" [2], " cm"^-2, " h"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );rd_plot_haplotype
```

### Plot: Ik  

View by means
```{r}
Ik_plot_haplotype<-master %>%
  filter(!is.na(Ik)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ik)%>%
  filter(species=="Pocillopora")%>%
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ik, na.rm=TRUE), se=sd(Ik, na.rm=TRUE)/sqrt(length(Ik)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 800)+
  
    xlab("") + 
    ylab(expression(bold(paste("I" [K], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ik_plot_haplotype
```

### Plot: Ic  

View by means
```{r}
Ic_plot_haplotype<-master %>%
  filter(!is.na(Ic)) %>%
  filter(!is.na(species))%>%
  select(colony_id_corr, species, haplotype, month, Ic)%>%
  filter(species=="Pocillopora")%>%
  filter(Ic<1000)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(Ic, na.rm=TRUE), se=sd(Ic, na.rm=TRUE)/sqrt(length(Ic)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 800)+
  
    xlab("") + 
    ylab(expression(bold(paste("I" [C], " (PAR)"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );Ic_plot_haplotype
```


## Calcification    

### Plot: Calcification umol cm2 

View by means
```{r}
calcplot_cm2_haplotype<-master %>%
  filter(!is.na(calc.umol.cm2.hr)) %>%
  filter(calc.umol.cm2.hr>0)%>%
  filter(!is.na(species))%>%
  filter(species=="Pocillopora")%>%
  select(colony_id_corr, species, haplotype, month, calc.umol.cm2.hr)%>%
  
  group_by(species, haplotype, month)%>%
  summarise(mean=mean(calc.umol.cm2.hr, na.rm=TRUE), se=sd(calc.umol.cm2.hr, na.rm=TRUE)/sqrt(length(calc.umol.cm2.hr)))%>%
  
  ggplot(., aes(x = month, y = mean, fill = haplotype, group=interaction(haplotype,month))) +
    geom_line(aes(group=haplotype, colour=haplotype), position=position_dodge(0.1), size=2, show.legend=FALSE)+
    geom_point(pch = 21, size=4, position = position_dodge(0.1)) + 
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0, linewidth=0.8, position=position_dodge(0.1))+
  
    labs(fill = "Haplotype") +
  
    scale_colour_manual(values=c("black", "darkgray")) +
    scale_fill_manual(values=c("black", "darkgray")) + 
  
    ylim(0, 2)+
  
    xlab("") + 
    ylab(expression(bold(paste("Calcification (", mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1, ")"))))+
    theme_classic() + 
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1, vjust=1),
      strip.text.x=element_text(face="italic", size=14), 
      panel.border = element_rect(fill=NA) 
      );calcplot_cm2_haplotype
```


## Generating univariate response panels  

First, specify symbiont and holobiont responses for plotting in separate panels.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ratio_AFDW.mg.cm2", "Ik", "Ic") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

Generate a panel of responses for symbiont metrics.  

```{r}
#build a plot for the legend
legend_plot_haplotype<-chlplot_total_cell_haplotype+theme(legend.position="bottom")

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1,1,1,1), legend.position="right")
)

Ic_plot<-Ic_plot_haplotype + theme(legend.position="none")

symbplot_afdw_haplotype<-symbplot_afdw_haplotype+theme(axis.text.x=element_blank())
sAFDWplot_haplotype<-sAFDWplot_haplotype+theme(axis.text.x=element_blank())
am_plot_haplotype<-am_plot_haplotype+theme(axis.text.x=element_blank())
aqy_plot_haplotype<-aqy_plot_haplotype+theme(axis.text.x=element_blank())
Ik_plot_haplotype<-Ik_plot_haplotype+theme(axis.text.x=element_blank())
Ic_plot_haplotype<-Ic_plot_haplotype+theme(axis.text.x=element_blank())

symbiont_univ_haplotype<-plot_grid(symbplot_afdw_haplotype, sAFDWplot_haplotype, am_plot_haplotype, aqy_plot_haplotype, Ik_plot_haplotype, Ic_plot_haplotype, chlplot_total_haplotype, chlplot_total_cell_haplotype, ratioAFDWplot_haplotype, ncol=3, nrow=3, rel_widths = c(1,1,1), rel_heights = c(0.8, 0.8, 1), align="v", label_y=1, labels=c("A", "B", "C", "D", "E", "F", "G", "H", "I"), label_size=20)

symbiont_univ2_haplotype<-plot_grid(symbiont_univ_haplotype, legend_haplotype, ncol=2, nrow=1, rel_widths = c(1, .25))

ggsave(filename="Figures/Univariate/Symbiont_Univariate_Panel_HaplotypePOC.png", plot=symbiont_univ2_haplotype, dpi=150, width=13, height=14, units="in")
```
 
Generate a panel of responses for host/holobiont metrics.  

```{r}
#build a plot for the legend
legend_plot_haplotype<-calcplot_cm2_haplotype+theme(legend.position="bottom")
rd_plot_haplotype<-rd_plot_haplotype+theme(legend.position="none")
calcplot_cm2_haplotype<-calcplot_cm2_haplotype+theme(legend.position="none") 

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1.5,1.5,1.5,1.5), legend.position="right")
)

tacplot_afdw_haplotype<-tacplot_afdw_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
hAFDWplot_haplotype<-hAFDWplot_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
protplot_afdw_haplotype<-protplot_afdw_haplotype+theme(axis.text.x=element_blank()) + theme(axis.title.x=element_blank()) + theme(axis.title.y=element_text(size=12))
rd_plot_haplotype<-rd_plot_haplotype+theme(axis.title.x=element_blank())+ theme(axis.title.y=element_text(size=12))
calcplot_cm2_haplotype<-calcplot_cm2_haplotype+ theme(axis.title.y=element_text(size=12))
 
host_univ_haplotype<-plot_grid(tacplot_afdw_haplotype, hAFDWplot_haplotype, protplot_afdw_haplotype, rd_plot_haplotype, calcplot_cm2_haplotype, ncol=3, nrow=2, align="vh", label_y=1, labels=c("A", "B", "C", "D", "E"), label_size=20)

host_univ2_haplotype<-plot_grid(host_univ_haplotype, legend_haplotype, ncol=2, nrow=1, rel_widths = c(1, .5))

ggsave(filename="Figures/Univariate/Host_Univariate_Panel_HaplotypePOC.png", plot=host_univ2_haplotype, dpi=150, width=18, height=11, units="in")
```

# Correlations  

Generate a correlation matrix to examine relationships between responses of interest. Use normalizations by afdw for all responses.     
 
Generating network plot with corrr package.  
```{r}
master%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  mutate(across(everything(), ~log(. + 1)))%>%# log transform all data 
  corrr::correlate(., method="spearman", diagonal=1)%>% #generate correlation matrix
  rearrange(.)%>%
  #shave(.)%>% #remove upper triangle
  network_plot(min_cor=.2)
```

Generating matrix with ggcorrplot for all colonies of all species.    

```{r}
p<-master%>%
  select(species, cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(!is.na(species))%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(calc.umol.mgAFDW.hr>-.06)%>%
  mutate(across(c(2:17), ~log(. + 1)))%>%# log transform all data 
  ggpairs(., columns=2:17, ggplot2::aes(colour=species)) #try to bold significant stats

#change colors
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("orange", "gray", "purple")) +
        scale_color_manual(values=c("orange", "gray", "purple")) +
        theme_classic()+
        theme(text=element_text(size=14, face="bold"))+
        theme(strip.text.x=element_text(face="bold", size=14))
  }
}

p

ggsave(plot=p, "Figures/Univariate/Correlation_Matrix.pdf", width=20, height=12, unit="in", dpi=300)
```
 
Acropora: 
```{r}
p<-master%>%
  filter(species=="Acropora")%>%
  select(site, cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(calc.umol.mgAFDW.hr>-.06)%>%
  mutate(across(c(2:17), ~log(. + 1)))%>%# log transform all data 
  ggpairs(., columns=2:17, ggplot2::aes(colour=site)) 

#change colors
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        scale_color_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        theme_classic()+
        theme(text=element_text(size=14, face="bold"))+
        theme(strip.text.x=element_text(face="bold", size=14))
  }
}

p

ggsave(plot=p, "Figures/Univariate/Correlation_Matrix_Acropora.pdf", width=20, height=12, unit="in", dpi=300)
```

Pocillopora
```{r}
p<-master%>%
  filter(species=="Pocillopora")%>%
  select(site, cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(calc.umol.mgAFDW.hr>-.06)%>%
  mutate(across(c(2:17), ~log(. + 1)))%>%# log transform all data 
  ggpairs(., columns=2:17, ggplot2::aes(colour=site)) 

#change colors
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        scale_color_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        theme_classic()+
        theme(text=element_text(size=14, face="bold"))+
        theme(strip.text.x=element_text(face="bold", size=14))
  }
}

p

ggsave(plot=p, "Figures/Univariate/Correlation_Matrix_Pocillopora.pdf", width=20, height=12, unit="in", dpi=300)
```

Porites 
```{r}
p<-master%>%
  filter(species=="Porites")%>%
  select(site, cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(calc.umol.mgAFDW.hr>-.06)%>%
  mutate(across(c(2:17), ~log(. + 1)))%>%# log transform all data 
  ggpairs(., columns=2:17, ggplot2::aes(colour=site)) 

#change colors
for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        scale_color_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
        theme_classic()+
        theme(text=element_text(size=14, face="bold"))+
        theme(strip.text.x=element_text(face="bold", size=14))
  }
}

p

ggsave(plot=p, "Figures/Univariate/Correlation_Matrix_Porites.pdf", width=20, height=12, unit="in", dpi=300)
```

Plot the most highly correlated variables across all species to determine which variables to plot over time.

View top 10 correlated variables.  
```{r}
p<-master%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  mutate(across(everything(), ~log(. + 1)))# log transform all data 

p<-na.omit(p)

z<-cor(p)
  
z[lower.tri(z,diag=TRUE)]=NA  #Prepare to drop duplicates and meaningless information
z=as.data.frame(as.table(z))  #Turn into a 3-column table
z=na.omit(z)  #Get rid of the junk we flagged above
z=z[order(-abs(z$Freq)),]    #Sort by highest correlation (whether +ve or -ve)

head(z, 10)
```

## Plotting correlation network 

First view correlation network for a low correlation threshold (r=0.3). 
```{r}
p<-master%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  mutate(across(everything(), ~log(. + 1)))# log transform all data 

p<-na.omit(p)

p %>% correlate() %>%
  network_plot(min_cor=0.3, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)

```

Variables clustered together are more highly correlated. Wider paths indicate a stronger correlation. The color represents the direction of the correlation. 

View correlation network for a medium correlation threshold (r=0.5). 
```{r}
p %>% correlate() %>%
  network_plot(min_cor=0.5, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)
```

View correlation network for a high correlation threshold (r=0.7). 
```{r}
p %>% correlate() %>%
  network_plot(min_cor=0.7, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)
```

The strongest correlations are between metabolic rates and Host/Symbiont biomass and antioxidant capacity. There is also a strong correlation between chlorophyll content and cell density. 

Plot medium threshold for each species independently. 

Acropora
```{r}
p<-master%>%
  filter(species=="Acropora")%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  mutate(across(everything(), ~log(. + 1)))# log transform all data 

p<-na.omit(p)

p %>% correlate() %>%
  network_plot(min_cor=0.5, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)

```

Pocillopora
```{r}
p<-master%>%
  filter(species=="Pocillopora")%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  mutate(across(everything(), ~log(. + 1)))# log transform all data 

p<-na.omit(p)

p %>% correlate() %>%
  network_plot(min_cor=0.5, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)
```

Porites
```{r}
p<-master%>%
  filter(species=="Porites")%>%
  select(cre.umol.mgafdw, prot_mg.mgafdw, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chlc2.ug.mgAFDW, chla.ug.mgAFDW, cells.mgAFDW, chla.ug.cell, chlc2.ug.cell, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  mutate(across(everything(), ~log(. + 1)))# log transform all data 

p<-na.omit(p)

p %>% correlate() %>%
  network_plot(min_cor=0.5, colours = c("skyblue1", "white", "indianred2"), repel = TRUE)
```

# Plotting relative change over time in each species  

## Over time    

Prepare data frame and calculate relative change in each metric compared to January 2020 timepoint. Not including calcification, relative change is far off the scale of other metrics. 

```{r}
timeseries<-master%>%
  select(species, site, month, colony_id_corr, timepoint, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2,  cells.mgAFDW, Total_Chl_cell, Total_Chl, prot_mg.mgafdw, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, Ratio_AFDW.mg.cm2, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #mutate(across(c(6:19), ~log(. + 1)))%>%# log transform all data 
  gather(key="Metric", value="Value", -species, -site, -month, -colony_id_corr, -timepoint)%>%
  group_by(Metric, colony_id_corr, timepoint)%>%
  arrange(colony_id_corr, .by_group = TRUE)%>%
  group_by(colony_id_corr, Metric)%>%
  mutate(rel_change=100*((Value-first(Value))/abs(first(Value))))
  #mutate(impact=(max(Value)/min(Value))*first(Value))

test<-master%>%
    select(species, site, month, colony_id_corr, timepoint, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2,  cells.mgAFDW, Total_Chl_cell, Total_Chl, prot_mg.mgafdw, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, Ratio_AFDW.mg.cm2, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #mutate(across(c(6:19), ~log(. + 1)))%>%# log transform all data 
  gather(key="Metric", value="Value", -species, -site, -month, -colony_id_corr, -timepoint)%>%
  
  mutate(month = as.character(month)) %>%           # ensure comparisons work
  group_by(colony_id_corr, Metric, species) %>%
  # pick baseline at March 2020 (first match if there are duplicates)
  mutate(init_val = Value[month == "March 2020"][1]) %>%
  # drop groups without a March 2020 measurement or with a zero baseline
  filter(!is.na(init_val) & init_val != 0) %>%
  # compute relative deviation from March 2020 for each row
  mutate(rel_dev = (Value - init_val) / init_val) %>%
  summarise(
    init_val     = first(init_val),
    max_dev      = max(rel_dev, na.rm = TRUE),     # largest proportional increase
    min_dev      = min(rel_dev, na.rm = TRUE),     # largest proportional decrease (negative)
    impact_max   = init_val * max_dev,             # absolute cells change for largest gain
    impact_min   = init_val * min_dev,             # absolute cells change for largest loss
    impact_score = max(abs(impact_max), abs(impact_min)), # single-number biggest swing (absolute)
    .groups = "drop"
  )

head(test)
```

Make a color palette. 
```{r}
nb.cols <- 13
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)
```

Plot relative change in each metric over time for all species together.  
```{r}
timeseries$month<-factor(timeseries$month, levels = c("January 2020", "March 2020", "September 2020", "November 2020"))

time_plot1<-timeseries%>%
  filter(!Metric=="calc.umol.cm2.hr")%>%
  group_by(Metric, month)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE))%>%
  
  ggplot(., aes(x=month, y=rel_change, colour=Metric))+
  scale_colour_manual(values=mycolors)+
  geom_line(aes(group=Metric), linewidth=1)+
  geom_hline(yintercept=0)+
  ylab("Percent Change")+
  xlab("Time Point")+
  #geom_smooth(inherit.aes=T, method="lm", aes(group = Metric), se=FALSE)+
  theme_classic(); time_plot1

ggsave(plot=time_plot1, "Figures/Univariate/Relative_Change_AllSpecies.png", width=9, height=6, unit="in", dpi=300)
```

Plot by species.   
```{r}
time_plot2<-timeseries%>%
  filter(!Metric=="calc.umol.cm2.hr")%>%
  group_by(Metric, month, species)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE))%>%
  
  ggplot(., aes(x=month, y=rel_change,  colour=Metric))+
  facet_wrap(~species) +
  scale_colour_manual(values=mycolors)+
  geom_line(aes(group=Metric), linewidth=1)+
  geom_hline(yintercept=0)+
  ylab("Percent Change")+
  xlab("Time Point")+
  #geom_smooth(inherit.aes=T, method="loess", aes(group = Metric), se=FALSE)+
  theme_classic(); time_plot2

ggsave(plot=time_plot2, "Figures/Univariate/Relative_Change_EachSpecies.png", width=15, height=6, unit="in", dpi=300)
```

View percent change. 
```{r}
change_metrics<-timeseries%>%
  filter(!Metric=="calc.umol.cm2.hr")%>%
  group_by(Metric, month, species)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE))%>%
  pivot_wider(names_from=Metric, values_from=rel_change)%>%
  arrange(species, month)%>%
  filter(!month=="January 2020");change_metrics
```

Calcification is not displayed due to relative change being beyond the scale we can display here. 

Display calcification separately here. 

```{r}
time_plot3<-timeseries%>%
  filter(Metric=="calc.umol.cm2.hr")%>%
  filter(!rel_change=="Inf")%>%
  filter(!rel_change=="NaN")%>%
  group_by(Metric, month, species)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE))%>%
  
  ggplot(., aes(x=month, y=rel_change, colour=species))+
  scale_color_manual(values=c("orange", "gray", "purple")) +
  geom_hline(yintercept=0)+
  geom_line(aes(group=species), linewidth=1)+
  ylab("Percent Change")+
  xlab("Time Point")+
  #geom_smooth(inherit.aes=T, method="loess", aes(group = species), se=FALSE)+
  theme_classic(); time_plot3

ggsave(plot=time_plot3, "Figures/Univariate/Relative_Change_Calcification.png", width=9, height=6, unit="in", dpi=300)
```

Run for specific metrics 

```{r}
time_plot4<-timeseries%>%
  filter(Metric %in% c("Host_AFDW.mg.cm2", "Ratio_AFDW.mg.cm2", "Sym_AFDW.mg.cm2", "cells.mgAFDW", "Am"))%>%
  filter(!rel_change=="Inf")%>%
  filter(!rel_change=="NaN")%>%
  group_by(Metric, month, species)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE))%>%
  
  ggplot(., aes(x=month, y=rel_change, colour=species))+
  facet_wrap(~Metric, scales="free")+
  scale_color_manual(values=c("orange", "gray", "purple")) +
  geom_hline(yintercept=0)+
  geom_line(aes(group=species), linewidth=1)+
  ylab("Percent Change")+
  xlab("Time Point")+
  #geom_smooth(inherit.aes=T, method="loess", aes(group = species), se=FALSE)+
  theme_classic(); time_plot4

ggsave(plot=time_plot4, "Figures/Univariate/Relative_Change_Symbiosis.png", width=9, height=6, unit="in", dpi=300)

perc_change<-timeseries%>%
  filter(Metric %in% c("Host_AFDW.mg.cm2", "Ratio_AFDW.mg.cm2", "Sym_AFDW.mg.cm2", "cells.mgAFDW", "Am"))%>%
  filter(!rel_change=="Inf")%>%
  filter(!rel_change=="NaN")%>%
  group_by(Metric, month, species)%>%
  summarise(rel_change=mean(rel_change, na.rm=TRUE));perc_change

write.csv(perc_change, "Output/percent_change_symbiosis_metrics.csv")
```

Plot impact score 

```{r}
impact1<-test%>%
  filter(Metric %in% c("Host_AFDW.mg.cm2", "Ratio_AFDW.mg.cm2", "Sym_AFDW.mg.cm2", "cells.mgAFDW", "Am"))%>%
  filter(!impact_score=="Inf")%>%
  filter(!impact_score=="NaN")%>%
  #group_by(Metric, species)%>%
  #summarise(impact_mean=mean(impact_score, na.rm=TRUE))%>%
  
  ggplot(., aes(x=species, y=impact_score, colour=species))+
  facet_wrap(~Metric, scales="free")+
  scale_color_manual(values=c("orange", "gray", "purple")) +
  geom_boxplot()+
  geom_point()+
  #geom_hline(yintercept=0)+
  #geom_line(aes(group=species), linewidth=1)+
  ylab("Percent Change")+
  xlab("Time Point")+
  #geom_smooth(inherit.aes=T, method="loess", aes(group = species), se=FALSE)+
  theme_classic(); impact1
```


# Run and summarize models for all metrics and generate output table 

```{r}
# Assuming 'master' is your dataframe

# Create a list to store model summaries
model_summaries <- list()

# Loop through each numeric column
for (col_name in names(master)) {
  if (is.numeric(master[[col_name]])) {
    # Perform log+1 transformation
    transformed_data <- log1p(master[[col_name]])
    
    # Fit a linear mixed-effects model
    model <- lmer(transformed_data ~ timepoint * site * species + (1|species/haplotype/colony_id_corr), data = master)  # Change 'Group' to your grouping variable
    
    # Store model summary
    model_summary <- anova(model, type="III")
    model_summaries[[col_name]] <- model_summary
  }
}

# Print or process the model summaries
for (col_name in names(model_summaries)) {
  cat("Summary for", col_name, ":\n")
  print(model_summaries[[col_name]])
  cat("\n")
}

model_summaries

write.csv(model_summaries, file="Output/univariate_model_results.csv")
```

# Plot respiration, biomass, and protein for time series molecular A. pulchra testing 

Make a list of samples used in molecular time series. 

```{r}
list<-c("ACR-225", "ACR-229", "ACR-237", "ACR-244", "AcR-265", "ACR-139", "ACR-145", "ACR-150", "ACR-173", "ACR-186")

apulchra<-master%>%
  filter(colony_id_corr %in% list)
```

Plot biomass. 

```{r}
plot_biomass<-apulchra%>%
  
  ggplot(aes(x=timepoint, y=Host_AFDW.mg.cm2))+
  geom_boxplot()+
  geom_point()+
  theme_classic()+
  ggtitle("Acropora pulchra Molecular Samples")+
  ylim(0, 5)+
  ylab(expression(paste("Host Biomass (mg AFDW cm"^-2, ")")));plot_biomass

ggsave(plot_biomass, filename="Figures/Univariate/apulchra_molecular_biomass.png", width=5, height=4)
```

Plot protein 

```{r}
plot_respiration<-apulchra%>%
  
  ggplot(aes(x=timepoint, y=Rd))+
  geom_boxplot()+
  geom_point()+
  theme_classic()+
  ggtitle("Acropora pulchra Molecular Samples")+
  ylim(0, 2)+
  ylab(expression(paste("Host Respiration (mol O2 cm"^-2, "hr" ^-1)));plot_respiration

ggsave(plot_respiration, filename="Figures/Univariate/apulchra_molecular_respiration.png", width=5, height=4)
```


# Plot calcification for time series molecular project separately 

Bring in a list of colonies for E5 time series molecular project 

```{r}
molecular<-read_csv(file="https://raw.githubusercontent.com/urol-e5/timeseries_molecular/refs/heads/main/master-molecular-metadata.csv")%>%pull(colony)%>%unique()

length(molecular)
```

```{r}
molecular<-master%>%
  filter(colony_id_corr %in% molecular)

length(unique(molecular$colony_id_corr))
```

Plot 

```{r}
calc_molecular<-molecular %>%
  filter(!is.na(calc.umol.cm2.hr)) %>% 
  filter(!is.na(haplotype))%>%
  select(colony_id_corr, haplotype, site, timepoint, calc.umol.cm2.hr)%>%
  
  ggplot(., aes(x = timepoint, y = log(1+calc.umol.cm2.hr), fill = haplotype, colour=haplotype, group=haplotype)) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    xlab("Time Point") + 
  scale_fill_manual(values=c("darkgray", "orange", "purple"))+
  scale_colour_manual(values=c("darkgray", "orange", "purple"))+
    scale_x_discrete(labels=c("timepoint1"="January 2020", "timepoint2"="March 2020", "timepoint3"="September 2020", "timepoint4"="November 2020"))+
    ylab(expression(bold(paste("log Calcification (", mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1, ")"))))+
  geom_smooth(aes(group=haplotype), method="loess")+
    theme_classic() + 
    theme(
      legend.position="right",
      legend.title=element_blank(),
      legend.text=element_text(size=12),
      axis.title=element_text(face="bold", size=12),
      axis.text=element_text(size=10, color="black"), 
      axis.text.x=element_text(size=10, color="black", angle=45, hjust=1), 
      strip.text.x=element_text(face="italic", size=12)
      );calc_molecular

ggsave(calc_molecular, filename="Figures/Univariate/calcification_timeseries_molecular.png", width=6, height=5)
```

Analysis 

Build a mixed model for univariate analysis and examine data distribution of surface area normalized antioxidant capacity. 

```{r}
calc_model<-lmer(log(1+calc.umol.cm2.hr)~timepoint*site*species+(1|species/haplotype/colony_id_corr), na.action=na.omit, data=molecular)
qqPlot(residuals(calc_model))
```

Generate a Type III Anova of model.    

```{r}
anova(calc_model, type="III")
summary(calc_model)

emm<-emmeans(calc_model, ~ species|timepoint)
pairs(emm)
```


