---
title: "Examining normalizers and preliminary PCA plots"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "03/15/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("pairwiseAdonis")) install.packages("pairwiseAdonis")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(pairwiseAdonis)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

# Examine multivariate profiles for each normalizer    

## All Responses (host and symbiont)    
### PCA's of each normalizer  
#### PCAs: Surface Area normalizers  

##### Combined PCA's   
 
Plot PCA using prcomp to center and scale.     
 
```{r}
pca_data_cm2<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.cm2, chlc2.ug.cm2, prot_mg.cm2, cells.cm2, cre.umol.cm2, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)

pca_data_cm2<-pca_data_cm2[complete.cases(pca_data_cm2), ]

scaled_data_cm2<-prcomp(pca_data_cm2[c(5:18)], scale=TRUE, center=TRUE) 
```

Groupings by species:  
```{r}
pcaSpecies_cm2<-ggplot2::autoplot(scaled_data_cm2, data=pca_data_cm2, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("orange", "darkgray", "purple")) +
  scale_fill_manual(values=c("orange", "darkgray", "purple")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaSpecies_cm2

```

PERMANOVA of the effect of species  
```{r}
# scale data
vegan_cm2 <- scale(pca_data_cm2[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ species, data = pca_data_cm2, method='eu')
```


Groupings by timepoint:  
```{r}
pcaTimepoint_cm2<-ggplot2::autoplot(scaled_data_cm2, data=pca_data_cm2, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaTimepoint_cm2
```

PERMANOVA of the effect of timepoint  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ timepoint, data = pca_data_cm2, method='eu')
```

Groupings by site:  
```{r}
pcaSite_cm2<-ggplot2::autoplot(scaled_data_cm2, data=pca_data_cm2, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaSite_cm2

```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ site_code, data = pca_data_cm2, method='eu')
```

Generate a plot of all PCA's for surface area normalization.  

```{r}
pcaplots_cm2<-plot_grid(pcaSpecies_cm2, pcaSite_cm2, pcaTimepoint_cm2, labels = c("Species: Surface Area", "Site: Surface Area", "Timepoint: Surface Area"), ncol=3, nrow=1, rel_widths = c(1, 1, 1), label_size = 20, label_x = 0.01)

```  

##### Species-specific PCA's   

Next generate PCA for site and timepoint for each species separately.  

Separate by species.  
```{r}
#porites
por_data_cm2<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.cm2, chlc2.ug.cm2, prot_mg.cm2, cells.cm2, cre.umol.cm2, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Porites")

por_data_cm2<-por_data_cm2[complete.cases(por_data_cm2), ]

scaled_por_cm2<-prcomp(por_data_cm2[c(5:18)], scale=TRUE, center=TRUE) 

#acropora
acr_data_cm2<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.cm2, chlc2.ug.cm2, prot_mg.cm2, cells.cm2, cre.umol.cm2, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")

acr_data_cm2<-acr_data_cm2[complete.cases(acr_data_cm2), ]

scaled_acr_cm2<-prcomp(acr_data_cm2[c(5:18)], scale=TRUE, center=TRUE) 

#pocillopora
poc_data_cm2<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.cm2, chlc2.ug.cm2, prot_mg.cm2, cells.cm2, cre.umol.cm2, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")

poc_data_cm2<-poc_data_cm2[complete.cases(poc_data_cm2), ]

scaled_poc_cm2<-prcomp(poc_data_cm2[c(5:18)], scale=TRUE, center=TRUE) 
```

###### Porites PCA's   

Groupings by timepoint:  
```{r}
porTimepoint_cm2<-ggplot2::autoplot(scaled_por_cm2, data=por_data_cm2, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.7,0.5)+
  #ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porTimepoint_cm2
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_cm2 <- scale(por_data_cm2[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ timepoint, data = por_data_cm2, method='eu')
```

Groupings by site:  
```{r}
porSite_cm2<-ggplot2::autoplot(scaled_por_cm2, data=por_data_cm2, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.7,0.5)+
  #ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porSite_cm2
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ site_code, data = por_data_cm2, method='eu')
```

###### Acropora PCA's  
 
Groupings by timepoint:  
```{r}
acrTimepoint_cm2<-ggplot2::autoplot(scaled_acr_cm2, data=acr_data_cm2, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrTimepoint_cm2
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_cm2 <- scale(acr_data_cm2[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ timepoint, data = acr_data_cm2, method='eu')
```

Groupings by site:  
```{r}
acrSite_cm2<-ggplot2::autoplot(scaled_acr_cm2, data=acr_data_cm2, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrSite_cm2
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ site_code, data = acr_data_cm2, method='eu')
```

###### Pocillopora PCA's  

Groupings by timepoint:   
```{r}
pocTimepoint_cm2<-ggplot2::autoplot(scaled_poc_cm2, data=poc_data_cm2, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.6,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocTimepoint_cm2
```
 
PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_cm2 <- scale(poc_data_cm2[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ timepoint, data = poc_data_cm2, method='eu')
```

Groupings by site:  
```{r}
pocSite_cm2<-ggplot2::autoplot(scaled_poc_cm2, data=poc_data_cm2, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocSite_cm2
```

PERMANOVA of the effect of site  
```{r}

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_cm2 ~ site_code, data = poc_data_cm2, method='eu')
```

Generate a plot of all PCA's for all Species    

```{r}
species_plots_cm2<-plot_grid(porSite_cm2, porTimepoint_cm2, acrSite_cm2, acrTimepoint_cm2, pocSite_cm2, pocTimepoint_cm2, labels = c("Porites: Surface Area", "Porites: Surface Area", "Acropora: Surface Area", "Acropora: Surface Area", "Pocillopora: Surface Area", "Pocillopora: Surface Area"), ncol=2, nrow=3, rel_widths = c(1, 1, 1, 1, 1, 1), label_size = 20, label_x = 0.01, label_y=1.0)

``` 

#### PCAs: Protein normalizers  

##### Combined PCA's  

Plot PCA using prcomp to center and scale.    
 
```{r}
pca_data_prot<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgprot, chlc2.ug.mgprot, prot_mg.cm2, cells.mgprot, cre.umol.mgprot, Am, AQY, Rd, Ik, Ic, calc.umol.mgprot.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)

pca_data_prot<-pca_data_prot[complete.cases(pca_data_prot), ]

scaled_data_prot<-prcomp(pca_data_prot[c(5:18)], scale=TRUE, center=TRUE) 
```

Groupings by species:  
```{r}
pcaSpecies_prot<-ggplot2::autoplot(scaled_data_prot, data=pca_data_prot, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("orange", "darkgray", "purple")) +
  scale_fill_manual(values=c("orange", "darkgray", "purple")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaSpecies_prot

```

PERMANOVA of the effect of species  
```{r}
# scale data
vegan_prot <- scale(pca_data_prot[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ species, data = pca_data_prot, method='eu')
```


Groupings by timepoint:  
```{r}
pcaTimepoint_prot<-ggplot2::autoplot(scaled_data_prot, data=pca_data_prot, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaTimepoint_prot
```

PERMANOVA of the effect of timepoint  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ timepoint, data = pca_data_prot, method='eu')
```

Groupings by site:  
```{r}
pcaSite_prot<-ggplot2::autoplot(scaled_data_prot, data=pca_data_prot, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaSite_prot

```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ site_code, data = pca_data_prot, method='eu')
```

Generate a plot of all PCA's for protein normalization.  

```{r}
pcaplots_prot<-plot_grid(pcaSpecies_prot, pcaSite_prot, pcaTimepoint_prot, labels = c("Species: Protein", "Site: Protein", "Timepoint: Protein"), ncol=3, nrow=1, rel_widths = c(1, 1, 1), label_size = 20, label_x = 0.02)

```  

##### Species-specific PCA's   

Next generate PCA for site and timepoint for each species separately.  

Separate by species.  
```{r}
#porites
por_data_prot<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgprot, chlc2.ug.mgprot, prot_mg.cm2, cells.mgprot, cre.umol.mgprot, Am, AQY, Rd, Ik, Ic, calc.umol.mgprot.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  filter(species=="Porites")

por_data_prot<-por_data_prot[complete.cases(por_data_prot), ]

scaled_por_prot<-prcomp(por_data_prot[c(5:18)], scale=TRUE, center=TRUE) 

#acropora
acr_data_prot<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgprot, chlc2.ug.mgprot, prot_mg.cm2, cells.mgprot, cre.umol.mgprot, Am, AQY, Rd, Ik, Ic, calc.umol.mgprot.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")

acr_data_prot<-acr_data_prot[complete.cases(acr_data_prot), ]

scaled_acr_prot<-prcomp(acr_data_prot[c(5:18)], scale=TRUE, center=TRUE) 

#pocillopora
poc_data_prot<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgprot, chlc2.ug.mgprot, prot_mg.cm2, cells.mgprot, cre.umol.mgprot, Am, AQY, Rd, Ik, Ic, calc.umol.mgprot.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")

poc_data_prot<-poc_data_prot[complete.cases(poc_data_prot), ]

scaled_poc_prot<-prcomp(poc_data_prot[c(5:18)], scale=TRUE, center=TRUE) 
```

###### Porites PCA's   

Groupings by timepoint:  
```{r}
porTimepoint_prot<-ggplot2::autoplot(scaled_por_prot, data=por_data_prot, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.7,0.5)+
  #ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porTimepoint_prot
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_prot <- scale(por_data_prot[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ timepoint, data = por_data_prot, method='eu')
```

Groupings by site:  
```{r}
porSite_prot<-ggplot2::autoplot(scaled_por_prot, data=por_data_prot, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.7,0.5)+
  #ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porSite_prot
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ site_code, data = por_data_prot, method='eu')
```

###### Acropora PCA's  
 
Groupings by timepoint:  
```{r}
acrTimepoint_prot<-ggplot2::autoplot(scaled_acr_prot, data=acr_data_prot, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrTimepoint_prot
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_prot <- scale(acr_data_prot[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ timepoint, data = acr_data_prot, method='eu')
```

Groupings by site:  
```{r}
acrSite_prot<-ggplot2::autoplot(scaled_acr_prot, data=acr_data_prot, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrSite_prot
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ site_code, data = acr_data_prot, method='eu')
```

###### Pocillopora PCA's  
 
Groupings by timepoint:   
```{r}
pocTimepoint_prot<-ggplot2::autoplot(scaled_poc_prot, data=poc_data_prot, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.6,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocTimepoint_prot
```
 
PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_prot <- scale(poc_data_prot[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ timepoint, data = poc_data_prot, method='eu')
```

Groupings by site:  
```{r}
pocSite_prot<-ggplot2::autoplot(scaled_poc_prot, data=poc_data_prot, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocSite_prot
```

PERMANOVA of the effect of site  
```{r}

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_prot ~ site_code, data = poc_data_prot, method='eu')
```

Generate a plot of all PCA's for all Species    

```{r}
species_plots_prot<-plot_grid(porSite_prot, porTimepoint_prot, acrSite_prot, acrTimepoint_prot, pocSite_prot, pocTimepoint_prot, labels = c("Porites: Protein", "Porites: Protein", "Acropora: Protein", "Acropora: Protein", "Pocillopora: Protein", "Pocillopora: Protein"), ncol=2, nrow=3, rel_widths = c(1, 1, 1, 1, 1, 1), label_size = 20, label_x = 0.01, label_y=1.0)

``` 

#### PCAs: Biomass normalizers  

##### Combined PCA's  

Plot PCA using prcomp to center and scale.    
  
```{r}
pca_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgAFDW, chlc2.ug.mgAFDW, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.mgAFDW.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(prot_mg.mgafdw<1.5) #remove outliers

pca_data_afdw<-pca_data_afdw[complete.cases(pca_data_afdw), ]

scaled_data_afdw<-prcomp(pca_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 
```

```{r}
# scale data
vegan <- scale(pca_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ species, data = pca_data_afdw, method='eu')
z_species<-permanova
z_species
``` 
 
Groupings by species:  
```{r}
pcaSpecies_afdw<-ggplot2::autoplot(scaled_data_afdw, data=pca_data_afdw, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.5) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  geom_text(x=0.15, y=-0.1, label=paste("p(Species)=", z_species$`Pr(>F)`[1]), size=4, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  #geom_text(x=0.3, y=-0.15, label=paste("p(Species)=", z_species$`Pr(>F)`[1]), size=6, color="black") + 
  xlab("PC1")+
  ylab("PC2")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw
```

PERMANOVA of the effect of species  
```{r}
# scale data
vegan_afdw <- scale(pca_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ species, data = pca_data_afdw, method='eu')
```
 
Groupings by timepoint:  
```{r}
pcaTimepoint_afdw<-ggplot2::autoplot(scaled_data_afdw, data=pca_data_afdw, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaTimepoint_afdw
```

PERMANOVA of the effect of timepoint  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ timepoint, data = pca_data_afdw, method='eu')
```

Groupings by site:  
```{r}
pcaSite_afdw<-ggplot2::autoplot(scaled_data_afdw, data=pca_data_afdw, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pcaSite_afdw

```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ site_code, data = pca_data_afdw, method='eu')
```

Generate a plot of all PCA's for afdw normalization.  

```{r}
pcaplots_afdw<-plot_grid(pcaSpecies_afdw, pcaSite_afdw, pcaTimepoint_afdw, labels = c("Species: Biomass", "Site: Biomass", "Timepoint: Biomass"), ncol=3, nrow=1, rel_widths = c(1, 1, 1), label_size = 20, label_x = 0.02)
```  

##### Species-specific PCA's   

Next generate PCA for site and timepoint for each species separately.  
 
Separate by species.  
```{r}
#porites
por_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgAFDW, chlc2.ug.mgAFDW, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.mgAFDW.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Porites")

por_data_afdw<-por_data_afdw[complete.cases(por_data_afdw), ]

scaled_por_afdw<-prcomp(por_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 

#acropora
acr_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgAFDW, chlc2.ug.mgAFDW, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.mgAFDW.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")

acr_data_afdw<-acr_data_afdw[complete.cases(acr_data_afdw), ]

scaled_acr_afdw<-prcomp(acr_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 

#pocillopora
poc_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, chla.ug.mgAFDW, chlc2.ug.mgAFDW, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.mgAFDW.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")

poc_data_afdw<-poc_data_afdw[complete.cases(poc_data_afdw), ]

scaled_poc_afdw<-prcomp(poc_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 
```

###### Porites PCA's   

Groupings by timepoint:  
```{r}
porTimepoint_afdw<-ggplot2::autoplot(scaled_por_afdw, data=por_data_afdw, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.7,0.5)+
  #ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porTimepoint_afdw
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_afdw <- scale(por_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ timepoint, data = por_data_afdw, method='eu')
```

Groupings by site:  
```{r}
porSite_afdw<-ggplot2::autoplot(scaled_por_afdw, data=por_data_afdw, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
 # xlim(-0.7,0.5)+
 # ylim(-0.7, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));porSite_afdw
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ site_code, data = por_data_afdw, method='eu')
```

###### Acropora PCA's  
 
Groupings by timepoint:  
```{r}
acrTimepoint_afdw<-ggplot2::autoplot(scaled_acr_afdw, data=acr_data_afdw, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrTimepoint_afdw
```

PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_afdw <- scale(acr_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ timepoint, data = acr_data_afdw, method='eu')
```

Groupings by site:  
```{r}
acrSite_afdw<-ggplot2::autoplot(scaled_acr_afdw, data=acr_data_afdw, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));acrSite_afdw
```

PERMANOVA of the effect of site  
```{r}
# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ site_code, data = acr_data_afdw, method='eu')
```

###### Pocillopora PCA's  

Groupings by timepoint:   
```{r}
pocTimepoint_afdw<-ggplot2::autoplot(scaled_poc_afdw, data=poc_data_afdw, frame.colour="timepoint", loadings=FALSE,  colour="timepoint", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) +
  scale_fill_manual(values=c("darkturquoise", "darkgray", "mediumvioletred", "black"), labels=c("Jan2020", "March2020", "Sept2020", "Nov2020")) + 
  theme_classic()+
 # xlim(-0.6,0.5)+
  #ylim(-0.5, 0.8)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocTimepoint_afdw
```
 
PERMANOVA of the effect of timepoint  
```{r}
# scale data
vegan_afdw <- scale(poc_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ timepoint, data = poc_data_afdw, method='eu')
```

Groupings by site:  
```{r}
pocSite_afdw<-ggplot2::autoplot(scaled_poc_afdw, data=poc_data_afdw, frame.colour="site_code", loadings=FALSE,  colour="site_code", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm') + 
  scale_colour_manual(values=c("blue4", "darkgray", "springgreen3")) +
  scale_fill_manual(values=c("blue4", "darkgray", "springgreen3")) + 
  theme_classic()+
  #xlim(-0.5,0.5)+
  #ylim(-0.5, 0.5)+
   theme(legend.text = element_text(size=18), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_text(size=18, face="bold"), 
        axis.text = element_text(size=18), 
        axis.title = element_text(size=18,  face="bold"));pocSite_afdw
```

PERMANOVA of the effect of site  
```{r}

# PerMANOVA - partitioning the euclidean distance matrix by species
adonis2(vegan_afdw ~ site_code, data = poc_data_afdw, method='eu')
```

Generate a plot of all PCA's for all Species    

```{r}
species_plots_afdw<-plot_grid(porSite_afdw, porTimepoint_afdw, acrSite_afdw, acrTimepoint_afdw, pocSite_afdw, pocTimepoint_afdw, labels = c("Porites: Biomass", "Porites: Biomass", "Acropora: Biomass", "Acropora: Biomass", "Pocillopora: Biomass", "Pocillopora: Biomass"), ncol=2, nrow=3, rel_widths = c(1, 1, 1, 1, 1, 1), label_size = 20, label_x = 0.01, label_y=1.0)

``` 

#### Assemble all PCA plots  

Generate plot of all combined PCA's

```{r}
All_PCAs<-plot_grid(pcaSpecies_prot, pcaSite_prot, pcaTimepoint_prot, pcaSpecies_afdw, pcaSite_afdw, pcaTimepoint_afdw, pcaSpecies_cm2, pcaSite_cm2, pcaTimepoint_cm2, labels = c("Protein", "Protein", "Protein", "Biomass", "Biomass", "Biomass","Surface Area", "Surface Area", "Surface Area"), ncol=3, nrow=3, rel_widths = c(1, 1, 1, 1, 1, 1, 1, 1, 1), label_size = 20, label_x = 0.06, label_y=1.0)

ggsave(filename="Figures/NormalizerPCA/All_PCAs.pdf", plot=All_PCAs, dpi=300, width=24, height=15, units="in")
``` 

Generate plot of all species specific PCA's

```{r}
All_PCAs_species<-plot_grid(porSite_prot, porTimepoint_prot, porSite_afdw, porTimepoint_afdw, porSite_cm2, porTimepoint_cm2, 
                    acrSite_prot, acrTimepoint_prot, acrSite_afdw, acrTimepoint_afdw, acrSite_cm2, acrTimepoint_cm2,
                    pocSite_prot, pocTimepoint_prot, pocSite_afdw, pocTimepoint_afdw, pocSite_cm2, pocTimepoint_cm2,
                    labels = c("Porites: Protein", "Porites: Protein", "Porites: Biomass", "Porites: Biomass", "Porites: SA", "Porites: SA",
                               "Acropora: Protein", "Acropora: Protein", "Acropora: Biomass", "Acropora: Biomass", "Acropora: SA", "Acropora: SA",
                               "Pocillopora: Protein", "Pocillopora: Protein", "Pocillopora: Biomass", "Pocillopora: Biomass", "Pocillopora: SA", "Pocillopora: SA"), ncol=6, nrow=3, rel_widths = c(1), label_size = 20, label_x = 0.06, label_y=1.0)

ggsave(filename="Figures/NormalizerPCA/All_Species_PCAs.pdf", plot=All_PCAs_species, dpi=300, width=48, height=15, units="in")
``` 

# Final species PCA of selected normalizers  

## Holobiont 

Plot PCA using prcomp to center and scale using biomass normalized tissue responses and surface area normalized metabolic rates and calcification rates. 
  
```{r}
pca_data_species<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>%
  filter(Total_Chl<14)%>%
  filter(Total_Chl_cell<6e-05)%>%
  filter(Ic<1000)#remove outliers

pca_data_species<-pca_data_species[complete.cases(pca_data_species), ]

scaled_data_species<-prcomp(pca_data_species[c(5:18)], scale=TRUE, center=TRUE) 
```

```{r}
# scale data
vegan <- scale(pca_data_species[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ species, data = pca_data_species, method='eu')
z_species<-permanova
z_species

#adonis2(formula = vegan ~ species, data = pca_data_species, method = "eu")
#          Df SumOfSqs      R2      F Pr(>F)    
#species    2   1872.9 0.36059 104.05  0.001 ***
#Residual 369   3321.1 0.63941                  
#Total    371   5194.0 1.00000  

#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=pca_data_species[ ,5:18],factors=pca_data_species$species, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#                    pairs Df    SumsOfSqs  F.Model        R2 p.value p.adjusted
#1 Acropora vs Pocillopora  1 4.536182e+12 181.3893 0.4387857   0.001      0.003
#2     Acropora vs Porites  1 1.998949e+12  72.5289 0.2335657   0.001      0.003
#3  Pocillopora vs Porites  1 6.239125e+11 100.7953 0.2733095   0.001      0.003
#  sig
#1   *
#2   *
#3   *
``` 
 
Groupings by species:  
```{r}
pcaSpecies_afdw<-ggplot2::autoplot(scaled_data_species, data=pca_data_species, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.5) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  geom_text(x=-0.04, y=0.3, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw

ggsave(filename="Figures/NormalizerPCA/Species_Overall_PCA.png", plot=pcaSpecies_afdw, dpi=300, width=7, height=5, units="in")
ggsave(filename="Figures/NormalizerPCA/Species_Overall_PCA.jpeg", plot=pcaSpecies_afdw, dpi=300, width=7, height=5, units="in")
ggsave(filename="Figures/NormalizerPCA/Species_Overall_PCA.pdf", plot=pcaSpecies_afdw, dpi=300, width=7, height=5, units="in")
```

Generate a biplot for this PCA. 
```{r}
pcaSpecies_afdw_biplot<-ggplot2::autoplot(scaled_data_species, data=pca_data_species, frame.colour="species", loadings=TRUE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.0) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  #geom_text(x=-0.04, y=0.3, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw_biplot

ggsave(filename="Figures/NormalizerPCA/Species_Overall_BiPlot.png", plot=pcaSpecies_afdw_biplot, dpi=300, width=5, height=5, units="in")
ggsave(filename="Figures/NormalizerPCA/Species_Overall_BiPlot.jpeg", plot=pcaSpecies_afdw_biplot, dpi=300, width=5, height=5, units="in")
ggsave(filename="Figures/NormalizerPCA/Species_Overall_BiPlot.pdf", plot=pcaSpecies_afdw_biplot, dpi=300, width=5, height=5, units="in")
```

Species specific PCAs and PERMANOVAs are in the next script, `4_multivariate_analysis.RmD`.  


## Host 

Plot PCA using prcomp to center and scale using biomass normalized tissue responses and surface area normalized metabolic rates and calcification rates. 
  
```{r}
pca_data_species_host<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, prot_mg.mgafdw, cre.umol.mgafdw, Rd, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5) #%>%
  #filter(Total_Chl<14)%>%
  #filter(Total_Chl_cell<6e-05)%>%
  #filter(Ic<1000)#remove outliers

pca_data_species_host<-pca_data_species_host[complete.cases(pca_data_species_host), ]

scaled_data_species_host<-prcomp(pca_data_species_host[c(5:9)], scale=TRUE, center=TRUE) 
```

```{r}
# scale data
vegan <- scale(pca_data_species_host[ ,5:9])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ species, data = pca_data_species_host, method='eu')
z_species<-permanova
z_species

#adonis2(formula = vegan ~ species, data = pca_data_species_host, method = "eu")
#          Df SumOfSqs     R2      F Pr(>F)    
#species    2   835.18 0.4283 145.34  0.001 ***
#Residual 388  1114.82 0.5717                  
#Total    390  1950.00 1.0000                  
``` 
 
Groupings by species:  
```{r}
pcaSpecies_afdw_host<-ggplot2::autoplot(scaled_data_species_host, data=pca_data_species_host, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.5) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  geom_text(x=-0.04, y=0.3, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
  ggtitle("A. Host Responses")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw_host

ggsave(filename="Figures/NormalizerPCA/Species_Host_PCA.jpeg", plot=pcaSpecies_afdw_host, dpi=300, width=5, height=5, units="in")
```

Generate a biplot for this PCA. 
```{r}
pcaSpecies_afdw_host_biplot<-ggplot2::autoplot(scaled_data_species_host, data=pca_data_species_host, frame.colour="species", loadings=TRUE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.0) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  #geom_text(x=-0.04, y=0.3, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw_host_biplot

ggsave(filename="Figures/NormalizerPCA/Species_Host_BiPlot.jpeg", plot=pcaSpecies_afdw_host_biplot, dpi=300, width=5, height=5, units="in")
```

Species specific PCAs and PERMANOVAs are in the next script, `4_multivariate_analysis.RmD`.  


## Symbiont 

Plot PCA using prcomp to center and scale using biomass normalized tissue responses and surface area normalized metabolic rates and calcification rates. 
  
```{r}
pca_data_species_sym<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, cells.mgAFDW, Am, AQY, Ik, Ic)%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(prot_mg.mgafdw<1.5)%>%
  filter(Total_Chl<14)%>%
  filter(Total_Chl_cell<6e-05)%>%
  filter(Ic<1000)#remove outliers

pca_data_species_sym<-pca_data_species_sym[complete.cases(pca_data_species_sym), ]

scaled_data_species_sym<-prcomp(pca_data_species_sym[c(5:13)], scale=TRUE, center=TRUE) 
```

```{r}
# scale data
vegan <- scale(pca_data_species_sym[ ,5:13])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ species, data = pca_data_species_sym, method='eu')
z_species<-permanova
z_species

#adonis2(formula = vegan ~ species, data = pca_data_species_sym, method = "eu")
#          Df SumOfSqs      R2      F Pr(>F)    
#species    2   1174.5 0.31829 95.248  0.001 ***
#Residual 408   2515.5 0.68171                  
#Total    410   3690.0 1.00000                
``` 
 
Groupings by species:  
```{r}
pcaSpecies_afdw_sym<-ggplot2::autoplot(scaled_data_species_sym, data=pca_data_species_sym, frame.colour="species", loadings=FALSE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=FALSE, frame=TRUE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.5) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  geom_text(x=-0.11, y=0.1, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
  ggtitle("B. Symbiont Responses")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="right",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw_sym

ggsave(filename="Figures/NormalizerPCA/Species_Symbiont_PCA.jpeg", plot=pcaSpecies_afdw_sym, dpi=300, width=7, height=5, units="in")
```

Generate a biplot for this PCA. 
```{r}
pcaSpecies_afdw_sym_biplot<-ggplot2::autoplot(scaled_data_species_sym, data=pca_data_species_sym, frame.colour="species", loadings=TRUE,  colour="species", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=5, loadings.label.vjust=-1, size=5, frame.type = 'norm', alpha=0.0) + 
  scale_colour_manual(values=c("darkgray", "orange", "purple")) +
  scale_fill_manual(values=c("darkgray", "orange", "purple")) + 
  theme_classic()+
  #geom_text(x=-0.04, y=0.3, label=paste("p =", z_species$`Pr(>F)`[1]), size=6, color=ifelse(z_species$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #xlim(-0.3,0.5)+
  #ylim(-0.3, 0.5)+
  xlab("PC1")+
  ylab("PC2")+
   theme(legend.text = element_text(size=18, face="italic"), 
         legend.position="none",
        plot.background = element_blank(),
        legend.title = element_blank(), 
        axis.text = element_text(size=18, color="black"), 
        axis.title = element_text(size=18, color="black"));pcaSpecies_afdw_sym_biplot

ggsave(filename="Figures/NormalizerPCA/Species_Sym_BiPlot.jpeg", plot=pcaSpecies_afdw_sym_biplot, dpi=300, width=5, height=5, units="in")
```

Species specific PCAs and PERMANOVAs are in the next script, `4_multivariate_analysis.RmD`.  

