---
title: "Multivariate analysis of E5 time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "08/1/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("pairwiseAdonis")) install.packages("pairwiseAdonis")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(pairwiseAdonis)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")

master<-master%>%
  select(!code)
```

Set seed for permutation tests (randomly picked). 
```{r}
set.seed(2534)
```

# Build trajectory PCA plots  

## Holobiont Responses 

### PCA's for centroid and trajectories  

#### Acropora  

Generate PCA using scaled (scaled_acr_afdw) from dataframe (acr_data_afdw).   
 
```{r}
acr_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Acropora")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

acr_data_afdw<-acr_data_afdw[complete.cases(acr_data_afdw), ]

acr_data_afdw[,-c(1:4)]<-log(acr_data_afdw[,-c(1:4)]+1)

scaled_acr_afdw<-prcomp(acr_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 

acr_info<-acr_data_afdw[c(2,4)]

acr_data<-scaled_acr_afdw%>%
  augment(acr_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

acr.centroids<-acr_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Examine PERMANOVA results.  
 
```{r}
# scale data
vegan <- scale(acr_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = acr_data_afdw, method='eu')
z_acr<-permanova
z_acr
``` 

adonis2(formula = vegan ~ timepoint * site, data = acr_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   423.98 0.29984 15.6680  0.001 ***
site             2    95.01 0.06719  5.2666  0.001 ***
timepoint:site   6    83.20 0.05884  1.5374  0.025 *  
Residual        90   811.81 0.57412                   
Total          101  1414.00 1.00000  

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:18],factors=acr_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(acr_data_afdw[ ,5:18], "euclidian"), acr_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:18],factors=acr_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(acr_data_afdw[ ,5:18], "euclidian"), acr_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is no difference in dispersion between time points, therefore differences are due to centroid location. 

Assemble plot with background points 
```{r}
#1. make plot with dots
acrPCA<-ggplot(acr_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Acropora")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_acr$`Pr(>F)`[1]), size=6, color=ifelse(z_acr$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_acr$`Pr(>F)`[2]), size=6, color=ifelse(z_acr$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_acr$`Pr(>F)`[3]), size=6, color=ifelse(z_acr$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.025", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"), 
         axis.title = element_text(size=18));acrPCA

```

Add centroids  

```{r}
#2. add centroids 
acrPCAcen<-acrPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=acr.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); acrPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-acr.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
acrPCAfull<-acrPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); acrPCAfull
```

Make bi plot with loadings  
```{r}
#1. make plot with dots
acrArrows<-ggplot2::autoplot(scaled_acr_afdw, data=acr_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0.5, loadings.label.hjust=-0.1, loadings.label.repel=FALSE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle(expression(italic("Acropora")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));acrArrows
```

#### Porites

Generate PCA using scaled (scaled_por_afdw) from dataframe (por_data_afdw).   
 
Calculate centroid locations for each site and timepoint and pull PC1 and PC2 locations.     
```{r}
por_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

por_data_afdw<-por_data_afdw[complete.cases(por_data_afdw), ]

por_data_afdw[,-c(1:4)]<-log(por_data_afdw[,-c(1:4)]+1)

scaled_por_afdw<-prcomp(por_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 

por_info<-por_data_afdw[c(2,4)]

por_data<-scaled_por_afdw%>%
  augment(por_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids<-por_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

```{r}
# scale data
vegan <- scale(por_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = por_data_afdw, method='eu')
z_por<-permanova
z_por
```

adonis2(formula = vegan ~ timepoint * site, data = por_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   337.22 0.17455 11.2136  0.001 ***
site             2   210.18 0.10879 10.4838  0.001 ***
timepoint:site   6   111.51 0.05772  1.8541  0.003 ** 
Residual       127  1273.08 0.65894                   
Total          138  1932.00 1.00000     

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:18],factors=por_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(por_data_afdw[ ,5:18], "euclidian"), por_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:18],factors=por_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(por_data_afdw[ ,5:18], "euclidian"), por_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is no difference in dispersion between time points, therefore differences are due to centroid location. 

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA<-ggplot(por_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  ylim(-6,6)+
  xlim(-6,6)+
  ggtitle(expression(italic("Porites")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_por$`Pr(>F)`[1]), size=6, color=ifelse(z_por$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_por$`Pr(>F)`[2]), size=6, color=ifelse(z_por$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_por$`Pr(>F)`[3]), size=6, color=ifelse(z_por$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.003", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position=c(1,0.3)); porPCAcen

#build a plot for the legend
legend_plot<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=TRUE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="bottom")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints<-por.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAfull<-porPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAfull
```

Add bi plot with loadings  
```{r}
#1. make plot with dots
porArrows<-ggplot2::autoplot(scaled_por_afdw, data=por_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=-0.1, loadings.label.hjust=0.1, loadings.label.repel=FALSE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.0)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-0.5,0.1)+
  #ylim(-0.5, 0.5)+
  ggtitle(expression(italic("Porites")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));porArrows
```

##### Generate POR panel by haplotype 

Calculate centroid locations for each site and timepoint and pull PC1 and PC2 locations.     
```{r}
por_data_afdw_haplotype<-master%>%
  select(colony_id_corr, timepoint, species, haplotype, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

por_data_afdw_haplotype<-por_data_afdw_haplotype[complete.cases(por_data_afdw_haplotype), ]

por_data_afdw_haplotype[,-c(1:4)]<-log(por_data_afdw_haplotype[,-c(1:4)]+1)

scaled_por_afdw_haplotype<-prcomp(por_data_afdw_haplotype[c(5:18)], scale=TRUE, center=TRUE) 

por_info_haplotype<-por_data_afdw_haplotype[c(2,4)]

por_data_haplotype<-scaled_por_afdw_haplotype%>%
  augment(por_info_haplotype)%>%
  group_by(timepoint, haplotype)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids_haplotype<-por_data_haplotype %>% 
  select(timepoint, haplotype, PC1.mean, PC2.mean)%>%
  group_by(timepoint, haplotype)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA_haplotype<-ggplot(por_data_haplotype, aes(.fittedPC1, .fittedPC2, color=haplotype)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("black", "darkgray")) +
  scale_fill_manual(values=c("black", "darkgray")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  #ylim(-6,6)+
  #xlim(-6,6)+
  ggtitle("Holobiont")+
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=18, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA_haplotype
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=FALSE) + 
  #scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="right"); porPCAcen_haplotype

#build a plot for the legend
legend_plot_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=FALSE) + 
  #scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="right")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints_haplotype<-por.centroids_haplotype%>%
  gather(variable, value, -(timepoint:haplotype)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAfull_haplotype<-porPCAcen_haplotype + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAfull_haplotype
```



#### Pocillopora   

Generate PCA using scaled (scaled_poc_afdw) from dataframe (poc_data_afdw). 
 
```{r}
poc_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

poc_data_afdw<-poc_data_afdw[complete.cases(poc_data_afdw), ]
 
poc_data_afdw[,-c(1:4)]<-log(poc_data_afdw[,-c(1:4)]+1)

scaled_poc_afdw<-prcomp(poc_data_afdw[c(5:18)], scale=TRUE, center=TRUE) 

poc_info<-poc_data_afdw[c(2,4)]

poc_data<-scaled_poc_afdw%>%
  augment(poc_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

poc.centroids<-poc_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
``` 
  
```{r}
# scale data
vegan <- scale(poc_data_afdw[ ,5:18])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = poc_data_afdw, method='eu')
z_poc<-permanova
z_poc
```

adonis2(formula = vegan ~ timepoint * site, data = poc_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   454.84 0.24613 15.4796  0.001 ***
site             2    69.20 0.03745  3.5327  0.001 ***
timepoint:site   6   138.82 0.07512  2.3622  0.001 ***
Residual       121  1185.13 0.64131                   
Total          132  1848.00 1.00000   

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:18],factors=poc_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(poc_data_afdw[ ,5:18], "euclidian"), poc_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is no difference in dispersion by site, so therefore differences are due to centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:18],factors=poc_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(poc_data_afdw[ ,5:18], "euclidian"), poc_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is a significant difference in dispersion by timepoint, so differences are likely due to both dispersion and centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
pocPCA<-ggplot(poc_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Pocillopora")))+
  #geom_text(x=3, y=5.5, label=paste("T=", z_poc$`Pr(>F)`[1]), size=6, color=ifelse(z_poc$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=", z_poc$`Pr(>F)`[2]), size=6, color=ifelse(z_poc$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=", z_poc$`Pr(>F)`[3]), size=6, color=ifelse(z_poc$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.001", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         plot.margin = margin(1, 1, 1, 1, "cm"),
         legend.title = element_blank(), 
         title = element_text(size=25, face="bold"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18,  face="bold"));pocPCA

```

Add centroids  

```{r}
#2. add centroids 
pocPCAcen<-pocPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=poc.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); pocPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-poc.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
pocPCAfull<-pocPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); pocPCAfull
```

Add bi plot with loadings  
```{r}

pocArrows<-ggplot2::autoplot(scaled_poc_afdw, data=poc_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0, loadings.label.hjust=-0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle(expression(italic("Pocillopora")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));pocArrows
```
 
#### Assemble all plots    

Assemble trajectory PCAs
```{r}
PCA_full_panel<-plot_grid(acrPCAfull, pocPCAfull, porPCAfull, labels = c("A", "B", "C"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
PCA_full_panel_legend<-plot_grid(PCA_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Holobiont.png", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Holobiont.jpg", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Holobiont.pdf", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
```  

Assemble biplots  
```{r}
biplot_full_panel<-plot_grid(acrArrows, pocArrows, porArrows, labels = c("A", "B", "C"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
biplot_full_panel_legend<-plot_grid(biplot_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Biplots_Holobiont.png", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Holobiont.jpg", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Holobiont.pdf", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
```  

### Matrix of PCA's by species and time  

Generate biplot showing groupings by site for each species and time point.  

#### Pocillopora  

```{r}
poc_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

poc_tp1_data<-poc_tp1_data[complete.cases(poc_tp1_data), ]

poc_tp1_data[,-c(1:4)]<-log(poc_tp1_data[,-c(1:4)]+1)
 
poc_tp1_scaled<-prcomp(poc_tp1_data[c(5:18)], scale=TRUE, center=TRUE) 

pocTP1<-ggplot2::autoplot(poc_tp1_scaled, data=poc_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP1
```

```{r}
poc_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
poc_tp2_data<-poc_tp2_data[complete.cases(poc_tp2_data), ]

poc_tp2_data[,-c(1:4)]<-log(poc_tp2_data[,-c(1:4)]+1)
 
poc_tp2_scaled<-prcomp(poc_tp2_data[c(5:18)], scale=TRUE, center=TRUE) 

pocTP2<-ggplot2::autoplot(poc_tp2_scaled, data=poc_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP2
```

```{r}
poc_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
poc_tp3_data<-poc_tp3_data[complete.cases(poc_tp3_data), ]

poc_tp3_data[,-c(1:4)]<-log(poc_tp3_data[,-c(1:4)]+1)
 
poc_tp3_scaled<-prcomp(poc_tp3_data[c(5:18)], scale=TRUE, center=TRUE) 

pocTP3<-ggplot2::autoplot(poc_tp3_scaled, data=poc_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP3
```

```{r}
poc_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Pocillopora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
poc_tp4_data<-poc_tp4_data[complete.cases(poc_tp4_data), ]

poc_tp4_data[,-c(1:4)]<-log(poc_tp4_data[,-c(1:4)]+1)
 
poc_tp4_scaled<-prcomp(poc_tp4_data[c(5:18)], scale=TRUE, center=TRUE) 

pocTP4<-ggplot2::autoplot(poc_tp4_scaled, data=poc_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP4
```

#### Acropora  

```{r}
acr_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Acropora")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

acr_tp1_data<-acr_tp1_data[complete.cases(acr_tp1_data), ]

acr_tp1_data[,-c(1:4)]<-log(acr_tp1_data[,-c(1:4)]+1)
 
acr_tp1_scaled<-prcomp(acr_tp1_data[c(5:18)], scale=TRUE, center=TRUE) 

acrTP1<-ggplot2::autoplot(acr_tp1_scaled, data=acr_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Acropora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP1
```

```{r}
acr_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Acropora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
acr_tp2_data<-acr_tp2_data[complete.cases(acr_tp2_data), ]

acr_tp2_data[,-c(1:4)]<-log(acr_tp2_data[,-c(1:4)]+1)
 
acr_tp2_scaled<-prcomp(acr_tp2_data[c(5:18)], scale=TRUE, center=TRUE) 

acrTP2<-ggplot2::autoplot(acr_tp2_scaled, data=acr_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP2
```

```{r}
acr_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Acropora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
acr_tp3_data<-acr_tp3_data[complete.cases(acr_tp3_data), ]

acr_tp3_data[,-c(1:4)]<-log(acr_tp3_data[,-c(1:4)]+1)
 
acr_tp3_scaled<-prcomp(acr_tp3_data[c(5:18)], scale=TRUE, center=TRUE) 

acrTP3<-ggplot2::autoplot(acr_tp3_scaled, data=acr_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP3
```

```{r}
acr_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Acropora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
acr_tp4_data<-acr_tp4_data[complete.cases(acr_tp4_data), ]

acr_tp4_data[,-c(1:4)]<-log(acr_tp4_data[,-c(1:4)]+1)
 
acr_tp4_scaled<-prcomp(acr_tp4_data[c(5:18)], scale=TRUE, center=TRUE) 

acrTP4<-ggplot2::autoplot(acr_tp4_scaled, data=acr_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-7, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP4
```


#### Porites  

```{r}
por_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

por_tp1_data<-por_tp1_data[complete.cases(por_tp1_data), ]

por_tp1_data[,-c(1:4)]<-log(por_tp1_data[,-c(1:4)]+1)
 
por_tp1_scaled<-prcomp(por_tp1_data[c(5:18)], scale=TRUE, center=TRUE) 

porTP1<-ggplot2::autoplot(por_tp1_scaled, data=por_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Porites)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP1
```

```{r}
por_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Porites")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
por_tp2_data<-por_tp2_data[complete.cases(por_tp2_data), ]

por_tp2_data[,-c(1:4)]<-log(por_tp2_data[,-c(1:4)]+1)
 
por_tp2_scaled<-prcomp(por_tp2_data[c(5:18)], scale=TRUE, center=TRUE) 

porTP2<-ggplot2::autoplot(por_tp2_scaled, data=por_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP2
```

```{r}
por_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Porites")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
por_tp3_data<-por_tp3_data[complete.cases(por_tp3_data), ]

por_tp3_data[,-c(1:4)]<-log(por_tp3_data[,-c(1:4)]+1)
 
por_tp3_scaled<-prcomp(por_tp3_data[c(5:18)], scale=TRUE, center=TRUE) 

porTP3<-ggplot2::autoplot(por_tp3_scaled, data=por_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP3
```

```{r}
por_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, Total_Chl, Total_Chl_cell, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(chla.ug.mgAFDW<2.2)%>%
  filter(species=="Porites")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
por_tp4_data<-por_tp4_data[complete.cases(por_tp4_data), ]

por_tp4_data[,-c(1:4)]<-log(por_tp4_data[,-c(1:4)]+1)
 
por_tp4_scaled<-prcomp(por_tp4_data[c(5:18)], scale=TRUE, center=TRUE) 

porTP4<-ggplot2::autoplot(por_tp4_scaled, data=por_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
 # ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP4
```

#### Assemble grid plot of all metrics by species and time  

```{r}
all_grid<-plot_grid(acrTP1, acrTP2, acrTP3, acrTP4, pocTP1, pocTP2, pocTP3, pocTP4, porTP1, porTP2, porTP3, porTP4, ncol=4, nrow=3)
all_grid2<-plot_grid(all_grid, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Matrix_HolobiontResponses.png", plot=all_grid2, dpi=500, width=20, height=20, units="in")

```


## Host Responses  

Generate a list of symbiont and host responses.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

### PCA's for centroid and trajectories  
 
#### Acropora  

Generate PCA using scaled (scaled_acr_afdw) from dataframe (acr_data_afdw).   
 
```{r}
acr_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Acropora")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5) %>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Respiration=Rd, Calc=calc.umol.cm2.hr)

acr_data_afdw<-acr_data_afdw[complete.cases(acr_data_afdw), ]

acr_data_afdw[,-c(1:4)]<-log(acr_data_afdw[,-c(1:4)]+1)

scaled_acr_afdw<-prcomp(acr_data_afdw[c(5:9)], scale=TRUE, center=TRUE) 

acr_info<-acr_data_afdw[c(2,4)]

acr_data<-scaled_acr_afdw%>%
  augment(acr_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

acr.centroids<-acr_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Examine PERMANOVA results.  
 
```{r}
# scale data
vegan <- scale(acr_data_afdw[ ,5:9])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = acr_data_afdw, method='eu')
z_acr<-permanova
z_acr
``` 
 
adonis2(formula = vegan ~ timepoint * site, data = acr_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   196.82 0.38217 23.6506  0.001 ***
site             2    35.68 0.06928  6.4312  0.001 ***
timepoint:site   6    27.30 0.05300  1.6400  0.033 *  
Residual        92   255.21 0.49554                   
Total          103   515.00 1.00000  

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:9],factors=acr_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(acr_data_afdw[ ,5:9], "euclidian"), acr_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:9],factors=acr_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(acr_data_afdw[ ,5:9], "euclidian"), acr_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is no significant difference in dispersion by timepoint, so differences are likely due to only centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
acrPCA<-ggplot(acr_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Acropora")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_acr$`Pr(>F)`[1]), size=6, color=ifelse(z_acr$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_acr$`Pr(>F)`[2]), size=6, color=ifelse(z_acr$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_acr$`Pr(>F)`[3]), size=6, color=ifelse(z_acr$`Pr(>F)`[3] < 0.05, "black", "darkgray")) +
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.035", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"), 
         axis.title = element_text(size=18));acrPCA

```

Add centroids  

```{r}
#2. add centroids 
acrPCAcen<-acrPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=acr.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); acrPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-acr.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
acrPCAfull<-acrPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); acrPCAfull
```

Add bi plot with loadings  
```{r}
#1. make plot with dots
acrArrows<-ggplot2::autoplot(scaled_acr_afdw, data=acr_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0.5, loadings.label.hjust=-0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
    ggtitle(expression(italic("Acropora")))+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));acrArrows
```

Assemble final figure with inset biplot.  
```{r}
#acrFullPCA<-ggdraw(acrPCAfull) + #theme_half_open(12)) +
  #draw_plot(acrArrows, .5, .5, .5, .5); acrFullPCA #x, y, w, h

#ggsave(filename="Figures/FullPCA_Acropora.pdf", plot=acrFullPCA, dpi=300, width=12, height=8, units="in")
```

#### Porites

Generate PCA using scaled (scaled_por_afdw) from dataframe (por_data_afdw).   
 
Calculate centroid locations for each site and timepoint and pull PC1 and PC2 locations.      
```{r}
por_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Respiration=Rd, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

por_data_afdw<-por_data_afdw[complete.cases(por_data_afdw), ]

por_data_afdw[,-c(1:4)]<-log(por_data_afdw[,-c(1:4)]+1)

scaled_por_afdw<-prcomp(por_data_afdw[c(5:9)], scale=TRUE, center=TRUE) 

por_info<-por_data_afdw[c(2,4)]

por_data<-scaled_por_afdw%>%
  augment(por_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids<-por_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

```{r}
# scale data
vegan <- scale(por_data_afdw[ ,5:9])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = por_data_afdw, method='eu')
z_por<-permanova
z_por
```

adonis2(formula = vegan ~ timepoint * site, data = por_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   153.82 0.21974 15.2898  0.001 ***
site             2    80.06 0.11438 11.9373  0.001 ***
timepoint:site   6    33.52 0.04789  1.6659  0.026 *  
Residual       129   432.60 0.61799                   
Total          140   700.00 1.00000 

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:9],factors=por_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(por_data_afdw[ ,5:9], "euclidian"), por_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is no difference in dispersion by site, so therefore differences are likely due to  centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:9],factors=por_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(por_data_afdw[ ,5:9], "euclidian"), por_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is no significant difference in dispersion by timepoint (rounds to p=0.05), so differences are likely due to only centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA<-ggplot(por_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  ylim(-6,6)+
  xlim(-6,6)+
  ggtitle(expression(italic("Porites")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_por$`Pr(>F)`[1]), size=6, color=ifelse(z_por$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_por$`Pr(>F)`[2]), size=6, color=ifelse(z_por$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_por$`Pr(>F)`[3]), size=6, color=ifelse(z_por$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.027", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position=c(1,0.3)); porPCAcen

#build a plot for the legend
legend_plot<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=TRUE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="bottom")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints<-por.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAfull<-porPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAfull
```

Add bi plot with loadings  
```{r}
#1. make plot with dots
porArrows<-ggplot2::autoplot(scaled_por_afdw, data=por_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=-0.1, loadings.label.hjust=0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-0.5,0.1)+
  #ylim(-0.5, 0.5)+
    ggtitle(expression(italic("Porites")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));porArrows
```

##### Generate plot for POR haplotypes 

Calculate centroid locations for each haplotype and timepoint and pull PC1 and PC2 locations.      
```{r}
por_data_afdw_haplotype<-master%>%
  select(colony_id_corr, timepoint, species, haplotype, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Respiration=Rd, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

por_data_afdw_haplotype<-por_data_afdw_haplotype[complete.cases(por_data_afdw_haplotype), ]

por_data_afdw_haplotype[,-c(1:4)]<-log(por_data_afdw_haplotype[,-c(1:4)]+1)

scaled_por_afdw_haplotype<-prcomp(por_data_afdw_haplotype[c(5:9)], scale=TRUE, center=TRUE) 

por_info_haplotype<-por_data_afdw_haplotype[c(2,4)]

por_data_haplotype<-scaled_por_afdw_haplotype%>%
  augment(por_info_haplotype)%>%
  group_by(timepoint, haplotype)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids_haplotype<-por_data_haplotype %>% 
  select(timepoint, haplotype, PC1.mean, PC2.mean)%>%
  group_by(timepoint, haplotype)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA_haplotype<-ggplot(por_data_haplotype, aes(.fittedPC1, .fittedPC2, color=haplotype)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("black", "darkgray")) +
  scale_fill_manual(values=c("black", "darkgray")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  ylim(-6,6)+
  xlim(-6,6)+
  ggtitle("Host")+
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=18, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA_haplotype
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=FALSE) + 
  #scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position=c(1,0.3)); porPCAcen_haplotype

#build a plot for the legend
legend_plot_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=TRUE) + 
  #scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="bottom")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints_haplotype<-por.centroids_haplotype%>%
  gather(variable, value, -(timepoint:haplotype)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAhost_haplotype<-porPCAcen_haplotype + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAhost_haplotype
```

#### Pocillopora   

Generate PCA using scaled (scaled_poc_afdw) from dataframe (poc_data_afdw). 
 
```{r}
poc_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Respiration=Rd, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

poc_data_afdw<-poc_data_afdw[complete.cases(poc_data_afdw), ]

poc_data_afdw[,-c(1:4)]<-log(poc_data_afdw[,-c(1:4)]+1)
 
scaled_poc_afdw<-prcomp(poc_data_afdw[c(5:9)], scale=TRUE, center=TRUE) 

poc_info<-poc_data_afdw[c(2,4)]

poc_data<-scaled_poc_afdw%>%
  augment(poc_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

poc.centroids<-poc_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
``` 
  
```{r}
# scale data
vegan <- scale(poc_data_afdw[ ,5:9])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = poc_data_afdw, method='eu')
z_poc<-permanova
z_poc
```
adonis2(formula = vegan ~ timepoint * site, data = poc_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   175.68 0.24232 17.0879  0.001 ***
site             2    28.47 0.03927  4.1534  0.001 ***
timepoint:site   6    61.63 0.08500  2.9971  0.001 ***
Residual       134   459.22 0.63341                   
Total          145   725.00 1.00000 

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:9],factors=poc_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(poc_data_afdw[ ,5:9], "euclidian"), poc_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is no difference in dispersion by site, so therefore differences are likely due to  centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:9],factors=poc_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(poc_data_afdw[ ,5:9], "euclidian"), poc_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is a significant difference in dispersion by timepoint, so differences are likely due to dispersion and centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
pocPCA<-ggplot(poc_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Pocillopora")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_poc$`Pr(>F)`[1]), size=6, color=ifelse(z_poc$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_poc$`Pr(>F)`[2]), size=6, color=ifelse(z_poc$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_poc$`Pr(>F)`[3]), size=6, color=ifelse(z_poc$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.001", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         plot.margin = margin(1, 1, 1, 1, "cm"),
         legend.title = element_blank(), 
         title = element_text(size=25, face="bold"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18,  face="bold"));pocPCA

```

Add centroids  

```{r}
#2. add centroids 
pocPCAcen<-pocPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=poc.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); pocPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-poc.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
pocPCAfull<-pocPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); pocPCAfull
```

Add bi plot with loadings  
```{r}
pocArrows<-ggplot2::autoplot(scaled_poc_afdw, data=poc_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0, loadings.label.hjust=-0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
    ggtitle(expression(italic("Pocillopora")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));pocArrows
```

#### Assemble all plots    

```{r}
PCA_full_panel<-plot_grid(acrPCAfull, pocPCAfull, porPCAfull, labels = c("D", "E", "F"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
PCA_full_panel_legend<-plot_grid(PCA_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Host.png", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Host.jpg", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Host.pdf", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
```  

Assemble biplots  
```{r}
biplot_full_panel<-plot_grid(acrArrows, pocArrows, porArrows, labels = c("D", "E", "F"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
biplot_full_panel_legend<-plot_grid(biplot_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Biplots_Host.png", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Host.pdf", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Host.jpg", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
``` 


### Matrix of PCA's by species and time  

Generate biplot showing groupings by site for each species and time point.  

#### Pocillopora  

```{r}
poc_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

poc_tp1_data<-poc_tp1_data[complete.cases(poc_tp1_data), ]

poc_tp1_data[,-c(1:4)]<-log(poc_tp1_data[,-c(1:4)]+1)
 
poc_tp1_scaled<-prcomp(poc_tp1_data[c(5:9)], scale=TRUE, center=TRUE) 

pocTP1<-ggplot2::autoplot(poc_tp1_scaled, data=poc_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP1
```

```{r}
poc_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
poc_tp2_data<-poc_tp2_data[complete.cases(poc_tp2_data), ]

poc_tp2_data[,-c(1:4)]<-log(poc_tp2_data[,-c(1:4)]+1)
 
poc_tp2_scaled<-prcomp(poc_tp2_data[c(5:9)], scale=TRUE, center=TRUE) 

pocTP2<-ggplot2::autoplot(poc_tp2_scaled, data=poc_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP2
```

```{r}
poc_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Pocillopora")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
poc_tp3_data<-poc_tp3_data[complete.cases(poc_tp3_data), ]

poc_tp3_data[,-c(1:4)]<-log(poc_tp3_data[,-c(1:4)]+1)
 
poc_tp3_scaled<-prcomp(poc_tp3_data[c(5:9)], scale=TRUE, center=TRUE) 

pocTP3<-ggplot2::autoplot(poc_tp3_scaled, data=poc_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP3
```

```{r}
poc_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Pocillopora")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
poc_tp4_data<-poc_tp4_data[complete.cases(poc_tp4_data), ]

poc_tp4_data[,-c(1:4)]<-log(poc_tp4_data[,-c(1:4)]+1)
 
poc_tp4_scaled<-prcomp(poc_tp4_data[c(5:9)], scale=TRUE, center=TRUE) 

pocTP4<-ggplot2::autoplot(poc_tp4_scaled, data=poc_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP4
```


#### Acropora  

```{r}
acr_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Acropora")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

acr_tp1_data<-acr_tp1_data[complete.cases(acr_tp1_data), ]

acr_tp1_data[,-c(1:4)]<-log(acr_tp1_data[,-c(1:4)]+1)
 
acr_tp1_scaled<-prcomp(acr_tp1_data[c(5:9)], scale=TRUE, center=TRUE) 

acrTP1<-ggplot2::autoplot(acr_tp1_scaled, data=acr_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Acropora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP1
```

```{r}
acr_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Acropora")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
acr_tp2_data<-acr_tp2_data[complete.cases(acr_tp2_data), ]

acr_tp2_data[,-c(1:4)]<-log(acr_tp2_data[,-c(1:4)]+1)
 
acr_tp2_scaled<-prcomp(acr_tp2_data[c(5:9)], scale=TRUE, center=TRUE) 

acrTP2<-ggplot2::autoplot(acr_tp2_scaled, data=acr_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP2
```

```{r}
acr_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Acropora")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
acr_tp3_data<-acr_tp3_data[complete.cases(acr_tp3_data), ]

acr_tp3_data[,-c(1:4)]<-log(acr_tp3_data[,-c(1:4)]+1)
 
acr_tp3_scaled<-prcomp(acr_tp3_data[c(5:9)], scale=TRUE, center=TRUE) 

acrTP3<-ggplot2::autoplot(acr_tp3_scaled, data=acr_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP3
```

```{r}
acr_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Acropora")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
acr_tp4_data<-acr_tp4_data[complete.cases(acr_tp4_data), ]

acr_tp4_data[,-c(1:4)]<-log(acr_tp4_data[,-c(1:4)]+1)
 
acr_tp4_scaled<-prcomp(acr_tp4_data[c(5:9)], scale=TRUE, center=TRUE) 

acrTP4<-ggplot2::autoplot(acr_tp4_scaled, data=acr_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP4
```



#### Porites  

```{r}
por_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint1")

por_tp1_data<-por_tp1_data[complete.cases(por_tp1_data), ]

por_tp1_data[,-c(1:4)]<-log(por_tp1_data[,-c(1:4)]+1)
 
por_tp1_scaled<-prcomp(por_tp1_data[c(5:9)], scale=TRUE, center=TRUE) 

porTP1<-ggplot2::autoplot(por_tp1_scaled, data=por_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Porites)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP1
```

```{r}
por_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint2")
 
por_tp2_data<-por_tp2_data[complete.cases(por_tp2_data), ]

por_tp2_data[,-c(1:4)]<-log(por_tp2_data[,-c(1:4)]+1)
 
por_tp2_scaled<-prcomp(por_tp2_data[c(5:9)], scale=TRUE, center=TRUE) 

porTP2<-ggplot2::autoplot(por_tp2_scaled, data=por_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP2
```

```{r}
por_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint3")
 
por_tp3_data<-por_tp3_data[complete.cases(por_tp3_data), ]

por_tp3_data[,-c(1:4)]<-log(por_tp3_data[,-c(1:4)]+1)
 
por_tp3_scaled<-prcomp(por_tp3_data[c(5:9)], scale=TRUE, center=TRUE) 

porTP3<-ggplot2::autoplot(por_tp3_scaled, data=por_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP3
```

```{r}
por_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(species=="Porites")%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)%>%
  filter(timepoint=="timepoint4")
 
por_tp4_data<-por_tp4_data[complete.cases(por_tp4_data), ]

por_tp4_data[,-c(1:4)]<-log(por_tp4_data[,-c(1:4)]+1)
 
por_tp4_scaled<-prcomp(por_tp4_data[c(5:9)], scale=TRUE, center=TRUE) 

porTP4<-ggplot2::autoplot(por_tp4_scaled, data=por_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
 # ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP4
```

#### Assemble grid plot of all metrics by species and time  
 
```{r}
all_grid<-plot_grid(acrTP1, acrTP2, acrTP3, acrTP4, pocTP1, pocTP2, pocTP3, pocTP4, porTP1, porTP2, porTP3, porTP4, ncol=4, nrow=3)
all_grid2<-plot_grid(all_grid, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Matrix_Host_Responses.png", plot=all_grid2, dpi=500, width=20, height=20, units="in")

```

## Symbiont Responses  

### PCA's for centroid and trajectories  

#### Acropora  

Generate PCA using scaled (scaled_acr_afdw) from dataframe (acr_data_afdw).   
  
```{r}
acr_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell, SH_Ratio=Ratio_AFDW.mg.cm2)

acr_data_afdw<-acr_data_afdw[complete.cases(acr_data_afdw), ]

acr_data_afdw[,-c(1:4)]<-log(acr_data_afdw[,-c(1:4)]+1)

scaled_acr_afdw<-prcomp(acr_data_afdw[c(5:13)], scale=TRUE, center=TRUE) 

acr_info<-acr_data_afdw[c(2,4)]

acr_data<-scaled_acr_afdw%>%
  augment(acr_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

acr.centroids<-acr_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Examine PERMANOVA results.  
 
```{r}
# scale data
vegan <- scale(acr_data_afdw[ ,5:13])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = acr_data_afdw, method='eu')
z_acr<-permanova
z_acr
``` 
 
adonis2(formula = vegan ~ timepoint * site, data = acr_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   255.03 0.25529 13.6811  0.001 ***
site             2    60.94 0.06100  4.9037  0.001 ***
timepoint:site   6    61.66 0.06172  1.6539  0.022 *  
Residual       100   621.37 0.62199                   
Total          111   999.00 1.00000 

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:13],factors=acr_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(acr_data_afdw[ ,5:13], "euclidian"), acr_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to  both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=acr_data_afdw[ ,5:13],factors=acr_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(acr_data_afdw[ ,5:13], "euclidian"), acr_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is a significant difference in dispersion by timepoint, so differences are likely due to dispersion and centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
acrPCA<-ggplot(acr_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Acropora")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_acr$`Pr(>F)`[1]), size=6, color=ifelse(z_acr$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_acr$`Pr(>F)`[2]), size=6, color=ifelse(z_acr$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_acr$`Pr(>F)`[3]), size=6, color=ifelse(z_acr$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.021", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"), 
         axis.title = element_text(size=18));acrPCA

```

Add centroids  

```{r}
#2. add centroids 
acrPCAcen<-acrPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=acr.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); acrPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-acr.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
acrPCAfull<-acrPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); acrPCAfull
```

Add bi plot with loadings  
```{r}
#1. make plot with dots
acrArrows<-ggplot2::autoplot(scaled_acr_afdw, data=acr_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0.5, loadings.label.hjust=-0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
    ggtitle(expression(italic("Acropora")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));acrArrows
```

#### Porites

Generate PCA using scaled (scaled_por_afdw) from dataframe (por_data_afdw).   
 
Calculate centroid locations for each site and timepoint and pull PC1 and PC2 locations.
 
```{r}
por_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Porites")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell, SH_Ratio=Ratio_AFDW.mg.cm2)

por_data_afdw<-por_data_afdw[complete.cases(por_data_afdw), ]

por_data_afdw[,-c(1:4)]<-log(por_data_afdw[,-c(1:4)]+1)

scaled_por_afdw<-prcomp(por_data_afdw[c(5:13)], scale=TRUE, center=TRUE) 

por_info<-por_data_afdw[c(2,4)]

por_data<-scaled_por_afdw%>%
  augment(por_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids<-por_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

```{r}
# scale data
vegan <- scale(por_data_afdw[ ,5:13])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = por_data_afdw, method='eu')
z_por<-permanova
z_por
```

adonis2(formula = vegan ~ timepoint * site, data = por_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   210.14 0.14967 10.5910  0.001 ***
site             2   155.16 0.11051 11.7296  0.001 ***
timepoint:site   6    79.69 0.05676  2.0081  0.003 ** 
Residual       145   959.01 0.68306                   
Total          156  1404.00 1.00000     

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:13],factors=por_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(por_data_afdw[ ,5:13], "euclidian"), por_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to  both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=por_data_afdw[ ,5:13],factors=por_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(por_data_afdw[ ,5:13], "euclidian"), por_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is no significant difference in dispersion by timepoint, so differences are likely due to only centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA<-ggplot(por_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  ylim(-6,6)+
  xlim(-6,6)+
  ggtitle(expression(italic("Porites")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_por$`Pr(>F)`[1]), size=6, color=ifelse(z_por$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_por$`Pr(>F)`[2]), size=6, color=ifelse(z_por$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_por$`Pr(>F)`[3]), size=6, color=ifelse(z_por$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.001", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position=c(1,0.3)); porPCAcen

#build a plot for the legend
legend_plot<-porPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=por.centroids, size=3, show.legend=TRUE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="bottom")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend <- get_legend(
  # create some space to the left of the legend
  legend_plot + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints<-por.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAfull<-porPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAfull
```

Add bi plot with loadings  
```{r}
#1. make plot with dots
porArrows<-ggplot2::autoplot(scaled_por_afdw, data=por_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=-0.1, loadings.label.hjust=0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-0.5,0.1)+
  #ylim(-0.5, 0.5)+
    ggtitle(expression(italic("Porites")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));porArrows
```

Assemble final figure with inset biplot.  
```{r}
#porFullPCA<-ggdraw(porPCAfull)+ #+ theme_half_open(12)) +
 # draw_plot(porArrows, .5, .5, .5, .5); porFullPCA #x, y, w, h

#ggsave(filename="Figures/FullPCA_Porites.pdf", plot=porFullPCA, dpi=300, width=12, height=8, units="in")
```

##### Plot POR haplotypes 

```{r}
por_data_afdw_haplotype<-master%>%
  select(colony_id_corr, timepoint, species, haplotype, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Porites")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell, SH_Ratio=Ratio_AFDW.mg.cm2)

por_data_afdw_haplotype<-por_data_afdw_haplotype[complete.cases(por_data_afdw_haplotype), ]

por_data_afdw_haplotype[,-c(1:4)]<-log(por_data_afdw_haplotype[,-c(1:4)]+1)

scaled_por_afdw_haplotype<-prcomp(por_data_afdw_haplotype[c(5:13)], scale=TRUE, center=TRUE) 

por_info_haplotype<-por_data_afdw_haplotype[c(2,4)]

por_data_haplotype<-scaled_por_afdw_haplotype%>%
  augment(por_info_haplotype)%>%
  group_by(timepoint, haplotype)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

por.centroids_haplotype<-por_data_haplotype %>% 
  select(timepoint, haplotype, PC1.mean, PC2.mean)%>%
  group_by(timepoint, haplotype)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
```

Assemble plot with background points
```{r}
#1. make plot with dots
porPCA_haplotype<-ggplot(por_data_haplotype, aes(.fittedPC1, .fittedPC2, color=haplotype)) + 
  geom_point(size = 4, alpha=0.2, show.legend=TRUE)+
  scale_colour_manual(values=c("black", "darkgray")) +
  scale_fill_manual(values=c("black", "darkgray")) + 
  theme_classic()+
  ylab("PC2")+
  xlab("PC1")+
  #ylim(-6,6)+
  #xlim(-6,6)+
  ggtitle("Symbiont")+
  theme(legend.text = element_text(size=18), 
         legend.position="right",
         plot.background = element_blank(),
         legend.title = element_blank(), 
         axis.text = element_text(size=18), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         title = element_text(size=18, face="bold"),
         axis.title = element_text(size=18,  face="bold"));porPCA_haplotype
```

Add centroids  

```{r}
#2. add centroids 
porPCAcen_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=TRUE); porPCAcen_haplotype

#build a plot for the legend
legend_plot_haplotype<-porPCA_haplotype + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=haplotype), data=por.centroids_haplotype, size=3, show.legend=TRUE) + 
  scale_colour_manual(values=c("black", "darkgray"))+
  theme(legend.position="bottom")+theme(legend.text=element_text(size=18))

# extract the legend from one of the plots
legend_haplotype <- get_legend(
  # create some space to the left of the legend
  legend_plot_haplotype + theme(legend.box.margin = margin(1,1,1,1))
)

```

Add segments

```{r}
#3. add segments
segpoints_haplotype<-por.centroids_haplotype%>%
  gather(variable, value, -(timepoint:haplotype)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
porPCAsym_haplotype<-porPCAcen_haplotype + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = haplotype), data = segpoints_haplotype, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); porPCAsym_haplotype
```

#### Pocillopora   

Generate PCA using scaled (scaled_poc_afdw) from dataframe (poc_data_afdw). 
 
```{r}
poc_data_afdw<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Pocillopora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell, SH_Ratio=Ratio_AFDW.mg.cm2)

poc_data_afdw<-poc_data_afdw[complete.cases(poc_data_afdw), ]

poc_data_afdw[,-c(1:4)]<-log(poc_data_afdw[,-c(1:4)]+1)

poc_data_afdw<-poc_data_afdw[complete.cases(poc_data_afdw), ]
 
scaled_poc_afdw<-prcomp(poc_data_afdw[c(5:13)], scale=TRUE, center=TRUE) 

poc_info<-poc_data_afdw[c(2,4)]

poc_data<-scaled_poc_afdw%>%
  augment(poc_info)%>%
  group_by(timepoint, site)%>%
  mutate(PC1.mean = mean(.fittedPC1),
         PC2.mean = mean(.fittedPC2))

poc.centroids<-poc_data %>% 
  select(timepoint, site, PC1.mean, PC2.mean)%>%
  group_by(timepoint, site)%>%
  summarise(PC1.mean = mean(PC1.mean),
         PC2.mean = mean(PC2.mean))
``` 
  
```{r}
# scale data
vegan <- scale(poc_data_afdw[ ,5:13])

# PerMANOVA - partitioning the euclidean distance matrix by species
permanova<-adonis2(vegan ~ timepoint*site, data = poc_data_afdw, method='eu')
z_poc<-permanova
z_poc
```
adonis2(formula = vegan ~ timepoint * site, data = poc_data_afdw, method = "eu")
                Df SumOfSqs      R2       F Pr(>F)    
timepoint        3   285.86 0.21905 14.4893  0.001 ***
site             2    39.15 0.03000  2.9764  0.003 ** 
timepoint:site   6    98.78 0.07569  2.5034  0.001 ***
Residual       134   881.22 0.67526                   
Total          145  1305.00 1.00000 

Examine dispersion for this model by site 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:13],factors=poc_data_afdw$site, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_site<-betadisper(vegdist(poc_data_afdw[ ,5:13], "euclidian"), poc_data_afdw$site)
anova(beta_site)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_site)
```
There is a difference in dispersion by site, so therefore differences are likely due to  both dispersion and centroid location. 

Examine dispersion for this model by season 
```{r}
#conduct pairwise adonis to look at pairwise permanova comparisons
pairwise.adonis(x=poc_data_afdw[ ,5:13],factors=poc_data_afdw$timepoint, sim.function="vegdist",  sim.method='euclidean', p.adjust.m='bonferroni')

#evaluate dispersion
beta_time<-betadisper(vegdist(poc_data_afdw[ ,5:13], "euclidian"), poc_data_afdw$timepoint)
anova(beta_time)

#conduct posthoc analysis of betadispersion
TukeyHSD(beta_time)
```
There is a significant difference in dispersion by timepoint, so differences are likely due to dispersion and centroid location.

Assemble plot with background points
```{r}
#1. make plot with dots
pocPCA<-ggplot(poc_data, aes(.fittedPC1, .fittedPC2, color=site)) + 
  geom_point(size = 4, alpha=0.2, show.legend=FALSE)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  ylim(-6,6)+
  xlim(-6,6)+
  ylab("PC2")+
  xlab("PC1")+
  ggtitle(expression(italic("Pocillopora")))+
  #geom_text(x=3, y=5.5, label=paste("T=",z_poc$`Pr(>F)`[1]), size=6, color=ifelse(z_poc$`Pr(>F)`[1] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4.75, label=paste("S=",z_poc$`Pr(>F)`[2]), size=6, color=ifelse(z_poc$`Pr(>F)`[2] < 0.05, "black", "darkgray")) + 
  #geom_text(x=3, y=4, label=paste("TxS=",z_poc$`Pr(>F)`[3]), size=6, color=ifelse(z_poc$`Pr(>F)`[3] < 0.05, "black", "darkgray")) + 
  geom_text(x=3, y=5.5, label="T=0.001", size=6, color="black") + 
  geom_text(x=3, y=4.75, label="S=0.001", size=6, color="black") + 
  geom_text(x=3, y=4, label="TxS=0.001", size=6, color="black") + 
  theme(legend.text = element_text(size=18), 
         legend.position="none",
         plot.background = element_blank(),
         plot.margin = margin(1, 1, 1, 1, "cm"),
         legend.title = element_blank(), 
         title = element_text(size=25, face="bold"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18,  face="bold"));pocPCA

```

Add centroids  

```{r}
#2. add centroids 
pocPCAcen<-pocPCA + geom_point(aes(x=PC1.mean, y=PC2.mean, colour=site), data=poc.centroids, size=3, show.legend=FALSE) + 
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633"))+
  theme(legend.position="none"); pocPCAcen

```

Add segments

```{r}
#3. add segments
segpoints<-poc.centroids%>%
  gather(variable, value, -(timepoint:site)) %>%
  unite(temp, timepoint, variable) %>%
  spread(temp, value)
  
pocPCAfull<-pocPCAcen + 
  geom_segment(aes(x = timepoint1_PC1.mean, y = timepoint1_PC2.mean, xend = timepoint2_PC1.mean, yend = timepoint2_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint2_PC1.mean, y = timepoint2_PC2.mean, xend = timepoint3_PC1.mean, yend = timepoint3_PC2.mean, colour = site), data = segpoints, size=2, show.legend=FALSE) +
  geom_segment(aes(x = timepoint3_PC1.mean, y = timepoint3_PC2.mean, xend = timepoint4_PC1.mean, yend = timepoint4_PC2.mean, colour = site), data = segpoints, size=2, arrow = arrow(length=unit(0.5,"cm")), show.legend=FALSE); pocPCAfull
```

Add bi plot with loadings  
```{r}

pocArrows<-ggplot2::autoplot(scaled_poc_afdw, data=poc_data_afdw, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, loadings.label.size=3.5, loadings.label.vjust=0, loadings.label.hjust=-0.1, loadings.label.repel=TRUE, size=4, alpha=0.0) + 
  #geom_point(aes(fill=site, colour=site), alpha=0.2)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
    ggtitle(expression(italic("Pocillopora")))+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         #axis.text.x=element_blank(),
         #axis.text.y=element_blank(),
         plot.background = element_blank(),
         legend.title = element_blank(), 
         plot.margin = margin(1, 1, 1, 1, "cm"),
         axis.text = element_text(size=18), 
         title = element_text(size=25, face="bold"),
         axis.title = element_text(size=20));pocArrows
```

#### Assemble all plots    

```{r}
PCA_full_panel<-plot_grid(acrPCAfull, pocPCAfull, porPCAfull, labels = c("G", "H", "I"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
PCA_full_panel_legend<-plot_grid(PCA_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Symbiont.png", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Symbiont.pdf", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Panel_PCAs_Symbiont.jpg", plot=PCA_full_panel_legend, dpi=500, width=20, height=6, units="in")
```  

Assemble biplots  
```{r}
biplot_full_panel<-plot_grid(acrArrows, pocArrows, porArrows, labels = c("G", "H", "I"), ncol=3, nrow=1, rel_heights= c(1,1,1), rel_widths = c(1,1,1), label_size = 20, label_y=0.85, align="vh")

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
biplot_full_panel_legend<-plot_grid(biplot_full_panel, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Full_Biplots_Symbiont.png", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Symbiont.pdf", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
ggsave(filename="Figures/Multivariate/Full_Biplots_Symbiont.jpg", plot=biplot_full_panel_legend, dpi=500, width=20, height=6, units="in")
``` 

### Matrix of PCA's by species and time  

Generate biplot showing groupings by site for each species and time point.  

#### Pocillopora  

```{r}
poc_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Pocillopora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2)%>%
  filter(timepoint=="timepoint1")

poc_tp1_data<-poc_tp1_data[complete.cases(poc_tp1_data), ]

poc_tp1_data[,-c(1:4)]<-log(poc_tp1_data[,-c(1:4)]+1)
 
poc_tp1_scaled<-prcomp(poc_tp1_data[c(5:13)], scale=TRUE, center=TRUE) 

pocTP1<-ggplot2::autoplot(poc_tp1_scaled, data=poc_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP1
```

```{r}
poc_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Pocillopora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint2")
 
poc_tp2_data<-poc_tp2_data[complete.cases(poc_tp2_data), ]

poc_tp2_data[,-c(1:4)]<-log(poc_tp2_data[,-c(1:4)]+1)
 
poc_tp2_scaled<-prcomp(poc_tp2_data[c(5:13)], scale=TRUE, center=TRUE) 

pocTP2<-ggplot2::autoplot(poc_tp2_scaled, data=poc_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP2
```

```{r}
poc_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Pocillopora")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint3")
 
poc_tp3_data<-poc_tp3_data[complete.cases(poc_tp3_data), ]

poc_tp3_data[,-c(1:4)]<-log(poc_tp3_data[,-c(1:4)]+1)
 
poc_tp3_scaled<-prcomp(poc_tp3_data[c(5:13)], scale=TRUE, center=TRUE) 

pocTP3<-ggplot2::autoplot(poc_tp3_scaled, data=poc_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP3
```

```{r}
poc_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(species=="Pocillopora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint4")
 
poc_tp4_data<-poc_tp4_data[complete.cases(poc_tp4_data), ]

poc_tp4_data[,-c(1:4)]<-log(poc_tp4_data[,-c(1:4)]+1)
 
poc_tp4_scaled<-prcomp(poc_tp4_data[c(5:13)], scale=TRUE, center=TRUE) 

pocTP4<-ggplot2::autoplot(poc_tp4_scaled, data=poc_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));pocTP4
```


#### Acropora  

```{r}
acr_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint1")

acr_tp1_data<-acr_tp1_data[complete.cases(acr_tp1_data), ]

acr_tp1_data[,-c(1:4)]<-log(acr_tp1_data[,-c(1:4)]+1)
 
acr_tp1_scaled<-prcomp(acr_tp1_data[c(5:13)], scale=TRUE, center=TRUE) 

acrTP1<-ggplot2::autoplot(acr_tp1_scaled, data=acr_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Acropora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP1
```

```{r}
acr_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint2")
 
acr_tp2_data<-acr_tp2_data[complete.cases(acr_tp2_data), ]

acr_tp2_data[,-c(1:4)]<-log(acr_tp2_data[,-c(1:4)]+1)
 
acr_tp2_scaled<-prcomp(acr_tp2_data[c(5:13)], scale=TRUE, center=TRUE) 

acrTP2<-ggplot2::autoplot(acr_tp2_scaled, data=acr_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP2
```

```{r}
acr_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Total_Chl<13)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  filter(species=="Acropora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint3")
 
acr_tp3_data<-acr_tp3_data[complete.cases(acr_tp3_data), ]

acr_tp3_data[,-c(1:4)]<-log(acr_tp3_data[,-c(1:4)]+1)
 
acr_tp3_scaled<-prcomp(acr_tp3_data[c(5:13)], scale=TRUE, center=TRUE) 

acrTP3<-ggplot2::autoplot(acr_tp3_scaled, data=acr_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP3
```

```{r}
acr_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Acropora")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint4")
 
acr_tp4_data<-acr_tp4_data[complete.cases(acr_tp4_data), ]

acr_tp4_data[,-c(1:4)]<-log(acr_tp4_data[,-c(1:4)]+1)
 
acr_tp4_scaled<-prcomp(acr_tp4_data[c(5:13)], scale=TRUE, center=TRUE) 

acrTP4<-ggplot2::autoplot(acr_tp4_scaled, data=acr_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));acrTP4
```



#### Porites  

```{r}
por_tp1_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Porites")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint1")

por_tp1_data<-por_tp1_data[complete.cases(por_tp1_data), ]

por_tp1_data[,-c(1:4)]<-log(por_tp1_data[,-c(1:4)]+1)
 
por_tp1_scaled<-prcomp(por_tp1_data[c(5:13)], scale=TRUE, center=TRUE) 

porTP1<-ggplot2::autoplot(por_tp1_scaled, data=por_tp1_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0, loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("January 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  annotate("text", x=-16, y=0, label="italic(Porites)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 2, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP1
```

```{r}
por_tp2_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Porites")%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint2")
 
por_tp2_data<-por_tp2_data[complete.cases(por_tp2_data), ]
 
por_tp2_data[,-c(1:4)]<-log(por_tp2_data[,-c(1:4)]+1)

por_tp2_scaled<-prcomp(por_tp2_data[c(5:13)], scale=TRUE, center=TRUE) 

porTP2<-ggplot2::autoplot(por_tp2_scaled, data=por_tp2_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("March 2020")+
  coord_cartesian(xlim=c(-8, 8), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP2
```

```{r}
por_tp3_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Porites")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint3")
 
por_tp3_data<-por_tp3_data[complete.cases(por_tp3_data), ]

por_tp3_data[,-c(1:4)]<-log(por_tp3_data[,-c(1:4)]+1)
 
por_tp3_scaled<-prcomp(por_tp3_data[c(5:13)], scale=TRUE, center=TRUE) 

porTP3<-ggplot2::autoplot(por_tp3_scaled, data=por_tp3_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
  #ggtitle("September 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP3
```

```{r}
por_tp4_data<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Ic<1000)%>%
  #filter(Total_Chl<13)%>%
  filter(species=="Porites")%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW, Total_Chl=Total_Chl, Total_Chl_Cell=Total_Chl_cell)%>%
  filter(timepoint=="timepoint4")
 
por_tp4_data<-por_tp4_data[complete.cases(por_tp4_data), ]

por_tp4_data[,-c(1:4)]<-log(por_tp4_data[,-c(1:4)]+1)
 
por_tp4_scaled<-prcomp(por_tp4_data[c(5:13)], scale=TRUE, center=TRUE) 

porTP4<-ggplot2::autoplot(por_tp4_scaled, data=por_tp4_data, loadings=TRUE,  colour="site", loadings.label.colour="black", loadings.colour="black", loadings.label=TRUE, frame=FALSE, frame.type="convex", frame.level=0.95, loadings.label.size=5, loadings.label.repel=TRUE, size=4, alpha=1.0,   loadings.arrow = grid::arrow(length = grid::unit(10, "points")), scale=0) + 
  #geom_point(aes(colour=site), size=3)+
  scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  theme_classic()+
  #xlim(-.3,.6)+
  #ylim(-.3, .3)+
 # ggtitle("November 2020")+
  coord_cartesian(xlim=c(-8, 9), ylim=c(-5, 5), clip = "off")+
  #annotate("text", x=-8, y=0, label="italic(Pocillopora)", fontface="bold", parse = TRUE, angle=90, size=10)+
   theme(legend.text = element_text(size=18), 
         legend.position="none",
         axis.text.x=element_text(colour="black"),
         axis.text.y=element_text(colour="black"),
         plot.background = element_blank(),
         plot.title = element_text(size=22, face="bold", hjust = 0.5), 
         plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
         axis.text = element_text(size=18), 
         axis.title = element_text(size=18));porTP4
```

#### Assemble grid plot of all metrics by species and time  

```{r}
all_grid<-plot_grid(acrTP1, acrTP2, acrTP3, acrTP4, pocTP1, pocTP2, pocTP3, pocTP4, porTP1, porTP2, porTP3, porTP4, ncol=4, nrow=3)
all_grid2<-plot_grid(all_grid, legend, rel_heights = c(4, .2), ncol=1, nrow=2)

ggsave(filename="Figures/Multivariate/Matrix_Symbiont_Responses.png", plot=all_grid2, dpi=500, width=20, height=20, units="in")

```




## Build panel for haplotype plots 

porPCAfull_haplotype = holobiont
porPCAhost_haplotype = host 

```{r}
porPCAsym_haplotype 
porPCAfull_haplotype
porPCAhost_haplotype

haplotype_grid<-plot_grid(porPCAfull_haplotype, porPCAhost_haplotype, porPCAsym_haplotype, ncol=3, rel_widths=c(0.7, 0.7, 1.1))

ggsave(filename="Figures/Multivariate/Haplotype_POR_Panel.pdf", haplotype_grid, width=18, height=5)
```


