---
title: "Plasticity analysis of E5 time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "08/22/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 
Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up     

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("reshape2")) install.packages("reshape2")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(reshape2)
```

# Load dataframe  

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

# Holobiont responses   

Calculate plasticity as total euclidean distance between multivariate physiology positions generated in a PCA between time points for all metrics.  

```{r}
data_all<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

head(data_all) 

data_all<-data_all[complete.cases(data_all), ]

data_all[,-c(1:4)]<-log(data_all[,-c(1:4)]+1)

scaled_all<-prcomp(data_all[c(5:18)], scale=TRUE, center=TRUE) 

all_info<-data_all[c(1:4)]

#generate a euclidean distance matrix between all samples (colony x time point)
all_data<-scaled_all%>%
  augment(all_info)%>%
  group_by(timepoint, site_code, species, colony_id_corr)%>%
  mutate(code=paste(timepoint, colony_id_corr))

all_data<-as.data.frame(all_data)

row.names(all_data)<-all_data$code

all_data<-all_data[6:19]
head(all_data)

dist_all<-dist(all_data, diag = FALSE, upper = FALSE, method="euclidean") #calculate euclidian distance

dist_all <- melt(as.matrix(dist_all), varnames = c("row", "col"))

head(dist_all)
```

Now calculate the dispersion (total distance traveled in euclidian distance) for each colony. We are calculating dispersion only for colony that were found at at least 3 timepoints so that we have a good representation of dispersion. Distances are calculated as the distance between timepoint 1 and 2, timepoint 2 and 3, and time point 3 and 4 summed together. If a colony was missing at one time point, the distance from the two adjancent time points was used. For example, if a colony was missing at time point 3, we will use distances from timepoint 1 to 2, timepoint 1 to 3, and timepoint 3 to 4.   

```{r}
#calculate distances between subsequent time points (T1-T2, T2-T3, T3-T4) and sum them together
dist_all_calc<-dist_all%>%
  separate(row, into=c("Timepoint1", "Colony1"), sep=" ")%>%
  separate(col, into=c("Timepoint2", "Colony2"), sep=" ")%>%
  filter(value>0.0)%>% #remove distances between the same points
  mutate(condition=if_else(Colony1==Colony2, "TRUE", "FALSE"))%>%
  filter(condition=="TRUE")%>% #keep within colony comparisons
  arrange(Colony1, Timepoint1, Timepoint2)%>% #arrange by time points
  select(Timepoint1, Timepoint2, Colony1, value)%>%
  mutate(comparison=paste(Timepoint1, "-", Timepoint2))%>% #add a comparison column
  filter(!duplicated(paste(Colony1, value)))%>% #remove duplicated/reciprocal timepoint matches 
  select(!Timepoint1)%>%
  select(!Timepoint2)%>%
  pivot_wider(names_from = comparison, values_from = value, id_cols=Colony1)%>%
  rename("T1_T2"="timepoint1 - timepoint2", 
         "T2_T3"="timepoint2 - timepoint3", 
         "T3-T4"="timepoint3 - timepoint4", 
         "T1_T3"="timepoint1 - timepoint3", 
         "T1_T4"="timepoint1 - timepoint4",
         "T3_T4"="timepoint3 - timepoint4", 
         "T2_T4"="timepoint2 - timepoint4")%>%
  mutate(sum_1=rowSums(.[ , c("T1_T2", "T2_T3", "T3_T4")]))%>% #summarize rows with complete timeseries data 
  mutate(sum_2=case_when(is.na(T1_T2) ~ rowSums(.[ , c("T1_T3", "T3_T4")])))%>% #If T2 is missing, instead sum T1-T3 and T3-T4
  mutate(sum_3=case_when(is.na(T1_T4) ~ rowSums(.[ , c("T1_T2", "T2_T3")])))%>% #If T4 is missing, instead sum T1-T2 and T2-T3
 mutate(sum_4=case_when(is.na(T1_T3) ~ rowSums(.[ , c("T1_T2", "T2_T4")])))%>% #If T3 is missing, instead sum T1-T2 and T2-T4
  mutate(sum_5=case_when(is.na(T1_T2) & is.na(T1_T3) & is.na(T1_T4) ~ rowSums(.[ , c("T2_T3", "T3_T4")])))%>% #If T1 is missing, instead sum T2-T3 and T3-T4 
  mutate(colony_dispersion=coalesce(sum_1, sum_2, sum_3, sum_4, sum_5))%>%
  select(Colony1, colony_dispersion)
  
dist_all_calc$site_code<-master$site_code[match(dist_all_calc$Colony1, master$colony_id_corr)]
dist_all_calc$species<-master$species[match(dist_all_calc$Colony1, master$colony_id_corr)]

head(dist_all_calc)
```

Plot as a dot plot with an overlay for means of each group.  
```{r}
#Calculate means to add to 
dist_all_means<-dist_all_calc%>%
  group_by(species, site_code)%>%
  summarise(mean=mean(colony_dispersion, na.rm=TRUE), sd=sd(colony_dispersion, na.rm=TRUE), N=sum(!is.na(colony_dispersion)), se=sd/sqrt(N))

dist_all_dotplot<-dist_all_calc%>%
  
    ggplot(aes(x = site_code, y = colony_dispersion)) +
  
    #add individual data points 
    geom_point(aes(fill=species, color=species, group=interaction(species, site_code)), pch = 21, size=4, position = position_jitterdodge(0.2), alpha=0.3) + 
  
    #add mean points and lines between groups
    geom_line(data=dist_all_means, aes(y=mean, x=site_code, group=species, color=species), position=position_dodge(0.2), size=2) +
    geom_point(data=dist_all_means, aes(y=mean, x=site_code, color=species, fill=species), pch = 21, size=5, position = position_dodge(0.2), alpha=0.8) + 
  
    #add stats
    #geom_text(x=2, y=20, size=4, label="p(Site)=0.140", color="darkgray")+
    #geom_text(x=2, y=19, size=4, label="p(Species)<0.001", color="black")+
    #geom_text(x=2, y=18, size=4, label="p(Site x Species)=0.535", color="darkgray")+
  
    #aesthetics 
    scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_color_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_x_discrete(labels=c("Mahana\nLow", "Hilton\nMedium", "Manava\nHigh"))+
    xlab("Site") + 
    ylim(0,20)+
    ylab(expression(bold("Multivariate Dispersion")))+
    theme_classic() + 
    ggtitle("Holobiont")+
    theme(
      legend.position="right",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14, face="italic"),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=14, color="black"), 
      strip.text.x=element_text(face="italic", size=14),
      title=element_text(face="bold", size=16)
      ); dist_all_dotplot

```

Run anova on these values.  
```{r}
dist_all_model<-aov(colony_dispersion~site_code*species, data=dist_all_calc)
summary(dist_all_model)
```

# Host responses  

Calculate distances for only host responses.  

Generate a list of symbiont and host responses.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

Calculate plasticity as total euclidean distance between multivariate physiology positions generated in a PCA between time points for all metrics.  

```{r}
data_host<-master%>%
  select(colony_id_corr, timepoint, species, site_code, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

data_host<-data_host[complete.cases(data_host), ]

data_host[,-c(1:4)]<-log(data_host[,-c(1:4)]+1)

scaled_host<-prcomp(data_host[c(5:9)], scale=TRUE, center=TRUE) 

host_info<-data_host[c(1:4)]

#generate a euclidean distance matrix between all samples (colony x time point)
host_data<-scaled_host%>%
  augment(host_info)%>%
  group_by(timepoint, site_code, species, colony_id_corr)%>%
  mutate(code=paste(timepoint, colony_id_corr))

host_data<-as.data.frame(host_data)

row.names(host_data)<-host_data$code

host_data<-host_data[6:10]
head(host_data)

dist_host<-dist(host_data, diag = FALSE, upper = FALSE, method="euclidean")

dist_host <- melt(as.matrix(dist_host), varnames = c("row", "col"))

head(dist_host)
```

Now calculate the dispersion (total distance traveled in euclidian distance) for each colony. We are calculating dispersion only for colony that were found at at least 3 timepoints so that we have a good representation of dispersion. Distances are calculated as the distance between timepoint 1 and 2, timepoint 2 and 3, and time point 3 and 4 summed together. If a colony was missing at one time point, the distance from the two adjancent time points was used. For example, if a colony was missing at time point 3, we will use distances from timepoint 1 to 2, timepoint 1 to 3, and timepoint 3 to 4.   

```{r}
#calculate distances between subsequent time points (T1-T2, T2-T3, T3-T4) and sum them together
dist_host_calc<-dist_host%>%
  separate(row, into=c("Timepoint1", "Colony1"), sep=" ")%>%
  separate(col, into=c("Timepoint2", "Colony2"), sep=" ")%>%
  filter(value>0.0)%>% #remove distances between the same points
  mutate(condition=if_else(Colony1==Colony2, "TRUE", "FALSE"))%>%
  filter(condition=="TRUE")%>% #keep within colony comparisons
  arrange(Colony1, Timepoint1, Timepoint2)%>% #arrange by time points
  select(Timepoint1, Timepoint2, Colony1, value)%>%
  mutate(comparison=paste(Timepoint1, "-", Timepoint2))%>% #add a comparison column
  filter(!duplicated(paste(Colony1, value)))%>% #remove duplicated/reciprocal timepoint matches 
  select(!Timepoint1)%>%
  select(!Timepoint2)%>%
  pivot_wider(names_from = comparison, values_from = value, id_cols=Colony1)%>%
  rename("T1_T2"="timepoint1 - timepoint2", 
         "T2_T3"="timepoint2 - timepoint3", 
         "T3-T4"="timepoint3 - timepoint4", 
         "T1_T3"="timepoint1 - timepoint3", 
         "T1_T4"="timepoint1 - timepoint4",
         "T3_T4"="timepoint3 - timepoint4", 
         "T2_T4"="timepoint2 - timepoint4")%>%
  mutate(sum_1=rowSums(.[ , c("T1_T2", "T2_T3", "T3_T4")]))%>% #summarize rows with complete timeseries data 
  mutate(sum_2=case_when(is.na(T1_T2) ~ rowSums(.[ , c("T1_T3", "T3_T4")])))%>% #If T2 is missing, instead sum T1-T3 and T3-T4
  mutate(sum_3=case_when(is.na(T1_T4) ~ rowSums(.[ , c("T1_T2", "T2_T3")])))%>% #If T4 is missing, instead sum T1-T2 and T2-T3
 mutate(sum_4=case_when(is.na(T1_T3) ~ rowSums(.[ , c("T1_T2", "T2_T4")])))%>% #If T3 is missing, instead sum T1-T2 and T2-T4
  mutate(sum_5=case_when(is.na(T1_T2) & is.na(T1_T3) & is.na(T1_T4) ~ rowSums(.[ , c("T2_T3", "T3_T4")])))%>% #If T1 is missing, instead sum T2-T3 and T3-T4 
  mutate(colony_dispersion=coalesce(sum_1, sum_2, sum_3, sum_4, sum_5))%>%
  select(Colony1, colony_dispersion)
  
dist_host_calc$site_code<-master$site_code[match(dist_host_calc$Colony1, master$colony_id_corr)]
dist_host_calc$species<-master$species[match(dist_host_calc$Colony1, master$colony_id_corr)]

head(dist_host_calc)
```

Plot as a dot plot with an overlay for means of each group.  
```{r}
#Calculate means to add to 
dist_host_means<-dist_host_calc%>%
  group_by(species, site_code)%>%
  summarise(mean=mean(colony_dispersion, na.rm=TRUE), sd=sd(colony_dispersion, na.rm=TRUE), N=sum(!is.na(colony_dispersion)), se=sd/sqrt(N))

dist_host_dotplot<-dist_host_calc%>%
  
    ggplot(aes(x = site_code, y = colony_dispersion)) +
  
    #add individual data points 
    geom_point(aes(fill=species, color=species, group=interaction(species, site_code)), pch = 21, size=4, position = position_jitterdodge(0.2), alpha=0.3) + 
  
    #add mean points and lines between groups
    geom_line(data=dist_host_means, aes(y=mean, x=site_code, group=species, color=species), position=position_dodge(0.2), size=2) +
    geom_point(data=dist_host_means, aes(y=mean, x=site_code, color=species, fill=species), pch = 21, size=5, position = position_dodge(0.2), alpha=0.8) + 
  
    #add stats
    #geom_text(x=2, y=20, size=4, label="p(Site)=0.007", color="black")+
    #geom_text(x=2, y=19, size=4, label="p(Species)<0.001", color="black")+
    #geom_text(x=2, y=18, size=4, label="p(Site x Species)=0.154", color="darkgray")+
  
    #aesthetics 
    scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_color_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_x_discrete(labels=c("Mahana\nLow", "Hilton\nMedium", "Manava\nHigh"))+
    xlab("Site") + 
    ylim(0,20)+
    ylab(expression(bold("Multivariate Dispersion")))+
    theme_classic() + 
    ggtitle("Host")+
    theme(
      legend.position="right",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14, face="italic"),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=14, color="black"), 
      strip.text.x=element_text(face="italic", size=14),
      title=element_text(face="bold", size=16)
      ); dist_host_dotplot

```

Run anova on these values.  
```{r}
dist_host_model<-aov(colony_dispersion~site_code*species, data=dist_host_calc)
summary(dist_host_model)
```


# Symbiont responses  

Calculate distances for only host responses.  

Generate a list of symbiont and host responses.  
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

Calculate plasticity as total euclidean distance between multivariate physiology positions generated in a PCA between time points for all metrics.  

```{r}
data_sym<-master%>%
  select(colony_id_corr, timepoint, species, site_code, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW)

data_sym<-data_sym[complete.cases(data_sym), ]

data_sym[,-c(1:4)]<-log(data_sym[,-c(1:4)]+1)

scaled_sym<-prcomp(data_sym[c(5:13)], scale=TRUE, center=TRUE) 

sym_info<-data_sym[c(1:4)]

#generate a euclidean distance matrix between all samples (colony x time point)
sym_data<-scaled_sym%>%
  augment(sym_info)%>%
  group_by(timepoint, site_code, species, colony_id_corr)%>%
  mutate(code=paste(timepoint, colony_id_corr))

sym_data<-as.data.frame(sym_data)

row.names(sym_data)<-sym_data$code

sym_data<-sym_data[6:14]
head(sym_data)

dist_sym<-dist(sym_data, diag = FALSE, upper = FALSE, method="euclidean")

dist_sym <- melt(as.matrix(dist_sym), varnames = c("row", "col"))

head(dist_sym)
```

Now calculate the dispersion (total distance traveled in euclidian distance) for each colony. We are calculating dispersion only for colony that were found at at least 3 timepoints so that we have a good representation of dispersion. Distances are calculated as the distance between timepoint 1 and 2, timepoint 2 and 3, and time point 3 and 4 summed together. If a colony was missing at one time point, the distance from the two adjancent time points was used. For example, if a colony was missing at time point 3, we will use distances from timepoint 1 to 2, timepoint 1 to 3, and timepoint 3 to 4.   

```{r}
#calculate distances between subsequent time points (T1-T2, T2-T3, T3-T4) and sum them together
dist_sym_calc<-dist_sym%>%
  separate(row, into=c("Timepoint1", "Colony1"), sep=" ")%>%
  separate(col, into=c("Timepoint2", "Colony2"), sep=" ")%>%
  filter(value>0.0)%>% #remove distances between the same points
  mutate(condition=if_else(Colony1==Colony2, "TRUE", "FALSE"))%>%
  filter(condition=="TRUE")%>% #keep within colony comparisons
  arrange(Colony1, Timepoint1, Timepoint2)%>% #arrange by time points
  select(Timepoint1, Timepoint2, Colony1, value)%>%
  mutate(comparison=paste(Timepoint1, "-", Timepoint2))%>% #add a comparison column
  filter(!duplicated(paste(Colony1, value)))%>% #remove duplicated/reciprocal timepoint matches 
  select(!Timepoint1)%>%
  select(!Timepoint2)%>%
  pivot_wider(names_from = comparison, values_from = value, id_cols=Colony1)%>%
  rename("T1_T2"="timepoint1 - timepoint2", 
         "T2_T3"="timepoint2 - timepoint3", 
         "T3-T4"="timepoint3 - timepoint4", 
         "T1_T3"="timepoint1 - timepoint3", 
         "T1_T4"="timepoint1 - timepoint4",
         "T3_T4"="timepoint3 - timepoint4", 
         "T2_T4"="timepoint2 - timepoint4")%>%
  mutate(sum_1=rowSums(.[ , c("T1_T2", "T2_T3", "T3_T4")]))%>% #summarize rows with complete timeseries data 
  mutate(sum_2=case_when(is.na(T1_T2) ~ rowSums(.[ , c("T1_T3", "T3_T4")])))%>% #If T2 is missing, instead sum T1-T3 and T3-T4
  mutate(sum_3=case_when(is.na(T1_T4) ~ rowSums(.[ , c("T1_T2", "T2_T3")])))%>% #If T4 is missing, instead sum T1-T2 and T2-T3
 mutate(sum_4=case_when(is.na(T1_T3) ~ rowSums(.[ , c("T1_T2", "T2_T4")])))%>% #If T3 is missing, instead sum T1-T2 and T2-T4
  mutate(sum_5=case_when(is.na(T1_T2) & is.na(T1_T3) & is.na(T1_T4) ~ rowSums(.[ , c("T2_T3", "T3_T4")])))%>% #If T1 is missing, instead sum T2-T3 and T3-T4 
  mutate(colony_dispersion=coalesce(sum_1, sum_2, sum_3, sum_4, sum_5))%>%
  select(Colony1, colony_dispersion)
  
dist_sym_calc$site_code<-master$site_code[match(dist_sym_calc$Colony1, master$colony_id_corr)]
dist_sym_calc$species<-master$species[match(dist_sym_calc$Colony1, master$colony_id_corr)]

head(dist_sym_calc)
```

Plot as a dot plot with an overlay for means of each group.  
```{r}
#Calculate means to add to 
dist_sym_means<-dist_sym_calc%>%
  group_by(species, site_code)%>%
  summarise(mean=mean(colony_dispersion, na.rm=TRUE), sd=sd(colony_dispersion, na.rm=TRUE), N=sum(!is.na(colony_dispersion)), se=sd/sqrt(N))

dist_sym_dotplot<-dist_sym_calc%>%
  
    ggplot(aes(x = site_code, y = colony_dispersion)) +
  
    #add individual data points 
    geom_point(aes(fill=species, color=species, group=interaction(species, site_code)), pch = 21, size=4, position = position_jitterdodge(0.2), alpha=0.3) + 
  
    #add mean points and lines between groups
    geom_line(data=dist_sym_means, aes(y=mean, x=site_code, group=species, color=species), position=position_dodge(0.2), size=2) +
    geom_point(data=dist_sym_means, aes(y=mean, x=site_code, color=species, fill=species), pch = 21, size=5, position = position_dodge(0.2), alpha=0.8) + 
  
    #add stats
    #geom_text(x=2, y=20, size=4, label="p(Site)=0.306", color="darkgray")+
    #geom_text(x=2, y=19, size=4, label="p(Species)=0.434", color="darkgray")+
    #geom_text(x=2, y=18, size=4, label="p(Site x Species)=0.024", color="black")+
  
    #aesthetics 
    scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_color_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    scale_x_discrete(labels=c("Mahana\nLow", "Hilton\nMedium", "Manava\nHigh"))+
    xlab("Site") + 
    ylim(0,20)+
    ylab(expression(bold("Multivariate Dispersion")))+
    theme_classic() + 
    ggtitle("Symbiont")+
    theme(
      legend.position="right",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14, face="italic"),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=14, color="black"), 
      strip.text.x=element_text(face="italic", size=14),
      title=element_text(face="bold", size=16)
      ); dist_sym_dotplot

```

Run anova on these values.  
```{r}
dist_sym_model<-aov(colony_dispersion~site_code*species, data=dist_sym_calc)
summary(dist_sym_model)
```


# Generate figure 

Generate a panel figure.  
```{r}
dist_all_dotplot2<-dist_all_dotplot+theme(legend.position="none")+xlab("")
dist_host_dotplot2<-dist_host_dotplot+theme(legend.position="none")+xlab("")
dist_sym_dotplot2<-dist_sym_dotplot+xlab("")

panel<-plot_grid(dist_all_dotplot2, dist_host_dotplot2, dist_sym_dotplot2, nrow=1, ncol=3, rel_widths = c(1,1,1.3), labels=c("A", "B", "C"), label_size=20, label_y=0.96)

ggsave(panel, file="Figures/Plasticity/Plasticity_Figure.png", width=13, height=5)
```


