---
title: "Modeling analysis of E5 time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
date: "03/21/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("Hmisc")) install.packages("Hmisc")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(Hmisc)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")
```

# Run models 

## All species 

Model the influence of physiological metrics on calcification. 
```{r}
df<-master%>%
  select(Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Am, AQY, Rd, calc.umol.mgAFDW.hr, cells.mgAFDW, prot_mg.mgafdw, cre.umol.mgafdw, Ratio_AFDW.mg.cm2, Total_Chl_cell, Total_Chl)

model1<-lm(calc.umol.mgAFDW.hr~ ., data=df)

summary(model1)
```

From this first look, Am, Rd, and protein.AFDW are most significant (<0.01) and antioxidant (cre.AFDW) and ratio host:symbiont AFDW are also significant (<0.05) as predictors of calcification.  

Show effect plot
```{r}
plot(allEffects(model1)) 
```

Next, look for covariates and simplify the model by removing highly correlated variables.  

```{r}
cor_all<-rcorr(as.matrix(df))
#combine all values 
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

cor_all_df<-flattenCorrMatrix(cor_all$r, cor_all$P)
```

List correlations with absolute value r>0.85.  
```{r}
cor_all_df%>%
  filter(abs(cor)>0.85)
```
Respiration and Am are > 0.85.  

List correlations with high significance p<0.01.  
```{r}
cor_all_df%>%
  filter(p<0.01)
```
Many correlations are highly significant.  

View the correlations between calcification and other physiological parameters.  
```{r}
cor_all_df%>%
  filter(p<0.01)%>%
  filter(row=="calc.umol.mgAFDW.hr")
```
There are significant, but weak positive correlations between calcification & symbiont cell density, protein, symbiont:host biomass ratio, and total chlorophyll. 


## Acropora 

Model the influence of physiological metrics on calcification. 
```{r}
df_acr<-master%>%
  filter(species=="Acropora")%>%
  select(Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Am, AQY, Rd, calc.umol.mgAFDW.hr, cells.mgAFDW, prot_mg.mgafdw, cre.umol.mgafdw, Ratio_AFDW.mg.cm2, Total_Chl_cell, Total_Chl)

model1_acr<-lm(calc.umol.mgAFDW.hr~ ., data=df_acr)

summary(model1_acr)
```

From this first look, protein.AFDW, total chlorophyll.AFDW and total chlorophyll per cell are significant (<0.01).    

Show effect plot
```{r}
plot(allEffects(model1_acr)) 
```

Next, look for covariates and simplify the model by removing highly correlated variables.  

```{r}
cor_acr<-rcorr(as.matrix(df_acr))
#combine all values 

cor_acr_df<-flattenCorrMatrix(cor_acr$r, cor_acr$P)
```

List correlations with absolute value r>0.9.  
```{r}
cor_acr_df%>%
  filter(abs(cor)>0.85)
```
No correlations > 0.85.  

List correlations with high significance p<0.01.  
```{r}
cor_acr_df%>%
  filter(p<0.01)
```
Many correlations are highly significant. 

View the correlations between calcification and other physiological parameters.  
```{r}
cor_acr_df%>%
  filter(p<0.01)%>%
  filter(row=="calc.umol.mgAFDW.hr")
```
There are significant positive correlations between calcification & protein, antioxidant capacity, host:symbiont biomass ratio, total chlorophyll, and total chlorophyll per symbiont cell. 

## Pocillopora 

Model the influence of physiological metrics on calcification. 
```{r}
df_poc<-master%>%
  filter(species=="Pocillopora")%>%
  select(Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Am, AQY, Rd, calc.umol.mgAFDW.hr, cells.mgAFDW, prot_mg.mgafdw, cre.umol.mgafdw, Ratio_AFDW.mg.cm2, Total_Chl_cell, Total_Chl)

model1_poc<-lm(calc.umol.mgAFDW.hr~ ., data=df_poc)

summary(model1_poc)
```

From this first look, antioxidant.AFDW is the only effect that is significant (<0.01).    

Show effect plot
```{r}
plot(allEffects(model1_poc)) 
```

Next, look for covariates and simplify the model by removing highly correlated variables.  

```{r}
cor_poc<-rcorr(as.matrix(df_poc))
#combine all values 

cor_poc_df<-flattenCorrMatrix(cor_poc$r, cor_poc$P)
```

List correlations with absolute value r>0.9.  
```{r}
cor_poc_df%>%
  filter(abs(cor)>0.85)
```
No correlations > 0.85.  

List correlations with high significance p<0.01.  
```{r}
cor_poc_df%>%
  filter(p<0.01)
```
Many correlations are highly significant. 

View the correlations between calcification and other physiological parameters.  
```{r}
cor_poc_df%>%
  filter(p<0.01)%>%
  filter(row=="calc.umol.mgAFDW.hr")
```
There is one significant negative correlation between calcification and antioxidant capacity.   

## Porites 

Model the influence of physiological metrics on calcification. 
```{r}
df_por<-master%>%
  filter(species=="Porites")%>%
  select(Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Am, AQY, Rd, calc.umol.mgAFDW.hr, cells.mgAFDW, prot_mg.mgafdw, cre.umol.mgafdw, Ratio_AFDW.mg.cm2, Total_Chl_cell, Total_Chl)

model1_por<-lm(calc.umol.mgAFDW.hr~ ., data=df_por)

summary(model1_por)
```

From this first look, Host AFDW is the only effect that is higly significant (<0.01) and Am is also significant (<0.05).      

Show effect plot
```{r}
plot(allEffects(model1_poc)) 
```

Next, look for covariates and simplify the model by removing highly correlated variables.  

```{r}
cor_por<-rcorr(as.matrix(df_por))
#combine all values 

cor_por_df<-flattenCorrMatrix(cor_por$r, cor_por$P)
```

List correlations with absolute value r>0.9.  
```{r}
cor_por_df%>%
  filter(abs(cor)>0.85)
```
No correlations > 0.85.  

List correlations with high significance p<0.01.  
```{r}
cor_por_df%>%
  filter(p<0.01)
```
Many correlations are highly significant. 

Based on correlation analysis there are no strongly correlated variables and we can leave all responses in the model.  

View the correlations between calcification and other physiological parameters.  
```{r}
cor_por_df%>%
  filter(p<0.01)%>%
  filter(row=="calc.umol.mgAFDW.hr")
```
There is  nothing that correlates with calcification in Porites.  

