---
title: "Variance partitioning analysis"
author: "Ariana S Huffmyer, Serena Hackerott, E5 RoL Team"
date: "05/09/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("vegan")) install.packages("vegan")
if (!require("ggplot2")) install.packages("ggplot2")

# load packages
library(tidyverse)
library(vegan)
library(ggplot2)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

#Phys matrix: master_timeseries.csv
#Temp matrix: Clean_Environmental_Temp_Means.csv


# Holobiont responses   
```{r}
data_all<-master%>%
  select(colony_id_corr, timepoint, species, site_code, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)
```

#Log+1 transform all responses
```{r}
data_all_log <-data_all
data_all_log[,-c(1:4)]<-log(data_all_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_all_log$timepoint<-as.factor(data_all_log$timepoint)

#change species to a factor 
data_all_log$species<-as.factor(data_all_log$species)
```


#Acropora  
```{r}
acr_data_afdw_log<-data_all_log%>%
  filter(species=="Acropora")

acr_data_afdw_log<-acr_data_afdw_log[complete.cases(acr_data_afdw_log), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_acr_afdw<-varpart(acr_data_afdw_log[,5:16], acr_data_afdw_log$timepoint, acr_data_afdw_log$site)

##Check variance explained by individual fractions
varpart_acr_afdw$part
#Individual fractions                                    
#[a] = X1|X2           3                 0.47384     TRUE
#[b] = X2|X1           2                 0.03215     TRUE
#[c]                   0                -0.00390    FALSE
#[d] = Residuals                         0.49791    FALSE

##Check model significance
anova(rda(acr_data_afdw_log[,5:16]~ acr_data_afdw_log$timepoint + acr_data_afdw_log$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(acr_data_afdw_log[,5:16]~ acr_data_afdw_log$timepoint + acr_data_afdw_log$site))$adj.r.squared*100
#The model explains 50.21% of the variation in coral physiology in Acropora

##Check effect of timepoint 
anova(rda(acr_data_afdw_log[,5:16], acr_data_afdw_log$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(acr_data_afdw_log[,5:16], acr_data_afdw_log$timepoint,  acr_data_afdw_log$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(acr_data_afdw_log[,5:16], acr_data_afdw_log$site))
#The effect of site with overlap with timepoint is significant p=0.035 
anova(rda(acr_data_afdw_log[,5:16], acr_data_afdw_log$site,  acr_data_afdw_log$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.003 
```

##Generate figure-Venn
```{r}
par(bty="n")

varpart_acr_afdw_parts<-data.frame(Parts=c("Season", "Seas&Site", "Site","Residuals"), 
                            Variance=c(varpart_acr_afdw$part$indfract$Adj.R.square))
varpart_acr_afdw_parts$PercVar<-paste(as.character(round(varpart_acr_afdw_parts$Variance*100,2)), "%", sep="")
varpart_acr_afdw_parts$PercVar[which(varpart_acr_afdw_parts$Variance<0)]<-NA
varpart_acr_afdw_parts$Species<-rep("Acropora", nrow(varpart_acr_afdw_parts))

showvarparts(parts=2, labels=varpart_acr_afdw_parts$PercVar, Xnames=NULL, 
             cex=c(1.4, 1.4, 1.4, 1, 1, 1.2), 
             bg=c("#4681F7FF", "#18DEC1FF"), alpha=125)+
text(-0.1, .4, "Season", cex=2)+
text(1, 0.4, "Site", cex=2)+
text(0, -.3, substitute(paste(italic('p = 0.001'))), cex=1.1)+
text(1, -.3, substitute(paste(italic('p = 0.003'))), cex=1.1)+
text(0.5, 1, "Acropora", cex=2)
```



#Porites
```{r}
por_data_afdw_log<-data_all_log%>%
  filter(species=="Porites")

por_data_afdw_log<-por_data_afdw_log[complete.cases(por_data_afdw_log), ]
```

##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_por_afdw<-varpart(por_data_afdw_log[,5:16], por_data_afdw_log$timepoint, por_data_afdw_log$site)

##Check variance explained by individual fractions
varpart_por_afdw$part
#Individual fractions                                    
#[a] = X1|X2           3                 0.21416     TRUE
#[b] = X2|X1           2                 0.13836     TRUE
#[c]                   0                 0.00191    FALSE
#[d] = Residuals                         0.64557    FALSE

##Check model significance
anova(rda(por_data_afdw_log[,5:16]~ por_data_afdw_log$timepoint + por_data_afdw_log$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(por_data_afdw_log[,5:16]~ por_data_afdw_log$timepoint + por_data_afdw_log$site))$adj.r.squared*100
#The model explains 35.44% of the variation in coral physiology in Acropora

##Check effect of timepoint 
anova(rda(por_data_afdw_log[,5:16], por_data_afdw_log$timepoint))
#The effect of timepoint with overlap with site_code is significant p=0.001 
anova(rda(por_data_afdw_log[,5:16], por_data_afdw_log$timepoint,  por_data_afdw_log$site))
#The effect of timepoint alone (controlling for site_code) is significant p=0.001 

##Check effect of site 
anova(rda(por_data_afdw_log[,5:16], por_data_afdw_log$site))
#The effect of site with overlap with timepoint is significant p=0.001
anova(rda(por_data_afdw_log[,5:16], por_data_afdw_log$site,  por_data_afdw_log$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.001 

```

##Generate figure- Venn
```{r}
par(bty="n")

varpart_por_afdw_parts<-data.frame(Parts=c("Season", "Seas&Site", "Site","Residuals"), 
                            Variance=c(varpart_por_afdw$part$indfract$Adj.R.square))
varpart_por_afdw_parts$PercVar<-paste(as.character(round(varpart_por_afdw_parts$Variance*100,2)), "%", sep="")
varpart_por_afdw_parts$PercVar[which(varpart_por_afdw_parts$Variance<0)]<-NA
varpart_por_afdw_parts$Species<-rep("Porites", nrow(varpart_por_afdw_parts))

showvarparts(parts=2, labels=varpart_por_afdw_parts$PercVar, Xnames=NULL, 
             cex=c(1.4, 1.4, 1.4, 1, 1, 1.2), 
             bg=c("#4681F7FF", "#18DEC1FF"), alpha=125)+
text(-0.1, .4, "Season", cex=2)+
text(1, 0.4, "Site", cex=2)+
text(0, -.3, substitute(paste(italic('p = 0.001'))), cex=1.1)+
text(1, -.3, substitute(paste(italic('p = 0.001'))), cex=1.1)+
text(0.5, 1, "Porites", cex=2)
```


#Pocillopora   
```{r}
poc_data_afdw_log<-data_all_log%>%
  filter(species=="Pocillopora")

poc_data_afdw_log<-poc_data_afdw_log[complete.cases(poc_data_afdw_log), ]
```

##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_poc_afdw<-varpart(poc_data_afdw_log[,5:16], poc_data_afdw_log$timepoint, poc_data_afdw_log$site)

##Check variance explained by individual fractions
varpart_poc_afdw$part
#[a] = X1|X2           3                 0.34959     TRUE
#[b] = X2|X1           2                 0.02326     TRUE
#[c]                   0                -0.00162    FALSE
#[d] = Residuals                         0.62877    FALSE

##Check model significance
anova(rda(poc_data_afdw_log[,5:16]~ poc_data_afdw_log$timepoint + poc_data_afdw_log$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(poc_data_afdw_log[,5:16]~ poc_data_afdw_log$timepoint + poc_data_afdw_log$site))$adj.r.squared*100
#The model explains 37.12% of the variation in coral physiology in Acropora

##Check effect of timepoint 
anova(rda(poc_data_afdw_log[,5:16], poc_data_afdw_log$timepoint))
#The effect of timepoint with overlap with site_code is significant p=0.001 
anova(rda(poc_data_afdw_log[,5:16], poc_data_afdw_log$timepoint,  poc_data_afdw_log$site))
#The effect of timepoint alone (controlling for site_code) is significant p=0.001 

##Check effect of site 
anova(rda(poc_data_afdw_log[,5:16], poc_data_afdw_log$site))
#The effect of site with overlap with timepoint is not significant p=0.051 .
anova(rda(poc_data_afdw_log[,5:16], poc_data_afdw_log$site,  poc_data_afdw_log$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.011 

```

##Generate figure- Venn
```{r}
par(bty="n")

varpart_poc_afdw_parts<-data.frame(Parts=c("Season", "Seas&Site", "Site","Residuals"), 
                            Variance=c(varpart_poc_afdw$part$indfract$Adj.R.square))
varpart_poc_afdw_parts$PercVar<-paste(as.character(round(varpart_poc_afdw_parts$Variance*100,2)), "%", sep="")
varpart_poc_afdw_parts$PercVar[which(varpart_poc_afdw_parts$Variance<0)]<-NA
varpart_poc_afdw_parts$Species<-rep("Pocillopora", nrow(varpart_poc_afdw_parts))

showvarparts(parts=2, labels=varpart_poc_afdw_parts$PercVar, Xnames=NULL, 
             cex=c(1.4, 1.4, 1.4, 1, 1, 1.2), 
             bg=c("#4681F7FF", "#18DEC1FF"), alpha=125)+
text(-0.1, .4, "Season", cex=2)+
text(1, 0.4, "Site", cex=2)+
text(0, -.3, substitute(paste(italic('p = 0.001'))), cex=1.1)+
text(1, -.3, substitute(paste(italic('p = 0.008'))), cex=1.1)+
text(0.5, 1, "Pocillopora", cex=2)
```


#Generate figure- Barplot all Species Holobiont

Note: Ariana added manual name for "Site", since it was coming up as NA
```{r}
##Combine VarPart results for all Species
varpart_afdw_parts_all<-rbind(varpart_acr_afdw_parts, varpart_por_afdw_parts, varpart_poc_afdw_parts)
varpart_afdw_parts_all<-na.omit(varpart_afdw_parts_all)
varpart_afdw_parts_all<-varpart_afdw_parts_all%>%filter(!Parts=="Site")
varpart_afdw_parts_all$Response<-rep("Holobiont", nrow(varpart_afdw_parts_all))
varpart_afdw_parts_all$Percent<-round(varpart_afdw_parts_all$Variance*100,2)
varpart_afdw_parts_all$Parts<-factor(varpart_afdw_parts_all$Parts, levels=c("Seas&Site", "Season", "Residuals"), ordered=TRUE)

names(varpart_afdw_parts_all)

varpartBarplot_all.holo<-ggplot(varpart_afdw_parts_all, aes(x=Species, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"));varpartBarplot_all.holo

#ggsave(varpartBarplot_all.holo, file="Figures/Multivariate/VarPart_All_Holobiont.png", width=6, height=6)

```


# Generate a list of symbiont and host responses.  Holobiont = all responses
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ratio_AFDW.mg.cm2") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

# Host responses
```{r}
data_host<-master%>%
  select(colony_id_corr, timepoint, species, site_code, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

data_host<-data_host[complete.cases(data_host), ]
```


#Log+1 transform all responses
```{r}
data_host_log <-data_host
data_host_log[,-c(1:4)]<-log(data_host_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_host_log$timepoint<-as.factor(data_host_log$timepoint)

#change species to a factor 
data_host_log$species<-as.factor(data_host_log$species)
```


#Acropora- Host  
```{r}
acr_data_afdw_log_host<-data_host_log%>%
  filter(species=="Acropora")

acr_data_afdw_log_host<-acr_data_afdw_log_host[complete.cases(acr_data_afdw_log_host), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_acr_afdw_host<-varpart(acr_data_afdw_log_host[,5:9], acr_data_afdw_log_host$timepoint, acr_data_afdw_log_host$site)

##Check variance explained by individual fractions
varpart_acr_afdw_host$part
# Individual fractions                                    
#[a] = X1|X2           3                 0.26728     TRUE
#[b] = X2|X1           2                 0.08019     TRUE
#[c]                   0                -0.01420    FALSE
#[d] = Residuals                         0.66673    FALSE

##Check model significance
anova(rda(acr_data_afdw_log_host[,5:9]~ acr_data_afdw_log_host$timepoint + acr_data_afdw_log_host$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(acr_data_afdw_log_host[,5:9]~ acr_data_afdw_log_host$timepoint + acr_data_afdw_log_host$site))$adj.r.squared*100
#The model explains 33.33% of the variation in coral physiology in Acropora Host

##Check effect of timepoint 
anova(rda(acr_data_afdw_log_host[,5:9], acr_data_afdw_log_host$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(acr_data_afdw_log_host[,5:9], acr_data_afdw_log_host$timepoint,  acr_data_afdw_log_host$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(acr_data_afdw_log_host[,5:9], acr_data_afdw_log_host$site))
#The effect of site with overlap with timepoint is significant p=0.003 
anova(rda(acr_data_afdw_log_host[,5:9], acr_data_afdw_log_host$site,  acr_data_afdw_log_host$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.001 
```


#Porites- Host  
```{r}
por_data_afdw_log_host<-data_host_log%>%
  filter(species=="Porites")

por_data_afdw_log_host<-por_data_afdw_log_host[complete.cases(por_data_afdw_log_host), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_por_afdw_host<-varpart(por_data_afdw_log_host[,5:9], por_data_afdw_log_host$timepoint, por_data_afdw_log_host$site)

##Check variance explained by individual fractions
varpart_por_afdw_host$part
# Individual fractions                                    
#[a] = X1|X2           3                 0.27738     TRUE
#[b] = X2|X1           2                 0.04781     TRUE
#[c]                   0                -0.00921    FALSE
#[d] = Residuals                         0.68402    FALSE

##Check model significance
anova(rda(por_data_afdw_log_host[,5:9]~ por_data_afdw_log_host$timepoint + por_data_afdw_log_host$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(por_data_afdw_log_host[,5:9]~ por_data_afdw_log_host$timepoint + por_data_afdw_log_host$site))$adj.r.squared*100
#The model explains 31.60% of the variation in coral physiology in Porites Host

##Check effect of timepoint 
anova(rda(por_data_afdw_log_host[,5:9], por_data_afdw_log_host$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(por_data_afdw_log_host[,5:9], por_data_afdw_log_host$timepoint,  por_data_afdw_log_host$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(por_data_afdw_log_host[,5:9], por_data_afdw_log_host$site))
#The effect of site with overlap with timepoint is significant p=0.002 
anova(rda(por_data_afdw_log_host[,5:9], por_data_afdw_log_host$site,  por_data_afdw_log_host$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.001 
```


#Pocillopora- Host  
```{r}
poc_data_afdw_log_host<-data_host_log%>%
  filter(species=="Pocillopora")

poc_data_afdw_log_host<-poc_data_afdw_log_host[complete.cases(poc_data_afdw_log_host), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_poc_afdw_host<-varpart(poc_data_afdw_log_host[,5:9], poc_data_afdw_log_host$timepoint, poc_data_afdw_log_host$site)

##Check variance explained by individual fractions
varpart_poc_afdw_host$part
# Individual fractions                                    
#[a] = X1|X2           3                 0.18242     TRUE
#[b] = X2|X1           2                 0.04315     TRUE
#[c]                   0                 0.00323    FALSE
#[d] = Residuals                         0.77121    FALSE

##Check model significance
anova(rda(poc_data_afdw_log_host[,5:9]~ poc_data_afdw_log_host$timepoint + poc_data_afdw_log_host$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(poc_data_afdw_log_host[,5:9]~ poc_data_afdw_log_host$timepoint + poc_data_afdw_log_host$site))$adj.r.squared*100
#The model explains 22.87% of the variation in coral physiology in Pocillopora Host

##Check effect of timepoint 
anova(rda(poc_data_afdw_log_host[,5:9], poc_data_afdw_log_host$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(poc_data_afdw_log_host[,5:9], poc_data_afdw_log_host$timepoint,  poc_data_afdw_log_host$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(poc_data_afdw_log_host[,5:9], poc_data_afdw_log_host$site))
#The effect of site with overlap with timepoint is significant p=0.005 
anova(rda(poc_data_afdw_log_host[,5:9], poc_data_afdw_log_host$site,  poc_data_afdw_log_host$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.001 
```

# Symbiont responses
```{r}
data_sym<-master%>%
  select(colony_id_corr, timepoint, species, site_code, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW)

data_sym<-data_sym[complete.cases(data_sym), ]
```


#Log+1 transform all responses
```{r}
data_sym_log <-data_sym
data_sym_log[,-c(1:4)]<-log(data_sym_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_sym_log$timepoint<-as.factor(data_sym_log$timepoint)

#change species to a factor 
data_sym_log$species<-as.factor(data_sym_log$species)
```

#Acropora- Sym  
```{r}
acr_data_afdw_log_sym<-data_sym_log%>%
  filter(species=="Acropora")

acr_data_afdw_log_sym<-acr_data_afdw_log_sym[complete.cases(acr_data_afdw_log_sym), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_acr_afdw_sym<-varpart(acr_data_afdw_log_sym[,5:11], acr_data_afdw_log_sym$timepoint, acr_data_afdw_log_sym$site)

##Check variance explained by individual fractions
varpart_acr_afdw_sym$part
# Individual fractions                                    
#[a] = X1|X2           3                 0.48852     TRUE
#[b] = X2|X1           2                 0.02296     TRUE
#[c]                   0                -0.00565    FALSE
#[d] = Residuals                         0.49417    FALSE

##Check model significance
anova(rda(acr_data_afdw_log_sym[,5:11]~ acr_data_afdw_log_sym$timepoint + acr_data_afdw_log_sym$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(acr_data_afdw_log_sym[,5:11]~ acr_data_afdw_log_sym$timepoint + acr_data_afdw_log_sym$site))$adj.r.squared*100
#The model explains 50.58% of the variation in coral physiology in Acropora Symbiont

##Check effect of timepoint 
anova(rda(acr_data_afdw_log_sym[,5:11], acr_data_afdw_log_sym$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(acr_data_afdw_log_sym[,5:11], acr_data_afdw_log_sym$timepoint,  acr_data_afdw_log_sym$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(acr_data_afdw_log_sym[,5:11], acr_data_afdw_log_sym$site))
#The effect of site with overlap with timepoint is not significant p=0.116
anova(rda(acr_data_afdw_log_sym[,5:11], acr_data_afdw_log_sym$site,  acr_data_afdw_log_sym$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.011 
```


#Porites- Sym  
```{r}
por_data_afdw_log_sym<-data_sym_log%>%
  filter(species=="Porites")

por_data_afdw_log_sym<-por_data_afdw_log_sym[complete.cases(por_data_afdw_log_sym), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_por_afdw_sym<-varpart(por_data_afdw_log_sym[,5:11], por_data_afdw_log_sym$timepoint, por_data_afdw_log_sym$site)

##Check variance explained by individual fractions
varpart_por_afdw_sym$part
# Individual fractions                                    
#[a] = X1|X2           3                 0.19581     TRUE
#[b] = X2|X1           2                 0.18111     TRUE
#[c]                   0                 0.00505    FALSE
#[d] = Residuals                         0.61804    FALSE

##Check model significance
anova(rda(por_data_afdw_log_sym[,5:11]~ por_data_afdw_log_sym$timepoint + por_data_afdw_log_sym$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(por_data_afdw_log_sym[,5:11]~ por_data_afdw_log_sym$timepoint + por_data_afdw_log_sym$site))$adj.r.squared*100
#The model explains 38.19% of the variation in coral physiology in Porites Symbiont

##Check effect of timepoint 
anova(rda(por_data_afdw_log_sym[,5:11], por_data_afdw_log_sym$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(por_data_afdw_log_sym[,5:11], por_data_afdw_log_sym$timepoint,  por_data_afdw_log_sym$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(por_data_afdw_log_sym[,5:11], por_data_afdw_log_sym$site))
#The effect of site with overlap with timepoint is significant p=0.001 
anova(rda(por_data_afdw_log_sym[,5:11], por_data_afdw_log_sym$site,  por_data_afdw_log_sym$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.001 
```


#Pocillopora- Sym  
```{r}
poc_data_afdw_log_sym<-data_sym_log%>%
  filter(species=="Pocillopora")

poc_data_afdw_log_sym<-poc_data_afdw_log_sym[complete.cases(poc_data_afdw_log_sym), ]
```


##Variance Partitioning
```{r}
##Run varpart function with responses and explanatory variables of interest
varpart_poc_afdw_sym<-varpart(poc_data_afdw_log_sym[,5:11], poc_data_afdw_log_sym$timepoint, poc_data_afdw_log_sym$site)

##Check variance explained by individual fractions
varpart_poc_afdw_sym$part
#Individual fractions                                    
#[a] = X1|X2           3                 0.35986     TRUE
#[b] = X2|X1           2                 0.01613     TRUE
#[c]                   0                -0.00278    FALSE
#[d] = Residuals                         0.62679    FALSE

##Check model significance
anova(rda(poc_data_afdw_log_sym[,5:11]~ poc_data_afdw_log_sym$timepoint + poc_data_afdw_log_sym$site))
#The model is significant p=0.001

##Check variance explained by model
RsquareAdj(rda(poc_data_afdw_log_sym[,5:11]~ poc_data_afdw_log_sym$timepoint + poc_data_afdw_log_sym$site))$adj.r.squared*100
#The model explains 37.32% of the variation in coral physiology in Pocillopora Symbiont

##Check effect of timepoint 
anova(rda(poc_data_afdw_log_sym[,5:11], poc_data_afdw_log_sym$timepoint))
#The effect of timepoint with overlap with site is significant p=0.001 
anova(rda(poc_data_afdw_log_sym[,5:11], poc_data_afdw_log_sym$timepoint,  poc_data_afdw_log_sym$site))
#The effect of timepoint alone (controlling for site) is significant p=0.001 

##Check effect of site 
anova(rda(poc_data_afdw_log_sym[,5:11], poc_data_afdw_log_sym$site))
#The effect of site with overlap with timepoint is not significant p=0.111 
anova(rda(poc_data_afdw_log_sym[,5:11], poc_data_afdw_log_sym$site,  poc_data_afdw_log_sym$timepoint))
#The effect of site alone (controlling for timepoint) is significant p=0.031 
```


#Organize Variance Partitioning Results

```{r}
##Acropora
#Holobiont
varpart_acr_afdw_parts.hol<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Holobiont"), Species=c("Acropora"), Variance=c(varpart_acr_afdw$part$indfract$Adj.R.square))

#Host
varpart_acr_afdw_parts.host<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Host"), Species=c("Acropora"), Variance=c(varpart_acr_afdw_host$part$indfract$Adj.R.square))

#Symbiont
varpart_acr_afdw_parts.sym<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Symbiont"), Species=c("Acropora"), Variance=c(varpart_acr_afdw_sym$part$indfract$Adj.R.square))


##Porites
#Holobiont
varpart_por_afdw_parts.hol<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Holobiont"), Species=c("Porites"), Variance=c(varpart_por_afdw$part$indfract$Adj.R.square))

#Host
varpart_por_afdw_parts.host<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Host"), Species=c("Porites"), Variance=c(varpart_por_afdw_host$part$indfract$Adj.R.square))

#Symbiont
varpart_por_afdw_parts.sym<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Symbiont"), Species=c("Porites"), Variance=c(varpart_por_afdw_sym$part$indfract$Adj.R.square))


##Pocillopora
#Holobiont
varpart_poc_afdw_parts.hol<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Holobiont"), Species=c("Pocillopora"), Variance=c(varpart_poc_afdw$part$indfract$Adj.R.square))

#Host
varpart_poc_afdw_parts.host<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Host"), Species=c("Pocillopora"), Variance=c(varpart_poc_afdw_host$part$indfract$Adj.R.square))

#Symbiont
varpart_poc_afdw_parts.sym<-data.frame(Parts=c("Season", "Season&Site", "Site", "Residuals"), Fraction=c("Symbiont"), Species=c("Pocillopora"), Variance=c(varpart_poc_afdw_sym$part$indfract$Adj.R.square))


varpart_afdw_parts<-rbind(varpart_acr_afdw_parts.hol, varpart_acr_afdw_parts.host, varpart_acr_afdw_parts.sym, varpart_por_afdw_parts.hol, varpart_por_afdw_parts.host, 
                          varpart_por_afdw_parts.sym, varpart_poc_afdw_parts.hol, 
                          varpart_poc_afdw_parts.host, varpart_poc_afdw_parts.sym)

##Percent Variance Explained
varpart_afdw_parts$Percent<-round(varpart_afdw_parts$Variance*100,2)

##Remove Site EnvFactor
varpart_afdw_parts<-varpart_afdw_parts[-c(which(varpart_afdw_parts$Parts=="Site")),]

##Ordered Factors
varpart_afdw_parts$Parts<-factor(varpart_afdw_parts$Parts, levels=c("Season&Site", "Season", "Residuals"), ordered=TRUE)

```


#Generate figures Option 1 Barplot Split by Species

```{r}

varpartBarplot_Species<-ggplot(varpart_afdw_parts, aes(x=Fraction, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5.5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"), 
        strip.text = element_text(size=16))+
  facet_wrap(vars(Species), ncol=3);varpartBarplot_Species

ggsave(varpartBarplot_Species, file="Figures/Multivariate/VarPart_Species.png", width=12, height=8)

```


#Generate figures Option 2 Barplot Split by Fraction

```{r}

varpartBarplot_Fraction<-ggplot(varpart_afdw_parts, aes(x=Species, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5.5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"), 
        strip.text = element_text(size=16))+
  facet_wrap(vars(Fraction), ncol=3);varpartBarplot_Fraction

ggsave(varpartBarplot_Fraction, file="Figures/Multivariate/VarPart_Fraction.png", width=12, height=8)

```

Plot individual fractions for panels in final multivariate plot. 
```{r}
varpartBarplot_holobiont<-varpart_afdw_parts%>%
  filter(Fraction=="Holobiont")%>%
  
  ggplot(aes(x=Species, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5.5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"), 
        strip.text = element_text(size=16));varpartBarplot_holobiont

ggsave(varpartBarplot_holobiont, file="Figures/Multivariate/VarPart_Holobiont.png", width=6, height=6)

```

```{r}
varpartBarplot_host<-varpart_afdw_parts%>%
  filter(Fraction=="Host")%>%
  
  ggplot(aes(x=Species, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5.5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"), 
        strip.text = element_text(size=16));varpartBarplot_host

ggsave(varpartBarplot_host, file="Figures/Multivariate/VarPart_Host.png", width=6, height=6)

```

```{r}
varpartBarplot_sym<-varpart_afdw_parts%>%
  filter(Fraction=="Symbiont")%>%
  
  ggplot(aes(x=Species, y=Percent, fill=Parts)) + 
  geom_bar(position="stack", stat="identity", colour="black")+
  scale_fill_manual(name="Predictors", values = c("white", "lightgray", "darkgray"), labels = c("Site", "Season", "Residuals"))+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), 
        axis.text.x=element_text(size=14, colour="black"), axis.text.y=element_text(size=14, colour="black"), 
        legend.text=element_text(size=12), legend.title=element_text(size=15))+
  ylab("% Variance Explained")+
  geom_text(aes(label=paste0(sprintf("%1.1f",Percent),"%"), fontface = "bold"),
            position=position_stack(vjust=0.5), size=5.5)+
  theme(plot.title=element_text(hjust=0.5), 
        axis.text.x=element_text(face="italic", color="black"), 
        axis.text.y=element_text(color="black"), 
        axis.title = element_text(face="bold"), 
        legend.title=element_text(face="bold"), 
        strip.text = element_text(size=16));varpartBarplot_sym

ggsave(varpartBarplot_sym, file="Figures/Multivariate/VarPart_Sym.png", width=6, height=6)

```
