---
title: "Plasticity analysis centroid distances"
author: "Ariana S Huffmyer, Serena Hackerott, E5 RoL Team"
date: "06/12/2023"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
--- 

Definitions: 

- Genus (Pocillopora, Acropora, and Porites) are coded as "species"
- Haplotype are coded as "haplotype". In the manuscript document, haplotype is referred to as "holobiont" because the host genetic haplotype and the associated symbiont communities are exclusive. 
- Biological levels are categorized in the code as all host and symbiont metrics together "holobiont" (i.e., "combined responses" in the manuscript to avoid confusion with above mentioned definition of holobiont), host metrics as "host", and symbiont metrics as "symbiont". 
- Time point is coded for seasonal sampling (January, March, September, and November) as "timepoint"
- Site is coded as "site" 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("vegan")) install.packages("vegan")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("ggpubr")) install.packages("ggpubr")

# load packages
library(tidyverse)
library(vegan)
library(ggplot2)
library(ggpubr)

```


#Site colors

```{r}
site.cols<-c("#374d7c", "#00cccc", "#ff6633")
#"Mahana", "Hilton", "Manava")
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("Output/master_timeseries.csv")

#reorder site levels 
master$site_code<-as.factor(master$site_code)
master$site_code<-fct_relevel(master$site_code, "Mahana Low", "Hilton Medium", "Manava High")

master$site<-as.factor(master$site)
master$site<-fct_relevel(master$site, "Mahana", "Hilton", "Manava")
```

# Holobiont responses   
```{r}
data_all<-master%>%
  select(colony_id_corr, timepoint, species, site, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Ratio_AFDW.mg.cm2, prot_mg.mgafdw, cells.mgAFDW, cre.umol.mgafdw, Am, AQY, Rd, Ik, Ic, calc.umol.cm2.hr, Total_Chl, Total_Chl_cell)%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  filter(Ic<1000)%>%
  rename(Host_Biomass=Host_AFDW.mg.cm2, Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Symbiont_Density=cells.mgAFDW, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)
```

##Log+1 transform all responses
```{r}
data_all_log <-data_all
data_all_log[,-c(1:4)]<-log(data_all_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_all_log$timepoint<-as.factor(data_all_log$timepoint)

#change species to a factor 
data_all_log$species<-as.factor(data_all_log$species)
```

##Update site naming
```{r}
data_all_log.comp<-data_all_log[complete.cases(data_all_log), ]

#data_all_log.comp<-separate(data_all_log.comp, "site_code", into=c("site", "nutrient"), remove=FALSE)
#data_all_log.comp$siteID<-paste(data_all_log.comp$site, data_all_log.comp$nutrient, sep="")
```

##Differences in dispersion between species
```{r}
species.disp<-betadisper(vegdist(data_all_log.comp[,5:18], "euclidean"), data_all_log.comp$species)
anova(species.disp)
#Analysis of Variance Table

#Response: Distances
#           Df Sum Sq Mean Sq F value  Pr(>F)  
#Groups      2  0.796 0.39788  3.1331 0.04474 *
#Residuals 371 47.114 0.12699     
#Slightly significant difference in dispersion between Species (may not be able to attribute significant effect from PERMANOVA to difference in centroids alone) 

#Difference in "plasticity" between species 

##Pairwise comparisons of dispersion between Species
TukeyHSD(species.disp)
#                            diff          lwr        upr     p adj
#Pocillopora-Acropora -0.04187698 -0.152245940 0.06849198 0.6451594
#Porites-Acropora      0.06547386 -0.043856295 0.17480401 0.3371537
#Porites-Pocillopora   0.10735084  0.005634634 0.20906704 0.0357894

#porites and pocillopora are significantly different in dispersion
```


##Centroid distances
```{r}
##Obtain centroid coordinates of each Sample Set (Same Species, Site, Season) with betadisper type="centroid"
data_all_log.comp$Set<-paste(data_all_log.comp$species, data_all_log.comp$site, data_all_log.comp$timepoint, sep=".")

all.disp.grp<-betadisper(vegdist(data_all_log.comp[,5:18], "euclidean"), data_all_log.comp$Set, type="centroid")

##Calculate the variance of Physiology explained by the first two PCoA axes
all.disp.grp$eig[1]/sum(all.disp.grp$eig[which(all.disp.grp$eig>0)])*100
all.disp.grp$eig[2]/sum(all.disp.grp$eig[which(all.disp.grp$eig>0)])*100
sum(all.disp.grp$eig[1:2])/sum(all.disp.grp$eig[which(all.disp.grp$eig>0)])*100
#70.73% explained by first two axes 

##Calculate the Euclidean distance between centroids using coordinates of the first 2 PCoA axes
centdist.all<-as.matrix(dist(all.disp.grp$centroids[,c(1:2)], method = "euclidean"))

##Create a dataframe of distances between centroids of Sample Sets
cents.all<-data.frame(SetA=colnames(centdist.all)[col(centdist.all)], SetB=rownames(centdist.all)[row(centdist.all)], Cent.Dist=c(centdist.all))

##Remove rows comparing self-pairs of Sample Sets 
cents.all<-cents.all[-c(which(cents.all$SetA==cents.all$SetB)),]

##Remove repeats (Ex: Set 1 vs Set 2 and Set 2 vs Set 1)
cents.all<-cents.all[!duplicated(cents.all$Cent.Dist),]

##Split Sets column to add variables of interest
cents.all<-separate(cents.all, "SetA", into=c("species.A", "site.A", "timepoint.A"), remove=FALSE)
cents.all<-separate(cents.all, "SetB", into=c("species.B", "site.B", "timepoint.B"), remove=FALSE)

##Filter to only keep rows comparing the same Species
cents.all<-cents.all[c(which(cents.all$species.A==cents.all$species.B)),]

##Species as factor
cents.all$species<-factor(cents.all$species.A, levels=c("Acropora", "Pocillopora", "Porites"))

##Dataframe with distances between Seasons by only keeping rows comparing the same Site
cents.all.seas<-cents.all[c(which(cents.all$site.A==cents.all$site.B)),]

##Dataframe with distances between Sites by only keeping rows comparing the same Season
cents.all.site<-cents.all[c(which(cents.all$timepoint.A==cents.all$timepoint.B)),]

##Comparisons of interest
spp_comparisons <- list( c("Acropora", "Pocillopora"),  c( "Pocillopora", "Porites"), c("Acropora", "Porites") )

```

##Run ANOVA on plasticity values 

Run anova on seasonal plasticity between species.
```{r}
summary(aov(Cent.Dist~species*site.A, data=cents.all.seas))
#               Df Sum Sq Mean Sq F value Pr(>F)  
#species         2  0.714  0.3572   2.587 0.0864 .
#site.A          2  0.307  0.1537   1.113 0.3374  
#species:site.A  4  0.389  0.0972   0.704 0.5935  
#Residuals      45  6.212  0.1380    
#No significant difference in seasonal plasticity between species
#The effect of season does not differ between sites

compare_means(Cent.Dist~species, data=cents.all.seas, method = "anova")
#   .y.           p p.adj p.format p.signif method
#   <chr>     <dbl> <dbl> <chr>    <chr>    <chr> 
# Cent.Dist 0.0813 0.081 0.081    ns       Anova 
```

Run anova on spatial plasticity between species.
```{r}
summary(aov(Cent.Dist~species*timepoint.A, data=cents.all.site))
#                    Df Sum Sq Mean Sq F value   Pr(>F)    
#species              2 0.6495  0.3248   9.475 0.000927 ***
#timepoint.A          3 0.4352  0.1451   4.232 0.015511 *  
#species:timepoint.A  6 0.1685  0.0281   0.819 0.565779    
R#esiduals           24 0.8226  0.0343                
#Significant difference in spatial plasticity between species
#The effect of site does differ between seasons 

compare_means(Cent.Dist~species, data=cents.all.site, method = "anova")
#   .y.              p   p.adj p.format p.signif method
#   <chr>        <dbl>   <dbl> <chr>    <chr>    <chr> 
# 1 Cent.Dist 0.00205 0.002 0.002    **       Anova 

```

##Generate figures- Seasonal and Spatial Plasticity
```{r}

season.plasticity.plot<-ggplot(cents.all.seas, aes(x=species.A, y=Cent.Dist, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
  stat_compare_means(method = "anova", label.x=1.75, size=6)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,1.55)+
  ylab("Holobiont Seasonal Plasticity")+
  xlab("Species");season.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_season.png", season.plasticity.plot, width=4, height=4, dpi=300)


spatial.plasticity.plot<-ggplot(cents.all.site, aes(x=species.A, y=Cent.Dist, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
    stat_compare_means(method = "anova", label.x=1.75, size=6)+
  stat_compare_means(comparisons = spp_comparisons, size=5)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,1.55)+
  ylab("Holobiont Spatial Plasticity")+
  xlab("Species");spatial.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_spatial.png", spatial.plasticity.plot, width=4, height=4, dpi=300)
```


# Generate a list of symbiont and host responses.  Holobiont = all responses
```{r}
symbiont_responses<-c("cells.mgAFDW", "Sym_AFDW.mg.cm2", "Total_Chl", "Total_Chl_cell", "Am", "AQY", "Ik", "Ic", "Ratio_AFDW.mg.cm2") 
host_responses<-c("cre.umol.mgafdw", "Host_AFDW.mg.cm2", "prot_mg.mgafdw", "Rd", "calc.umol.cm2.hr")
```

# Host responses
```{r}
data_host<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(host_responses))%>%
  filter(Host_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2>0)%>%
  #filter(Ratio_AFDW.mg.cm2>0)%>%
  #filter(Sym_AFDW.mg.cm2<20)%>%
  filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Host_Biomass=Host_AFDW.mg.cm2, Host_Protein=prot_mg.mgafdw, Antiox_Capacity=cre.umol.mgafdw, Calc=calc.umol.cm2.hr)

data_host<-data_host[complete.cases(data_host), ]
```


##Log+1 transform all responses
```{r}
data_host_log <-data_host
data_host_log[,-c(1:4)]<-log(data_host_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_host_log$timepoint<-as.factor(data_host_log$timepoint)

#change species to a factor 
data_host_log$species<-as.factor(data_host_log$species)
```

##Update site naming
```{r}
data_all_host.comp<-data_host_log

#data_all_host.comp<-separate(data_all_host.comp, "site_code", into=c("site", "nutrient"), remove=FALSE)
#data_all_host.comp$siteID<-paste(data_all_host.comp$site, data_all_host.comp$nutrient, sep="")
```


##Differences in dispersion between species
```{r}
species.disp_host<-betadisper(vegdist(data_all_host.comp[,5:9], "euclidean"), data_all_host.comp$species)
anova(species.disp_host)
#           Df Sum Sq Mean Sq F value    Pr(>F)    
#Groups      2 3.2558 1.62788  63.361 < 2.2e-16 ***
#Residuals 388 9.9685 0.02569       
#Significant difference in dispersion of host physiology between Species (cannot attribute significant effect from PERMANOVA to difference in centroids alone) 
#Difference in "plasticity" of host between species 

##Pairwise comparisons of dispersion of host physiology between Species
TukeyHSD(species.disp_host)
#                            diff         lwr         upr     p adj
#Pocillopora-Acropora -0.04189572 -0.09028505 0.006493617 0.1048045
#Porites-Acropora      0.16243124  0.11368629 0.211176178 0.0000000
#Porites-Pocillopora   0.20432695  0.15979952 0.248854392 0.0000000

#Porites is different from other species
```


##Centroid distances
```{r}
##Obtain centroid coordinates of each Sample Set (Same Species, Site, Season) with betadisper type="centroid"
data_all_host.comp$Set<-paste(data_all_host.comp$species, data_all_host.comp$site, data_all_host.comp$timepoint, sep=".")

host.disp.grp<-betadisper(vegdist(data_all_host.comp[,5:9], "euclidean"), data_all_host.comp$Set, type="centroid")

##Calculate the variance of Physiology explained by the first two PCoA axes
host.disp.grp$eig[1]/sum(host.disp.grp$eig[which(host.disp.grp$eig>0)])*100
host.disp.grp$eig[2]/sum(host.disp.grp$eig[which(host.disp.grp$eig>0)])*100
sum(host.disp.grp$eig[1:2])/sum(host.disp.grp$eig[which(host.disp.grp$eig>0)])*100
#94.16% explained by first two axes 

##Calculate the Euclidean distance between centroids using coordinates of the first 2 PCoA axes
centdist.host<-as.matrix(dist(host.disp.grp$centroids[,c(1:2)], method = "euclidean"))

##Create a dataframe of distances between centroids of Sample Sets
cents.host<-data.frame(SetA=colnames(centdist.host)[col(centdist.host)], SetB=rownames(centdist.host)[row(centdist.host)], Cent.Dist.Host=c(centdist.host))

##Remove rows comparing self-pairs of Sample Sets 
cents.host<-cents.host[-c(which(cents.host$SetA==cents.host$SetB)),]

##Remove repeats (Ex: Set 1 vs Set 2 and Set 2 vs Set 1)
cents.host<-cents.host[!duplicated(cents.host$Cent.Dist),]

##Split Sets column to add variables of interest
cents.host<-separate(cents.host, "SetA", into=c("species.A", "site.A", "timepoint.A"), remove=FALSE)
cents.host<-separate(cents.host, "SetB", into=c("species.B", "site.B", "timepoint.B"), remove=FALSE)

##Filter to only keep rows comparing the same Species
cents.host<-cents.host[c(which(cents.host$species.A==cents.host$species.B)),]

##Add variable of Pair for merging
cents.host$Pair<-paste(cents.host$SetA, cents.host$SetB, sep="_")

##Species as factor
cents.host$species<-factor(cents.host$species.A, levels=c("Acropora", "Pocillopora", "Porites"))

##Dataframe with distances between Seasons by only keeping rows comparing the same Site
cents.host.seas<-cents.host[c(which(cents.host$site.A==cents.host$site.B)),]

##Dataframe with distances between Sites by only keeping rows comparing the same Season
cents.host.site<-cents.host[c(which(cents.host$timepoint.A==cents.host$timepoint.B)),]

```

##Run ANOVA on plasticity values 

Run anova on seasonal plasticity between species.
```{r}
summary(aov(Cent.Dist.Host~species*site.A, data=cents.host.seas))
#               Df Sum Sq Mean Sq F value   Pr(>F)    
#species         2 0.4221 0.21104   9.702 0.000314 ***
#site.A          2 0.0251 0.01254   0.576 0.565970    
#species:site.A  4 0.1314 0.03286   1.511 0.215163    
#Residuals      45 0.9788 0.02175  
#Significant difference in seasonal plasticity of host between species
#The effect of season does not differ between sites

compare_means(Cent.Dist.Host~species, data=cents.host.seas, method = "anova")
#   .y.                   p   p.adj p.format p.signif method
#   <chr>             <dbl>   <dbl> <chr>    <chr>    <chr> 
# 1 Cent.Dist.Host 0.000316 0.00032 0.00032  ***      Anova 
  
```

Run anova on spatial plasticity between species.
```{r}
summary(aov(Cent.Dist.Host~species*timepoint.A, data=cents.host.site))
#                    Df  Sum Sq  Mean Sq F value Pr(>F)  
#species              2 0.05497 0.027483   2.949 0.0716 .
#timepoint.A          3 0.00093 0.000312   0.033 0.9916  
#species:timepoint.A  6 0.07374 0.012290   1.319 0.2871  
#Residuals           24 0.22366 0.009319     
#No difference in spatial plasticity of host between species

compare_means(Cent.Dist.Host~species, data=cents.host.site, method = "anova")
#   .y.                 p p.adj p.format p.signif method
#   <chr>           <dbl> <dbl> <chr>    <chr>    <chr> 
# 1 Cent.Dist.Host 0.0614 0.061 0.061    ns       Anova 
```

##Generate figures- Seasonal and Spatial Plasticity
```{r}

host.season.plasticity.plot<-ggplot(cents.host.seas, aes(x=species.A, y=Cent.Dist.Host, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
  stat_compare_means(method = "anova", label.x=1.75, size=6)+
  stat_compare_means(comparisons = spp_comparisons, size=5)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,.85)+
  ylab("Host Seasonal Plasticity")+
  xlab("Species");host.season.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_season_host.png", host.season.plasticity.plot, width=4, height=4, dpi=300)


host.spatial.plasticity.plot<-ggplot(cents.host.site, aes(x=species.A, y=Cent.Dist.Host, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
  stat_compare_means(method = "anova", label.x=1.75, size=6)+
  stat_compare_means(comparisons = spp_comparisons, size=5)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,.85)+
  ylab("Host Spatial Plasticity")+
  xlab("Species");host.spatial.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_spatial_host.png", host.spatial.plasticity.plot, width=4, height=4, dpi=300)
```

# Symbiont responses
```{r}
data_sym<-master%>%
  select(colony_id_corr, timepoint, species, site, all_of(symbiont_responses))%>%
  #filter(Host_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2>0)%>%
  filter(Ratio_AFDW.mg.cm2>0)%>%
  filter(Sym_AFDW.mg.cm2<20)%>%
  filter(Ic<1000)%>%
  #filter(prot_mg.mgafdw<1.5)%>% #remove outliers
  rename(Symbiont_Biomass=Sym_AFDW.mg.cm2, S_H_Biomass_Ratio=Ratio_AFDW.mg.cm2, Symbiont_Density=cells.mgAFDW)

data_sym<-data_sym[complete.cases(data_sym), ]
```


##Log+1 transform all responses
```{r}
data_sym_log <-data_sym
data_sym_log[,-c(1:4)]<-log(data_sym_log[,-c(1:4)]+1)

#change timepoint to a factor 
data_sym_log$timepoint<-as.factor(data_sym_log$timepoint)

#change species to a factor 
data_sym_log$species<-as.factor(data_sym_log$species)
```

##Update site naming
```{r}
data_all_sym.comp<-data_sym_log

#data_all_sym.comp<-separate(data_all_sym.comp, "site_code", into=c("site", "nutrient"), remove=FALSE)
#data_all_sym.comp$siteID<-paste(data_all_sym.comp$site, data_all_sym.comp$nutrient, sep="")
```


##Differences in dispersion between species
```{r}
species.disp_sym<-betadisper(vegdist(data_all_sym.comp[,5:13], "euclidean"), data_all_sym.comp$species)
anova(species.disp_sym)
#           Df Sum Sq  Mean Sq F value Pr(>F)
#Groups      2  0.149 0.074543  0.5552 0.5744
#Residuals 412 55.317 0.134265    
#No significant difference in dispersion of symbiont physiology between Species 
```


##Centroid distances
```{r}
##Obtain centroid coordinates of each Sample Set (Same Species, Site, Season) with betadisper type="centroid"
data_all_sym.comp$Set<-paste(data_all_sym.comp$species, data_all_sym.comp$site, data_all_sym.comp$timepoint, sep=".")

sym.disp.grp<-betadisper(vegdist(data_all_sym.comp[,5:13], "euclidean"), data_all_sym.comp$Set, type="centroid")

##Calculate the variance of Physiology explained by the first two PCoA axes
sym.disp.grp$eig[1]/sum(sym.disp.grp$eig[which(sym.disp.grp$eig>0)])*100
sym.disp.grp$eig[2]/sum(sym.disp.grp$eig[which(sym.disp.grp$eig>0)])*100
sum(sym.disp.grp$eig[1:2])/sum(sym.disp.grp$eig[which(sym.disp.grp$eig>0)])*100
#71.47% explained by first two axes 

##Calculate the Euclidean distance between centroids using coordinates of the first 2 PCoA axes
centdist.sym<-as.matrix(dist(sym.disp.grp$centroids[,c(1:2)], method = "euclidean"))

##Create a dataframe of distances between centroids of Sample Sets
cents.sym<-data.frame(SetA=colnames(centdist.sym)[col(centdist.sym)], SetB=rownames(centdist.sym)[row(centdist.sym)], Cent.Dist.Sym=c(centdist.sym))

##Remove rows comparing self-pairs of Sample Sets 
cents.sym<-cents.sym[-c(which(cents.sym$SetA==cents.sym$SetB)),]

##Remove repeats (Ex: Set 1 vs Set 2 and Set 2 vs Set 1)
cents.sym<-cents.sym[!duplicated(cents.sym$Cent.Dist),]

##Split Sets column to add variables of interest
cents.sym<-separate(cents.sym, "SetA", into=c("species.A", "site.A", "timepoint.A"), remove=FALSE)
cents.sym<-separate(cents.sym, "SetB", into=c("species.B", "site.B", "timepoint.B"), remove=FALSE)

##Filter to only keep rows comparing the same Species
cents.sym<-cents.sym[c(which(cents.sym$species.A==cents.sym$species.B)),]

##Add variable of Pair for merging
cents.sym$Pair<-paste(cents.sym$SetA, cents.sym$SetB, sep="_")

##Species as factor
cents.sym$species<-factor(cents.sym$species.A, levels=c("Acropora", "Pocillopora", "Porites"))

##Dataframe with distances between Seasons by only keeping rows comparing the same Site
cents.sym.seas<-cents.sym[c(which(cents.sym$site.A==cents.sym$site.B)),]

##Dataframe with distances between Sites by only keeping rows comparing the same Season
cents.sym.site<-cents.sym[c(which(cents.sym$timepoint.A==cents.sym$timepoint.B)),]

```

##Run ANOVA on plasticity values 

Run anova on seasonal plasticity between species.
```{r}
summary(aov(Cent.Dist.Sym~species*site.A, data=cents.sym.seas))
#               Df Sum Sq Mean Sq F value Pr(>F)  
#species         2  0.814  0.4068   2.691 0.0787 .
#site.A          2  0.670  0.3352   2.218 0.1206  
#species:site.A  4  0.183  0.0456   0.302 0.8751  
#Residuals      45  6.801  0.1511      
#No significant difference in seasonal plasticity of symbiont between species
#The effect of season does not differ between sites

compare_means(Cent.Dist.Sym~species, data=cents.sym.seas, method = "anova")
#   .y.                p p.adj p.format p.signif method
#   <chr>          <dbl> <dbl> <chr>    <chr>    <chr> 
# 1 Cent.Dist.Sym 0.0761 0.076 0.076    ns       Anova  
```

Run anova on spatial plasticity between species.
```{r}
summary(aov(Cent.Dist.Sym~species*timepoint.A, data=cents.sym.site))
#                    Df Sum Sq Mean Sq F value   Pr(>F)    
#species              2 0.2791 0.13957   3.830 0.036015 *  
#timepoint.A          3 0.8297 0.27656   7.589 0.000975 ***
#species:timepoint.A  6 0.1446 0.02410   0.661 0.681178    
#Residuals           24 0.8746 0.03644  
#Significant difference in spatial plasticity of symbiont between species
#Significant difference in spatial plasticity of symbiont between seasons

compare_means(Cent.Dist.Sym~species, data=cents.sym.site, method = "anova")
#   .y.                 p  p.adj p.format p.signif method
#   <chr>           <dbl>  <dbl> <chr>    <chr>    <chr> 
# 1 Cent.Dist.Sym 0.0983 0.098 0.098    ns       Anova 
```

##Generate figures- Seasonal and Spatial Plasticity
```{r}

sym.season.plasticity.plot<-ggplot(cents.sym.seas, aes(x=species.A, y=Cent.Dist.Sym, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
  stat_compare_means(method = "anova", label.x=1.75, size=6)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,1.7)+
  ylab("Symbiont Seasonal Plasticity")+
  xlab("Species");sym.season.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_season_sym.png", sym.season.plasticity.plot, width=4, height=4, dpi=300)


sym.spatial.plasticity.plot<-ggplot(cents.sym.site, aes(x=species.A, y=Cent.Dist.Sym, fill=species.A)) + 
  geom_boxplot(alpha=0.8, outlier.shape = NA)+
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.5, size=3)+
  scale_fill_manual(name="Species", values = c("darkgray", "orange", "purple"))+
  stat_compare_means(method = "anova", label.x=1.75, size=6)+
  stat_compare_means(comparisons = spp_comparisons, size=5)+
  theme_classic()+
  theme(axis.title.x = element_text(size = 18), axis.title.y = element_text(size = 18), axis.text.x = element_text(size=14, face="italic"), axis.text.y = element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=15), legend.position="none")+
  ylim(0,1.7)+
  ylab("Symbiont Spatial Plasticity")+
  xlab("Species");sym.spatial.plasticity.plot

ggsave("Figures/Plasticity/dispersion_plasticity_spatial_sym.png", sym.spatial.plasticity.plot, width=4, height=4, dpi=300)
```


#Correlation between Host and Symbiont Seasonal Plasticity 
```{r}

##Merge Host and Symbiont Seasonal plasticity dataframes
cent.dists.all.seas<-merge(cents.host.seas, cents.sym.seas)

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.all.seas$Cent.Dist.Host, cent.dists.all.seas$Cent.Dist.Sym, method="spearman")
#data:  cent.dists.all.seas$Cent.Dist.Host and cent.dists.all.seas$Cent.Dist.Sym
#S = 33652, p-value = 0.03868
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.2827139 


####Acropora
cent.dists.acr.seas<-cent.dists.all.seas%>%
  filter(species.A=="Acropora")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.acr.seas$Cent.Dist.Host, cent.dists.acr.seas$Cent.Dist.Sym, method="spearman")
#S = 1116, p-value = 0.5467
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.1517028 


####Porites
cent.dists.por.seas<-cent.dists.all.seas%>%
  filter(species.A=="Porites")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.por.seas$Cent.Dist.Host, cent.dists.por.seas$Cent.Dist.Sym, method="spearman")
#S = 1256, p-value = 0.232
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.2961816 

####Pocillopora
cent.dists.poc.seas<-cent.dists.all.seas%>%
  filter(species.A=="Pocillopora")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.poc.seas$Cent.Dist.Host, cent.dists.poc.seas$Cent.Dist.Sym, method="spearman")
#S = 966, p-value = 0.9934
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#        rho 
#0.003095975
```


#Correlation between Host and Symbiont Spatial Plasticity 
```{r}

##Merge Host and Symbiont Spatial plasticity dataframes
cent.dists.all.site<-merge(cents.host.site, cents.sym.site)

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.all.site$Cent.Dist.Host, cent.dists.all.site$Cent.Dist.Sym, method="spearman")
#S = 8970, p-value = 0.367
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.1544402 


####Acropora
cent.dists.acr.site<-cent.dists.all.site%>%
  filter(species.A=="Acropora")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.acr.site$Cent.Dist.Host, cent.dists.acr.site$Cent.Dist.Sym, method="spearman")
#S = 370, p-value = 0.3543
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.2937063 


####Porites
cent.dists.por.site<-cent.dists.all.site%>%
  filter(species.A=="Porites")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.por.site$Cent.Dist.Host, cent.dists.por.site$Cent.Dist.Sym, method="spearman")
#S = 386, p-value = 0.266
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.3496503 

####Pocillopora
cent.dists.poc.site<-cent.dists.all.site%>%
  filter(species.A=="Pocillopora")

##Correlation between Plasticity of Host and Symbiont across Seasons
cor.test(cent.dists.poc.site$Cent.Dist.Host, cent.dists.poc.site$Cent.Dist.Sym, method="spearman")
#S = 406, p-value = 0.1766
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#       rho 
#-0.4195804 
```
