---
title: "Integrating E5 time series environmental data"
author: "Ariana S Huffmyer"
date: "11/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

AH needs to recalculate temperature metrics - how do we determine temperatures from a time point? Are these ranges between time points or do we want it to be 2 weeks around each time point? 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```
 
```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("scales")) install.packages("scales")
if (!require("ggiraphExtra")) install.packages("ggiraphExtra")
#if (!require("ggradar")) remotes::install_github("ricardo-bion/ggradar")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(scales)
#library(ggradar)
library(ggiraphExtra)
```
 
# Load and manipulate data  

Load all .csv files from output of all timepoints for temp/pH, hobo temp, and light measurements.        
 
```{r}
temp_pH_files<-list.files("../", pattern = "pH_temp.csv", recursive=T, full.names=T)
light_files<-list.files("../", pattern = "light.csv", recursive=T, full.names=T)
hobo_files<-list.files("../", pattern = "hobo.csv", recursive=T, full.names=T)
```
 

## Read data files 
 
Load all data frames.  

```{r}
#temp and pH 
temp_pH_dataset <- data.frame()

for (i in 1:length(temp_pH_files)){
  temp_pH_df <- read.csv(temp_pH_files[i]) #each file will be read in
  temp_pH_dataset <- rbind(temp_pH_dataset, temp_pH_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#light 
light_dataset <- data.frame()

for (i in 1:length(light_files)){
  light_df <- read.csv(light_files[i]) #each file will be read in
  light_dataset <- rbind(light_dataset, light_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#hobo temp 
hobo_dataset <- data.frame()

for (i in 1:length(hobo_files)){
  hobo_df <- read.csv(hobo_files[i]) #each file will be read in
  hobo_dataset <- rbind(hobo_dataset, hobo_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#remove files that are not needed from loops
rm(list = ls(pattern = "*_df"))
```

## Generate master data frame    

Format dataframes for date time.    

```{r}
light_dataset$Date.Time<-as.POSIXct(light_dataset$Date.Time, "%y/%m/%d %H:%M:%S")
temp_pH_dataset$Date.Time<-as.POSIXct(temp_pH_dataset$Date.Time, "%y/%m/%d %H:%M:%S")
hobo_dataset$Date.Time<-as.POSIXct(hobo_dataset$Date.Time, "%y/%m/%d %H:%M:%S")
```

Join all data frames together.  

```{r}
#can also use "join_all"
master <- full_join(full_join(
  temp_pH_dataset,
  light_dataset, by=c("site", "timepoint", "Date.Time")),
  hobo_dataset, by=c("site", "timepoint", "Date.Time"))

head(master)
```

# Plot data over time series with raw data  

Plot temperature from pH probes.    

```{r}
temp_plot<-master%>%
  ggplot(aes(x=Date.Time, y=Temp, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=Temp,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Temperature (°C)")+
  xlab("Date");temp_plot
```

Plot temperature from Hobo loggers    

```{r}
hobo_plot<-master%>%
  ggplot(aes(x=Date.Time, y=Hobo, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=Hobo,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Hobo Temperature (°C)")+
  xlab("Date");hobo_plot
```

Plot pH   

```{r}
pH_plot<-master%>%
  ggplot(aes(x=Date.Time, y=pH, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=pH,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("pH")+
  xlab("Date");pH_plot
```

Plot light   

```{r}
light_plot<-master%>%
  ggplot(aes(x=Date.Time, y=light, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=light,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Light (PAR)")+
  xlab("Date");light_plot
```

Generate a panel of these three plots.  

```{r}
environmental_fig<-plot_grid(temp_plot,hobo_plot, pH_plot,light_plot,rel_widths=c(1,1,1,1), ncol=1, nrow=4, rel_heights= c(1,1,1,1),  label_y=1, align="vh")

ggsave(filename="Figures/Environmental/Environmental_Timeseries_Raw.jpg", plot=environmental_fig, dpi=300, width=10, height=8, units="in")
```

# Generate table of raw data summary statistics  

```{r}
summary_stats<-master%>%
  dplyr::select(Date.Time, site, timepoint, Temp, pH, light, Hobo)%>%
  group_by(site)%>%
  summarise(across(Temp:Hobo, list(min=min, max=max, mean=mean, stdev=sd), na.rm=TRUE));summary_stats #%>%
  #mutate(Temp30=sum(with(., Temp_max>30)));summary_stats
  #write_csv("Output/Environmental_Summary_Stats_Raw.csv")
```


# QC and plotting clean data  

## Temperature from pH probes  

Trim data out for temperature for data from site 3 after there appears to be a probe malfunction.  

```{r}
master_qc<-master%>%
  mutate(Temp = if_else(site=="site1" & Date.Time > '2020-03-01 00:00:00', "NA", as.character(Temp)))%>%
  mutate(Temp=as.numeric(Temp))
```

Trim out data from probe errors (likely from read outs).  

```{r}
master_qc<-master_qc%>%
  mutate(Temp = if_else(Date.Time > '2019-12-30 00:00:00' & Date.Time < '2020-01-14 00:00:00', "NA", as.character(Temp)))%>%
  mutate(Temp=as.numeric(Temp))%>%
  mutate(Temp = if_else(Date.Time > '2020-02-29 00:00:00' & Date.Time < '2020-03-07 00:00:00', "NA", as.character(Temp)))%>%
  mutate(Temp=as.numeric(Temp))
```

Trim out data from start of series from deployment.  

```{r}
master_qc<-master_qc%>%
  mutate(Temp = if_else(Date.Time < '2019-10-30 00:00:00', "NA", as.character(Temp)))%>%
  mutate(Temp=as.numeric(Temp))
```

View temperature plot again.  

```{r}
temp_plotQC<-master_qc%>%
  ggplot(aes(x=Date.Time, y=Temp, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=Temp,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Temperature (°C)")+
  xlab("Date");temp_plotQC
```

## pH 

Remove pH value from site 1 due to probe malfunction as above.    
```{r}
master_qc<-master_qc%>%
  mutate(pH = if_else(site=="site1" & Date.Time > '2020-03-01 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))
```

Trim out data from probe errors (likely from read outs).  

```{r}
master_qc<-master_qc%>%
  mutate(pH = if_else(Date.Time > '2019-12-30 00:00:00' & Date.Time < '2020-01-14 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))%>%
  mutate(pH = if_else(Date.Time > '2020-02-29 00:00:00' & Date.Time < '2020-03-07 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))
```

Trim out data from start of series from deployment.  

```{r}
master_qc<-master_qc%>%
  mutate(pH = if_else(Date.Time < '2019-10-30 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))
```

Remove pH probe malfunction at site 2.  

```{r}
master_qc<-master_qc%>%
  mutate(pH = if_else(Date.Time > '2020-03-17 00:00:00' & Date.Time < '2020-05-16 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))
```

Remove malfunction probe at site 2 for TP3 series.  
```{r}
master_qc<-master_qc%>%
  mutate(pH = if_else(site=="site2" & Date.Time > '2020-05-16 00:00:00' & Date.Time < '2020-08-20 00:00:00', "NA", as.character(pH)))%>%
  mutate(pH=as.numeric(pH))
```


View pH plot again.    

```{r}
pH_plotQC<-master_qc%>%
  ggplot(aes(x=Date.Time, y=pH, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=pH,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("pH")+
  xlab("Date");pH_plotQC
```

Need to see if the lower site 2 values are valid after 2020-05-16.  

## Light  

Keep only the first couple weeks of each deployment due to fouling.  
```{r}
master_qc<-master_qc%>%
  mutate(light = if_else(Date.Time > '2019-11-18 00:00:00' & Date.Time < '2020-01-15 00:00:00', "NA", as.character(light)))%>%
  mutate(light=as.numeric(light))
```

```{r}
master_qc<-master_qc%>%
  mutate(light = if_else(Date.Time > '2020-03-10 00:00:00' & Date.Time < '2020-09-29 00:00:00', "NA", as.character(light)))%>%
  mutate(light=as.numeric(light))
```

View light plot again.  

```{r}
light_plotQC<-master_qc%>%
  ggplot(aes(x=Date.Time, y=light, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=light,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Light (PAR)")+
  xlab("Date");light_plotQC
```

## Hobo temperature data, from loggers  

Trim out data from start of series from deployment.  

```{r}
master_qc<-master_qc%>%
  mutate(Hobo = if_else(Date.Time < '2019-10-30 00:00:00', "NA", as.character(Hobo)))%>%
  mutate(Hobo=as.numeric(Hobo))
```

Trim out data from probe errors (likely from read outs).  

```{r}
master_qc<-master_qc%>%
  mutate(Hobo = if_else(Date.Time > '2019-12-30 00:00:00' & Date.Time < '2020-01-14 00:00:00', "NA", as.character(Hobo)))%>%
  mutate(Hobo=as.numeric(Hobo))%>%
  mutate(Hobo = if_else(Date.Time > '2020-02-29 00:00:00' & Date.Time < '2020-03-07 00:00:00', "NA", as.character(Hobo)))%>%
  mutate(Hobo=as.numeric(Hobo))
```

Trim out data from end of series from deployment.  

```{r}
master_qc<-master_qc%>%
  mutate(Hobo = if_else(Date.Time > '2020-11-07 00:00:00', "NA", as.character(Hobo)))%>%
  mutate(Hobo=as.numeric(Hobo))
```

View plot of Hobo temperature again.  

```{r}
hobo_plotQC<-master_qc%>%
  ggplot(aes(x=Date.Time, y=Hobo, color=site))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=Hobo,color=site),se=FALSE,span=0.15,size=0.5)+
  scale_x_datetime(date_breaks="20 days",minor_breaks=waiver(),labels=date_format("%m-%d"))+
  theme_classic()+
  ylab("Hobo Temperature (°C)")+
  xlab("Date");hobo_plotQC
```


## Now assemble a revised plot after QC.  

```{r}
qc_fig<-plot_grid(temp_plotQC,hobo_plotQC,pH_plotQC,light_plotQC,rel_widths=c(1,1,1), ncol=1, nrow=4, rel_heights= c(1,1,1),  label_y=1, align="vh")

ggsave(filename="Figures/Environmental/Clean_Environmental_Timeseries.jpg", plot=qc_fig, dpi=300, width=10, height=8, units="in")
```

# Combine temperature data     

Plot correlation between data loggers.  
```{r}
corr_temp_plot<-master_qc%>%
  ggplot(aes(x=Hobo, y=Temp))+
  geom_point(size=0.5, alpha=0.3)+
  #geom_smooth(aes(x=Date.Time,y=Temp,color=site),se=FALSE,span=0.15,size=0.5)+
  theme_classic()+
  ylab("Temp/pH logger temperature (°C)")+
  xlab("Hobo logger temperature (°C)");corr_temp_plot

ggsave(filename="Figures/Environmental/Hobo_Temp_Correlation.jpg", plot=corr_temp_plot, dpi=300, width=6, height=6, units="in")
```

Very tight correlation between temperature probes, so we can use the full hobo time series filled in with temperature from pH probe wherever data is missing (site 1 at the beginning of the time series). 

Combine temperature datasets to have complete time series.  

```{r}
#if hobo is na, fill with data from the pH temperature logger 
master_qc<-master_qc%>%
  mutate(FullTemp=if_else(is.na(Hobo), Temp, Hobo))
```


# Generate table of clean summary statistics  

```{r}
summary_stats_qc<-master_qc%>%
  dplyr::select(Date.Time, site, timepoint, FullTemp, pH, light)%>%
  group_by(site)%>%
  summarise(across(FullTemp:light, list(min=min, max=max, mean=mean, stdev=sd), na.rm=TRUE))%>%
  mutate(FullTemp_range=(FullTemp_max-FullTemp_min), na.rm=TRUE)%>%
  mutate(pH_range=(pH_max-pH_min), na.rm=TRUE)%>%
  write_csv("Output/Clean_Environmental_Summary_Stats.csv")
```

# Generate table of temperature statistics to correlate responses to  

```{r}
summary_stats_temp<-master_qc%>%
  dplyr::select(Date.Time, site, timepoint, FullTemp)%>%
  group_by(timepoint, site)%>%
  summarise(across(FullTemp, list(min=min, max=max, mean=mean, stdev=sd), na.rm=TRUE))%>%
  mutate(FullTemp_range=(FullTemp_max-FullTemp_min))%>%
  write_csv("Output/Clean_Environmental_Temp_Means.csv"); summary_stats_temp
```


# Output cleaned and qc dataframe  

```{r}
write_csv(master_qc, "Output/environmental_timeseries.csv")
```

# Plot radar plot of multivariate data  

## Radar plot  

To do this, we need to first summarize data by a unit of time because the loggers differ by seconds/minutes, producing a lot of na's. We can first summarize by day for temperature (using Hobo here), pH, and light. 

Calculate additional metrics of daily min/max, range, and variability.  

Temperature
```{r}
radar_data_temp<-master_qc %>%
    mutate(Date=as.factor(as.Date(Date.Time)))%>%
    select(Date, site, FullTemp)%>%
    group_by(Date, site)%>%
    summarise(across(FullTemp, list(mean=mean, min=min, max=max, sd=sd), na.rm=TRUE))%>%
    mutate(Over30=if_else(FullTemp_max>30, 1, 0))%>%
    mutate(FullTemp_range=(FullTemp_max-FullTemp_min), na.rm=TRUE)

radar_data_temp[mapply(is.infinite, radar_data_temp)] <- NA
radar_data_temp[mapply(is.nan, radar_data_temp)] <- NA

#radar plot for temperature

radar<-radar_data_temp%>%
  select(site, FullTemp_mean, FullTemp_sd, FullTemp_max, FullTemp_min, FullTemp_range, Over30)%>%
  mutate_each(funs(rescale), -site) %>%
  drop_na()%>%
  ungroup()%>%
  droplevels()%>%
  select(!Date)%>%
  rename(Mean=FullTemp_mean, Variability=FullTemp_sd, Max=FullTemp_max, Min=FullTemp_min, Range=FullTemp_range, DaysOver30=Over30)%>%
    mutate(site = if_else(site == "site3", "Hilton", 
                            if_else(site == "site2", "Mahana", 
                                    if_else(site == "site1", "Manava", "NA"))))

png(filename="Figures/Environmental/Environmental_Radar_Temperature.png", width=2280, heigh=1880, units="px", res=300)
ggRadar(radar, aes(colour=site), scales="fixed")+
  ggplot2::theme_minimal()+
  ggplot2::scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633")) + 
  ggplot2::theme(text=element_text(color="black", size=11))+
  ggplot2::theme(legend.position="top")
dev.off()

```

pH and light 
```{r}
radar_data_pH<-master_qc %>%
    mutate(Date=as.factor(as.Date(Date.Time)))%>%
    select(Date, site, FullTemp, pH, light)%>%
    group_by(Date, site)%>%
    summarise(across(pH:light, list(mean=mean, min=min, max=max, sd=sd), na.rm=TRUE))%>%
    mutate(pH_range=(pH_max-pH_min), na.rm=TRUE)

radar_data_pH[mapply(is.infinite, radar_data_pH)] <- NA
radar_data_pH[mapply(is.nan, radar_data_pH)] <- NA

#radar plot for ph/light

radar<-radar_data_pH%>%
  select(site, pH_mean, light_mean, pH_sd, pH_max, pH_min, pH_range, light_max)%>%
  mutate_each(funs(rescale), -site) %>%
  drop_na()%>%
  ungroup()%>%
  droplevels()%>%
  select(!Date)%>%
  rename(pH_Mean=pH_mean, pH_Variability=pH_sd, pH_Max=pH_max, pH_Min=pH_min, pH_Range=pH_range, Light_Mean=light_mean, Light_Max=light_max)%>%
  mutate(site = if_else(site == "site3", "Hilton", 
                            if_else(site == "site2", "Mahana", 
                                    if_else(site == "site1", "Manava", "NA"))))

png(filename="Figures/Environmental/Environmental_Radar_pH_light.png", width=2280, heigh=1880, units="px", res=300)
ggRadar(radar, aes(colour=site), scales="fixed")+
  ggplot2::theme_minimal()+
  ggplot2::theme(text=element_text(color="black", size=11))+
  ggplot2::theme(legend.position="top")+
  ggplot2::scale_colour_manual(values=c("#374d7c", "#00cccc", "#ff6633")) +
  ggplot2::scale_fill_manual(values=c("#374d7c", "#00cccc", "#ff6633"))
dev.off()

```

## Temperature summary plot 

```{r}
temp_summary_plot <- master_qc %>%
  ggplot(., aes(x=Date.Time, y=FullTemp)) + theme_classic() + 
  geom_line(size=0.3)

ggsave(file="Figures/Environmental/Environmental_Temperature_summary.png", temp_summary_plot, width = 7, height = 4, units = c("in"))
```


## Statistical tests  

Conduct ANOVA across sites to look at differences in these environmental characteristics.  

*Temperature variables*
```{r}
summary(aov(FullTemp_mean~site, radar_data_temp))
summary(aov(FullTemp_sd~site, radar_data_temp))
summary(aov(FullTemp_max~site, radar_data_temp))
summary(aov(FullTemp_min~site, radar_data_temp))
summary(aov(FullTemp_range~site, radar_data_temp))
```


*Light variables*
```{r}
summary(aov(light_mean~site, radar_data_pH))
summary(aov(light_max~site, radar_data_pH))
```

Mean daily light and max daily light are different by site.  

*pH variables*  
```{r}
summary(aov(pH_mean~site, radar_data_pH))
summary(aov(pH_sd~site, radar_data_pH))
summary(aov(pH_max~site, radar_data_pH))
summary(aov(pH_min~site, radar_data_pH))
summary(aov(pH_range~site, radar_data_pH))
```

Mean daily pH, daily pH variability (stdev), max daily pH, min daily pH, and pH daily range are different by site.  

# Plot boxplots of data for each timepoint as a summary.    

Summarize data.  

```{r}
boxplot_data<-master_qc %>%
    mutate(Date=as.factor(as.Date(Date.Time)))%>%
    select(Date, timepoint, site, FullTemp, pH, light)%>%
    group_by(Date, site, timepoint)%>%
    summarise(across(FullTemp:light, list(mean=mean, min=min, max=max, sd=sd), na.rm=TRUE))%>%
    mutate(FullTemp_range=(FullTemp_max-FullTemp_min), na.rm=TRUE)%>%
    mutate(pH_range=(pH_max-pH_min), na.rm=TRUE)

boxplot_data[mapply(is.infinite, boxplot_data)] <- NA
boxplot_data[mapply(is.nan, boxplot_data)] <- NA
```

Temperature
```{r}
#mean temp
temp_plot1<-boxplot_data %>%
  filter(!is.na(FullTemp_mean))%>%

  ggplot(., aes(x = site, y = FullTemp_mean, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(25,32)+
    ylab("Daily Mean Temperature °C")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );temp_plot1

#min temp
temp_plot2<-boxplot_data %>%
  filter(!is.na(FullTemp_min))%>%

  ggplot(., aes(x = site, y = FullTemp_min, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(25,32)+
    ylab("Daily Minimum Temperature °C")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );temp_plot2

#max temp
temp_plot3<-boxplot_data %>%
  filter(!is.na(FullTemp_max))%>%

  ggplot(., aes(x = site, y = FullTemp_max, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(25,32)+
    ylab("Daily Maximum Temperature °C")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );temp_plot3

#variability temp
temp_plot4<-boxplot_data %>%
  filter(!is.na(FullTemp_sd))%>%

  ggplot(., aes(x = site, y = FullTemp_sd, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
   #ylim(25,32)+
    ylab("Daily Temperature Variability (stdev; °C)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );temp_plot4

#range temp
temp_plot5<-boxplot_data %>%
  filter(!is.na(FullTemp_range))%>%

  ggplot(., aes(x = site, y = FullTemp_range, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
   #ylim(25,32)+
    ylab("Daily Temperature Range (°C)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );temp_plot5

#assemble plot
temp_fig<-plot_grid(temp_plot1, temp_plot2, temp_plot3,temp_plot4,temp_plot5,rel_widths=c(1,1,1,1,1), ncol=5, nrow=1, rel_heights= c(1,1,1,1,1),  label_y=1, align="vh")

ggsave(filename="Figures/Environmental/temp_boxplots.jpg", plot=temp_fig, dpi=200, width=24, height=6, units="in")

```

pH
```{r}
#mean pH
pH_plot1<-boxplot_data %>%
  filter(!is.na(pH_mean))%>%

  ggplot(., aes(x = site, y = pH_mean, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(7.9,8.7)+
    ylab("Daily Mean pH")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );pH_plot1

#min pH
pH_plot2<-boxplot_data %>%
  filter(!is.na(pH_min))%>%

  ggplot(., aes(x = site, y = pH_min, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(7.9,8.7)+
    ylab("Daily Minimum pH")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );pH_plot2

#max pH
pH_plot3<-boxplot_data %>%
  filter(!is.na(pH_max))%>%

  ggplot(., aes(x = site, y = pH_max, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(7.9,8.7)+
    ylab("Daily Maximum pH")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );pH_plot3

#variability pH
pH_plot4<-boxplot_data %>%
  filter(!is.na(pH_sd))%>%

  ggplot(., aes(x = site, y = pH_sd, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(0,0.2)+
    ylab("Daily pH Variability (stdev)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );pH_plot4

#range pH
pH_plot5<-boxplot_data %>%
  filter(!is.na(pH_range))%>%

  ggplot(., aes(x = site, y = pH_range, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(0,0.8)+
    ylab("Daily pH Range")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );pH_plot5

#assemble plot
pH_fig<-plot_grid(pH_plot1, pH_plot2, pH_plot3,pH_plot4,pH_plot5,rel_widths=c(1,1,1,1,1), ncol=5, nrow=1, rel_heights= c(1,1,1,1,1),  label_y=1, align="vh")

ggsave(filename="Figures/Environmental/pH_boxplots.jpg", plot=pH_fig, dpi=200, width=24, height=6, units="in")
```

Light
```{r}
#mean light
light_plot1<-boxplot_data %>%
  filter(!is.na(light_mean))%>%
   filter(!timepoint=="timepoint3")%>%
  droplevels()%>%

  ggplot(., aes(x = site, y = light_mean, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(0,2000)+
    ylab("Daily Mean Light (PAR)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );light_plot1

#max light
light_plot2<-boxplot_data %>%
  filter(!is.na(light_max))%>%
  filter(!timepoint=="timepoint3")%>%
  droplevels()%>%

  ggplot(., aes(x = site, y = light_max, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(0,5000)+
    ylab("Daily Maximum Light (PAR)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );light_plot2

#light sd
light_plot3<-boxplot_data %>%
  filter(!is.na(light_sd))%>%
  filter(!timepoint=="timepoint3")%>%
  droplevels()%>%

  ggplot(., aes(x = site, y = light_sd, fill = site, colour=site)) +
    geom_boxplot(fill="white", size=1) +
    geom_point(colour="black",pch = 21, size=1, alpha=0.5, position = position_jitterdodge(0.4)) + 
    facet_grid(~timepoint) +
    xlab("Site") + 
    theme_classic() + 
    ylim(0,2000)+
    ylab("Daily Light Variability (PAR)")+
    theme(
      legend.position="none",
      legend.title=element_text(face="bold", size=14),
      legend.text=element_text(size=14),
      axis.title=element_text(face="bold", size=14),
      axis.text=element_text(size=10, color="black"), 
      strip.text.x=element_text(face="italic", size=14)
      );light_plot3

#assemble plot
light_fig<-plot_grid(light_plot1, light_plot2, light_plot3,rel_widths=c(1,1,1), ncol=3, nrow=1, rel_heights= c(1,1,1),  label_y=1, align="vh")

ggsave(filename="Figures/Environmental/light_boxplots.jpg", plot=light_fig, dpi=200, width=16, height=6, units="in")
```
