---
title: "Integrating E5 time series biological data"
author: "Ariana S Huffmyer"
date: "1/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data files 

Load all .csv files from output of all timepoints for each biological response

```{r}
all_files<-list.files(path="../", pattern = "*.csv", recursive=T)

biomass_files<-list.files("../", pattern = "biomass.csv", recursive=T)
antioxidant_files<-list.files("../", pattern = "antioxidant_capacity.csv", recursive=T)
pi_curve_rate_files<-list.files("../", pattern = "pi_curve_rates.csv", recursive=T)
surface_area_files<-list.files("../", pattern = "surface_area.csv", recursive=T)
pi_curve_pars_files<-list.files("../", pattern = "pi_curve_pars.csv", recursive=T)
protein_files<-list.files("../", pattern = "protein.csv", recursive=T)
symb_densities_files<-list.files("../", pattern = "symb_densities.csv", recursive=T)
chlorophyll_files<-list.files("../", pattern = "chlorophyll.csv", recursive=T)

#need to add calcification
```

## Correct, control for colony id  

## Generate master data frame  















2. Generate master data file for each or all parameters that include a column for timepoint  

3. Plot parameters over time and output plots  

4. Conduct statistical analyses within and between time points and generate summary information for all parameters 
