---
title: "Integrating E5 time series biological data"
author: "Ariana S Huffmyer"
date: "1/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")

# load packages
library(tidyverse)
library(ggplot2)
```

## Loading data files 

Load all .csv files from output of all timepoints for each biological response

```{r}
biomass_files<-list.files("../", pattern = "biomass.csv", recursive=T, full.names=T)
antioxidant_files<-list.files("../", pattern = "antioxidant_capacity.csv", recursive=T, full.names=T)
pi_curve_rate_files<-list.files("../", pattern = "pi_curve_rates.csv", recursive=T, full.names=T)
surface_area_files<-list.files("../", pattern = "surface_area.csv", recursive=T, full.names=T)
pi_curve_pars_files<-list.files("../", pattern = "pi_curve_pars.csv", recursive=T, full.names=T)
protein_files<-list.files("../", pattern = "protein.csv", recursive=T, full.names=T)
symb_densities_files<-list.files("../", pattern = "symb_densities.csv", recursive=T, full.names=T)
chlorophyll_files<-list.files("../", pattern = "chlorophyll.csv", recursive=T, full.names=T)

#need to add calcification
```

## Correct and control for colony id that changed between time points due to relabeling  

```{r}
#Load tag metadata sheet 
tags<-read.csv("../metadata/coral_id_metadata.csv")

```

Correct for tags  
Add site   
Change columns to TP1,2,3,4


## Read files and generate master data frame  

Load all data frames.  

```{r}
#biomass WORKS
biomass_dataset <- data.frame()

for (i in 1:length(biomass_files)){
  biomass_df <- read.csv(biomass_files[i]) #each file will be read in
  biomass_dataset <- rbind(biomass_dataset, biomass_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#antioxidant capacity WORKS
antioxidant_dataset <- data.frame()

for (i in 1:length(antioxidant_files)){
  antioxidant_df <- read.csv(antioxidant_files[i]) #each file will be read in
  antioxidant_dataset <- rbind(antioxidant_dataset, antioxidant_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#PI curve rates- NOT READY
#pi_curve_rate_dataset <- data.frame()

#for (i in 1:length(pi_curve_rate_files)){
  #pi_curve_rate_df <- read.csv(pi_curve_rate_files[i]) #each file will be read in
  #pi_curve_rate_dataset <- rbind(pi_curve_rate_dataset, pi_curve_rate_df) #for each iteration, bind the new data to the building dataset
#}
```

```{r}
#surface area WORKS
surface_area_dataset <- data.frame()

for (i in 1:length(surface_area_files)){
  surface_area_df <- read.csv(surface_area_files[i]) #each file will be read in
  surface_area_dataset <- rbind(surface_area_dataset, surface_area_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#PI Curve parameters - NOT READY
#pi_curve_pars_dataset <- data.frame()

#for (i in 1:length(pi_curve_pars_files)){
  #pi_curve_pars_df <- read.csv(pi_curve_pars_files[i]) #each file will be read in
  #pi_curve_pars_dataset <- rbind(pi_curve_pars_dataset, pi_curve_pars_df) #for each iteration, bind the new data to the building dataset
#}
```

```{r}
#protein WORKS
protein_dataset <- data.frame()

for (i in 1:length(protein_files)){
  protein_df <- read.csv(protein_files[i]) #each file will be read in
  protein_dataset <- rbind(protein_dataset, protein_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#symb densities WORKS, NEEDS TP4
symb_densities_dataset <- data.frame()

for (i in 1:length(symb_densities_files)){
  symb_densities_df <- read.csv(symb_densities_files[i]) #each file will be read in
  symb_densities_dataset <- rbind(symb_densities_dataset, symb_densities_df) #for each iteration, bind the new data to the building dataset
}
```

```{r}
#chlorophyll files WORKS
chlorophyll_dataset <- data.frame()

for (i in 1:length(chlorophyll_files)){
  chlorophyll_df <- read.csv(chlorophyll_files[i]) #each file will be read in
  chlorophyll_dataset <- rbind(chlorophyll_dataset, chlorophyll_df) #for each iteration, bind the new data to the building dataset
}

```

```{r}
#remove files that are not needed from loops
rm(list = ls(pattern = "*_df"))
```


## Antioxidant capacity  

View by site and timepoint for each species.  
```{r}
tacplot<-antioxidant_dataset %>%
  filter(!is.na(species)) %>%
  ggplot(., aes(x = site, y = cre.umol.mgprot, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, position = position_jitterdodge()) + 
    scale_fill_brewer(palette = "YlGnBu")+
    facet_wrap(~species)+
    theme_classic(); tacplot
```

## Symbiont Densities  

View by site and timepoint for each species.  
```{r}
symbplot<-symb_densities_dataset %>%
  filter(!is.na(species)) %>%
  ggplot(., aes(x = site, y = cells.cm2, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, position = position_jitterdodge()) + 
    scale_fill_brewer(palette = "YlGnBu")+
    facet_wrap(~species)+
    theme_classic(); symbplot
```


## Biomass

View by site and timepoint for each species in the host. *Need to address negative values in dataset by removing outliers.  
```{r}
biomassHplot<-biomass_dataset %>%
  filter(!is.na(species)) %>%
  filter(partner=="Host") %>%
  ggplot(., aes(x = site, y = AFDW.mg.cm2, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, position = position_jitterdodge()) + 
    ylim(0,15)+
    scale_fill_brewer(palette = "YlGnBu")+
    facet_wrap(~species)+
    ggtitle("Host Biomass") +
    theme_classic(); biomassHplot
```

View by site and timepoint for each species in the symbiont *Need to address negative values in dataset by removing outliers.  
```{r}
biomassSplot<-biomass_dataset %>%
  filter(!is.na(species)) %>%
  filter(partner=="Sym") %>%
  ggplot(., aes(x = site, y = AFDW.mg.cm2, fill = timepoint, group=interaction(site,timepoint))) +
    geom_boxplot(outlier.size = 0) +
    geom_point(pch = 21, position = position_jitterdodge()) + 
    ylim(0,15)+
    scale_fill_brewer(palette = "YlGnBu")+
    facet_wrap(~species)+
    ggtitle("Symbiont Biomass") +
    theme_classic(); biomassSplot
```