---
title: "E5 time series environmental data for RDA analyses"
author: "Ariana S Huffmyer"
date: "2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```
 
```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("scales")) install.packages("scales")
if (!require("ggiraphExtra")) install.packages("ggiraphExtra")
#if (!require("ggradar")) remotes::install_github("ricardo-bion/ggradar")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(scales)
#library(ggradar)
library(ggiraphExtra)
```
 
# MCR data for other enviornmental characteristics across seasons 

Moorea Coral Reef LTER, L. Washburn, and A. Brooks. 2024. MCR LTER: Coral Reef: Gump Station Meteorological Data, ongoing since 2006 ver 50. Environmental Data Initiative. https://doi.org/10.6073/pasta/c41e5ef3131a9328ccee7097a4221f7a (Accessed 2024-11-26). Load in data using provided R script at the citation.  

```{r}
# Package ID: knb-lter-mcr.9.50 Cataloging System:https://pasta.edirepository.org.
# Data set title: MCR LTER: Coral Reef: Gump Station Meteorological Data, ongoing since 2006.
# Data set creator:    - Moorea Coral Reef LTER 
# Data set creator:  Libe Washburn - Moorea Coral Reef LTER 
# Data set creator:  Andrew Brooks - Moorea Coral Reef LTER 
# Metadata Provider:  Hillary Krumbholz - MCR LTER 
# Contact:  Hillary Krumbholz - Information Manager MCR LTER  - hkrumbholz@ucsb.edu
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/6abebe52e6b0245fe74cf0be7191e379" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
 dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "time_local",     
                    "time_utc",     
                    "air_temp_C",     
                    "data_coverage_air_temp_C",     
                    "wind_stress_pa",     
                    "data_coverage_wind_stress_pa",     
                    "wind_direction_degtrue",     
                    "data_coverage_wind_direction_degtrue",     
                    "wind_U_mps",     
                    "data_coverage_wind_U_mps",     
                    "wind_V_mps",     
                    "data_coverage_wind_V_mps",     
                    "mean_solar_rad_kwpm2",     
                    "data_coverage_mean_solar_rad_kwpm2",     
                    "atm_press_mbar",     
                    "data_coverage_atm_press_mbar",     
                    "cumulative_rainfall_mm",     
                    "data_coverage_cumulative_rainfall_mm"    ), check.names=TRUE)
               
unlink(infile1)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                                                   
# attempting to convert dt1$time_local dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp1time_local<-as.POSIXct(dt1$time_local,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt1[dt1$time_local != "",]) == length(tmp1time_local[!is.na(tmp1time_local)])){dt1$time_local <- tmp1time_local } else {print("Date conversion failed for dt1$time_local. Please inspect the data and do the date conversion yourself.")}                                                                    
                                                                   
# attempting to convert dt1$time_utc dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp1time_utc<-as.POSIXct(dt1$time_utc,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt1[dt1$time_utc != "",]) == length(tmp1time_utc[!is.na(tmp1time_utc)])){dt1$time_utc <- tmp1time_utc } else {print("Date conversion failed for dt1$time_utc. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt1$air_temp_C)=="factor") dt1$air_temp_C <-as.numeric(levels(dt1$air_temp_C))[as.integer(dt1$air_temp_C) ]               
if (class(dt1$air_temp_C)=="character") dt1$air_temp_C <-as.numeric(dt1$air_temp_C)
if (class(dt1$data_coverage_air_temp_C)=="factor") dt1$data_coverage_air_temp_C <-as.numeric(levels(dt1$data_coverage_air_temp_C))[as.integer(dt1$data_coverage_air_temp_C) ]               
if (class(dt1$data_coverage_air_temp_C)=="character") dt1$data_coverage_air_temp_C <-as.numeric(dt1$data_coverage_air_temp_C)
if (class(dt1$wind_stress_pa)=="factor") dt1$wind_stress_pa <-as.numeric(levels(dt1$wind_stress_pa))[as.integer(dt1$wind_stress_pa) ]               
if (class(dt1$wind_stress_pa)=="character") dt1$wind_stress_pa <-as.numeric(dt1$wind_stress_pa)
if (class(dt1$data_coverage_wind_stress_pa)=="factor") dt1$data_coverage_wind_stress_pa <-as.numeric(levels(dt1$data_coverage_wind_stress_pa))[as.integer(dt1$data_coverage_wind_stress_pa) ]               
if (class(dt1$data_coverage_wind_stress_pa)=="character") dt1$data_coverage_wind_stress_pa <-as.numeric(dt1$data_coverage_wind_stress_pa)
if (class(dt1$wind_direction_degtrue)=="factor") dt1$wind_direction_degtrue <-as.numeric(levels(dt1$wind_direction_degtrue))[as.integer(dt1$wind_direction_degtrue) ]               
if (class(dt1$wind_direction_degtrue)=="character") dt1$wind_direction_degtrue <-as.numeric(dt1$wind_direction_degtrue)
if (class(dt1$data_coverage_wind_direction_degtrue)=="factor") dt1$data_coverage_wind_direction_degtrue <-as.numeric(levels(dt1$data_coverage_wind_direction_degtrue))[as.integer(dt1$data_coverage_wind_direction_degtrue) ]               
if (class(dt1$data_coverage_wind_direction_degtrue)=="character") dt1$data_coverage_wind_direction_degtrue <-as.numeric(dt1$data_coverage_wind_direction_degtrue)
if (class(dt1$wind_U_mps)=="factor") dt1$wind_U_mps <-as.numeric(levels(dt1$wind_U_mps))[as.integer(dt1$wind_U_mps) ]               
if (class(dt1$wind_U_mps)=="character") dt1$wind_U_mps <-as.numeric(dt1$wind_U_mps)
if (class(dt1$data_coverage_wind_U_mps)=="factor") dt1$data_coverage_wind_U_mps <-as.numeric(levels(dt1$data_coverage_wind_U_mps))[as.integer(dt1$data_coverage_wind_U_mps) ]               
if (class(dt1$data_coverage_wind_U_mps)=="character") dt1$data_coverage_wind_U_mps <-as.numeric(dt1$data_coverage_wind_U_mps)
if (class(dt1$wind_V_mps)=="factor") dt1$wind_V_mps <-as.numeric(levels(dt1$wind_V_mps))[as.integer(dt1$wind_V_mps) ]               
if (class(dt1$wind_V_mps)=="character") dt1$wind_V_mps <-as.numeric(dt1$wind_V_mps)
if (class(dt1$data_coverage_wind_V_mps)=="factor") dt1$data_coverage_wind_V_mps <-as.numeric(levels(dt1$data_coverage_wind_V_mps))[as.integer(dt1$data_coverage_wind_V_mps) ]               
if (class(dt1$data_coverage_wind_V_mps)=="character") dt1$data_coverage_wind_V_mps <-as.numeric(dt1$data_coverage_wind_V_mps)
if (class(dt1$mean_solar_rad_kwpm2)=="factor") dt1$mean_solar_rad_kwpm2 <-as.numeric(levels(dt1$mean_solar_rad_kwpm2))[as.integer(dt1$mean_solar_rad_kwpm2) ]               
if (class(dt1$mean_solar_rad_kwpm2)=="character") dt1$mean_solar_rad_kwpm2 <-as.numeric(dt1$mean_solar_rad_kwpm2)
if (class(dt1$data_coverage_mean_solar_rad_kwpm2)=="factor") dt1$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(levels(dt1$data_coverage_mean_solar_rad_kwpm2))[as.integer(dt1$data_coverage_mean_solar_rad_kwpm2) ]               
if (class(dt1$data_coverage_mean_solar_rad_kwpm2)=="character") dt1$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(dt1$data_coverage_mean_solar_rad_kwpm2)
if (class(dt1$atm_press_mbar)=="factor") dt1$atm_press_mbar <-as.numeric(levels(dt1$atm_press_mbar))[as.integer(dt1$atm_press_mbar) ]               
if (class(dt1$atm_press_mbar)=="character") dt1$atm_press_mbar <-as.numeric(dt1$atm_press_mbar)
if (class(dt1$data_coverage_atm_press_mbar)=="factor") dt1$data_coverage_atm_press_mbar <-as.numeric(levels(dt1$data_coverage_atm_press_mbar))[as.integer(dt1$data_coverage_atm_press_mbar) ]               
if (class(dt1$data_coverage_atm_press_mbar)=="character") dt1$data_coverage_atm_press_mbar <-as.numeric(dt1$data_coverage_atm_press_mbar)
if (class(dt1$cumulative_rainfall_mm)=="factor") dt1$cumulative_rainfall_mm <-as.numeric(levels(dt1$cumulative_rainfall_mm))[as.integer(dt1$cumulative_rainfall_mm) ]               
if (class(dt1$cumulative_rainfall_mm)=="character") dt1$cumulative_rainfall_mm <-as.numeric(dt1$cumulative_rainfall_mm)
if (class(dt1$data_coverage_cumulative_rainfall_mm)=="factor") dt1$data_coverage_cumulative_rainfall_mm <-as.numeric(levels(dt1$data_coverage_cumulative_rainfall_mm))[as.integer(dt1$data_coverage_cumulative_rainfall_mm) ]               
if (class(dt1$data_coverage_cumulative_rainfall_mm)=="character") dt1$data_coverage_cumulative_rainfall_mm <-as.numeric(dt1$data_coverage_cumulative_rainfall_mm)
                
# Convert Missing Values to NA for non-dates
                
dt1$air_temp_C <- ifelse((trimws(as.character(dt1$air_temp_C))==trimws("-999")),NA,dt1$air_temp_C)               
suppressWarnings(dt1$air_temp_C <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$air_temp_C))==as.character(as.numeric("-999"))),NA,dt1$air_temp_C))
dt1$wind_stress_pa <- ifelse((trimws(as.character(dt1$wind_stress_pa))==trimws("-999")),NA,dt1$wind_stress_pa)               
suppressWarnings(dt1$wind_stress_pa <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$wind_stress_pa))==as.character(as.numeric("-999"))),NA,dt1$wind_stress_pa))
dt1$wind_direction_degtrue <- ifelse((trimws(as.character(dt1$wind_direction_degtrue))==trimws("-999")),NA,dt1$wind_direction_degtrue)               
suppressWarnings(dt1$wind_direction_degtrue <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$wind_direction_degtrue))==as.character(as.numeric("-999"))),NA,dt1$wind_direction_degtrue))
dt1$wind_U_mps <- ifelse((trimws(as.character(dt1$wind_U_mps))==trimws("-999")),NA,dt1$wind_U_mps)               
suppressWarnings(dt1$wind_U_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$wind_U_mps))==as.character(as.numeric("-999"))),NA,dt1$wind_U_mps))
dt1$wind_V_mps <- ifelse((trimws(as.character(dt1$wind_V_mps))==trimws("-999")),NA,dt1$wind_V_mps)               
suppressWarnings(dt1$wind_V_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$wind_V_mps))==as.character(as.numeric("-999"))),NA,dt1$wind_V_mps))
dt1$mean_solar_rad_kwpm2 <- ifelse((trimws(as.character(dt1$mean_solar_rad_kwpm2))==trimws("-999")),NA,dt1$mean_solar_rad_kwpm2)               
suppressWarnings(dt1$mean_solar_rad_kwpm2 <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$mean_solar_rad_kwpm2))==as.character(as.numeric("-999"))),NA,dt1$mean_solar_rad_kwpm2))
dt1$atm_press_mbar <- ifelse((trimws(as.character(dt1$atm_press_mbar))==trimws("-999")),NA,dt1$atm_press_mbar)               
suppressWarnings(dt1$atm_press_mbar <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$atm_press_mbar))==as.character(as.numeric("-999"))),NA,dt1$atm_press_mbar))
dt1$cumulative_rainfall_mm <- ifelse((trimws(as.character(dt1$cumulative_rainfall_mm))==trimws("-999")),NA,dt1$cumulative_rainfall_mm)               
suppressWarnings(dt1$cumulative_rainfall_mm <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt1$cumulative_rainfall_mm))==as.character(as.numeric("-999"))),NA,dt1$cumulative_rainfall_mm))


# Here is the structure of the input data frame:
str(dt1)                            
attach(dt1)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(time_local)
summary(time_utc)
summary(air_temp_C)
summary(data_coverage_air_temp_C)
summary(wind_stress_pa)
summary(data_coverage_wind_stress_pa)
summary(wind_direction_degtrue)
summary(data_coverage_wind_direction_degtrue)
summary(wind_U_mps)
summary(data_coverage_wind_U_mps)
summary(wind_V_mps)
summary(data_coverage_wind_V_mps)
summary(mean_solar_rad_kwpm2)
summary(data_coverage_mean_solar_rad_kwpm2)
summary(atm_press_mbar)
summary(data_coverage_atm_press_mbar)
summary(cumulative_rainfall_mm)
summary(data_coverage_cumulative_rainfall_mm) 
                # Get more details on character variables
                
detach(dt1)               
         

inUrl2  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/b414775cd20e9185f4b428639d53fc79" 
infile2 <- tempfile()
try(download.file(inUrl2,infile2,method="curl"))
if (is.na(file.size(infile2))) download.file(inUrl2,infile2,method="auto")

                   
 dt2 <-read.csv(infile2,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "airtc_avg",     
                    "batt_volt_min",     
                    "bp_mbar_avg",     
                    "rain_mm_tot",     
                    "record",     
                    "relhum_avg",     
                    "slrkw_avg",     
                    "slrkw_max",     
                    "slrkw_min",     
                    "slrkw_std",     
                    "winddir_d1_wvt",     
                    "winddir_sd1_wvt",     
                    "ws_ms_max",     
                    "ws_ms_min",     
                    "ws_ms_s_wvt",     
                    "ws_ms_std",     
                    "tstamp"    ), check.names=TRUE)
               
unlink(infile2)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(dt2$airtc_avg)=="factor") dt2$airtc_avg <-as.numeric(levels(dt2$airtc_avg))[as.integer(dt2$airtc_avg) ]               
if (class(dt2$airtc_avg)=="character") dt2$airtc_avg <-as.numeric(dt2$airtc_avg)
if (class(dt2$batt_volt_min)=="factor") dt2$batt_volt_min <-as.numeric(levels(dt2$batt_volt_min))[as.integer(dt2$batt_volt_min) ]               
if (class(dt2$batt_volt_min)=="character") dt2$batt_volt_min <-as.numeric(dt2$batt_volt_min)
if (class(dt2$bp_mbar_avg)=="factor") dt2$bp_mbar_avg <-as.numeric(levels(dt2$bp_mbar_avg))[as.integer(dt2$bp_mbar_avg) ]               
if (class(dt2$bp_mbar_avg)=="character") dt2$bp_mbar_avg <-as.numeric(dt2$bp_mbar_avg)
if (class(dt2$rain_mm_tot)=="factor") dt2$rain_mm_tot <-as.numeric(levels(dt2$rain_mm_tot))[as.integer(dt2$rain_mm_tot) ]               
if (class(dt2$rain_mm_tot)=="character") dt2$rain_mm_tot <-as.numeric(dt2$rain_mm_tot)
if (class(dt2$record)=="factor") dt2$record <-as.numeric(levels(dt2$record))[as.integer(dt2$record) ]               
if (class(dt2$record)=="character") dt2$record <-as.numeric(dt2$record)
if (class(dt2$relhum_avg)=="factor") dt2$relhum_avg <-as.numeric(levels(dt2$relhum_avg))[as.integer(dt2$relhum_avg) ]               
if (class(dt2$relhum_avg)=="character") dt2$relhum_avg <-as.numeric(dt2$relhum_avg)
if (class(dt2$slrkw_avg)=="factor") dt2$slrkw_avg <-as.numeric(levels(dt2$slrkw_avg))[as.integer(dt2$slrkw_avg) ]               
if (class(dt2$slrkw_avg)=="character") dt2$slrkw_avg <-as.numeric(dt2$slrkw_avg)
if (class(dt2$slrkw_max)=="factor") dt2$slrkw_max <-as.numeric(levels(dt2$slrkw_max))[as.integer(dt2$slrkw_max) ]               
if (class(dt2$slrkw_max)=="character") dt2$slrkw_max <-as.numeric(dt2$slrkw_max)
if (class(dt2$slrkw_min)=="factor") dt2$slrkw_min <-as.numeric(levels(dt2$slrkw_min))[as.integer(dt2$slrkw_min) ]               
if (class(dt2$slrkw_min)=="character") dt2$slrkw_min <-as.numeric(dt2$slrkw_min)
if (class(dt2$slrkw_std)=="factor") dt2$slrkw_std <-as.numeric(levels(dt2$slrkw_std))[as.integer(dt2$slrkw_std) ]               
if (class(dt2$slrkw_std)=="character") dt2$slrkw_std <-as.numeric(dt2$slrkw_std)
if (class(dt2$winddir_d1_wvt)=="factor") dt2$winddir_d1_wvt <-as.numeric(levels(dt2$winddir_d1_wvt))[as.integer(dt2$winddir_d1_wvt) ]               
if (class(dt2$winddir_d1_wvt)=="character") dt2$winddir_d1_wvt <-as.numeric(dt2$winddir_d1_wvt)
if (class(dt2$winddir_sd1_wvt)=="factor") dt2$winddir_sd1_wvt <-as.numeric(levels(dt2$winddir_sd1_wvt))[as.integer(dt2$winddir_sd1_wvt) ]               
if (class(dt2$winddir_sd1_wvt)=="character") dt2$winddir_sd1_wvt <-as.numeric(dt2$winddir_sd1_wvt)
if (class(dt2$ws_ms_max)=="factor") dt2$ws_ms_max <-as.numeric(levels(dt2$ws_ms_max))[as.integer(dt2$ws_ms_max) ]               
if (class(dt2$ws_ms_max)=="character") dt2$ws_ms_max <-as.numeric(dt2$ws_ms_max)
if (class(dt2$ws_ms_min)=="factor") dt2$ws_ms_min <-as.numeric(levels(dt2$ws_ms_min))[as.integer(dt2$ws_ms_min) ]               
if (class(dt2$ws_ms_min)=="character") dt2$ws_ms_min <-as.numeric(dt2$ws_ms_min)
if (class(dt2$ws_ms_s_wvt)=="factor") dt2$ws_ms_s_wvt <-as.numeric(levels(dt2$ws_ms_s_wvt))[as.integer(dt2$ws_ms_s_wvt) ]               
if (class(dt2$ws_ms_s_wvt)=="character") dt2$ws_ms_s_wvt <-as.numeric(dt2$ws_ms_s_wvt)
if (class(dt2$ws_ms_std)=="factor") dt2$ws_ms_std <-as.numeric(levels(dt2$ws_ms_std))[as.integer(dt2$ws_ms_std) ]               
if (class(dt2$ws_ms_std)=="character") dt2$ws_ms_std <-as.numeric(dt2$ws_ms_std)                                   
# attempting to convert dt2$tstamp dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp2tstamp<-as.POSIXct(dt2$tstamp,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt2[dt2$tstamp != "",]) == length(tmp2tstamp[!is.na(tmp2tstamp)])){dt2$tstamp <- tmp2tstamp } else {print("Date conversion failed for dt2$tstamp. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
                
# Convert Missing Values to NA for non-dates
                
dt2$airtc_avg <- ifelse((trimws(as.character(dt2$airtc_avg))==trimws("NaN")),NA,dt2$airtc_avg)               
suppressWarnings(dt2$airtc_avg <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$airtc_avg))==as.character(as.numeric("NaN"))),NA,dt2$airtc_avg))
dt2$batt_volt_min <- ifelse((trimws(as.character(dt2$batt_volt_min))==trimws("NaN")),NA,dt2$batt_volt_min)               
suppressWarnings(dt2$batt_volt_min <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$batt_volt_min))==as.character(as.numeric("NaN"))),NA,dt2$batt_volt_min))
dt2$bp_mbar_avg <- ifelse((trimws(as.character(dt2$bp_mbar_avg))==trimws("NaN")),NA,dt2$bp_mbar_avg)               
suppressWarnings(dt2$bp_mbar_avg <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$bp_mbar_avg))==as.character(as.numeric("NaN"))),NA,dt2$bp_mbar_avg))
dt2$rain_mm_tot <- ifelse((trimws(as.character(dt2$rain_mm_tot))==trimws("NaN")),NA,dt2$rain_mm_tot)               
suppressWarnings(dt2$rain_mm_tot <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$rain_mm_tot))==as.character(as.numeric("NaN"))),NA,dt2$rain_mm_tot))
dt2$record <- ifelse((trimws(as.character(dt2$record))==trimws("NaN")),NA,dt2$record)               
suppressWarnings(dt2$record <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$record))==as.character(as.numeric("NaN"))),NA,dt2$record))
dt2$relhum_avg <- ifelse((trimws(as.character(dt2$relhum_avg))==trimws("NaN")),NA,dt2$relhum_avg)               
suppressWarnings(dt2$relhum_avg <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$relhum_avg))==as.character(as.numeric("NaN"))),NA,dt2$relhum_avg))
dt2$slrkw_avg <- ifelse((trimws(as.character(dt2$slrkw_avg))==trimws("NaN")),NA,dt2$slrkw_avg)               
suppressWarnings(dt2$slrkw_avg <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$slrkw_avg))==as.character(as.numeric("NaN"))),NA,dt2$slrkw_avg))
dt2$slrkw_max <- ifelse((trimws(as.character(dt2$slrkw_max))==trimws("NaN")),NA,dt2$slrkw_max)               
suppressWarnings(dt2$slrkw_max <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$slrkw_max))==as.character(as.numeric("NaN"))),NA,dt2$slrkw_max))
dt2$slrkw_min <- ifelse((trimws(as.character(dt2$slrkw_min))==trimws("NaN")),NA,dt2$slrkw_min)               
suppressWarnings(dt2$slrkw_min <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$slrkw_min))==as.character(as.numeric("NaN"))),NA,dt2$slrkw_min))
dt2$slrkw_std <- ifelse((trimws(as.character(dt2$slrkw_std))==trimws("NaN")),NA,dt2$slrkw_std)               
suppressWarnings(dt2$slrkw_std <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$slrkw_std))==as.character(as.numeric("NaN"))),NA,dt2$slrkw_std))
dt2$winddir_d1_wvt <- ifelse((trimws(as.character(dt2$winddir_d1_wvt))==trimws("NaN")),NA,dt2$winddir_d1_wvt)               
suppressWarnings(dt2$winddir_d1_wvt <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$winddir_d1_wvt))==as.character(as.numeric("NaN"))),NA,dt2$winddir_d1_wvt))
dt2$winddir_sd1_wvt <- ifelse((trimws(as.character(dt2$winddir_sd1_wvt))==trimws("NaN")),NA,dt2$winddir_sd1_wvt)               
suppressWarnings(dt2$winddir_sd1_wvt <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$winddir_sd1_wvt))==as.character(as.numeric("NaN"))),NA,dt2$winddir_sd1_wvt))
dt2$ws_ms_max <- ifelse((trimws(as.character(dt2$ws_ms_max))==trimws("NaN")),NA,dt2$ws_ms_max)               
suppressWarnings(dt2$ws_ms_max <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$ws_ms_max))==as.character(as.numeric("NaN"))),NA,dt2$ws_ms_max))
dt2$ws_ms_min <- ifelse((trimws(as.character(dt2$ws_ms_min))==trimws("NaN")),NA,dt2$ws_ms_min)               
suppressWarnings(dt2$ws_ms_min <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$ws_ms_min))==as.character(as.numeric("NaN"))),NA,dt2$ws_ms_min))
dt2$ws_ms_s_wvt <- ifelse((trimws(as.character(dt2$ws_ms_s_wvt))==trimws("NaN")),NA,dt2$ws_ms_s_wvt)               
suppressWarnings(dt2$ws_ms_s_wvt <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$ws_ms_s_wvt))==as.character(as.numeric("NaN"))),NA,dt2$ws_ms_s_wvt))
dt2$ws_ms_std <- ifelse((trimws(as.character(dt2$ws_ms_std))==trimws("NaN")),NA,dt2$ws_ms_std)               
suppressWarnings(dt2$ws_ms_std <- ifelse(!is.na(as.numeric("NaN")) & (trimws(as.character(dt2$ws_ms_std))==as.character(as.numeric("NaN"))),NA,dt2$ws_ms_std))


# Here is the structure of the input data frame:
str(dt2)                            
attach(dt2)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(airtc_avg)
summary(batt_volt_min)
summary(bp_mbar_avg)
summary(rain_mm_tot)
summary(record)
summary(relhum_avg)
summary(slrkw_avg)
summary(slrkw_max)
summary(slrkw_min)
summary(slrkw_std)
summary(winddir_d1_wvt)
summary(winddir_sd1_wvt)
summary(ws_ms_max)
summary(ws_ms_min)
summary(ws_ms_s_wvt)
summary(ws_ms_std)
summary(tstamp) 
                # Get more details on character variables
                
detach(dt2)               
         

inUrl3  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/b3d5412513c2bda192bb29034c49f1d0" 
infile3 <- tempfile()
try(download.file(inUrl3,infile3,method="curl"))
if (is.na(file.size(infile3))) download.file(inUrl3,infile3,method="auto")

                   
 dt3 <-read.csv(infile3,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "time_local",     
                    "time_utc",     
                    "air_temp_C",     
                    "data_coverage_air_temp_C",     
                    "wind_stress_pa",     
                    "data_coverage_wind_stress_pa",     
                    "wind_direction_degtrue",     
                    "data_coverage_wind_direction_degtrue",     
                    "wind_U_mps",     
                    "data_coverage_wind_U_mps",     
                    "wind_V_mps",     
                    "data_coverage_wind_V_mps",     
                    "mean_solar_rad_kwpm2",     
                    "data_coverage_mean_solar_rad_kwpm2",     
                    "atm_press_mbar",     
                    "data_coverage_atm_press_mbar",     
                    "cumulative_rainfall_mm",     
                    "data_coverage_cumulative_rainfall_mm"    ), check.names=TRUE)
               
unlink(infile3)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                                                   
# attempting to convert dt3$time_local dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp3time_local<-as.POSIXct(dt3$time_local,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt3[dt3$time_local != "",]) == length(tmp3time_local[!is.na(tmp3time_local)])){dt3$time_local <- tmp3time_local } else {print("Date conversion failed for dt3$time_local. Please inspect the data and do the date conversion yourself.")}                                                                    
                                                                   
# attempting to convert dt3$time_utc dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp3time_utc<-as.POSIXct(dt3$time_utc,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt3[dt3$time_utc != "",]) == length(tmp3time_utc[!is.na(tmp3time_utc)])){dt3$time_utc <- tmp3time_utc } else {print("Date conversion failed for dt3$time_utc. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt3$air_temp_C)=="factor") dt3$air_temp_C <-as.numeric(levels(dt3$air_temp_C))[as.integer(dt3$air_temp_C) ]               
if (class(dt3$air_temp_C)=="character") dt3$air_temp_C <-as.numeric(dt3$air_temp_C)
if (class(dt3$data_coverage_air_temp_C)=="factor") dt3$data_coverage_air_temp_C <-as.numeric(levels(dt3$data_coverage_air_temp_C))[as.integer(dt3$data_coverage_air_temp_C) ]               
if (class(dt3$data_coverage_air_temp_C)=="character") dt3$data_coverage_air_temp_C <-as.numeric(dt3$data_coverage_air_temp_C)
if (class(dt3$wind_stress_pa)=="factor") dt3$wind_stress_pa <-as.numeric(levels(dt3$wind_stress_pa))[as.integer(dt3$wind_stress_pa) ]               
if (class(dt3$wind_stress_pa)=="character") dt3$wind_stress_pa <-as.numeric(dt3$wind_stress_pa)
if (class(dt3$data_coverage_wind_stress_pa)=="factor") dt3$data_coverage_wind_stress_pa <-as.numeric(levels(dt3$data_coverage_wind_stress_pa))[as.integer(dt3$data_coverage_wind_stress_pa) ]               
if (class(dt3$data_coverage_wind_stress_pa)=="character") dt3$data_coverage_wind_stress_pa <-as.numeric(dt3$data_coverage_wind_stress_pa)
if (class(dt3$wind_direction_degtrue)=="factor") dt3$wind_direction_degtrue <-as.numeric(levels(dt3$wind_direction_degtrue))[as.integer(dt3$wind_direction_degtrue) ]               
if (class(dt3$wind_direction_degtrue)=="character") dt3$wind_direction_degtrue <-as.numeric(dt3$wind_direction_degtrue)
if (class(dt3$data_coverage_wind_direction_degtrue)=="factor") dt3$data_coverage_wind_direction_degtrue <-as.numeric(levels(dt3$data_coverage_wind_direction_degtrue))[as.integer(dt3$data_coverage_wind_direction_degtrue) ]               
if (class(dt3$data_coverage_wind_direction_degtrue)=="character") dt3$data_coverage_wind_direction_degtrue <-as.numeric(dt3$data_coverage_wind_direction_degtrue)
if (class(dt3$wind_U_mps)=="factor") dt3$wind_U_mps <-as.numeric(levels(dt3$wind_U_mps))[as.integer(dt3$wind_U_mps) ]               
if (class(dt3$wind_U_mps)=="character") dt3$wind_U_mps <-as.numeric(dt3$wind_U_mps)
if (class(dt3$data_coverage_wind_U_mps)=="factor") dt3$data_coverage_wind_U_mps <-as.numeric(levels(dt3$data_coverage_wind_U_mps))[as.integer(dt3$data_coverage_wind_U_mps) ]               
if (class(dt3$data_coverage_wind_U_mps)=="character") dt3$data_coverage_wind_U_mps <-as.numeric(dt3$data_coverage_wind_U_mps)
if (class(dt3$wind_V_mps)=="factor") dt3$wind_V_mps <-as.numeric(levels(dt3$wind_V_mps))[as.integer(dt3$wind_V_mps) ]               
if (class(dt3$wind_V_mps)=="character") dt3$wind_V_mps <-as.numeric(dt3$wind_V_mps)
if (class(dt3$data_coverage_wind_V_mps)=="factor") dt3$data_coverage_wind_V_mps <-as.numeric(levels(dt3$data_coverage_wind_V_mps))[as.integer(dt3$data_coverage_wind_V_mps) ]               
if (class(dt3$data_coverage_wind_V_mps)=="character") dt3$data_coverage_wind_V_mps <-as.numeric(dt3$data_coverage_wind_V_mps)
if (class(dt3$mean_solar_rad_kwpm2)=="factor") dt3$mean_solar_rad_kwpm2 <-as.numeric(levels(dt3$mean_solar_rad_kwpm2))[as.integer(dt3$mean_solar_rad_kwpm2) ]               
if (class(dt3$mean_solar_rad_kwpm2)=="character") dt3$mean_solar_rad_kwpm2 <-as.numeric(dt3$mean_solar_rad_kwpm2)
if (class(dt3$data_coverage_mean_solar_rad_kwpm2)=="factor") dt3$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(levels(dt3$data_coverage_mean_solar_rad_kwpm2))[as.integer(dt3$data_coverage_mean_solar_rad_kwpm2) ]               
if (class(dt3$data_coverage_mean_solar_rad_kwpm2)=="character") dt3$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(dt3$data_coverage_mean_solar_rad_kwpm2)
if (class(dt3$atm_press_mbar)=="factor") dt3$atm_press_mbar <-as.numeric(levels(dt3$atm_press_mbar))[as.integer(dt3$atm_press_mbar) ]               
if (class(dt3$atm_press_mbar)=="character") dt3$atm_press_mbar <-as.numeric(dt3$atm_press_mbar)
if (class(dt3$data_coverage_atm_press_mbar)=="factor") dt3$data_coverage_atm_press_mbar <-as.numeric(levels(dt3$data_coverage_atm_press_mbar))[as.integer(dt3$data_coverage_atm_press_mbar) ]               
if (class(dt3$data_coverage_atm_press_mbar)=="character") dt3$data_coverage_atm_press_mbar <-as.numeric(dt3$data_coverage_atm_press_mbar)
if (class(dt3$cumulative_rainfall_mm)=="factor") dt3$cumulative_rainfall_mm <-as.numeric(levels(dt3$cumulative_rainfall_mm))[as.integer(dt3$cumulative_rainfall_mm) ]               
if (class(dt3$cumulative_rainfall_mm)=="character") dt3$cumulative_rainfall_mm <-as.numeric(dt3$cumulative_rainfall_mm)
if (class(dt3$data_coverage_cumulative_rainfall_mm)=="factor") dt3$data_coverage_cumulative_rainfall_mm <-as.numeric(levels(dt3$data_coverage_cumulative_rainfall_mm))[as.integer(dt3$data_coverage_cumulative_rainfall_mm) ]               
if (class(dt3$data_coverage_cumulative_rainfall_mm)=="character") dt3$data_coverage_cumulative_rainfall_mm <-as.numeric(dt3$data_coverage_cumulative_rainfall_mm)
                
# Convert Missing Values to NA for non-dates
                
dt3$air_temp_C <- ifelse((trimws(as.character(dt3$air_temp_C))==trimws("-999")),NA,dt3$air_temp_C)               
suppressWarnings(dt3$air_temp_C <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$air_temp_C))==as.character(as.numeric("-999"))),NA,dt3$air_temp_C))
dt3$wind_stress_pa <- ifelse((trimws(as.character(dt3$wind_stress_pa))==trimws("-999")),NA,dt3$wind_stress_pa)               
suppressWarnings(dt3$wind_stress_pa <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$wind_stress_pa))==as.character(as.numeric("-999"))),NA,dt3$wind_stress_pa))
dt3$wind_direction_degtrue <- ifelse((trimws(as.character(dt3$wind_direction_degtrue))==trimws("-999")),NA,dt3$wind_direction_degtrue)               
suppressWarnings(dt3$wind_direction_degtrue <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$wind_direction_degtrue))==as.character(as.numeric("-999"))),NA,dt3$wind_direction_degtrue))
dt3$wind_U_mps <- ifelse((trimws(as.character(dt3$wind_U_mps))==trimws("-999")),NA,dt3$wind_U_mps)               
suppressWarnings(dt3$wind_U_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$wind_U_mps))==as.character(as.numeric("-999"))),NA,dt3$wind_U_mps))
dt3$wind_V_mps <- ifelse((trimws(as.character(dt3$wind_V_mps))==trimws("-999")),NA,dt3$wind_V_mps)               
suppressWarnings(dt3$wind_V_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$wind_V_mps))==as.character(as.numeric("-999"))),NA,dt3$wind_V_mps))
dt3$mean_solar_rad_kwpm2 <- ifelse((trimws(as.character(dt3$mean_solar_rad_kwpm2))==trimws("-999")),NA,dt3$mean_solar_rad_kwpm2)               
suppressWarnings(dt3$mean_solar_rad_kwpm2 <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$mean_solar_rad_kwpm2))==as.character(as.numeric("-999"))),NA,dt3$mean_solar_rad_kwpm2))
dt3$atm_press_mbar <- ifelse((trimws(as.character(dt3$atm_press_mbar))==trimws("-999")),NA,dt3$atm_press_mbar)               
suppressWarnings(dt3$atm_press_mbar <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$atm_press_mbar))==as.character(as.numeric("-999"))),NA,dt3$atm_press_mbar))
dt3$cumulative_rainfall_mm <- ifelse((trimws(as.character(dt3$cumulative_rainfall_mm))==trimws("-999")),NA,dt3$cumulative_rainfall_mm)               
suppressWarnings(dt3$cumulative_rainfall_mm <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt3$cumulative_rainfall_mm))==as.character(as.numeric("-999"))),NA,dt3$cumulative_rainfall_mm))


# Here is the structure of the input data frame:
str(dt3)                            
attach(dt3)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(time_local)
summary(time_utc)
summary(air_temp_C)
summary(data_coverage_air_temp_C)
summary(wind_stress_pa)
summary(data_coverage_wind_stress_pa)
summary(wind_direction_degtrue)
summary(data_coverage_wind_direction_degtrue)
summary(wind_U_mps)
summary(data_coverage_wind_U_mps)
summary(wind_V_mps)
summary(data_coverage_wind_V_mps)
summary(mean_solar_rad_kwpm2)
summary(data_coverage_mean_solar_rad_kwpm2)
summary(atm_press_mbar)
summary(data_coverage_atm_press_mbar)
summary(cumulative_rainfall_mm)
summary(data_coverage_cumulative_rainfall_mm) 
                # Get more details on character variables
                
detach(dt3)               
         

inUrl4  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/7165fb592255fad3f693c1f5217267d9" 
infile4 <- tempfile()
try(download.file(inUrl4,infile4,method="curl"))
if (is.na(file.size(infile4))) download.file(inUrl4,infile4,method="auto")

                   
 dt4 <-read.csv(infile4,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "time_local",     
                    "time_utc",     
                    "air_temp_C",     
                    "data_coverage_air_temp_C",     
                    "wind_stress_pa",     
                    "data_coverage_wind_stress_pa",     
                    "wind_direction_degtrue",     
                    "data_coverage_wind_direction_degtrue",     
                    "wind_U_mps",     
                    "data_coverage_wind_U_mps",     
                    "wind_V_mps",     
                    "data_coverage_wind_V_mps",     
                    "mean_solar_rad_kwpm2",     
                    "data_coverage_mean_solar_rad_kwpm2",     
                    "atm_press_mbar",     
                    "data_coverage_atm_press_mbar",     
                    "cumulative_rainfall_mm",     
                    "data_coverage_cumulative_rainfall_mm"    ), check.names=TRUE)
               
unlink(infile4)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                                                   
# attempting to convert dt4$time_local dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp4time_local<-as.POSIXct(dt4$time_local,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt4[dt4$time_local != "",]) == length(tmp4time_local[!is.na(tmp4time_local)])){dt4$time_local <- tmp4time_local } else {print("Date conversion failed for dt4$time_local. Please inspect the data and do the date conversion yourself.")}                                                                    
                                                                   
# attempting to convert dt4$time_utc dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp4time_utc<-as.POSIXct(dt4$time_utc,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt4[dt4$time_utc != "",]) == length(tmp4time_utc[!is.na(tmp4time_utc)])){dt4$time_utc <- tmp4time_utc } else {print("Date conversion failed for dt4$time_utc. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt4$air_temp_C)=="factor") dt4$air_temp_C <-as.numeric(levels(dt4$air_temp_C))[as.integer(dt4$air_temp_C) ]               
if (class(dt4$air_temp_C)=="character") dt4$air_temp_C <-as.numeric(dt4$air_temp_C)
if (class(dt4$data_coverage_air_temp_C)=="factor") dt4$data_coverage_air_temp_C <-as.numeric(levels(dt4$data_coverage_air_temp_C))[as.integer(dt4$data_coverage_air_temp_C) ]               
if (class(dt4$data_coverage_air_temp_C)=="character") dt4$data_coverage_air_temp_C <-as.numeric(dt4$data_coverage_air_temp_C)
if (class(dt4$wind_stress_pa)=="factor") dt4$wind_stress_pa <-as.numeric(levels(dt4$wind_stress_pa))[as.integer(dt4$wind_stress_pa) ]               
if (class(dt4$wind_stress_pa)=="character") dt4$wind_stress_pa <-as.numeric(dt4$wind_stress_pa)
if (class(dt4$data_coverage_wind_stress_pa)=="factor") dt4$data_coverage_wind_stress_pa <-as.numeric(levels(dt4$data_coverage_wind_stress_pa))[as.integer(dt4$data_coverage_wind_stress_pa) ]               
if (class(dt4$data_coverage_wind_stress_pa)=="character") dt4$data_coverage_wind_stress_pa <-as.numeric(dt4$data_coverage_wind_stress_pa)
if (class(dt4$wind_direction_degtrue)=="factor") dt4$wind_direction_degtrue <-as.numeric(levels(dt4$wind_direction_degtrue))[as.integer(dt4$wind_direction_degtrue) ]               
if (class(dt4$wind_direction_degtrue)=="character") dt4$wind_direction_degtrue <-as.numeric(dt4$wind_direction_degtrue)
if (class(dt4$data_coverage_wind_direction_degtrue)=="factor") dt4$data_coverage_wind_direction_degtrue <-as.numeric(levels(dt4$data_coverage_wind_direction_degtrue))[as.integer(dt4$data_coverage_wind_direction_degtrue) ]               
if (class(dt4$data_coverage_wind_direction_degtrue)=="character") dt4$data_coverage_wind_direction_degtrue <-as.numeric(dt4$data_coverage_wind_direction_degtrue)
if (class(dt4$wind_U_mps)=="factor") dt4$wind_U_mps <-as.numeric(levels(dt4$wind_U_mps))[as.integer(dt4$wind_U_mps) ]               
if (class(dt4$wind_U_mps)=="character") dt4$wind_U_mps <-as.numeric(dt4$wind_U_mps)
if (class(dt4$data_coverage_wind_U_mps)=="factor") dt4$data_coverage_wind_U_mps <-as.numeric(levels(dt4$data_coverage_wind_U_mps))[as.integer(dt4$data_coverage_wind_U_mps) ]               
if (class(dt4$data_coverage_wind_U_mps)=="character") dt4$data_coverage_wind_U_mps <-as.numeric(dt4$data_coverage_wind_U_mps)
if (class(dt4$wind_V_mps)=="factor") dt4$wind_V_mps <-as.numeric(levels(dt4$wind_V_mps))[as.integer(dt4$wind_V_mps) ]               
if (class(dt4$wind_V_mps)=="character") dt4$wind_V_mps <-as.numeric(dt4$wind_V_mps)
if (class(dt4$data_coverage_wind_V_mps)=="factor") dt4$data_coverage_wind_V_mps <-as.numeric(levels(dt4$data_coverage_wind_V_mps))[as.integer(dt4$data_coverage_wind_V_mps) ]               
if (class(dt4$data_coverage_wind_V_mps)=="character") dt4$data_coverage_wind_V_mps <-as.numeric(dt4$data_coverage_wind_V_mps)
if (class(dt4$mean_solar_rad_kwpm2)=="factor") dt4$mean_solar_rad_kwpm2 <-as.numeric(levels(dt4$mean_solar_rad_kwpm2))[as.integer(dt4$mean_solar_rad_kwpm2) ]               
if (class(dt4$mean_solar_rad_kwpm2)=="character") dt4$mean_solar_rad_kwpm2 <-as.numeric(dt4$mean_solar_rad_kwpm2)
if (class(dt4$data_coverage_mean_solar_rad_kwpm2)=="factor") dt4$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(levels(dt4$data_coverage_mean_solar_rad_kwpm2))[as.integer(dt4$data_coverage_mean_solar_rad_kwpm2) ]               
if (class(dt4$data_coverage_mean_solar_rad_kwpm2)=="character") dt4$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(dt4$data_coverage_mean_solar_rad_kwpm2)
if (class(dt4$atm_press_mbar)=="factor") dt4$atm_press_mbar <-as.numeric(levels(dt4$atm_press_mbar))[as.integer(dt4$atm_press_mbar) ]               
if (class(dt4$atm_press_mbar)=="character") dt4$atm_press_mbar <-as.numeric(dt4$atm_press_mbar)
if (class(dt4$data_coverage_atm_press_mbar)=="factor") dt4$data_coverage_atm_press_mbar <-as.numeric(levels(dt4$data_coverage_atm_press_mbar))[as.integer(dt4$data_coverage_atm_press_mbar) ]               
if (class(dt4$data_coverage_atm_press_mbar)=="character") dt4$data_coverage_atm_press_mbar <-as.numeric(dt4$data_coverage_atm_press_mbar)
if (class(dt4$cumulative_rainfall_mm)=="factor") dt4$cumulative_rainfall_mm <-as.numeric(levels(dt4$cumulative_rainfall_mm))[as.integer(dt4$cumulative_rainfall_mm) ]               
if (class(dt4$cumulative_rainfall_mm)=="character") dt4$cumulative_rainfall_mm <-as.numeric(dt4$cumulative_rainfall_mm)
if (class(dt4$data_coverage_cumulative_rainfall_mm)=="factor") dt4$data_coverage_cumulative_rainfall_mm <-as.numeric(levels(dt4$data_coverage_cumulative_rainfall_mm))[as.integer(dt4$data_coverage_cumulative_rainfall_mm) ]               
if (class(dt4$data_coverage_cumulative_rainfall_mm)=="character") dt4$data_coverage_cumulative_rainfall_mm <-as.numeric(dt4$data_coverage_cumulative_rainfall_mm)
                
# Convert Missing Values to NA for non-dates
                
dt4$air_temp_C <- ifelse((trimws(as.character(dt4$air_temp_C))==trimws("-999")),NA,dt4$air_temp_C)               
suppressWarnings(dt4$air_temp_C <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$air_temp_C))==as.character(as.numeric("-999"))),NA,dt4$air_temp_C))
dt4$wind_stress_pa <- ifelse((trimws(as.character(dt4$wind_stress_pa))==trimws("-999")),NA,dt4$wind_stress_pa)               
suppressWarnings(dt4$wind_stress_pa <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$wind_stress_pa))==as.character(as.numeric("-999"))),NA,dt4$wind_stress_pa))
dt4$wind_direction_degtrue <- ifelse((trimws(as.character(dt4$wind_direction_degtrue))==trimws("-999")),NA,dt4$wind_direction_degtrue)               
suppressWarnings(dt4$wind_direction_degtrue <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$wind_direction_degtrue))==as.character(as.numeric("-999"))),NA,dt4$wind_direction_degtrue))
dt4$wind_U_mps <- ifelse((trimws(as.character(dt4$wind_U_mps))==trimws("-999")),NA,dt4$wind_U_mps)               
suppressWarnings(dt4$wind_U_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$wind_U_mps))==as.character(as.numeric("-999"))),NA,dt4$wind_U_mps))
dt4$wind_V_mps <- ifelse((trimws(as.character(dt4$wind_V_mps))==trimws("-999")),NA,dt4$wind_V_mps)               
suppressWarnings(dt4$wind_V_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$wind_V_mps))==as.character(as.numeric("-999"))),NA,dt4$wind_V_mps))
dt4$mean_solar_rad_kwpm2 <- ifelse((trimws(as.character(dt4$mean_solar_rad_kwpm2))==trimws("-999")),NA,dt4$mean_solar_rad_kwpm2)               
suppressWarnings(dt4$mean_solar_rad_kwpm2 <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$mean_solar_rad_kwpm2))==as.character(as.numeric("-999"))),NA,dt4$mean_solar_rad_kwpm2))
dt4$atm_press_mbar <- ifelse((trimws(as.character(dt4$atm_press_mbar))==trimws("-999")),NA,dt4$atm_press_mbar)               
suppressWarnings(dt4$atm_press_mbar <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$atm_press_mbar))==as.character(as.numeric("-999"))),NA,dt4$atm_press_mbar))
dt4$cumulative_rainfall_mm <- ifelse((trimws(as.character(dt4$cumulative_rainfall_mm))==trimws("-999")),NA,dt4$cumulative_rainfall_mm)               
suppressWarnings(dt4$cumulative_rainfall_mm <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt4$cumulative_rainfall_mm))==as.character(as.numeric("-999"))),NA,dt4$cumulative_rainfall_mm))


# Here is the structure of the input data frame:
str(dt4)                            
attach(dt4)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(time_local)
summary(time_utc)
summary(air_temp_C)
summary(data_coverage_air_temp_C)
summary(wind_stress_pa)
summary(data_coverage_wind_stress_pa)
summary(wind_direction_degtrue)
summary(data_coverage_wind_direction_degtrue)
summary(wind_U_mps)
summary(data_coverage_wind_U_mps)
summary(wind_V_mps)
summary(data_coverage_wind_V_mps)
summary(mean_solar_rad_kwpm2)
summary(data_coverage_mean_solar_rad_kwpm2)
summary(atm_press_mbar)
summary(data_coverage_atm_press_mbar)
summary(cumulative_rainfall_mm)
summary(data_coverage_cumulative_rainfall_mm) 
                # Get more details on character variables
                
detach(dt4)               
         

inUrl5  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/a2479a853f484b0a53478176a03ba29d" 
infile5 <- tempfile()
try(download.file(inUrl5,infile5,method="curl"))
if (is.na(file.size(infile5))) download.file(inUrl5,infile5,method="auto")

                   
 dt5 <-read.csv(infile5,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "time_local",     
                    "time_utc",     
                    "air_temp_C",     
                    "data_coverage_air_temp_C",     
                    "wind_stress_pa",     
                    "data_coverage_wind_stress_pa",     
                    "wind_direction_degtrue",     
                    "data_coverage_wind_direction_degtrue",     
                    "wind_U_mps",     
                    "data_coverage_wind_U_mps",     
                    "wind_V_mps",     
                    "data_coverage_wind_V_mps",     
                    "mean_solar_rad_kwpm2",     
                    "data_coverage_mean_solar_rad_kwpm2",     
                    "atm_press_mbar",     
                    "data_coverage_atm_press_mbar",     
                    "cumulative_rainfall_mm",     
                    "data_coverage_cumulative_rainfall_mm"    ), check.names=TRUE)
               
unlink(infile5)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                                                   
# attempting to convert dt5$time_local dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp5time_local<-as.POSIXct(dt5$time_local,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt5[dt5$time_local != "",]) == length(tmp5time_local[!is.na(tmp5time_local)])){dt5$time_local <- tmp5time_local } else {print("Date conversion failed for dt5$time_local. Please inspect the data and do the date conversion yourself.")}                                                                    
                                                                   
# attempting to convert dt5$time_utc dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp5time_utc<-as.POSIXct(dt5$time_utc,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt5[dt5$time_utc != "",]) == length(tmp5time_utc[!is.na(tmp5time_utc)])){dt5$time_utc <- tmp5time_utc } else {print("Date conversion failed for dt5$time_utc. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt5$air_temp_C)=="factor") dt5$air_temp_C <-as.numeric(levels(dt5$air_temp_C))[as.integer(dt5$air_temp_C) ]               
if (class(dt5$air_temp_C)=="character") dt5$air_temp_C <-as.numeric(dt5$air_temp_C)
if (class(dt5$data_coverage_air_temp_C)=="factor") dt5$data_coverage_air_temp_C <-as.numeric(levels(dt5$data_coverage_air_temp_C))[as.integer(dt5$data_coverage_air_temp_C) ]               
if (class(dt5$data_coverage_air_temp_C)=="character") dt5$data_coverage_air_temp_C <-as.numeric(dt5$data_coverage_air_temp_C)
if (class(dt5$wind_stress_pa)=="factor") dt5$wind_stress_pa <-as.numeric(levels(dt5$wind_stress_pa))[as.integer(dt5$wind_stress_pa) ]               
if (class(dt5$wind_stress_pa)=="character") dt5$wind_stress_pa <-as.numeric(dt5$wind_stress_pa)
if (class(dt5$data_coverage_wind_stress_pa)=="factor") dt5$data_coverage_wind_stress_pa <-as.numeric(levels(dt5$data_coverage_wind_stress_pa))[as.integer(dt5$data_coverage_wind_stress_pa) ]               
if (class(dt5$data_coverage_wind_stress_pa)=="character") dt5$data_coverage_wind_stress_pa <-as.numeric(dt5$data_coverage_wind_stress_pa)
if (class(dt5$wind_direction_degtrue)=="factor") dt5$wind_direction_degtrue <-as.numeric(levels(dt5$wind_direction_degtrue))[as.integer(dt5$wind_direction_degtrue) ]               
if (class(dt5$wind_direction_degtrue)=="character") dt5$wind_direction_degtrue <-as.numeric(dt5$wind_direction_degtrue)
if (class(dt5$data_coverage_wind_direction_degtrue)=="factor") dt5$data_coverage_wind_direction_degtrue <-as.numeric(levels(dt5$data_coverage_wind_direction_degtrue))[as.integer(dt5$data_coverage_wind_direction_degtrue) ]               
if (class(dt5$data_coverage_wind_direction_degtrue)=="character") dt5$data_coverage_wind_direction_degtrue <-as.numeric(dt5$data_coverage_wind_direction_degtrue)
if (class(dt5$wind_U_mps)=="factor") dt5$wind_U_mps <-as.numeric(levels(dt5$wind_U_mps))[as.integer(dt5$wind_U_mps) ]               
if (class(dt5$wind_U_mps)=="character") dt5$wind_U_mps <-as.numeric(dt5$wind_U_mps)
if (class(dt5$data_coverage_wind_U_mps)=="factor") dt5$data_coverage_wind_U_mps <-as.numeric(levels(dt5$data_coverage_wind_U_mps))[as.integer(dt5$data_coverage_wind_U_mps) ]               
if (class(dt5$data_coverage_wind_U_mps)=="character") dt5$data_coverage_wind_U_mps <-as.numeric(dt5$data_coverage_wind_U_mps)
if (class(dt5$wind_V_mps)=="factor") dt5$wind_V_mps <-as.numeric(levels(dt5$wind_V_mps))[as.integer(dt5$wind_V_mps) ]               
if (class(dt5$wind_V_mps)=="character") dt5$wind_V_mps <-as.numeric(dt5$wind_V_mps)
if (class(dt5$data_coverage_wind_V_mps)=="factor") dt5$data_coverage_wind_V_mps <-as.numeric(levels(dt5$data_coverage_wind_V_mps))[as.integer(dt5$data_coverage_wind_V_mps) ]               
if (class(dt5$data_coverage_wind_V_mps)=="character") dt5$data_coverage_wind_V_mps <-as.numeric(dt5$data_coverage_wind_V_mps)
if (class(dt5$mean_solar_rad_kwpm2)=="factor") dt5$mean_solar_rad_kwpm2 <-as.numeric(levels(dt5$mean_solar_rad_kwpm2))[as.integer(dt5$mean_solar_rad_kwpm2) ]               
if (class(dt5$mean_solar_rad_kwpm2)=="character") dt5$mean_solar_rad_kwpm2 <-as.numeric(dt5$mean_solar_rad_kwpm2)
if (class(dt5$data_coverage_mean_solar_rad_kwpm2)=="factor") dt5$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(levels(dt5$data_coverage_mean_solar_rad_kwpm2))[as.integer(dt5$data_coverage_mean_solar_rad_kwpm2) ]               
if (class(dt5$data_coverage_mean_solar_rad_kwpm2)=="character") dt5$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(dt5$data_coverage_mean_solar_rad_kwpm2)
if (class(dt5$atm_press_mbar)=="factor") dt5$atm_press_mbar <-as.numeric(levels(dt5$atm_press_mbar))[as.integer(dt5$atm_press_mbar) ]               
if (class(dt5$atm_press_mbar)=="character") dt5$atm_press_mbar <-as.numeric(dt5$atm_press_mbar)
if (class(dt5$data_coverage_atm_press_mbar)=="factor") dt5$data_coverage_atm_press_mbar <-as.numeric(levels(dt5$data_coverage_atm_press_mbar))[as.integer(dt5$data_coverage_atm_press_mbar) ]               
if (class(dt5$data_coverage_atm_press_mbar)=="character") dt5$data_coverage_atm_press_mbar <-as.numeric(dt5$data_coverage_atm_press_mbar)
if (class(dt5$cumulative_rainfall_mm)=="factor") dt5$cumulative_rainfall_mm <-as.numeric(levels(dt5$cumulative_rainfall_mm))[as.integer(dt5$cumulative_rainfall_mm) ]               
if (class(dt5$cumulative_rainfall_mm)=="character") dt5$cumulative_rainfall_mm <-as.numeric(dt5$cumulative_rainfall_mm)
if (class(dt5$data_coverage_cumulative_rainfall_mm)=="factor") dt5$data_coverage_cumulative_rainfall_mm <-as.numeric(levels(dt5$data_coverage_cumulative_rainfall_mm))[as.integer(dt5$data_coverage_cumulative_rainfall_mm) ]               
if (class(dt5$data_coverage_cumulative_rainfall_mm)=="character") dt5$data_coverage_cumulative_rainfall_mm <-as.numeric(dt5$data_coverage_cumulative_rainfall_mm)
                
# Convert Missing Values to NA for non-dates
                
dt5$air_temp_C <- ifelse((trimws(as.character(dt5$air_temp_C))==trimws("-999")),NA,dt5$air_temp_C)               
suppressWarnings(dt5$air_temp_C <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$air_temp_C))==as.character(as.numeric("-999"))),NA,dt5$air_temp_C))
dt5$wind_stress_pa <- ifelse((trimws(as.character(dt5$wind_stress_pa))==trimws("-999")),NA,dt5$wind_stress_pa)               
suppressWarnings(dt5$wind_stress_pa <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$wind_stress_pa))==as.character(as.numeric("-999"))),NA,dt5$wind_stress_pa))
dt5$wind_direction_degtrue <- ifelse((trimws(as.character(dt5$wind_direction_degtrue))==trimws("-999")),NA,dt5$wind_direction_degtrue)               
suppressWarnings(dt5$wind_direction_degtrue <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$wind_direction_degtrue))==as.character(as.numeric("-999"))),NA,dt5$wind_direction_degtrue))
dt5$wind_U_mps <- ifelse((trimws(as.character(dt5$wind_U_mps))==trimws("-999")),NA,dt5$wind_U_mps)               
suppressWarnings(dt5$wind_U_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$wind_U_mps))==as.character(as.numeric("-999"))),NA,dt5$wind_U_mps))
dt5$wind_V_mps <- ifelse((trimws(as.character(dt5$wind_V_mps))==trimws("-999")),NA,dt5$wind_V_mps)               
suppressWarnings(dt5$wind_V_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$wind_V_mps))==as.character(as.numeric("-999"))),NA,dt5$wind_V_mps))
dt5$mean_solar_rad_kwpm2 <- ifelse((trimws(as.character(dt5$mean_solar_rad_kwpm2))==trimws("-999")),NA,dt5$mean_solar_rad_kwpm2)               
suppressWarnings(dt5$mean_solar_rad_kwpm2 <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$mean_solar_rad_kwpm2))==as.character(as.numeric("-999"))),NA,dt5$mean_solar_rad_kwpm2))
dt5$atm_press_mbar <- ifelse((trimws(as.character(dt5$atm_press_mbar))==trimws("-999")),NA,dt5$atm_press_mbar)               
suppressWarnings(dt5$atm_press_mbar <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$atm_press_mbar))==as.character(as.numeric("-999"))),NA,dt5$atm_press_mbar))
dt5$cumulative_rainfall_mm <- ifelse((trimws(as.character(dt5$cumulative_rainfall_mm))==trimws("-999")),NA,dt5$cumulative_rainfall_mm)               
suppressWarnings(dt5$cumulative_rainfall_mm <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt5$cumulative_rainfall_mm))==as.character(as.numeric("-999"))),NA,dt5$cumulative_rainfall_mm))


# Here is the structure of the input data frame:
str(dt5)                            
attach(dt5)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(time_local)
summary(time_utc)
summary(air_temp_C)
summary(data_coverage_air_temp_C)
summary(wind_stress_pa)
summary(data_coverage_wind_stress_pa)
summary(wind_direction_degtrue)
summary(data_coverage_wind_direction_degtrue)
summary(wind_U_mps)
summary(data_coverage_wind_U_mps)
summary(wind_V_mps)
summary(data_coverage_wind_V_mps)
summary(mean_solar_rad_kwpm2)
summary(data_coverage_mean_solar_rad_kwpm2)
summary(atm_press_mbar)
summary(data_coverage_atm_press_mbar)
summary(cumulative_rainfall_mm)
summary(data_coverage_cumulative_rainfall_mm) 
                # Get more details on character variables
                
detach(dt5)               
         

inUrl6  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-mcr/9/50/2738f2c96013f0ffdb00eea519233b56" 
infile6 <- tempfile()
try(download.file(inUrl6,infile6,method="curl"))
if (is.na(file.size(infile6))) download.file(inUrl6,infile6,method="auto")

                   
 dt6 <-read.csv(infile6,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "time_local",     
                    "time_utc",     
                    "air_temp_C",     
                    "data_coverage_air_temp_C",     
                    "wind_stress_pa",     
                    "data_coverage_wind_stress_pa",     
                    "wind_direction_degtrue",     
                    "data_coverage_wind_direction_degtrue",     
                    "wind_U_mps",     
                    "data_coverage_wind_U_mps",     
                    "wind_V_mps",     
                    "data_coverage_wind_V_mps",     
                    "mean_solar_rad_kwpm2",     
                    "data_coverage_mean_solar_rad_kwpm2",     
                    "atm_press_mbar",     
                    "data_coverage_atm_press_mbar",     
                    "cumulative_rainfall_mm",     
                    "data_coverage_cumulative_rainfall_mm"    ), check.names=TRUE)
               
unlink(infile6)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                                                   
# attempting to convert dt6$time_local dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp6time_local<-as.POSIXct(dt6$time_local,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt6[dt6$time_local != "",]) == length(tmp6time_local[!is.na(tmp6time_local)])){dt6$time_local <- tmp6time_local } else {print("Date conversion failed for dt6$time_local. Please inspect the data and do the date conversion yourself.")}                                                                    
                                                                   
# attempting to convert dt6$time_utc dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d %H:%M:%S" 
tmp6time_utc<-as.POSIXct(dt6$time_utc,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(nrow(dt6[dt6$time_utc != "",]) == length(tmp6time_utc[!is.na(tmp6time_utc)])){dt6$time_utc <- tmp6time_utc } else {print("Date conversion failed for dt6$time_utc. Please inspect the data and do the date conversion yourself.")}                                                                    
                                
if (class(dt6$air_temp_C)=="factor") dt6$air_temp_C <-as.numeric(levels(dt6$air_temp_C))[as.integer(dt6$air_temp_C) ]               
if (class(dt6$air_temp_C)=="character") dt6$air_temp_C <-as.numeric(dt6$air_temp_C)
if (class(dt6$data_coverage_air_temp_C)=="factor") dt6$data_coverage_air_temp_C <-as.numeric(levels(dt6$data_coverage_air_temp_C))[as.integer(dt6$data_coverage_air_temp_C) ]               
if (class(dt6$data_coverage_air_temp_C)=="character") dt6$data_coverage_air_temp_C <-as.numeric(dt6$data_coverage_air_temp_C)
if (class(dt6$wind_stress_pa)=="factor") dt6$wind_stress_pa <-as.numeric(levels(dt6$wind_stress_pa))[as.integer(dt6$wind_stress_pa) ]               
if (class(dt6$wind_stress_pa)=="character") dt6$wind_stress_pa <-as.numeric(dt6$wind_stress_pa)
if (class(dt6$data_coverage_wind_stress_pa)=="factor") dt6$data_coverage_wind_stress_pa <-as.numeric(levels(dt6$data_coverage_wind_stress_pa))[as.integer(dt6$data_coverage_wind_stress_pa) ]               
if (class(dt6$data_coverage_wind_stress_pa)=="character") dt6$data_coverage_wind_stress_pa <-as.numeric(dt6$data_coverage_wind_stress_pa)
if (class(dt6$wind_direction_degtrue)=="factor") dt6$wind_direction_degtrue <-as.numeric(levels(dt6$wind_direction_degtrue))[as.integer(dt6$wind_direction_degtrue) ]               
if (class(dt6$wind_direction_degtrue)=="character") dt6$wind_direction_degtrue <-as.numeric(dt6$wind_direction_degtrue)
if (class(dt6$data_coverage_wind_direction_degtrue)=="factor") dt6$data_coverage_wind_direction_degtrue <-as.numeric(levels(dt6$data_coverage_wind_direction_degtrue))[as.integer(dt6$data_coverage_wind_direction_degtrue) ]               
if (class(dt6$data_coverage_wind_direction_degtrue)=="character") dt6$data_coverage_wind_direction_degtrue <-as.numeric(dt6$data_coverage_wind_direction_degtrue)
if (class(dt6$wind_U_mps)=="factor") dt6$wind_U_mps <-as.numeric(levels(dt6$wind_U_mps))[as.integer(dt6$wind_U_mps) ]               
if (class(dt6$wind_U_mps)=="character") dt6$wind_U_mps <-as.numeric(dt6$wind_U_mps)
if (class(dt6$data_coverage_wind_U_mps)=="factor") dt6$data_coverage_wind_U_mps <-as.numeric(levels(dt6$data_coverage_wind_U_mps))[as.integer(dt6$data_coverage_wind_U_mps) ]               
if (class(dt6$data_coverage_wind_U_mps)=="character") dt6$data_coverage_wind_U_mps <-as.numeric(dt6$data_coverage_wind_U_mps)
if (class(dt6$wind_V_mps)=="factor") dt6$wind_V_mps <-as.numeric(levels(dt6$wind_V_mps))[as.integer(dt6$wind_V_mps) ]               
if (class(dt6$wind_V_mps)=="character") dt6$wind_V_mps <-as.numeric(dt6$wind_V_mps)
if (class(dt6$data_coverage_wind_V_mps)=="factor") dt6$data_coverage_wind_V_mps <-as.numeric(levels(dt6$data_coverage_wind_V_mps))[as.integer(dt6$data_coverage_wind_V_mps) ]               
if (class(dt6$data_coverage_wind_V_mps)=="character") dt6$data_coverage_wind_V_mps <-as.numeric(dt6$data_coverage_wind_V_mps)
if (class(dt6$mean_solar_rad_kwpm2)=="factor") dt6$mean_solar_rad_kwpm2 <-as.numeric(levels(dt6$mean_solar_rad_kwpm2))[as.integer(dt6$mean_solar_rad_kwpm2) ]               
if (class(dt6$mean_solar_rad_kwpm2)=="character") dt6$mean_solar_rad_kwpm2 <-as.numeric(dt6$mean_solar_rad_kwpm2)
if (class(dt6$data_coverage_mean_solar_rad_kwpm2)=="factor") dt6$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(levels(dt6$data_coverage_mean_solar_rad_kwpm2))[as.integer(dt6$data_coverage_mean_solar_rad_kwpm2) ]               
if (class(dt6$data_coverage_mean_solar_rad_kwpm2)=="character") dt6$data_coverage_mean_solar_rad_kwpm2 <-as.numeric(dt6$data_coverage_mean_solar_rad_kwpm2)
if (class(dt6$atm_press_mbar)=="factor") dt6$atm_press_mbar <-as.numeric(levels(dt6$atm_press_mbar))[as.integer(dt6$atm_press_mbar) ]               
if (class(dt6$atm_press_mbar)=="character") dt6$atm_press_mbar <-as.numeric(dt6$atm_press_mbar)
if (class(dt6$data_coverage_atm_press_mbar)=="factor") dt6$data_coverage_atm_press_mbar <-as.numeric(levels(dt6$data_coverage_atm_press_mbar))[as.integer(dt6$data_coverage_atm_press_mbar) ]               
if (class(dt6$data_coverage_atm_press_mbar)=="character") dt6$data_coverage_atm_press_mbar <-as.numeric(dt6$data_coverage_atm_press_mbar)
if (class(dt6$cumulative_rainfall_mm)=="factor") dt6$cumulative_rainfall_mm <-as.numeric(levels(dt6$cumulative_rainfall_mm))[as.integer(dt6$cumulative_rainfall_mm) ]               
if (class(dt6$cumulative_rainfall_mm)=="character") dt6$cumulative_rainfall_mm <-as.numeric(dt6$cumulative_rainfall_mm)
if (class(dt6$data_coverage_cumulative_rainfall_mm)=="factor") dt6$data_coverage_cumulative_rainfall_mm <-as.numeric(levels(dt6$data_coverage_cumulative_rainfall_mm))[as.integer(dt6$data_coverage_cumulative_rainfall_mm) ]               
if (class(dt6$data_coverage_cumulative_rainfall_mm)=="character") dt6$data_coverage_cumulative_rainfall_mm <-as.numeric(dt6$data_coverage_cumulative_rainfall_mm)
                
# Convert Missing Values to NA for non-dates
                
dt6$air_temp_C <- ifelse((trimws(as.character(dt6$air_temp_C))==trimws("-999")),NA,dt6$air_temp_C)               
suppressWarnings(dt6$air_temp_C <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$air_temp_C))==as.character(as.numeric("-999"))),NA,dt6$air_temp_C))
dt6$wind_stress_pa <- ifelse((trimws(as.character(dt6$wind_stress_pa))==trimws("-999")),NA,dt6$wind_stress_pa)               
suppressWarnings(dt6$wind_stress_pa <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$wind_stress_pa))==as.character(as.numeric("-999"))),NA,dt6$wind_stress_pa))
dt6$wind_direction_degtrue <- ifelse((trimws(as.character(dt6$wind_direction_degtrue))==trimws("-999")),NA,dt6$wind_direction_degtrue)               
suppressWarnings(dt6$wind_direction_degtrue <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$wind_direction_degtrue))==as.character(as.numeric("-999"))),NA,dt6$wind_direction_degtrue))
dt6$wind_U_mps <- ifelse((trimws(as.character(dt6$wind_U_mps))==trimws("-999")),NA,dt6$wind_U_mps)               
suppressWarnings(dt6$wind_U_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$wind_U_mps))==as.character(as.numeric("-999"))),NA,dt6$wind_U_mps))
dt6$wind_V_mps <- ifelse((trimws(as.character(dt6$wind_V_mps))==trimws("-999")),NA,dt6$wind_V_mps)               
suppressWarnings(dt6$wind_V_mps <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$wind_V_mps))==as.character(as.numeric("-999"))),NA,dt6$wind_V_mps))
dt6$mean_solar_rad_kwpm2 <- ifelse((trimws(as.character(dt6$mean_solar_rad_kwpm2))==trimws("-999")),NA,dt6$mean_solar_rad_kwpm2)               
suppressWarnings(dt6$mean_solar_rad_kwpm2 <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$mean_solar_rad_kwpm2))==as.character(as.numeric("-999"))),NA,dt6$mean_solar_rad_kwpm2))
dt6$atm_press_mbar <- ifelse((trimws(as.character(dt6$atm_press_mbar))==trimws("-999")),NA,dt6$atm_press_mbar)               
suppressWarnings(dt6$atm_press_mbar <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$atm_press_mbar))==as.character(as.numeric("-999"))),NA,dt6$atm_press_mbar))
dt6$cumulative_rainfall_mm <- ifelse((trimws(as.character(dt6$cumulative_rainfall_mm))==trimws("-999")),NA,dt6$cumulative_rainfall_mm)               
suppressWarnings(dt6$cumulative_rainfall_mm <- ifelse(!is.na(as.numeric("-999")) & (trimws(as.character(dt6$cumulative_rainfall_mm))==as.character(as.numeric("-999"))),NA,dt6$cumulative_rainfall_mm))


# Here is the structure of the input data frame:
str(dt6)                            
attach(dt6)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(time_local)
summary(time_utc)
summary(air_temp_C)
summary(data_coverage_air_temp_C)
summary(wind_stress_pa)
summary(data_coverage_wind_stress_pa)
summary(wind_direction_degtrue)
summary(data_coverage_wind_direction_degtrue)
summary(wind_U_mps)
summary(data_coverage_wind_U_mps)
summary(wind_V_mps)
summary(data_coverage_wind_V_mps)
summary(mean_solar_rad_kwpm2)
summary(data_coverage_mean_solar_rad_kwpm2)
summary(atm_press_mbar)
summary(data_coverage_atm_press_mbar)
summary(cumulative_rainfall_mm)
summary(data_coverage_cumulative_rainfall_mm) 
                # Get more details on character variables
                
detach(dt6)               
```

dt4 contains daily summarized data, which we will use for this application. 
```{r}
str(dt4)
```

The metrics we will use are: 

- mean_solar_rad_kwpm2 (sunlight) - mean and standard deviation 
- cumulative_rainfall_mm (rainfall) - mean and standard deviation 

We will use mean and standard deviation of temperature from 

```{r}
data<-dt4%>%
  select(time_local, mean_solar_rad_kwpm2, cumulative_rainfall_mm)%>%
  filter(time_local > as.POSIXct("2019-12-01"))%>%
  filter(time_local < as.POSIXct("2021-01-01")) #select data from Dec 2019 to Dec 2020
```

Generate timepoint average light and rainfall as the 4 weeks leading up to the end of the sampling period. 

Timepoint 1 sampling = 28 Dec 2019 - 14 Jan 2020 [use data from 15 Dec-15 Jan]
Timepoint 2 sampling = 27 Feb - 14 March [use data from 15 Feb-14 Mar] - no data available for this range, using April 1-30 for light and rain fall, using correct range for temp 
Timepoint 3 sampling = 8 Sept - 28 Sept [use data from 29 Aug-28 Sept]
Timepoint 4 sampling = 30 Oct - 21 Nov [use data from 22 Oct-21 Nov]

Calculate mean and sd of each metric for the specified date ranges.  

```{r}
# Define time ranges
tp_ranges <- list(
  timepoint1 = as.POSIXct(c("2019-12-05", "2020-01-15")),
  timepoint2 = as.POSIXct(c("2020-02-15", "2020-04-30")),
  timepoint3 = as.POSIXct(c("2020-08-29", "2020-09-28")),
  timepoint4 = as.POSIXct(c("2020-10-22", "2020-11-21"))
)

# Initialize a list to store results
results <- list()

# Calculate mean and standard deviation for each time range
for (tp in names(tp_ranges)) {
  start_date <- tp_ranges[[tp]][1]
  end_date <- tp_ranges[[tp]][2]
  
  # Filter data for the current time range
  tp_data <- data %>%
    filter(time_local >= start_date & time_local <= end_date) %>%
    select(-time_local) # Remove the time_local column for calculations
  
  # Compute mean and standard deviation for all columns
  stats <- tp_data %>%
    summarise(across(everything(), 
                     list(mean = ~mean(.x, na.rm = TRUE), 
                          sd = ~sd(.x, na.rm = TRUE))))
  
  # Store results with time period label
  results[[tp]] <- stats
}

# Combine results into a single dataframe for easy viewing
final_results <- bind_rows(results, .id = "Time_Period")

# View results
print(final_results)

```

Calculate monthly average temperature from our data set for the same date ranges as above and add to the data frame.  

```{r}
temp<-read_csv("Output/environmental_timeseries.csv")%>%
  select(Date.Time, FullTemp)%>%
  filter(!is.na(FullTemp))
```

Summarize by day and then generate time period means. 

```{r}
# Summarize mean Temp for each day
daily_mean_temp <- temp %>%
  mutate(Date = as.Date(Date.Time)) %>%  # Extract the date from the Date.Time column
  group_by(Date) %>%                     # Group by day
  summarise(mean_Temp = mean(FullTemp, na.rm = TRUE))%>%  # Calculate mean Temp for each day
  mutate(Date = as.POSIXct(Date, tz="UTC")) 

str(daily_mean_temp)
daily_mean_temp$Date
```

Run a loop to calculate temperatures for each time period. 

```{r}
tp_ranges <- list(
  timepoint1 = as.POSIXct(c("2019-12-05", "2020-01-15")),
  timepoint2 = as.POSIXct(c("2020-02-15", "2020-03-14")),
  timepoint3 = as.POSIXct(c("2020-08-29", "2020-09-28")),
  timepoint4 = as.POSIXct(c("2020-10-22", "2020-11-21"))
)

# Initialize a list to store results
results2 <- list()

# Calculate mean and standard deviation for each time range
for (tp in names(tp_ranges)) {
  start_date <- tp_ranges[[tp]][1]
  end_date <- tp_ranges[[tp]][2]
  
  # Filter data for the current time range
  tp_data2 <- daily_mean_temp %>%
    filter(Date >= start_date & Date <= end_date) %>%
    select(-Date) # Remove the Date column for calculations
  
  # Compute mean and standard deviation for all columns
  stats2 <- tp_data2 %>%
    summarise(across(everything(), 
                     list(mean = ~mean(.x, na.rm = TRUE), 
                          sd = ~sd(.x, na.rm = TRUE))))
  
  # Store results with time period label
  results2[[tp]] <- stats2
}

# Combine results into a single dataframe for easy viewing
final_results2 <- bind_rows(results2, .id = "Time_Period")

# View results
print(final_results2)

```

Merge results. 

```{r}
results<-merge(final_results, final_results2)
```

Scale the data and output for RDA analyses. 
```{r}
str(results)

scaled_results<-results%>%select(!Time_Period)

scaled_results<-scale(scaled_results)%>%as.data.frame()%>%mutate(timepoint=c("timepoint1", "timepoint2", "timepont3", "timepoint4"))

#output
write_csv(scaled_results, "Output/environment_characteristics_RDA.csv")
```

# Display plot of characteristics 

```{r}
plot1<-results%>%
  rename(solar_mean=mean_solar_rad_kwpm2_mean, solar_sd=mean_solar_rad_kwpm2_sd, rain_mean=cumulative_rainfall_mm_mean, rain_sd=cumulative_rainfall_mm_sd, temp_mean=mean_Temp_mean, temp_sd=mean_Temp_sd)%>%
  
  ggplot(aes(x=Time_Period, y=solar_mean))+
    geom_point(size=4, shape=19)+
    geom_errorbar(aes(ymin=solar_mean-solar_sd, ymax=solar_mean+solar_sd), width=0.01)+
    ylab(expression(bold(paste("Solar Radiation (kWh m" ^-2, ")"))))+
    ylim(0,0.25)+
    xlab("Time Point")+
    ggtitle("C) Solar Radiation")+
    theme_classic()+
    theme(
      axis.text.x = element_blank(), 
      axis.title.y = element_text(face="bold", size=14), 
      axis.title.x = element_text(face="bold", size=14), 
      plot.title = element_text(face="bold", size=14), 
      axis.text.y=element_text(size=12)
      );plot1
```

```{r}
plot2<-results%>%
  rename(solar_mean=mean_solar_rad_kwpm2_mean, solar_sd=mean_solar_rad_kwpm2_sd, rain_mean=cumulative_rainfall_mm_mean, rain_sd=cumulative_rainfall_mm_sd, temp_mean=mean_Temp_mean, temp_sd=mean_Temp_sd)%>%
  
  ggplot(aes(x=Time_Period, y=rain_mean))+
    geom_point(size=4, shape=19)+
    geom_errorbar(aes(ymin=rain_mean-rain_sd, ymax=rain_mean+rain_sd), width=0.01)+
    ylab(expression(bold(paste("Average Cumulative Rainfall (mm)"))))+
    ylim(-10,200)+
    xlab("Time Point")+
    ggtitle("D) Rainfall")+
    theme_classic()+
    theme(
      axis.text.x = element_blank(), 
      axis.title.y = element_text(face="bold", size=14), 
      axis.title.x = element_text(face="bold", size=14), 
      plot.title = element_text(face="bold", size=14), 
      axis.text.y=element_text(size=12)
      );plot2
```

```{r}
plot3<-results%>%
  rename(solar_mean=mean_solar_rad_kwpm2_mean, solar_sd=mean_solar_rad_kwpm2_sd, rain_mean=cumulative_rainfall_mm_mean, rain_sd=cumulative_rainfall_mm_sd, temp_mean=mean_Temp_mean, temp_sd=mean_Temp_sd)%>%
  
  ggplot(aes(x=Time_Period, y=temp_mean))+
    geom_point(size=4, shape=19)+
    geom_errorbar(aes(ymin=temp_mean-temp_sd, ymax=temp_mean+temp_sd), width=0.01)+
    ylab(expression(bold(paste("Water Temperature (C)"))))+
    ylim(25,32)+
    xlab("Time Point")+
    ggtitle("B) Temperature")+
    theme_classic()+
    theme(
      axis.text.x = element_blank(), 
      axis.title.y = element_text(face="bold", size=14), 
      axis.title.x = element_text(face="bold", size=14), 
      plot.title = element_text(face="bold", size=14), 
      axis.text.y=element_text(size=12)
      );plot3
```

Arrange plots
```{r}
library(cowplot)

plots<-plot_grid(plot3, plot1, plot2, nrow=1, ncol=3, align="vh", rel_widths=c(1,1,1))

ggsave(plots, filename="Figures/Environmental/Seasonal_averages.pdf", width=10, height=4)
```



