---
title: "Calcification Total Alkalinity"
author: "Nyssa Silbiger, Danielle Becker, edited by Ariana S Huffmyer"
date: "7/9/2021"
output:
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Time point 1 - January 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

```{r load_packages}
## install packages if you dont already have them in your library
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 
if ("purrr" %in% rownames(installed.packages()) == 'FALSE') install.packages('purrr') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("nlstools" %in% rownames(installed.packages()) == 'FALSE') install.packages('nlstools')
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr') 

#Read in required libraries

library(broom)
library(purrr)
library(lubridate)
library(tidyverse)
library(nlstools)
library(stringr)
```

# Import data and format

```{r import_data}
#bring in calcification data file with TA and chamber pH, temp, salinity measurements
raw.data <- list.files(path = 'data/1_calcification/raw_files', pattern = ".csv", full.names = TRUE) %>% 
  set_names(.) %>% 
  map_dfr(read.table, .id = "titration.run", header=TRUE, sep=",") %>%
  filter(!SampleID == "JUNK 1") %>% filter(!SampleID == "JUNK 2") %>% filter(!SampleID == "CRM") %>% 
  rename(Salinity.lab = Salinity) %>%
  separate(SampleID, c("colony_id", "Date", "Run.Number"), sep = "_", convert = FALSE)

# load field data for calcification runs 
deltaTA <- read.csv("data/1_calcification/1_DeltaTA_metadata.csv", header = TRUE, sep = ",",
                    colClasses=c(rep('character', 11), rep('numeric', 5), rep('character', 3))) %>%
           rename(Salinity.chamber = Salinity) %>% select(-Surface.Area) %>% filter(!colony_id == "BK-5_2") %>%
           rename(Vol.L = Vol.ml)

# changing BK-5_1 to BK-5 because this is the only blank 5 used 
deltaTA$colony_id <- gsub("BK-5_1", "BK-5", deltaTA$colony_id)

# load surface area values 
SA <- read.csv("output/1_surface_area.csv", header = TRUE, sep = ",",
                    colClasses=c('character', 'numeric', 'character')) %>% select(colony_id, surface.area.cm2)

data <- full_join(deltaTA, raw.data, by = c("colony_id", "Date", "Run.Number")) %>% 
  mutate(sample.type = case_when(
    startsWith(colony_id, "Initial") ~ "Initial",
    startsWith(colony_id, "BK") ~ "Blank",
    startsWith(colony_id, "P") ~ "Sample", #this covers POC, POR
    startsWith(colony_id, "A") ~ "Sample")) # this covers ACR

data <- full_join(data, SA, by = "colony_id")
```

*Volume of blanks --- variations.* 
Some blank very low - especially chamber 10. 

## Normalize TA values to salinity and separate blanks from samples.  

```{r}
# Normalize data to the salinity measurement taken at the same time as titration
data <- data %>% mutate(Ta.norm = TA * Salinity.lab/36)

# Calculate initial TA norm averages for each run
initial <- data %>% subset(sample.type == "Initial") %>% dplyr::group_by(Run.Number) %>%
  mutate(Ta.norm_initial_avg = mean(Ta.norm)) %>% select(Run.Number, Ta.norm_initial_avg) %>% distinct()

# Rejoining initial data back to the full df   
calc.data <- full_join(data, initial, by = "Run.Number") %>% subset(!sample.type == "Initial")

# Calculating delta TA for the blank samples
blanks <- calc.data %>% subset(sample.type == "Blank") %>%
  mutate(delta.TA.blank = Ta.norm_initial_avg - Ta.norm) %>% select(Run.Number, delta.TA.blank)

# Rejoining blank data back to the full df 
calc.data <- full_join(calc.data, blanks, by = "Run.Number") %>% subset(!sample.type == "Blank")   
```

#### Looking at variation in samples across blanks, initials, and titration data source (CSUN vs URI)

```{r}
data %>% filter(!is.na(Ta.norm)) %>%
  ggplot(aes(x=Run.Number, y=Ta.norm, group = sample.type, color = sample.type)) + theme_bw() + geom_point()
  
data %>% subset(!sample.type == "Sample") %>%
  ggplot(aes(x=Run.Number, y=Ta.norm, group = sample.type, color = sample.type)) + theme_bw() + geom_point()
```

## Calculate Net Ecosystem Calcification (NEC) from total alkalinity method

1. Convert time from character to time for both start and stop time.  
2. Calculate deltaTA (for each coral) as the sample TA.norm subtracted from the initial from that run, then subtracting the blank from that run.  
3. Calculate the time difference between start and stop times for each chamber(coral).  
4. Calculate calcification rate according the Net Ecosystem Calcification rates equation, normalizing to the surface area of the coral.  
5. Make any negative values equal to zero.  
6. Add the time point for the exported df (to then bring into a larger df with all time points in a different script). 

```{r}
NEC <- calc.data %>%
  mutate(start.time = strptime(as.character(TA.Start.Time), "%H:%M")) %>% #convert time from character to time
  mutate(stop.time = strptime(as.character(TA.Stop.Time), "%H:%M")) %>%
  mutate(deltaTA = (Ta.norm_initial_avg - Ta.norm) - delta.TA.blank) %>% #calculate difference in TA corrected to blanks
  mutate(timediff = as.numeric(stop.time - start.time)) %>% #calculate time difference
  mutate(umol.cm2.hr = (deltaTA/2)*(1.023)*((Vol.L*1000)/surface.area.cm2)*(1/timediff)*(1/1000)) %>% # convert L to mL and calculate net ecosystem calcification rates
  mutate(umol.cm2.hr=if_else(umol.cm2.hr<0, 0, umol.cm2.hr)) %>% #make any negative values equal to 0
  mutate(timepoint="timepoint1") 

write.csv(NEC, 'output/1_calcification_rates.csv') 

NEC %>% filter(is.na(umol.cm2.hr)) # missing calc rates 
```

#### Check sample variation one more time post normalizing to surface area
#### TA as a function of chamber temperature, salinity, pH, initial measurements temperature and pH

Do we need to do any correction for this? 

```{r}
NEC %>% filter(!is.na(umol.cm2.hr)) %>%
  ggplot(aes(x=Run.Number, y=umol.cm2.hr, color = Species)) + theme_bw() + geom_point() + facet_grid(~Species)

NEC %>% filter(!is.na(umol.cm2.hr)) %>%
  ggplot(aes(x=Temp.C, y=umol.cm2.hr, color = Species)) + theme_bw() + geom_point() 

NEC %>% filter(!is.na(umol.cm2.hr)) %>%
  ggplot(aes(x=pH.mV, y=umol.cm2.hr, color = Species)) + theme_bw() + geom_point() 

NEC %>% filter(!is.na(umol.cm2.hr)) %>%
  ggplot(aes(x=Salinity.chamber, y=umol.cm2.hr, color = Species)) + theme_bw() + geom_point() 

NEC %>% filter(!is.na(umol.cm2.hr)) %>%
  ggplot(aes(x=Chamber, y=umol.cm2.hr, color = Species)) + theme_bw() + geom_point() # this accounts for chamber and channel # 
```

## Visuals 

```{r}
NEC %>%
  filter(!Run.Number=="NA")%>%
  filter(!Species=="NA")%>%
  filter(!Site=="NA")%>%
  ggplot(aes(x = as.factor(Site), y = umol.cm2.hr, color = Species), position=position_dodge(0.5)) +
  geom_boxplot() +
  geom_point(size=3, position=position_jitterdodge(0.6))+
  facet_wrap(~Species) +
  labs(x = "Site", y = expression(paste(mu,"mol Ca", CO[3], " cm"^-2, " hr"^-1))) +                                       # Plot all points
  theme_classic() 
```

