---
title: "Photographic_Bleaching"
author: "EL Strand"
date: "4/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear workspace and load required packages. 
```{r}
rm(list=ls()) #clears workspace 

if ("vegan" %in% rownames(installed.packages()) == 'FALSE') install.packages('vegan') 
if ("ggpubr" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggpubr') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("lsmeans" %in% rownames(installed.packages()) == 'FALSE') install.packages('lsmeans') 
if ("multcompView" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcompView') 

library("vegan")
library("ggpubr")
library("gridExtra")
library("plyr") 
library("lsmeans")
library("multcompView")
```

Load the bleaching color score datasheet. 
```{r}
E5_bleaching <- read.csv("data/Moorea_ImageJ_E5.csv", header=T, sep=",", na.string="NA") #read in data file
E5_bleaching <-na.omit(E5_bleaching) # takes out the rows with NA value
E5_bleaching$Red.Standard <- as.numeric(E5_bleaching$Red.Standard) # changing Red Standard values to numeric instead of factor
```

Normalize to color standard
= Coral color value / Standard color value
```{r}
E5_bleaching$Red.Norm.Coral <- E5_bleaching$Red.Coral/E5_bleaching$Red.Standard 
E5_bleaching$Green.Norm.Coral <- E5_bleaching$Green.Coral/E5_bleaching$Green.Standard
E5_bleaching$Blue.Norm.Coral <- E5_bleaching$Blue.Coral/E5_bleaching$Blue.Standard 
```

Creating a matrix of the normalized coral color score values
```{r}
blch.scor <- as.matrix(cbind(E5_bleaching$Red.Norm.Coral,E5_bleaching$Green.Norm.Coral,E5_bleaching$Blue.Norm.Coral))
rownames(blch.scor) <- E5_bleaching$Coral.ID # name rows in dataframe
```

Calculating a distance matrix based on color scores and running a principal components analysis (PCA)
```{r}
dist <- vegdist(blch.scor, method="euclidean") #calculate distance matrix of color scores
PCA.color <- princomp(dist) # PCA
summary(PCA.color) # view variance explained by PCs
```

Building a dataframe from PCA Eigenvalues. 
```{r}
Blch <- as.data.frame(PCA.color$scores[,1]) #extract PC1
Blch$Coral.ID <- rownames(blch.scor)
Blch  <- cbind(Blch, E5_bleaching$Species, E5_bleaching$Timepoint, E5_bleaching$Site) #make a dataframe of PC1 and experiment factors
colnames(Blch) <- c("Bleaching.Score", "Coral.ID", "Species", "Timepoint", "Site")
Blch$Group <- paste(Blch$Species, Blch$Timepoint, Blch$Site)
Blch$SpGroup <- paste(Blch$Timepoint, Blch$Site)
```

Exporting datatable of bleaching score values
```{r}
write.table(Blch, file = "data/E5_Bleaching_Score.csv", append = FALSE, quote = TRUE, sep = ",",
            eol = "\n", na = "NA", dec = ".", row.names = FALSE,
            col.names = TRUE, qmethod = c("escape", "double"),
            fileEncoding = "")
```

Visualzing data by species and Group (Species name, Timepoint, Site)
```{r}
x<- Blch #assign values to test data frame to look for outliers
par(mar=c(10,4,2,2)) #bottom, left, top and right margins respectively
boxplot(Bleaching.Score ~ Group, data = x, lwd = 1, ylab = 'PC1Color', las=2, cex=0.8) #plot boxplot of PC1 color score by Group: Species, Timepoint, Site 
```

Reading in the bleaching score dataframe created above.
```{r}
Bleaching.Score <- read.csv("data/E5_Bleaching_Score.csv", header=T, sep=",", na.string="NA") #read in data file
```

Calculating means of the groups.  
```{r}
All.Means <- ddply(Blch, c('Species','Timepoint', 'Site'), summarize,
                    mean= mean(Bleaching.Score, na.rm=T), #mean 
                    N = sum(!is.na(Bleaching.Score)), # sample size
                    se = sd(Bleaching.Score, na.rm=T)/sqrt(N)) #SE
All.Means
All.Means$se[is.na(All.Means$se)] <- 0
All.Means$Group <- paste(All.Means$Species, All.Means$Timepoint, All.Means$Site)
All.Means$SpGroup <- paste(All.Means$Timepoint, All.Means$Site)

All.Means$Timepoint <- as.factor(All.Means$Timepoint)
All.Means$Site <- as.factor(All.Means$Site)

```

Plotting the bleaching score. 
```{r}
bleaching <- ggplot(All.Means, aes(x=Timepoint, y=mean, group=Site)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="black", width=0.1, position = position_dodge(width = 0.1)) +
  geom_point(aes(colour=Site), size = 2, position = position_dodge(width = 0.1)) +
  facet_grid(Species ~ ., scales = "free_y") +
  theme_classic() + 
  xlab("Timepoint") +
  ylab(expression(paste("Bleaching Intensity"))) + 
  ggtitle("Mo'orea E5 2020") + theme(plot.title = element_text(hjust = 0.5)) + 
  geom_line(aes(colour=Site, group=Site), position = position_dodge(width = 0.1), alpha=0.4) # colour, group both depend on cond2
bleaching
#ggsave(file="output/2_Bleaching.pdf", bleaching, width = 6, height = 10, units = c("in"))
ggsave(file="output/2_Bleaching.jpeg", bleaching, width = 6, height = 10, units = c("in"))
```

## Statistics 

Summary Statistics 
```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggpubr" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggpubr') 
if ("rstatix" %in% rownames(installed.packages()) == 'FALSE') install.packages('rstatix') 
if ("xml2" %in% rownames(installed.packages()) == 'FALSE') install.packages('xml2') 
if ("lme4" %in% rownames(installed.packages()) == 'FALSE') install.packages('lme4') 
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car') 
if ("lmerTest" %in% rownames(installed.packages()) == 'FALSE') install.packages('lmerTest') 

library(tidyverse)
library(ggpubr)
library(rstatix)
library(xml2)
library(lme4)
library(car)
library(lmerTest)

Blch %>% group_by(Species, Site, Timepoint) %>% get_summary_stats(Bleaching.Score, type = "mean_sd")
## this can also be seen with the All.Means table created in previous steps
```


Checking assumptions 
https://www.datanovia.com/en/lessons/repeated-measures-anova-in-r/

Model bleaching in Acropora using mixed effect model: `ACR_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=ACR_blch)`
```{r}
## Outlier check
Blch %>% group_by(Species, Site, Timepoint) %>% identify_outliers(Bleaching.Score)

## Normality check (checking for a normal distribution) - Shapiro Wilk's Test of Normality
## Normal distribution is indictated by p>0.05
Blch %>% group_by(Species) %>% shapiro_test(Bleaching.Score)

ACR_blch <- subset(Blch, Species=="Acropora")
POR_blch <- subset(Blch, Species=="Porites")
POC_blch <- subset(Blch, Species=="Pocillopora")

## QQ plot 
## Normality is assumed when all of the points fall approximately on the reference line (aka don't want tails on either end)
ggqqplot(ACR_blch, "Bleaching.Score", ggtheme = theme_classic()) + facet_grid(Site ~ Timepoint, labeller = "label_both")
ggqqplot(POR_blch, "Bleaching.Score", ggtheme = theme_classic()) + facet_grid(Site ~ Timepoint, labeller = "label_both")
ggqqplot(POC_blch, "Bleaching.Score", ggtheme = theme_classic()) + facet_grid(Site ~ Timepoint, labeller = "label_both")

## Repeated measures two way ANOVA 
#ACR.aov <- anova_test(data = ACR_blch, dv=Bleaching.Score, wid = Coral.ID, within = c(Site, Timepoint))
#get_anova_table(ACR.aov)

#Using lmer mixed models rather than ANOVA
ACR_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=ACR_blch)
qqPlot(residuals(ACR_model))
anova(ACR_model, type=2)

```

Model bleaching in Porites using mixed effect model: `POR_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=POR_blch)`
```{r}
POR_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=POR_blch)
qqPlot(residuals(POR_model))
anova(POR_model, type=2)
```

Model bleaching in Pocillopora using mixed effect model: `POC_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=POC_blch)`
```{r}
POC_model<-lmer(Bleaching.Score~Site*Timepoint + (1|Coral.ID), data=POC_blch)
qqPlot(residuals(POC_model))
anova(POC_model, type=2)
```

There are upper end outliers that should be addressed in further statistical analyses.  













